---
title: "File loading and preparation"
author: "Rasmus Riemer Jakobsen"
date: "`r Sys.time()`"
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
link-citations: yes
---

```{r setup, include=F, message=F}
#Set working directory to script directory
#setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

dir.create("results", showWarnings = FALSE, recursive = TRUE)

source("Supplementary_scripts/package_load_install.R")

source("Phyloseq_import_CSS_16S_laca.R")

theme_set(theme_classic())

knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.fullwidth = TRUE
                      #root.dir = setwd(dirname(rstudioapi::getActiveDocumentContext()$path))  #Set working directory to script directory 
                      )

#Coulour pallettes
col_fil <- pal_jco("default")(10)

col_scale <- scale_color_jco()
```


# Manual removal and sample id correction

## Remove UMI16S_134_PRJ+UMI16S_135_PRJ as they appear failed and  as no controls are sequenced (therefore not possible to confirm run)

```{r, message = F}
#PSB.all <- subset_samples(PSB.all, !seq.run %in% c("UMI16S_134_PRJ","UMI16S_135_PRJ","UMI16S_181_MH04")) #
```

## Remove milk 1+2 re-runs,  as sample IDS seem wrong

```{r, message = F, eval = FALSE}
PSB.M2.1 <- subset_samples(PSB.all, seq.run %in% "UMI16S_180_MH03" & Run %in% "1")
PSB.M2.2 <- subset_samples(PSB.all, seq.run %in% "UMI16S_180_MH03" & Run %in% "2")

PSB.M2.1
PSB.M2.2
```

# Check B2 milk sized

```{r, message = F}
subset_samples(PSB.all)
```

## Remove UMI16S_179_MH02 - B3 Feacal swabs as positive controls seem to have a large shift

```{r, message = F}
#PSB.all <- subset_samples(PSB.all, !seq.run %in% c("UMI16S_179_MH02"))
```

## Remove 2nd half of batch 2 milk samples (UMI16S_180_MH03 B) - due to bad controls

```{r, message = F, eval = T}
#PSB.all <- subset_samples(PSB.all, !seq.run %in% "UMI16S_180_MH03")

# m2.1 <- subset_samples(PSB.all, seq.run %in% "UMI16S_180_MH03" & Run %in% "1")
# m2.2 <- subset_samples(PSB.all, seq.run %in% "UMI16S_180_MH03" & Run %in% "2")
# 
# m2.1
# m2.2
# 
# m2.1.meta <- as.data.frame(sample_data(m2.1))
# m2.2.meta <- as.data.frame(sample_data(m2.2))
#PSB.all <- subset_samples(PSB.all, !seq.run %in% "UMI16S_180_MH03" | !Run %in% "2")
```

## Correction of viginal/rectal swabs

These swabs appear to have been mislabeled in the birth clinic. Some vaginal swabs might not be mislabeled but instead heavily contaminated with faecal matter during birth.

# Replicate samples

##Add add column with unique duplicate identifiers

```{r, message = F}
## Add unique identifier numbers for for duplicate RCID within each group
map_mat <- as.data.frame(as.matrix(sample_data(PSB.all)))

map_mat <- map_mat %>% 
  tidyr::unite(type.id, c("Sample.type","RCID"), remove = F) %>% # Add combined sample type+RCID column
  mutate(type.id.uniq = make.unique(type.id)) %>% # Add unique ids for all duplicates
  mutate(unique = isUnique(type.id)) %>% # Add tag for duplicate RCIDs within each sample type 
  mutate(RCID.uniq = sub(".*?_", "", type.id.uniq)) #Remove sample type for uniqe RCID column

#map_mat$RCID.uniq
#map_mat$type.id.uniq

# Remove combined type-id columns
map_mat <- map_mat %>% dplyr::select(-c("type.id","type.id.uniq"))

ids <- subset(map_mat, unique == FALSE)$RCID.uniq
str_sort(ids,
numeric=TRUE
)[1:40]

#Add classes back to data frame
map_mat <- type.convert(map_mat, as.is = TRUE)

PSB.all@sam_data <-  sample_data(map_mat)
```

# Remove ASVs by taxonomy

## Remove ASVs assigned as eukaryotes

```{r, message = F}
#Removed reads
sum(sample_sums(subset_taxa(PSB.all, Kingdom %in% c("Eukaryota","Archaea"))))/sum(sample_sums(PSB.all))

#Remove unassigned taxa
PSB.all <- subset_taxa(PSB.all, !Kingdom %in% c("Eukaryota","Archaea"))
```

## Remove ASVs not assigned at phylum level - not run as none were detected

```{r, message = F,eval=T}
tax <- data.frame(PSB.all@tax_table)

#Removed reads
sum(sample_sums(subset_taxa(PSB.all, Phylum %in% c("","Unassigned") | Class %in% c("","Unassigned"))))/sum(sample_sums(PSB.all))

#Remove unassigned taxa
PSB.all <- subset_taxa(PSB.all, !Phylum %in% c("","Unassigned") | !Class %in% c("","Unassigned"))
```

4 percent of reads were removed as they were unable to be classified as bacteria at phylum level. Manual blast search revealed that most appeared to be amplified human genomic DNA.

# Remove low-abundance ASVs

Remove taxa not seen found in at least 3 samples with a total count of minimum 1000 reads. This protects against an OTU with small mean & trivially large C.V.

```{r, message = F}
prev <- 3/nrow(PSB.all@sam_data)
PSB.all.fil = phyloseq_filter_prevalence(PSB.all, prev.trh = prev, abund.trh = 1000, abund.type = "total", threshold_condition = "AND")

spec <- specnumber(as.matrix(otu_table(PSB.all.fil))) %>% sort(decreasing=FALSE)
head(spec)

sum(sample_sums(PSB.all.fil))/sum(sample_sums(PSB.all))

PSB.all <- PSB.all.fil

```

93% of reads were kept after filtering

```{r}
PSB.all
```


# Set class and order of factors

```{r, message = F}
PSB.all@sam_data$BMI.group <- factor(PSB.all@sam_data$BMI.group, levels = c("Normal","Overweight","Obese"))
PSB.all@sam_data$Batch <- as.factor(PSB.all@sam_data$Batch)
PSB.all@sam_data$Run <- as.factor(PSB.all@sam_data$Run)
```

# Update CSS object

## Remove filtered samples and ASVs

```{r, message = F}
PSB.CSS.all <- prune_samples(sample_names(PSB.all), PSB.CSS.all)

PSB.CSS.all <- prune_taxa(taxa_names(PSB.all), PSB.CSS.all)
```

# Update metadata for CSS normalised object

```{r, message = F}
PSB.CSS.all@sam_data <- PSB.all@sam_data
```

# Collapse at genus level

```{r}
PSB.all.genus <- tax_glom(PSB.all,taxrank = "Genus")
PSB.all.species <- tax_glom(PSB.all,taxrank = "Species")
```

# Save objects in workspace

```{r, message = F}
save.image("0.2_Filtered_and_formated.RData")

rm(list = ls())
```

# SessionInfo

```{r}
sessionInfo()
```

