---
title: "Quality control and decontamination"
author: "Rasmus Riemer Jakobsen"
date: "`r Sys.time()`"
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
link-citations: yes
---

```{r setup, include=F, message=F}
#Set working directory to script directory
#setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

dir.create("results", showWarnings = FALSE, recursive = TRUE)

#Load required packages and install if necessary
source("Supplementary_scripts/package_load_install.R")

theme_set(theme_classic())

knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.fullwidth = TRUE,
                      warning = FALSE,
                      root.dir = setwd(dirname(rstudioapi::getActiveDocumentContext()$path))) #Set working directory to script directory 

#Coulour pallettes
col_fil <- pal_jco("default")(10)

col_scale <- scale_color_jco()
```

# Load filtered and formated phyloseq objects

```{r, message = F}
load("0.2_Filtered_and_formated.RData")
```

# Validate data import

## Check correct sample mapping file - if empty samples are included 

```{r, message = F}
nas <- subset_samples(PSB.all, Sample.type %in% NA)
sample_sums(nas) 
```

Rectal swab BRK74_UMI16S133_RRJ has only 33 reads so is likely just light contamination

## Check 

```{r, message = F}
## Check each group
faec <- subset_samples(PSB.all,Sample.type == "Faecal sample")
levels(as.factor(faec@sam_data$Run))
levels(as.factor(faec@sam_data$Set))

milk <- subset_samples(PSB.all,Sample.type == "Milk")
levels(as.factor(milk@sam_data$Run))
levels(as.factor(milk@sam_data$Set))
#subset_samples(PSB.all,Sample.type == "Milk" & Batch == 3 & Set == 3 )@sam_data$RC.id
#subset_samples(PSB.all,Sample.type == "Milk" & Batch == 3 & Set == 3 )@sam_data %>% rownames()

rec <- subset_samples(PSB.all,Sample.type == "Rectal swab")
levels(as.factor(rec@sam_data$Run))
levels(as.factor(rec@sam_data$Set))

vag <- subset_samples(PSB.all,Sample.type == "Vaginal swab")
levels(as.factor(vag@sam_data$Run))
levels(as.factor(vag@sam_data$Set))

vag <- subset_samples(PSB.all,Sample.type == "Vaginal swab")
levels(as.factor(vag@sam_data$Run))
levels(as.factor(vag@sam_data$Set))

vag <- subset_samples(PSB.all,Sample.type == "Vaginal swab")
levels(as.factor(vag@sam_data$Run))
levels(as.factor(vag@sam_data$Set))
```

## Check each library

### Check each run

```{r, message = F}
#Make combined column for subsetting
PSB.all@sam_data$type.batch.run <- paste(PSB.all@sam_data$Sample.type,PSB.all@sam_data$Batch,PSB.all@sam_data$Run)

ps.sub <- subset_samples(PSB.all, Sample.type != "Control") %>%
  subset_taxa(Phylum != "")

#Collapse to genus level
bac.phyl <- tax_glom(ps.sub, "Genus", NArm = FALSE)

ps0 <- transform_sample_counts(bac.phyl, function(x) x / sum(x))

ps0<- merge_samples(ps0, "type.batch.run")
ps0 <- transform_sample_counts(ps0, function(x) x / sum(x))

#Create melted dataframe
df <- psmelt(ps0)

#Select last non-empty taxonomic rank
df[df==""] <- NA
df$tax <- apply(df, 1, function(x) tail(na.omit(x), 1))

#Arrange samples by mean abundance
top <- df %>%
  group_by(tax) %>%
  dplyr::summarize(Mean = mean(Abundance)) %>%
  arrange(-Mean)
#Show top
#top

top10 <- top$tax[1:10]
df0 <- df %>%
  mutate(tax = fct_other(tax, c(as.matrix(top10))))

df0 <- df0[order(df0$Sample, decreasing = TRUE), ]

#Set order for samples
#df0$Treatment <- factor(df0$Treatment, levels = c("Control", "Amp", "FOP", "LM", "LM+Amp", "LM+FOP"))

#Set Hours as factor
#$Hours <- factor(df0$Hours, levels = c("3h", "6h", "24h"))

barplot.type.batch.run.col <- ggplot(df0, aes(Sample, Abundance, fill = fct_reorder(tax, Abundance, .fun = mean))) +
  geom_col(width = 0.8) +
  scale_fill_jco(name = "Taxonomy") +
  scale_fill_manual(values=rep(col_fil,5),name="Genus")+
  theme_classic() +
  theme(axis.line=element_line(size=0.5),
        #panel.border = element_blank(),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        axis.text.x=element_text(angle = -45, vjust = 0.8, hjust = 0),
        strip.text.x = element_text(angle = 0,  face = "bold"),
        legend.key.height = unit(0.5, 'cm'),
        legend.text = element_text(face = "italic"),
        text = element_text(family = 'serif', size = 12)
        #legend.position = "none")
  ) +
  ylab("Relative abundance")
  #facet_wrap(c("Sample.type","Batch","Run"),nrow=1,scales="free_x")


barplot.type.batch.run.col
```


```{r, message = F, eval = F}
subset_samples(PSB.all, seq.run == "UMI16S_132_RRJ")@sam_data #B1VS1SK1
subset_samples(PSB.all, seq.run == "UMI16S_134_PRJ")@sam_data
subset_samples(PSB.all, seq.run == "UMI16S048")@sam_data
subset_samples(PSB.all, seq.run == "UMI16S048")@sam_data
subset_samples(PSB.all, seq.run == "UMI16S048")@sam_data
subset_samples(PSB.all, seq.run == "UMI16S048")@sam_data
subset_samples(PSB.all, seq.run == "UMI16S048")@sam_data
subset_samples(PSB.all, seq.run == "UMI16S048")@sam_data
```

# Negative + postive controls

## Barplots

### All controls

```{r, message = F}
load("0.2_Filtered_and_formated.RData")

PSB.c <- subset_samples(PSB.all.genus, Sample.type == "Control")
PSB.c <- subset_taxa(PSB.c, Phylum != "")

ps0 <- transform_sample_counts(PSB.c, function(x) x / sum(x))

#ps0<- merge_samples(ps0, "Time")
#ps0 <- transform_sample_counts(ps0, function(x) x / sum(x))

#Create melted dataframe
df <- psmelt(ps0)

#Select last non-empty taxonomic rank
df[df==""] <- NA
df$tax <- apply(df, 1, function(x) tail(na.omit(x), 1))

#Arrange samples by mean abundance
top <- df %>%
  group_by(tax) %>%
  dplyr::summarize(Mean = mean(Abundance)) %>%
  arrange(-Mean)
#Show top
#top

top10 <- top$tax[1:10]
df0 <- df %>%
  mutate(tax = fct_other(tax, c(as.matrix(top10))))

df0 <- df0[order(df0$Sample, decreasing = TRUE), ]

#Set order for samples
#df0$Treatment <- factor(df0$Treatment, levels = c("Control", "Amp", "FOP", "LM", "LM+Amp", "LM+FOP"))

#Set Hours as factor
#$Hours <- factor(df0$Hours, levels = c("3h", "6h", "24h"))

barplot.controls <- ggplot(df0, aes(RC.id, Abundance, fill = fct_reorder(tax, Abundance, .fun = mean))) +
  geom_col(width = 0.8) +
  scale_fill_jco(name = "Taxonomy") +
  scale_fill_manual(values=rep(col_fil,5),name="Genus")+
  theme_classic() +
  theme(axis.line=element_line(size=0.5),
        #panel.border = element_blank(),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        axis.text.x=element_text(angle = -45, vjust = 0.8, hjust = 0),
        strip.text.x = element_text(angle = 0,  face = "bold"),
        legend.key.height = unit(0.5, 'cm'),
        legend.text = element_text(face = "italic"),
        text = element_text(family = 'serif', size = 12)
        #legend.position = "none")
  ) +
  ylab("Relative abundance")+
  facet_wrap("Batch",nrow=1,scales="free_x")


barplot.controls
```

### All controls by seq run

```{r, message = F}

load("0.2_Filtered_and_formated.RData")

bac.phyl <- subset_samples(PSB.all.genus, Sample.type == "Control")

ps0 <- transform_sample_counts(bac.phyl, function(x) x / sum(x))

#ps1<- merge_samples(ps0, "Time")

#ps1 <- transform_sample_counts(ps0, function(x) x / sum(x))


#Create melted dataframe
df <- psmelt(ps0)

#Select last non-empty taxonomic rank
df[df==""] <- NA
df$tax <- apply(df, 1, function(x) tail(na.omit(x), 1))

#Arrange samples by mean abundance
top <- df %>%
  group_by(tax) %>%
  dplyr::summarize(Mean = mean(Abundance)) %>%
  arrange(-Mean)
#Show top
#top  

top10 <- top$tax[1:10]
df0 <- df %>%
  mutate(tax = fct_other(tax, c(as.matrix(top10))))

df0 <- df0[order(df0$Sample, decreasing = TRUE), ]

#Set order for samples
#df0$Treatment <- factor(df0$Treatment, levels = c("Control", "Amp", "FOP", "LM", "LM+Amp", "LM+FOP"))

#Set Hours as factor
#$Hours <- factor(df0$Hours, levels = c("3h", "6h", "24h"))

barplot.controls.seq.run <- ggplot(df0, aes(RC.id, Abundance, fill = fct_reorder(tax, Abundance, .fun = mean))) +
  geom_col(width = 0.8) +
  scale_fill_jco(name = "Taxonomy") +
  scale_fill_manual(values=rep(col_fil,5),name="Genus")+
  theme_classic() +
  theme(axis.line=element_line(size=0.5),
        #panel.border = element_blank(),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        axis.text.x=element_text(angle = -45, vjust = 0.8, hjust = 0),
        strip.text.x = element_text(angle = 0,  face = "bold"),
        legend.key.height = unit(0.5, 'cm'),
        legend.text = element_text(face = "italic"),
        text = element_text(family = 'serif', size = 12)
        #legend.position = "none")
  ) +
  ylab("Relative abundance")+
  facet_wrap(c("seq.run","Sample.type"),nrow=2,scales="free_x")


barplot.controls.seq.run

ggsave("barplot.controls.seq.run.pdf",width=25,height=8)

```

### All samples by seq.run

```{r, message = F, fig.width=12, fig.height=30}
load("0.2_Filtered_and_formated.RData")

bac.phyl <- PSB.all.genus

ps0 <- transform_sample_counts(bac.phyl, function(x) x / sum(x))

#ps1<- merge_samples(ps0, "Time")

#ps1 <- transform_sample_counts(ps0, function(x) x / sum(x))


#Create melted dataframe
df <- psmelt(ps0)

#Select last non-empty taxonomic rank
df[df==""] <- NA
df$tax <- apply(df, 1, function(x) tail(na.omit(x), 1))

#Arrange samples by mean abundance
top <- df %>%
  group_by(tax) %>%
  dplyr::summarize(Mean = mean(Abundance)) %>%
  arrange(-Mean)
#Show top
#top

top10 <- top$tax[1:10]
df0 <- df %>%
  mutate(tax = fct_other(tax, c(as.matrix(top10))))

df0 <- df0[order(df0$Sample, decreasing = TRUE), ]

#Set order for samples
#df0$Treatment <- factor(df0$Treatment, levels = c("Control", "Amp", "FOP", "LM", "LM+Amp", "LM+FOP"))

#Set Hours as factor
#$Hours <- factor(df0$Hours, levels = c("3h", "6h", "24h"))

barplot.all.seq.run <- ggplot(df0, aes(RCID.uniq, Abundance, fill = fct_reorder(tax, Abundance, .fun = mean))) +
  geom_col(width = 0.8) +
  scale_fill_jco(name = "Taxonomy") +
  scale_fill_manual(values=rep(col_fil,5),name="Genus")+
  theme_classic() +
  theme(axis.line=element_line(size=0.5),
        #panel.border = element_blank(),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        axis.text.x=element_text(angle = -45, vjust = 0.8, hjust = 0),
        strip.text.x = element_text(angle = 0,  face = "bold"),
        legend.key.height = unit(0.5, 'cm'),
        legend.text = element_text(face = "italic"),
        text = element_text(family = 'serif', size = 12)
        #legend.position = "none")
  ) +
  ylab("Relative abundance")+
  facet_wrap(c("seq.run","Sample.type"),scales="free_x")


barplot.all.seq.run

ggsave("barplot.all.seq.run.pdf",barplot.all.seq.run,width=40,height=40,limitsize = FALSE)

```

### Subset only genera found in positive controls

```{r, message = F}
load("0.2_Filtered_and_formated.RData")

PSB.all.genus.R <- transform_sample_counts(PSB.all.genus, function(x) x / sum(x))

# PSB.mock.genera.R <- subset_taxa(PSB.all.genus.R, Genus %in% c("Akkermansia","Streptococcus","Bifidobacterium","Erysipelatoclostridium","Escherichia−Shigella", "Escherichia", "Faecalibacterium","Lacticaseibacillus","Bacteroides")
# )

PSB.mock.genera.R <- subset_taxa(PSB.all.genus.R, Genus %in% c("Akkermansia","Streptococcus","Bifidobacterium","Erysipelatoclostridium", "Lactobacillus", "Escherichia-Shigella", "Faecalibacterium","Lacticaseibacillus","Bacteroides")
)

tax <- as.data.frame(PSB.all.genus@tax_table)

ps0 <- PSB.mock.genera.R

#ps0 <- transform_sample_counts(bac.phyl, function(x) x / sum(x))

#ps1<- merge_samples(ps0, "Time")

#ps1 <- transform_sample_counts(ps0, function(x) x / sum(x))


#Create melted dataframe
df <- psmelt(ps0)

#Select last non-empty taxonomic rank
df[df==""] <- NA
df$tax <- apply(df, 1, function(x) tail(na.omit(x), 1))

#Arrange samples by mean abundance
top <- df %>%
  group_by(tax) %>%
  dplyr::summarize(Mean = mean(Abundance)) %>%
  arrange(-Mean)
#Show top
#top

top10 <- top$tax[1:10]
df0 <- df %>%
  mutate(tax = fct_other(tax, c(as.matrix(top10))))

df0 <- df0[order(df0$Sample, decreasing = TRUE), ]

#Set order for samples
#df0$Treatment <- factor(df0$Treatment, levels = c("Control", "Amp", "FOP", "LM", "LM+Amp", "LM+FOP"))

#Set Hours as factor
#$Hours <- factor(df0$Hours, levels = c("3h", "6h", "24h"))

barplot.all.seq.mock.genera <- ggplot(df0, aes(RCID.uniq, Abundance, fill = fct_reorder(tax, Abundance, .fun = mean))) +
  geom_col(width = 0.8) +
  scale_fill_jco(name = "Taxonomy") +
  scale_fill_manual(values=rep(col_fil,5),name="Genus")+
  theme_classic() +
  theme(
        #strip.background = element_rect(colour = "white", fill = "white"),
        axis.text.x=element_text(angle = -45, vjust = 0.8, hjust = 0),
        strip.text.x = element_text(angle = 0,  face = "bold"),
        legend.key.height = unit(0.5, 'cm'),
        legend.text = element_text(face = "italic"),
        text = element_text(family = 'serif', size = 12)
        #legend.position = "none")
  ) +
  ylab("Relative abundance")+
  facet_wrap(c("seq.run","Run"),
             scales="free_x")


barplot.all.seq.mock.genera


ggsave("barplot.all.seq.mock.genera.pdf",barplot.all.seq.mock.genera,width=50,height=50,limitsize = FALSE)
```

## Sample counts for postive, negative, PCR controls

```{r, message = F}
load("0.2_Filtered_and_formated.RData")

PSB.c <- subset_samples(PSB.all, Sample.type == "Control")

meta.c <- as.data.frame(as.matrix(sample_data(PSB.c)))
meta.c$read.count <- sample_sums(PSB.c)

ggplot(meta.c,aes(read.count, fill=Control)) +
  geom_histogram() +
  facet_wrap(c("seq.run","Run")) +
  scale_x_log10()
```

## PCoA plots for controls

```{r, eval = T, include = T, echo = F, massage = F, warning = F}
#Do PCoA ordination based on Bray-curtis distance
GP.ord <- ordinate(PSB.CSS.all, "PCoA", "bray")

#####################Run + Batch

beta.id = plot_ordination(PSB.CSS.all, GP.ord, color="Control") + 
  stat_ellipse(geom = "polygon", level = 0.95, fill = NA, size = 1) +
  #scale_fill_manual(values= col_fil) +
  #scale_color_manual(values = col_fil) +
  theme_classic() + 
  theme(text = element_text(size = 12, family = 'serif'),
        axis.line=element_line(size=1),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        #axis.text.x=element_text(angle = 45, hjust = 1),
        strip.text.x = element_text(size = 12, angle = 0),
        legend.text = element_text(size = 12)
        #legend.position = "none"
  ) +
  facet_wrap(~seq.run,scales="free")

beta.id
```


# Batch/run effects

## All

### Sample types

```{r, eval = T, include = T, echo = F, massage = F}
#Do PCoA ordination based on Bray-curtis distance
GP.ord <- ordinate(PSB.CSS.all, "PCoA", "bray")

#####################Run + Batch

#PSB.CSS@sam_data$Run <- as.factor(PSB.CSS@sam_data$Run)

PSB.CSS.all@sam_data$Batch <- as.factor(PSB.CSS.all@sam_data$Batch)

beta.id = plot_ordination(PSB.CSS.all, GP.ord, color="Batch") + 
  stat_ellipse(geom = "polygon", level = 0.95, fill = NA, size = 1) +
  #scale_fill_manual(values= col_fil) +
  #scale_color_manual(values = col_fil) +
  theme_classic() + 
  theme(text = element_text(size = 12, family = 'serif'),
        axis.line=element_line(size=1),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        #axis.text.x=element_text(angle = 45, hjust = 1),
        strip.text.x = element_text(size = 12, angle = 0),
        legend.text = element_text(size = 12)
        #legend.position = "none"
  ) +
  facet_wrap(~Sample.type,scales="free")

beta.id
```


### Runs by batch + sample type

```{r, eval = T, include = T, echo = F, massage = F}
#Do PCoA ordination based on Bray-curtis distance
PSB.CSS.all@sam_data$Set <- as.factor(PSB.CSS.all@sam_data$Set)
GP.ord <- ordinate(PSB.CSS.all, "PCoA", "bray")

#####################Run + Batch

#PSB.CSS@sam_data$Run <- as.factor(PSB.CSS@sam_data$Run)

PSB.CSS.all@sam_data$Run <- as.factor(PSB.CSS.all@sam_data$Run)

beta.type.batch = plot_ordination(PSB.CSS.all, GP.ord, color="Run") + 
  stat_ellipse(geom = "polygon", level = 0.95, fill = NA, size = 1) +
  #scale_fill_manual(values= col_fil) +
  #scale_color_manual(values = col_fil) +
  theme_classic() + 
  theme(text = element_text(size = 12, family = 'serif'),
        axis.line=element_line(size=1),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        #axis.text.x=element_text(angle = 45, hjust = 1),
        strip.text.x = element_text(size = 12, angle = 0),
        legend.text = element_text(size = 12)
        #legend.position = "none"
  ) +
  facet_wrap(~Sample.type+Batch,scales="free")

beta.type.batch
```

### Sets by batch + sample type

```{r, eval = T, include = T, echo = F, massage = F}
#Do PCoA ordination based on Bray-curtis distance
PSB.CSS.all@sam_data$Set <- as.factor(PSB.CSS.all@sam_data$Set)
GP.ord <- ordinate(PSB.CSS.all, "PCoA", "bray")

#####################Run + Batch

#PSB.CSS@sam_data$Run <- as.factor(PSB.CSS@sam_data$Run)

PSB.CSS.all@sam_data$Batch <- as.factor(PSB.CSS.all@sam_data$Batch)

beta.type.batch = plot_ordination(PSB.CSS.all, GP.ord, color="Set") + 
  stat_ellipse(geom = "polygon", level = 0.95, fill = NA, size = 1) +
  #scale_fill_manual(values= col_fil) +
  #scale_color_manual(values = col_fil) +
  theme_classic() + 
  theme(text = element_text(size = 12, family = 'serif'),
        axis.line=element_line(size=1),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        #axis.text.x=element_text(angle = 45, hjust = 1),
        strip.text.x = element_text(size = 12, angle = 0),
        legend.text = element_text(size = 12)
        #legend.position = "none"
  ) +
  facet_wrap(~Sample.type+Batch,scales="free")

beta.type.batch
```

### Faecal samples

#### Bray curtis

```{r, eval = T, include = T, echo = F, massage = F}
load("0.2_Filtered_and_formated.RData")

ps <- subset_samples(PSB.CSS.all, Sample.type == "Faecal sample")

#Do PCoA ordination based on Bray-curtis distance
GP.ord <- ordinate(ps, "PCoA", "bray")

#####################Run + Batch

#PSB.CSS@sam_data$Run <- as.factor(PSB.CSS@sam_data$Run)

ps@sam_data$ps <- as.factor(ps@sam_data$Batch)

beta.batch.faeces = plot_ordination(ps, GP.ord, color="Batch") + 
  stat_ellipse(geom = "polygon", level = 0.95, fill = NA, size = 1) +
  #scale_fill_manual(values= col_fil) +
  #scale_color_manual(values = col_fil) +
  theme_classic() + 
  theme(text = element_text(size = 12, family = 'serif'),
        axis.line=element_line(size=1),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        #axis.text.x=element_text(angle = 45, hjust = 1),
        strip.text.x = element_text(size = 12, angle = 0),
        legend.text = element_text(size = 12)
        #legend.position = "none"
  ) +
  ggtitle("Bray Curtis - non collapsed")

beta.batch.faeces
```

#### Bray curtis - species collapsed

```{r, eval = T, include = T, echo = F, massage = F}
load("0.2_Filtered_and_formated.RData")

ps <- subset_samples(
  PSB.CSS.all, Sample.type == "Faecal sample") %>% tax_glom(taxrank = "Species")


#Do PCoA ordination based on Bray-curtis distance
GP.ord <- ordinate(ps, "PCoA", "bray")

#####################Run + Batch

#PSB.CSS@sam_data$Run <- as.factor(PSB.CSS@sam_data$Run)

ps@sam_data$ps <- as.factor(ps@sam_data$Batch)

beta.batch.faeces.species = plot_ordination(ps, GP.ord, color="Batch") + 
  stat_ellipse(geom = "polygon", level = 0.95, fill = NA, size = 1) +
  #scale_fill_manual(values= col_fil) +
  #scale_color_manual(values = col_fil) +
  theme_classic() + 
  theme(text = element_text(size = 12, family = 'serif'),
        axis.line=element_line(size=1),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        #axis.text.x=element_text(angle = 45, hjust = 1),
        strip.text.x = element_text(size = 12, angle = 0),
        legend.text = element_text(size = 12)
        #legend.position = "none"
  )+
  ggtitle("Bray Curtis - species collapsed")

beta.batch.faeces.species
```

#### Bray curtis - genus collapsed

```{r, eval = T, include = T, echo = F, massage = F}
load("0.2_Filtered_and_formated.RData")

ps <- subset_samples(
  PSB.CSS.all, Sample.type == "Faecal sample") %>%
  tax_glom(taxrank = "Genus")


#Do PCoA ordination based on Bray-curtis distance
GP.ord <- ordinate(ps, "PCoA", "bray")

#####################Run + Batch

#PSB.CSS@sam_data$Run <- as.factor(PSB.CSS@sam_data$Run)

ps@sam_data$ps <- as.factor(ps@sam_data$Batch)

beta.batch.faeces.genus = plot_ordination(ps, GP.ord, color="Batch") + 
  stat_ellipse(geom = "polygon", level = 0.95, fill = NA, size = 1) +
  #scale_fill_manual(values= col_fil) +
  #scale_color_manual(values = col_fil) +
  theme_classic() + 
  theme(text = element_text(size = 12, family = 'serif'),
        axis.line=element_line(size=1),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        #axis.text.x=element_text(angle = 45, hjust = 1),
        strip.text.x = element_text(size = 12, angle = 0),
        legend.text = element_text(size = 12)
        #legend.position = "none"
  )+
  ggtitle("Bray Curtis - Genus")

beta.batch.faeces.genus
```


```{r, eval = T, include = T, echo = F, massage = F, warning = F}
ggarrange(beta.batch.faeces,beta.batch.faeces.species,beta.batch.faeces.genus,beta.batch.faeces.wunifrac)
```

### Feacal samples - decontaminated

#### Bray curtis

```{r, eval = T, include = T, echo = F, massage = F}
load("0.3_deduplicated_and_decontaminated.RData")

ps <- subset_samples(PSB.CSS, Sample.type == "Faecal sample")

#Do PCoA ordination based on Bray-curtis distance
GP.ord <- ordinate(ps, "PCoA", "bray")

#####################Run + Batch

#PSB.CSS@sam_data$Run <- as.factor(PSB.CSS@sam_data$Run)

ps@sam_data$ps <- as.factor(ps@sam_data$Batch)

beta.batch.faeces = plot_ordination(ps, GP.ord, color="Batch") + 
  stat_ellipse(geom = "polygon", level = 0.95, fill = NA, size = 1) +
  #scale_fill_manual(values= col_fil) +
  #scale_color_manual(values = col_fil) +
  theme_classic() + 
  theme(text = element_text(size = 12, family = 'serif'),
        axis.line=element_line(size=1),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        #axis.text.x=element_text(angle = 45, hjust = 1),
        strip.text.x = element_text(size = 12, angle = 0),
        legend.text = element_text(size = 12)
        #legend.position = "none"
  ) +
  ggtitle("Bray Curtis - non collapsed")

beta.batch.faeces
```

#### Bray curtis - species collapsed

```{r, eval = T, include = T, echo = F, massage = F}
load("0.3_deduplicated_and_decontaminated.RData")

ps <- subset_samples(
  PSB.CSS, Sample.type == "Faecal sample") %>% tax_glom(taxrank = "Species")


#Do PCoA ordination based on Bray-curtis distance
GP.ord <- ordinate(ps, "PCoA", "bray")

#####################Run + Batch

#PSB.CSS@sam_data$Run <- as.factor(PSB.CSS@sam_data$Run)

ps@sam_data$ps <- as.factor(ps@sam_data$Batch)

beta.batch.faeces.species = plot_ordination(ps, GP.ord, color="Batch") + 
  stat_ellipse(geom = "polygon", level = 0.95, fill = NA, size = 1) +
  #scale_fill_manual(values= col_fil) +
  #scale_color_manual(values = col_fil) +
  theme_classic() + 
  theme(text = element_text(size = 12, family = 'serif'),
        axis.line=element_line(size=1),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        #axis.text.x=element_text(angle = 45, hjust = 1),
        strip.text.x = element_text(size = 12, angle = 0),
        legend.text = element_text(size = 12)
        #legend.position = "none"
  )+
  ggtitle("Bray Curtis - species collapsed")

beta.batch.faeces.species
```

#### Bray curtis - genus collapsed

```{r, eval = T, include = T, echo = F, massage = F}
load("0.3_deduplicated_and_decontaminated.RData")

ps <- subset_samples(
  PSB.CSS, Sample.type == "Faecal sample") %>%
  tax_glom(taxrank = "Genus")


#Do PCoA ordination based on Bray-curtis distance
GP.ord <- ordinate(ps, "PCoA", "bray")

#####################Run + Batch

#PSB.CSS@sam_data$Run <- as.factor(PSB.CSS@sam_data$Run)

ps@sam_data$ps <- as.factor(ps@sam_data$Batch)

beta.batch.faeces.genus = plot_ordination(ps, GP.ord, color="Batch") + 
  stat_ellipse(geom = "polygon", level = 0.95, fill = NA, size = 1) +
  #scale_fill_manual(values= col_fil) +
  #scale_color_manual(values = col_fil) +
  theme_classic() + 
  theme(text = element_text(size = 12, family = 'serif'),
        axis.line=element_line(size=1),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        #axis.text.x=element_text(angle = 45, hjust = 1),
        strip.text.x = element_text(size = 12, angle = 0),
        legend.text = element_text(size = 12)
        #legend.position = "none"
  )+
  ggtitle("Bray Curtis - Genus")

beta.batch.faeces.genus
```


```{r, eval = T, include = T, echo = F, massage = F, warning = F}
ggarrange(beta.batch.faeces,beta.batch.faeces.species,beta.batch.faeces.genus,)
```

### Feacal batch 1

#### Barplots per run - genus

```{r, message = F}
#Make combined column for subsetting
PSB.all@sam_data$type.batch.run <- paste(PSB.all@sam_data$Sample.type,PSB.all@sam_data$Batch,PSB.all@sam_data$Run)

ps.sub <- subset_samples(PSB.all, Sample.type != "Control") %>%
  subset_samples(Sample.type == "Faecal sample" & Batch == 1)

#Collapse to genus level
bac.phyl <- tax_glom(ps.sub, "Genus", NArm = FALSE)

ps0 <- transform_sample_counts(bac.phyl, function(x) x / sum(x))

ps0<- merge_samples(ps0, "type.batch.run")
ps0 <- transform_sample_counts(ps0, function(x) x / sum(x))

#Create melted dataframe
df <- psmelt(ps0)

#Select last non-empty taxonomic rank
df[df==""] <- NA
df$tax <- apply(df, 1, function(x) tail(na.omit(x), 1))

#Arrange samples by mean abundance
top <- df %>%
  group_by(tax) %>%
  dplyr::summarize(Mean = mean(Abundance)) %>%
  arrange(-Mean)
#Show top
#top

top10 <- top$tax[1:10]
df0 <- df %>%
  mutate(tax = fct_other(tax, c(as.matrix(top10))))

df0 <- df0[order(df0$Sample, decreasing = TRUE), ]

#Set order for samples
#df0$Treatment <- factor(df0$Treatment, levels = c("Control", "Amp", "FOP", "LM", "LM+Amp", "LM+FOP"))

#Set Hours as factor
#$Hours <- factor(df0$Hours, levels = c("3h", "6h", "24h"))

barplot.type.batch.run.col <- ggplot(df0, aes(Sample, Abundance, fill = fct_reorder(tax, Abundance, .fun = mean))) +
  geom_col(width = 0.8) +
  scale_fill_jco(name = "Taxonomy") +
  scale_fill_manual(values=rep(col_fil,5),name="Genus")+
  theme_classic() +
  theme(axis.line=element_line(size=0.5),
        #panel.border = element_blank(),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        axis.text.x=element_text(angle = -45, vjust = 0.8, hjust = 0),
        strip.text.x = element_text(angle = 0,  face = "bold"),
        legend.key.height = unit(0.5, 'cm'),
        legend.text = element_text(face = "italic"),
        text = element_text(family = 'serif', size = 12)
        #legend.position = "none")
  ) +
  ylab("Relative abundance")
  #facet_wrap(c("Sample.type","Batch","Run"),nrow=1,scales="free_x")


barplot.type.batch.run.col
```

#### Barplots per run - species

```{r, message = F}
#Make combined column for subsetting
PSB.all@sam_data$type.batch.run <- paste(PSB.all@sam_data$Sample.type,PSB.all@sam_data$Batch,PSB.all@sam_data$Run)

ps.sub <- subset_samples(PSB.all, Sample.type != "Control") %>%
  subset_samples(Sample.type == "Faecal sample" & Batch == 1)

#Collapse to genus level
bac.phyl <- tax_glom(ps.sub, "Species", NArm = FALSE)

ps0 <- transform_sample_counts(bac.phyl, function(x) x / sum(x))

ps0<- merge_samples(ps0, "type.batch.run")
ps0 <- transform_sample_counts(ps0, function(x) x / sum(x))

#Create melted dataframe
df <- psmelt(ps0)

#Select last non-empty taxonomic rank
df[df==""] <- NA
df$tax <- apply(df, 1, function(x) tail(na.omit(x), 1))

#Arrange samples by mean abundance
top <- df %>%
  group_by(tax) %>%
  dplyr::summarize(Mean = mean(Abundance)) %>%
  arrange(-Mean)
#Show top
#top

top10 <- top$tax[1:10]
df0 <- df %>%
  mutate(tax = fct_other(tax, c(as.matrix(top10))))

df0 <- df0[order(df0$Sample, decreasing = TRUE), ]

#Set order for samples
#df0$Treatment <- factor(df0$Treatment, levels = c("Control", "Amp", "FOP", "LM", "LM+Amp", "LM+FOP"))

#Set Hours as factor
#$Hours <- factor(df0$Hours, levels = c("3h", "6h", "24h"))

barplot.type.batch.run.col <- ggplot(df0, aes(Sample, Abundance, fill = fct_reorder(tax, Abundance, .fun = mean))) +
  geom_col(width = 0.8) +
  scale_fill_jco(name = "Taxonomy") +
  scale_fill_manual(values=rep(col_fil,5),name="Genus")+
  theme_classic() +
  theme(axis.line=element_line(size=0.5),
        #panel.border = element_blank(),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        axis.text.x=element_text(angle = -45, vjust = 0.8, hjust = 0),
        strip.text.x = element_text(angle = 0,  face = "bold"),
        legend.key.height = unit(0.5, 'cm'),
        legend.text = element_text(face = "italic"),
        text = element_text(family = 'serif', size = 12)
        #legend.position = "none")
  ) +
  ylab("Relative abundance")
  #facet_wrap(c("Sample.type","Batch","Run"),nrow=1,scales="free_x")


barplot.type.batch.run.col
```

#### PCoA by run

```{r, eval = T, include = T, echo = F, massage = F}
load("0.3_deduplicated_and_decontaminated.RData")

ps <- subset_samples(PSB.CSS, Sample.type == "Faecal sample" & Batch == 1)

#Do PCoA ordination based on Bray-curtis distance
GP.ord <- ordinate(ps, "PCoA", "bray")

#####################Run + Batch

#PSB.CSS@sam_data$Run <- as.factor(PSB.CSS@sam_data$Run)

#ps@sam_data$set <- as.factor(ps@sam_data$Batch)

beta.batch.faeces = plot_ordination(ps, GP.ord, color="Run") + 
  stat_ellipse(geom = "polygon", level = 0.95, fill = NA, size = 1) +
  #scale_fill_manual(values= col_fil) +
  #scale_color_manual(values = col_fil) +
  theme_classic() + 
  theme(text = element_text(size = 12, family = 'serif'),
        axis.line=element_line(size=1),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        #axis.text.x=element_text(angle = 45, hjust = 1),
        strip.text.x = element_text(size = 12, angle = 0),
        legend.text = element_text(size = 12)
        #legend.position = "none"
  ) +
  ggtitle("Bray Curtis - non collapsed") 

beta.batch.faeces
```

#### PCoA by run - Genus level

```{r, eval = T, include = T, echo = F, massage = F}
load("0.3_deduplicated_and_decontaminated.RData")

ps <- subset_samples(PSB.CSS, Sample.type == "Faecal sample" & Batch == 1) %>%
  tax_glom(taxrank = "Genus")

#Do PCoA ordination based on Bray-curtis distance
GP.ord <- ordinate(ps, "PCoA", "bray")

#####################Run + Batch

#PSB.CSS@sam_data$Run <- as.factor(PSB.CSS@sam_data$Run)

#ps@sam_data$set <- as.factor(ps@sam_data$Batch)

beta.batch.faeces = plot_ordination(ps, GP.ord, color="Run") + 
  stat_ellipse(geom = "polygon", level = 0.95, fill = NA, size = 1) +
  #scale_fill_manual(values= col_fil) +
  #scale_color_manual(values = col_fil) +
  theme_classic() + 
  theme(text = element_text(size = 12, family = 'serif'),
        axis.line=element_line(size=1),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        #axis.text.x=element_text(angle = 45, hjust = 1),
        strip.text.x = element_text(size = 12, angle = 0),
        legend.text = element_text(size = 12)
        #legend.position = "none"
  ) +
  ggtitle("Bray Curtis - non collapsed") 

beta.batch.faeces
```


## Controls vs samples by sequencing run

### Milk

```{r, eval = T, include = T, echo = F, massage = F, warning = F}
ps <- subset_samples(PSB.all, seq.run %in% c("UMI16S062","UMI16S064","UMI16S065","UMI16S066","UMI16S_131_RRJ","UMI16S_178_MH01","UMI16S_180_MH03"))

#Do PCoA ordination based on Bray-curtis distance
GP.ord <- ordinate(ps, "PCoA", "bray")

#####################Run + Batch

ps@sam_data$Batch <- as.factor(ps@sam_data$Batch)
ps@sam_data$Run <- as.factor(ps@sam_data$Run)

beta.controls.seq.run.milk = plot_ordination(ps, GP.ord, color="Control") + 
  #stat_ellipse(geom = "polygon", level = 0.95, fill = NA, size = 1) +
  #scale_fill_manual(values= col_fil) +
  #scale_color_manual(values = col_fil) +
  theme_classic() + 
  theme(text = element_text(size = 12, family = 'serif'),
        axis.line=element_line(size=1),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        #axis.text.x=element_text(angle = 45, hjust = 1),
        strip.text.x = element_text(size = 12, angle = 0),
        legend.text = element_text(size = 12)
        #legend.position = "none"
  ) +
  facet_wrap("seq.run",scales="free")

beta.controls.seq.run.milk
```


## Milk

### Batch

```{r, eval = T, include = T, echo = F, massage = F}
ps <- subset_samples(PSB.all, Sample.type == "Milk")

#Do PCoA ordination based on Bray-curtis distance
GP.ord <- ordinate(ps, "PCoA", "bray")

#####################Run + Batch

ps@sam_data$Batch <- as.factor(ps@sam_data$Batch)
ps@sam_data$Run <- as.factor(ps@sam_data$Run)

beta.id = plot_ordination(ps, GP.ord, color="Batch") + 
  stat_ellipse(geom = "polygon", level = 0.95, fill = NA, size = 1) +
  #scale_fill_manual(values= col_fil) +
  #scale_color_manual(values = col_fil) +
  theme_classic() + 
  theme(text = element_text(size = 12, family = 'serif'),
        axis.line=element_line(size=1),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        #axis.text.x=element_text(angle = 45, hjust = 1),
        strip.text.x = element_text(size = 12, angle = 0),
        legend.text = element_text(size = 12)
        #legend.position = "none"
  ) 

beta.id
```

### Milk batch by Sequencing run

```{r, eval = T, include = T, echo = F, massage = F}
ps <- subset_samples(PSB.all, Sample.type == "Milk")

#####################Run + Batch
ps@sam_data$Batch <- as.factor(ps@sam_data$Batch)
ps@sam_data$Run <- as.factor(ps@sam_data$Run)
ps@sam_data$batch.run <- paste(ps@sam_data$Batch,ps@sam_data$Run)

#Do PCoA ordination based on Bray-curtis distance
GP.ord <- ordinate(ps, "PCoA", "bray")

beta.milk.batch.run = plot_ordination(ps, GP.ord, color="batch.run") + 
  stat_ellipse(geom = "polygon", level = 0.95, fill = NA, size = 1) +
  #scale_fill_manual(values= col_fil) +
  #scale_color_manual(values = col_fil) +
  theme_classic() + 
  theme(text = element_text(size = 12, family = 'serif'),
        axis.line=element_line(size=1),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        #axis.text.x=element_text(angle = 45, hjust = 1),
        strip.text.x = element_text(size = 12, angle = 0),
        legend.text = element_text(size = 12)
        #legend.position = "none"
  )

beta.milk.batch.run
```



### Milk runs per patch

```{r, eval = T, include = T, echo = F, massage = F}
ps <- subset_samples(PSB.all, Sample.type == "Milk" #& Batch == 1
                     )

#####################Run + Batch
ps@sam_data$Batch <- as.factor(ps@sam_data$Batch)
ps@sam_data$Run <- as.factor(ps@sam_data$Run)
ps@sam_data$batch.run <- paste(ps@sam_data$Batch,ps@sam_data$Run)
ps@sam_data$Set <- paste(ps@sam_data$Batch,ps@sam_data$Set)

#Do PCoA ordination based on Bray-curtis distance
GP.ord <- ordinate(ps, "PCoA", "bray")

beta.milk.batch1.sets = plot_ordination(ps, GP.ord, color="Set") + 
  stat_ellipse(geom = "polygon", level = 0.95, fill = NA, size = 1) +
  # scale_fill_manual(values= col_fil) +
  # scale_color_manual(values = col_fil) +
  theme_classic() + 
  theme(text = element_text(size = 12, family = 'serif'),
        axis.line=element_line(size=1),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        #axis.text.x=element_text(angle = 45, hjust = 1),
        strip.text.x = element_text(size = 12, angle = 0),
        legend.text = element_text(size = 12)
        #legend.position = "none"
  ) + facet_wrap(vars(Batch,Run))

beta.milk.batch1.sets
```

## Batch 1 vaginal vs Rectal swabs

### Batch

```{r, eval = T, include = T, echo = F, massage = F}
load("0.2_Filtered_and_formated.RData")

ps <- subset_samples(PSB.CSS.all, seq.run %in% c("UMI16S_131_RRJ","UMI16S133_RRJ") & Sample.type %in% c("Rectal swab","Vaginal swab","Milk"))

#Do PCoA ordination based on Bray-curtis distance
GP.ord <- ordinate(ps, "PCoA", "bray")

#####################Run + Batch

ps@sam_data$Run <- as.factor(ps@sam_data$Run)

ps@sam_data$Batch <- as.factor(ps@sam_data$Batch)

beta.id = plot_ordination(PSB.CSS.all, GP.ord, color="Sample.type") + 
  stat_ellipse(geom = "polygon", level = 0.95, fill = NA, size = 1) +
  #scale_fill_manual(values= col_fil) +
  #scale_color_manual(values = col_fil) +
  theme_classic() + 
  theme(text = element_text(size = 12, family = 'serif'),
        axis.line=element_line(size=1),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        #axis.text.x=element_text(angle = 45, hjust = 1),
        strip.text.x = element_text(size = 12, angle = 0),
        legend.text = element_text(size = 12)
        #legend.position = "none"
  ) 

beta.id
```

### Barplots

```{r, warning = F}
load("0.2_Filtered_and_formated.RData")


ps <- subset_samples(PSB.all.genus, seq.run %in% c("UMI16S_131_RRJ","UMI16S_132_RRJ","UMI16S133_RRJ") & Sample.type %in% c("Rectal swab","Vaginal swab"))

#Make column with concatenated unique rcid and sample type for easy comparision

ps@sam_data$id.unique.type <- paste(ps@sam_data$RCID.uniq,ps@sam_data$Sample.type)

ps0 <- transform_sample_counts(ps, function(x) x / sum(x))

#ps1<- merge_samples(ps0, "Time")

#ps1 <- transform_sample_counts(ps0, function(x) x / sum(x))

#Create melted dataframe
df <- psmelt(ps0)

#Select last non-empty taxonomic rank
df[df==""] <- NA
df$tax <- apply(df, 1, function(x) tail(na.omit(x), 1))

#Arrange samples by mean abundance
top <- df %>%
  group_by(tax) %>%
  dplyr::summarize(Mean = mean(Abundance)) %>%
  arrange(-Mean)
#Show top
#top

top10 <- top$tax[1:10]
df0 <- df %>%
  mutate(tax = fct_other(tax, c(as.matrix(top10))))

df0 <- df0[order(df0$Sample, decreasing = TRUE), ]

#Set order for samples
#df0$Treatment <- factor(df0$Treatment, levels = c("Control", "Amp", "FOP", "LM", "LM+Amp", "LM+FOP"))

#Set Hours as factor
#$Hours <- factor(df0$Hours, levels = c("3h", "6h", "24h"))

barplot.UMI16S_131 <- ggplot(df0, aes(id.unique.type, Abundance, fill = fct_reorder(tax, Abundance, .fun = mean))) +
  geom_col(width = 0.8) +
  scale_fill_jco(name = "Taxonomy") +
  scale_fill_manual(values=rep(col_fil,5),name="Genus")+
  theme_classic() +
  theme(axis.line=element_line(size=0.5),
        #panel.border = element_blank(),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        axis.text.x=element_text(angle = -45, vjust = 0.8, hjust = 0),
        strip.text.x = element_text(angle = 0,  face = "bold"),
        legend.key.height = unit(0.5, 'cm'),
        legend.text = element_text(face = "italic"),
        text = element_text(family = 'serif', size = 12)
        #legend.position = "none")
  ) +
  ylab("Relative abundance")+
  facet_wrap("RCID",scales="free_x")


barplot.UMI16S_131

ggsave("barplot.UMI16S_131.pdf",width=30,height=20,limitsize = FALSE)
```


# Decontamination

## All samples

### Inspect Library Sizes

```{r, message = F}
load("0.2_Filtered_and_formated.RData")

ps <- PSB.all

df <- as.data.frame(sample_data(ps)) # Put sample_data into a ggplot-friendly data.frame
df$LibrarySize <- sample_sums(ps)
df <- df[order(df$LibrarySize),]
df$Index <- seq(nrow(df))
ggplot(data=df, aes(x=Index, y=LibrarySize, color=Control)) + geom_point() + scale_y_log10()
```

```{r, message = F}
ggplot(data=df, aes(x=Index, y=LibrarySize, color=Control)) +
  geom_boxplot() +
  scale_y_log10()
```

### Identify Contaminants - Frequency

```{r, message = F}
load("0.2_Filtered_and_formated.RData")

ps <- PSB.all

ps@sam_data$DNA.conc <- as.numeric(ps@sam_data$DNA.conc) %>%
  pmax(0) + 0.000001 # add pseudocount

# Select only samples with measured DNA concentrations
ps <- subset_samples(ps, !is.na(DNA.conc))


contamdf.freq <- isContaminant(ps, method="frequency", conc="DNA.conc", threshold = 0.4)
head(contamdf.freq)

table(contamdf.freq$contaminant)

head(which(contamdf.freq$contaminant))
cont.rank <- head(which(contamdf.freq$contaminant))

contamdf.freq.all <- contamdf.freq


plot_frequency(ps, taxa_names(ps)[cont.rank], conc="DNA.conc") + 
  xlab("DNA Concentration (Qbit dsDNA high sensitivity kit)")
```

The could be some trend of frequncy being depandant on DNA concentration - but none of these show clear enough correlations or are found in enough low read samples to be considered contaminants.

###  Identify Contaminants - Prevalence

```{r, message = F}
load("0.2_Filtered_and_formated.RData")

ps <- PSB.all

sample_data(ps)$is.neg <- sample_data(ps)$Control == "Neg"
contamdf.prev <- isContaminant(ps, method="prevalence", neg="is.neg", threshold = 0.1)
table(contamdf.prev$contaminant)

head(which(contamdf.prev$contaminant))

# Make phyloseq object of presence-absence in negative controls and true samples
ps.pa <- transform_sample_counts(ps, function(abund) 1*(abund>0))
ps.pa.neg <- prune_samples(sample_data(ps.pa)$Control == "Neg", ps.pa)
ps.pa.pos <- prune_samples(sample_data(ps.pa)$Sample.type != "Control", ps.pa)
# Make data.frame of prevalence in positive and negative samples
df.pa <- data.frame(pa.pos=taxa_sums(ps.pa.pos), pa.neg=taxa_sums(ps.pa.neg),
                      contaminant=contamdf.prev$contaminant)
ggplot(data=df.pa, aes(x=pa.neg, y=pa.pos, color=contaminant)) + geom_point() +
  xlab("Prevalence (Negative Controls)") + ylab("Prevalence (True Samples)")
```

## Per Seq run

### Inspect Library Sizes

```{r, message = F}
load("0.2_Filtered_and_formated.RData")

ps <- PSB.all

df <- as.data.frame(sample_data(ps)) # Put sample_data into a ggplot-friendly data.frame
df$LibrarySize <- sample_sums(ps)
df <- df[order(df$LibrarySize),]
df$Index <- seq(nrow(df))
ggplot(data=df, aes(x=Index, y=LibrarySize, color=Control)) +
  geom_point() +
  scale_y_log10() +
  facet_wrap("seq.run")
```

```{r, message = F}
ggplot(data=df, aes(x=Index, y=LibrarySize, color=Control)) +
  geom_boxplot() +
  scale_y_log10() +
  facet_wrap("seq.run")
```

### Identify Contaminants - Frequency

```{r, message = F}
load("0.2_Filtered_and_formated.RData")

ps <- PSB.all

ps@sam_data$DNA.conc <- as.numeric(ps@sam_data$DNA.conc) %>%
  pmax(0) + 0.000001 # add pseudocount

# Select only samples with measured DNA concentrations
ps <- subset_samples(ps, !is.na(DNA.conc))

contamdf.freq <- isContaminant(ps, method="frequency", conc="DNA.conc",batch="seq.run" , threshold = 0.1)
head(contamdf.freq)

table(contamdf.freq$contaminant)

head(which(contamdf.freq$contaminant))

sum(sample_sums(prune_taxa(!contamdf.prev$contaminant, ps)))/sum(sample_sums(PSB.all))

cont.rank <- head(which(contamdf.freq$contaminant))
contamdf.freq.seq.run <- contamdf.freq

plot_frequency(ps, taxa_names(ps)[cont.rank], conc="DNA.conc") + 
  xlab("DNA Concentration (Qbit dsDNA high sensitivity kit)") 
```

###  Identify Contaminants - Prevalence

```{r, message = F}
load("0.2_Filtered_and_formated.RData")

ps <- PSB.all

sample_data(ps)$is.neg <- sample_data(ps)$Control == "Neg"
contamdf.prev <- isContaminant(ps, method="prevalence", neg="is.neg", threshold = 0.1, batch = "seq.run")
table(contamdf.prev$contaminant)

head(which(contamdf.prev$contaminant))

# Make phyloseq object of presence-absence in negative controls and true samples
ps.pa <- transform_sample_counts(ps, function(abund) 1*(abund>0))
ps.pa.neg <- prune_samples(sample_data(ps.pa)$Control == "Neg", ps.pa)
ps.pa.pos <- prune_samples(sample_data(ps.pa)$Sample.type != "Control", ps.pa)
# Make data.frame of prevalence in positive and negative samples
df.pa <- data.frame(pa.pos=taxa_sums(ps.pa.pos), pa.neg=taxa_sums(ps.pa.neg),
                      contaminant=contamdf.prev$contaminant)

ggplot(data=df.pa, aes(x=pa.neg, y=pa.pos, color=contaminant)) + geom_point() +
  xlab("Prevalence (Negative Controls)") + ylab("Prevalence (True Samples)")
```


## All samples by PCR plates

### Identify Contaminants - Frequency

```{r, message = F}
load("0.2_Filtered_and_formated.RData")

ps <- PSB.all

ps@sam_data$batch.run.type <- paste(ps@sam_data$batch,ps@sam_data$Run,ps@sam_data$Sample.type)

ps@sam_data$DNA.conc <- as.numeric(ps@sam_data$DNA.conc) %>%
  pmax(0) + 0.000001 # add pseudocount

# Select only samples with measured DNA concentrations
ps <- subset_samples(ps, !is.na(DNA.conc))

contamdf.freq <- isContaminant(ps, method="frequency", conc="DNA.conc",batch="batch.run.type" , threshold = 0.1)

table(contamdf.freq$contaminant)

head(which(contamdf.freq$contaminant))
cont.rank <- head(which(contamdf.freq$contaminant))

sum(sample_sums(prune_taxa(!contamdf.freq$contaminant, ps)))/sum(sample_sums(PSB.all))

contamdf.freq.pcr <- contamdf.freq

plot_frequency(ps, taxa_names(ps)[cont.rank], conc="DNA.conc") + 
  xlab("DNA Concentration (Qbit dsDNA high sensitivity kit)")
```


###  Identify Contaminants - Prevalence

```{r, message = F}
sample_data(ps)$is.neg <- sample_data(ps)$Control == "Neg"
contamdf.prev <- isContaminant(ps, method="prevalence", neg="is.neg", threshold = 0.1, batch = "batch.run.type")
table(contamdf.prev$contaminant)

head(which(contamdf.prev$contaminant))

sum(sample_sums(prune_taxa(!contamdf.prev$contaminant, ps)))/sum(sample_sums(PSB.all))

# Make phyloseq object of presence-absence in negative controls and true samples
ps.pa <- transform_sample_counts(ps, function(abund) 1*(abund>0))
ps.pa.neg <- prune_samples(sample_data(ps.pa)$Control == "Neg", ps.pa)
ps.pa.pos <- prune_samples(sample_data(ps.pa)$Sample.type != "Control", ps.pa)
# Make data.frame of prevalence in positive and negative samples
df.pa <- data.frame(pa.pos=taxa_sums(ps.pa.pos), pa.neg=taxa_sums(ps.pa.neg),
                      contaminant=contamdf.prev$contaminant)
ggplot(data=df.pa, aes(x=pa.neg, y=pa.pos, color=contaminant)) + geom_point() +
  xlab("Prevalence (Negative Controls)") + ylab("Prevalence (True Samples)")
```

## Show extent of decontaminated taxa

### Decontam across all

#### Remaining

```{r, message = F}
load("0.2_Filtered_and_formated.RData")

#Make combined column for subsetting
PSB.all@sam_data$type.batch.run <- paste(PSB.all@sam_data$Sample.type,PSB.all@sam_data$Batch,PSB.all@sam_data$Run)

ps0 <- transform_sample_counts(PSB.all, function(x) x / sum(x))

ps0<- merge_samples(ps0, "type.batch.run")

ps0 <- transform_sample_counts(ps0, function(x) x / sum(x))

ps0 <- prune_taxa(!contamdf.freq.all$contaminant, ps0)

ps0 <- tax_glom(ps0, taxrank = "Genus")
  
#Create melted dataframe
df <- psmelt(ps0)

#Select last non-empty taxonomic rank
df[df==""] <- NA
df$tax <- apply(df, 1, function(x) tail(na.omit(x), 1))

#Arrange samples by mean abundance
top <- df %>%
  group_by(tax) %>%
  dplyr::summarize(Mean = mean(Abundance)) %>%
  arrange(-Mean)
#Show top
#top

top10 <- top$tax[1:10]
df0 <- df %>%
  mutate(tax = fct_other(tax, c(as.matrix(top10))))

df0 <- df0[order(df0$Sample, decreasing = TRUE), ]

#Set order for samples
#df0$Treatment <- factor(df0$Treatment, levels = c("Control", "Amp", "FOP", "LM", "LM+Amp", "LM+FOP"))

#Set Hours as factor
#$Hours <- factor(df0$Hours, levels = c("3h", "6h", "24h"))

barplot.decontam.all <- ggplot(df0, aes(Sample, Abundance, fill = fct_reorder(tax, Abundance, .fun = mean))) +
  geom_col(width = 0.8) +
  scale_fill_jco(name = "Taxonomy") +
  scale_fill_manual(values=rep(col_fil,5),name="Genus")+
  theme_classic() +
  theme(axis.line=element_line(size=0.5),
        #panel.border = element_blank(),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        axis.text.x=element_text(angle = -45, vjust = 0.8, hjust = 0),
        strip.text.x = element_text(angle = 0,  face = "bold"),
        legend.key.height = unit(0.5, 'cm'),
        legend.text = element_text(face = "italic"),
        text = element_text(family = 'serif', size = 12)
        #legend.position = "none")
  ) +
  ylab("Relative abundance")
  #facet_wrap(c("Sample.type","Batch","Run"),nrow=1,scales="free_x")


barplot.decontam.all


```

#### Removed

```{r, message = F}
load("0.2_Filtered_and_formated.RData")

#Make combined column for subsetting
PSB.all@sam_data$type.batch.run <- paste(PSB.all@sam_data$Sample.type,PSB.all@sam_data$Batch,PSB.all@sam_data$Run)

ps0 <- transform_sample_counts(PSB.all, function(x) x / sum(x))

ps0<- merge_samples(ps0, "type.batch.run")

ps0 <- transform_sample_counts(ps0, function(x) x / sum(x))

ps0 <- prune_taxa(contamdf.freq.all$contaminant, ps0)

ps0 <- tax_glom(ps0, taxrank = "Genus")
  
#Create melted dataframe
df <- psmelt(ps0)

#Select last non-empty taxonomic rank
df[df==""] <- NA
df$tax <- apply(df, 1, function(x) tail(na.omit(x), 1))

#Arrange samples by mean abundance
top <- df %>%
  group_by(tax) %>%
  dplyr::summarize(Mean = mean(Abundance)) %>%
  arrange(-Mean)
#Show top
#top

top10 <- top$tax[1:10]
df0 <- df %>%
  mutate(tax = fct_other(tax, c(as.matrix(top10))))

df0 <- df0[order(df0$Sample, decreasing = TRUE), ]

#Set order for samples
#df0$Treatment <- factor(df0$Treatment, levels = c("Control", "Amp", "FOP", "LM", "LM+Amp", "LM+FOP"))

#Set Hours as factor
#$Hours <- factor(df0$Hours, levels = c("3h", "6h", "24h"))

barplot.decontam.all.removed <- ggplot(df0, aes(Sample, Abundance, fill = fct_reorder(tax, Abundance, .fun = mean))) +
  geom_col(width = 0.8) +
  scale_fill_jco(name = "Taxonomy") +
  scale_fill_manual(values=rep(col_fil,5),name="Genus")+
  theme_classic() +
  theme(axis.line=element_line(size=0.5),
        #panel.border = element_blank(),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        axis.text.x=element_text(angle = -45, vjust = 0.8, hjust = 0),
        strip.text.x = element_text(angle = 0,  face = "bold"),
        legend.key.height = unit(0.5, 'cm'),
        legend.text = element_text(face = "italic"),
        text = element_text(family = 'serif', size = 12)
        #legend.position = "none")
  ) +
  ylab("Relative abundance")+
  scale_y_continuous(limits = c(0,1))
  #facet_wrap(c("Sample.type","Batch","Run"),nrow=1,scales="free_x")


barplot.decontam.all.removed
```

### Decontam across seq runs

#### Remaining

```{r, message = F}
load("0.2_Filtered_and_formated.RData")

#Make combined column for subsetting
PSB.all@sam_data$type.batch.run <- paste(PSB.all@sam_data$Sample.type,PSB.all@sam_data$Batch,PSB.all@sam_data$Run)

ps0 <- transform_sample_counts(PSB.all, function(x) x / sum(x))

ps0<- merge_samples(ps0, "type.batch.run")

ps0 <- transform_sample_counts(ps0, function(x) x / sum(x))

ps0 <- prune_taxa(!contamdf.freq.seq.run$contaminant, ps0)

ps0 <- tax_glom(ps0, taxrank = "Genus")
  
#Create melted dataframe
df <- psmelt(ps0)

#Select last non-empty taxonomic rank
df[df==""] <- NA
df$tax <- apply(df, 1, function(x) tail(na.omit(x), 1))

#Arrange samples by mean abundance
top <- df %>%
  group_by(tax) %>%
  dplyr::summarize(Mean = mean(Abundance)) %>%
  arrange(-Mean)
#Show top
#top

top10 <- top$tax[1:10]
df0 <- df %>%
  mutate(tax = fct_other(tax, c(as.matrix(top10))))

df0 <- df0[order(df0$Sample, decreasing = TRUE), ]

#Set order for samples
#df0$Treatment <- factor(df0$Treatment, levels = c("Control", "Amp", "FOP", "LM", "LM+Amp", "LM+FOP"))

#Set Hours as factor
#$Hours <- factor(df0$Hours, levels = c("3h", "6h", "24h"))

barplot.decontam.seq.run <- ggplot(df0, aes(Sample, Abundance, fill = fct_reorder(tax, Abundance, .fun = mean))) +
  geom_col(width = 0.8) +
  scale_fill_jco(name = "Taxonomy") +
  scale_fill_manual(values=rep(col_fil,5),name="Genus")+
  theme_classic() +
  theme(axis.line=element_line(size=0.5),
        #panel.border = element_blank(),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        axis.text.x=element_text(angle = -45, vjust = 0.8, hjust = 0),
        strip.text.x = element_text(angle = 0,  face = "bold"),
        legend.key.height = unit(0.5, 'cm'),
        legend.text = element_text(face = "italic"),
        text = element_text(family = 'serif', size = 12)
        #legend.position = "none")
  ) +
  ylab("Relative abundance")
  #facet_wrap(c("Sample.type","Batch","Run"),nrow=1,scales="free_x")


barplot.decontam.seq.run
```

#### Removed

```{r, message = F}
load("0.2_Filtered_and_formated.RData")

#Make combined column for subsetting
PSB.all@sam_data$type.batch.run <- paste(PSB.all@sam_data$Sample.type,PSB.all@sam_data$Batch,PSB.all@sam_data$Run)

ps0 <- transform_sample_counts(PSB.all, function(x) x / sum(x))

ps0<- merge_samples(ps0, "type.batch.run")

ps0 <- transform_sample_counts(ps0, function(x) x / sum(x))

ps0 <- prune_taxa(contamdf.freq.seq.run$contaminant, ps0)

ps0 <- tax_glom(ps0, taxrank = "Genus")
  
#Create melted dataframe
df <- psmelt(ps0)

#Select last non-empty taxonomic rank
df[df==""] <- NA
df$tax <- apply(df, 1, function(x) tail(na.omit(x), 1))

#Arrange samples by mean abundance
top <- df %>%
  group_by(tax) %>%
  dplyr::summarize(Mean = mean(Abundance)) %>%
  arrange(-Mean)
#Show top
#top

top10 <- top$tax[1:10]
df0 <- df %>%
  mutate(tax = fct_other(tax, c(as.matrix(top10))))

df0 <- df0[order(df0$Sample, decreasing = TRUE), ]

#Set order for samples
#df0$Treatment <- factor(df0$Treatment, levels = c("Control", "Amp", "FOP", "LM", "LM+Amp", "LM+FOP"))

#Set Hours as factor
#$Hours <- factor(df0$Hours, levels = c("3h", "6h", "24h"))

barplot.decontam.seq.run.removed <- ggplot(df0, aes(Sample, Abundance, fill = fct_reorder(tax, Abundance, .fun = mean))) +
  geom_col(width = 0.8) +
  scale_fill_jco(name = "Taxonomy") +
  scale_fill_manual(values=rep(col_fil,5),name="Genus")+
  theme_classic() +
  theme(axis.line=element_line(size=0.5),
        #panel.border = element_blank(),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        axis.text.x=element_text(angle = -45, vjust = 0.8, hjust = 0),
        strip.text.x = element_text(angle = 0,  face = "bold"),
        legend.key.height = unit(0.5, 'cm'),
        legend.text = element_text(face = "italic"),
        text = element_text(family = 'serif', size = 12)
        #legend.position = "none")
  ) +
  ylab("Relative abundance")+
  scale_y_continuous(limits = c(0,1))
  #facet_wrap(c("Sample.type","Batch","Run"),nrow=1,scales="free_x")


barplot.decontam.seq.run.removed
```

### Decontam across pcr plates

```{r, message = F}
load("0.2_Filtered_and_formated.RData")

#Make combined column for subsetting
PSB.all@sam_data$type.batch.run <- paste(PSB.all@sam_data$Sample.type,PSB.all@sam_data$Batch,PSB.all@sam_data$Run)

ps0 <- transform_sample_counts(PSB.all, function(x) x / sum(x))

ps0<- merge_samples(ps0, "type.batch.run")

ps0 <- transform_sample_counts(ps0, function(x) x / sum(x))

ps0 <- prune_taxa(!contamdf.freq.pcr$contaminant, ps0)

ps0 <- tax_glom(ps0, taxrank = "Genus")
  
#Create melted dataframe
df <- psmelt(ps0)

#Select last non-empty taxonomic rank
df[df==""] <- NA
df$tax <- apply(df, 1, function(x) tail(na.omit(x), 1))

#Arrange samples by mean abundance
top <- df %>%
  group_by(tax) %>%
  dplyr::summarize(Mean = mean(Abundance)) %>%
  arrange(-Mean)
#Show top
#top

top10 <- top$tax[1:10]
df0 <- df %>%
  mutate(tax = fct_other(tax, c(as.matrix(top10))))

df0 <- df0[order(df0$Sample, decreasing = TRUE), ]

#Set order for samples
#df0$Treatment <- factor(df0$Treatment, levels = c("Control", "Amp", "FOP", "LM", "LM+Amp", "LM+FOP"))

#Set Hours as factor
#$Hours <- factor(df0$Hours, levels = c("3h", "6h", "24h"))

barplot.decontam.pcr <- ggplot(df0, aes(Sample, Abundance, fill = fct_reorder(tax, Abundance, .fun = mean))) +
  geom_col(width = 0.8) +
  scale_fill_jco(name = "Taxonomy") +
  scale_fill_manual(values=rep(col_fil,5),name="Genus")+
  theme_classic() +
  theme(axis.line=element_line(size=0.5),
        #panel.border = element_blank(),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        axis.text.x=element_text(angle = -45, vjust = 0.8, hjust = 0),
        strip.text.x = element_text(angle = 0,  face = "bold"),
        legend.key.height = unit(0.5, 'cm'),
        legend.text = element_text(face = "italic"),
        text = element_text(family = 'serif', size = 12)
        #legend.position = "none")
  ) +
  ylab("Relative abundance")
  #facet_wrap(c("Sample.type","Batch","Run"),nrow=1,scales="free_x")


barplot.decontam.pcr
```

## Per PCR individual plate - UMI16S056

### Identify Contaminants - Frequency

```{r, message = F}
load("0.2_Filtered_and_formated.RData")

ps <- subset_samples(PSB.all, seq.run == "UMI16S056")

ps@sam_data$batch.run.type <- paste(ps@sam_data$batch,ps@sam_data$Run,ps@sam_data$Sample.type)

ps@sam_data$DNA.conc <- as.numeric(ps@sam_data$DNA.conc) %>%
  pmax(0) + 0.000001 # add pseudocount

# Select only samples with measured DNA concentrations
ps <- subset_samples(ps, !is.na(DNA.conc))

##

contamdf.freq <- isContaminant(ps, method="frequency", conc="DNA.conc" , threshold = 0.5)
head(contamdf.freq)

table(contamdf.freq$contaminant)

head(which(contamdf.freq$contaminant))
cont.rank <- which(contamdf.freq$contaminant)

plot_frequency(ps, taxa_names(ps)[cont.rank], conc="DNA.conc") + 
  xlab("DNA Concentration (Qbit dsDNA high sensitivity kit)")
```


###  Identify Contaminants - Prevalence

Not run since no negative controls had sufficient reads to be included in this run

```{r, message = F,eval = F}
sample_data(ps)$is.neg <- sample_data(ps)$Control == "Neg"
contamdf.prev <- isContaminant(ps, method="prevalence", neg="is.neg", threshold = 0.1)
table(contamdf.prev$contaminant)

head(which(contamdf.prev$contaminant))

# Make phyloseq object of presence-absence in negative controls and true samples
ps.pa <- transform_sample_counts(ps, function(abund) 1*(abund>0))
ps.pa.neg <- prune_samples(sample_data(ps.pa)$Control == "Neg", ps.pa)
ps.pa.pos <- prune_samples(sample_data(ps.pa)$Sample.type != "Control", ps.pa)
# Make data.frame of prevalence in positive and negative samples
df.pa <- data.frame(pa.pos=taxa_sums(ps.pa.pos), pa.neg=taxa_sums(ps.pa.neg),
                      contaminant=contamdf.prev$contaminant)
ggplot(data=df.pa, aes(x=pa.neg, y=pa.pos, color=contaminant)) + geom_point() +
  xlab("Prevalence (Negative Controls)") + ylab("Prevalence (True Samples)")
```

## Per PCR individual plate - UMI16S048

### Identify Contaminants - Frequency

```{r, message = F}
load("0.2_Filtered_and_formated.RData")

ps <- subset_samples(PSB.all, seq.run == "UMI16S048")

ps@sam_data$batch.run.type <- paste(ps@sam_data$batch,ps@sam_data$Run,ps@sam_data$Sample.type)

ps@sam_data$DNA.conc <- as.numeric(ps@sam_data$DNA.conc) %>%
  pmax(0) + 0.000001 # add pseudocount

# Select only samples with measured DNA concentrations
ps <- subset_samples(ps, !is.na(DNA.conc))

##

contamdf.freq <- isContaminant(ps, method="frequency", conc="DNA.conc" , threshold = 0.5)
head(contamdf.freq)

table(contamdf.freq$contaminant)

head(which(contamdf.freq$contaminant))
cont.rank <- head(which(contamdf.freq$contaminant))

plot_frequency(ps, taxa_names(ps)[cont.rank], conc="DNA.conc") + 
  xlab("DNA Concentration (Qbit dsDNA high sensitivity kit)")
```



###  Identify Contaminants - Prevalence

Not run since no negative controls had sufficient reads to be included in this run

```{r, message = F,eval = F}
sample_data(ps)$is.neg <- sample_data(ps)$Control == "Neg"
contamdf.prev <- isContaminant(ps, method="prevalence", neg="is.neg", threshold = 0.1)
table(contamdf.prev$contaminant)

head(which(contamdf.prev$contaminant))

# Make phyloseq object of presence-absence in negative controls and true samples
ps.pa <- transform_sample_counts(ps, function(abund) 1*(abund>0))
ps.pa.neg <- prune_samples(sample_data(ps.pa)$Control == "Neg", ps.pa)
ps.pa.pos <- prune_samples(sample_data(ps.pa)$Sample.type != "Control", ps.pa)
# Make data.frame of prevalence in positive and negative samples
df.pa <- data.frame(pa.pos=taxa_sums(ps.pa.pos), pa.neg=taxa_sums(ps.pa.neg),
                      contaminant=contamdf.prev$contaminant)
ggplot(data=df.pa, aes(x=pa.neg, y=pa.pos, color=contaminant)) + geom_point() +
  xlab("Prevalence (Negative Controls)") + ylab("Prevalence (True Samples)")
```

# Replicate samples

## Barplots of replicates

### All

```{r, message = F}
load("0.2_Filtered_and_formated.RData")

ps <- subset_samples(PSB.all.genus, unique == FALSE)

ps0 <- transform_sample_counts(ps, function(x) x / sum(x))

#ps1<- merge_samples(ps0, "Time")

#ps1 <- transform_sample_counts(ps0, function(x) x / sum(x))

#Create melted dataframe
df <- psmelt(ps0)

#Select last non-empty taxonomic rank
df[df==""] <- NA
df$tax <- apply(df, 1, function(x) tail(na.omit(x), 1))

#Arrange samples by mean abundance
top <- df %>%
  group_by(tax) %>%
  dplyr::summarize(Mean = mean(Abundance)) %>%
  arrange(-Mean)
#Show top
#top

top10 <- top$tax[1:10]
df0 <- df %>%
  mutate(tax = fct_other(tax, c(as.matrix(top10))))

df0 <- df0[order(df0$Sample, decreasing = TRUE), ]

#Set order for samples
#df0$Treatment <- factor(df0$Treatment, levels = c("Control", "Amp", "FOP", "LM", "LM+Amp", "LM+FOP"))

#Set Hours as factor
#$Hours <- factor(df0$Hours, levels = c("3h", "6h", "24h"))

barplot.all.replicates <- ggplot(df0, aes(RCID.uniq, Abundance, fill = fct_reorder(tax, Abundance, .fun = mean))) +
  geom_col(width = 0.8) +
  scale_fill_jco(name = "Taxonomy") +
  scale_fill_manual(values=rep(col_fil,5),name="Genus")+
  theme_classic() +
  theme(axis.line=element_line(size=0.5),
        #panel.border = element_blank(),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        axis.text.x=element_text(angle = -45, vjust = 0.8, hjust = 0),
        strip.text.x = element_text(angle = 0,  face = "bold"),
        legend.key.height = unit(0.5, 'cm'),
        legend.text = element_text(face = "italic"),
        text = element_text(family = 'serif', size = 12)
        #legend.position = "none")
  ) +
  ylab("Relative abundance")+
  facet_wrap("Sample.type",scales="free_x")


barplot.all.replicates

ggsave("barplot.replicates.pdf",width=50,height=4,limitsize = FALSE)
```

### Milk

```{r, message = F}
load("0.2_Filtered_and_formated.RData")

ps <- subset_samples(PSB.all.genus, unique == FALSE & Sample.type == "Milk")

ps0 <- transform_sample_counts(ps, function(x) x / sum(x))

#ps1<- merge_samples(ps0, "Time")

#ps1 <- transform_sample_counts(ps0, function(x) x / sum(x))

#Create melted dataframe
df <- psmelt(ps0)

#Select last non-empty taxonomic rank
df[df==""] <- NA
df$tax <- apply(df, 1, function(x) tail(na.omit(x), 1))

#Arrange samples by mean abundance
top <- df %>%
  group_by(tax) %>%
  dplyr::summarize(Mean = mean(Abundance)) %>%
  arrange(-Mean)
#Show top
#top

top10 <- top$tax[1:10]
df0 <- df %>%
  mutate(tax = fct_other(tax, c(as.matrix(top10))))

df0 <- df0[order(df0$Sample, decreasing = TRUE), ]

#Set order for samples
#df0$Treatment <- factor(df0$Treatment, levels = c("Control", "Amp", "FOP", "LM", "LM+Amp", "LM+FOP"))

#Set Hours as factor
#$Hours <- factor(df0$Hours, levels = c("3h", "6h", "24h"))

barplot.milk.replicates <- ggplot(df0, aes(RCID.uniq, Abundance, fill = fct_reorder(tax, Abundance, .fun = mean))) +
  geom_col(width = 0.8) +
  scale_fill_jco(name = "Taxonomy") +
  scale_fill_manual(values=rep(col_fil,5),name="Genus")+
  theme_classic() +
  theme(axis.line=element_line(size=0.5),
        #panel.border = element_blank(),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        axis.text.x=element_text(angle = -45, vjust = 0.8, hjust = 0),
        strip.text.x = element_text(angle = 0,  face = "bold"),
        legend.key.height = unit(0.5, 'cm'),
        legend.text = element_text(face = "italic"),
        text = element_text(family = 'serif', size = 12)
        #legend.position = "none")
  ) +
  ylab("Relative abundance")+
  facet_wrap("Sample.type",scales="free_x")


barplot.milk.replicates

ggsave("barplot.replicates.milk.pdf",width=50,height=4,limitsize = FALSE)
```

#### Milk - low read samples removed

```{r, message = F}
load("0.2_Filtered_and_formated.RData")

ps <- subset_samples(PSB.all.genus, unique == FALSE & Sample.type == "Milk")
ps = prune_samples(sample_sums(ps)>=200, ps)

ps0 <- transform_sample_counts(ps, function(x) x / sum(x))

#ps1<- merge_samples(ps0, "Time")

#ps1 <- transform_sample_counts(ps0, function(x) x / sum(x))

#Create melted dataframe
df <- psmelt(ps0)

#Select last non-empty taxonomic rank
df[df==""] <- NA
df$tax <- apply(df, 1, function(x) tail(na.omit(x), 1))

#Arrange samples by mean abundance
top <- df %>%
  group_by(tax) %>%
  dplyr::summarize(Mean = mean(Abundance)) %>%
  arrange(-Mean)
#Show top
#top

top10 <- top$tax[1:10]
df0 <- df %>%
  mutate(tax = fct_other(tax, c(as.matrix(top10))))

df0 <- df0[order(df0$Sample, decreasing = TRUE), ]

#Set order for samples
#df0$Treatment <- factor(df0$Treatment, levels = c("Control", "Amp", "FOP", "LM", "LM+Amp", "LM+FOP"))

#Set Hours as factor
#$Hours <- factor(df0$Hours, levels = c("3h", "6h", "24h"))

barplot.milk.replicates.low.read.removed <- ggplot(df0, aes(RCID.uniq, Abundance, fill = fct_reorder(tax, Abundance, .fun = mean))) +
  geom_col(width = 0.8) +
  scale_fill_jco(name = "Taxonomy") +
  scale_fill_manual(values=rep(col_fil,5),name="Genus")+
  theme_classic() +
  theme(axis.line=element_line(size=0.5),
        #panel.border = element_blank(),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        axis.text.x=element_text(angle = -45, vjust = 0.8, hjust = 0),
        strip.text.x = element_text(angle = 0,  face = "bold"),
        legend.key.height = unit(0.5, 'cm'),
        legend.text = element_text(face = "italic"),
        text = element_text(family = 'serif', size = 12)
        #legend.position = "none")
  ) +
  ylab("Relative abundance")+
  facet_wrap("Sample.type",scales="free_x")


barplot.milk.replicates.low.read.removed

ggsave("barplot.replicates.milk.low.read.removed.pdf",width=50,height=4,limitsize = FALSE)
```

#### Milk - low read samples removed - by batch

```{r, message = F}
load("0.2_Filtered_and_formated.RData")

ps <- subset_samples(PSB.all.genus, unique == FALSE & Sample.type == "Milk")
ps = prune_samples(sample_sums(ps)>=200, ps)

ps0 <- transform_sample_counts(ps, function(x) x / sum(x))

#ps1<- merge_samples(ps0, "Time")

#ps1 <- transform_sample_counts(ps0, function(x) x / sum(x))

#Create melted dataframe
df <- psmelt(ps0)

#Select last non-empty taxonomic rank
df[df==""] <- NA
df$tax <- apply(df, 1, function(x) tail(na.omit(x), 1))

#Arrange samples by mean abundance
top <- df %>%
  group_by(tax) %>%
  dplyr::summarize(Mean = mean(Abundance)) %>%
  arrange(-Mean)
#Show top
#top

top10 <- top$tax[1:10]
df0 <- df %>%
  mutate(tax = fct_other(tax, c(as.matrix(top10))))

df0 <- df0[order(df0$Sample, decreasing = TRUE), ]

#Set order for samples
#df0$Treatment <- factor(df0$Treatment, levels = c("Control", "Amp", "FOP", "LM", "LM+Amp", "LM+FOP"))

#Set Hours as factor
#$Hours <- factor(df0$Hours, levels = c("3h", "6h", "24h"))

barplot.milk.replicates.low.read.removed.batch <- ggplot(df0, aes(RCID.uniq, Abundance, fill = fct_reorder(tax, Abundance, .fun = mean))) +
  geom_col(width = 0.8) +
  scale_fill_jco(name = "Taxonomy") +
  scale_fill_manual(values=rep(col_fil,5),name="Genus")+
  theme_classic() +
  theme(axis.line=element_line(size=0.5),
        #panel.border = element_blank(),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        axis.text.x=element_text(angle = -45, vjust = 0.8, hjust = 0),
        strip.text.x = element_text(angle = 0,  face = "bold"),
        legend.key.height = unit(0.5, 'cm'),
        legend.text = element_text(face = "italic"),
        text = element_text(family = 'serif', size = 12)
        #legend.position = "none")
  ) +
  ylab("Relative abundance")+
  facet_wrap(c("Batch","Run"),scales="free_x")


barplot.milk.replicates.low.read.removed.batch

ggsave("barplot.replicates.milk.low.read.removed.batch.pdf",width=30,height=20,limitsize = FALSE)
```

### Vaginal + rectal swabs

```{r, message = F}
load("0.2_Filtered_and_formated.RData")

ps <- subset_samples(PSB.all.genus, Sample.type %in% c("Vaginal swab", "Rectal swab"))

ps0 <- transform_sample_counts(ps, function(x) x / sum(x))

#ps1<- merge_samples(ps0, "Time")

#ps1 <- transform_sample_counts(ps0, function(x) x / sum(x))

#Create melted dataframe
df <- psmelt(ps0)

#Select last non-empty taxonomic rank
df[df==""] <- NA
df$tax <- apply(df, 1, function(x) tail(na.omit(x), 1))

#Arrange samples by mean abundance
top <- df %>%
  group_by(tax) %>%
  dplyr::summarize(Mean = mean(Abundance)) %>%
  arrange(-Mean)
#Show top
#top

top10 <- top$tax[1:10]
df0 <- df %>%
  mutate(tax = fct_other(tax, c(as.matrix(top10))))

df0 <- df0[order(df0$Sample, decreasing = TRUE), ]

#Set order for samples
#df0$Treatment <- factor(df0$Treatment, levels = c("Control", "Amp", "FOP", "LM", "LM+Amp", "LM+FOP"))

#Set Hours as factor
#$Hours <- factor(df0$Hours, levels = c("3h", "6h", "24h"))

barplot.vag.rek.replicates <- ggplot(df0, aes(RCID.uniq, Abundance, fill = fct_reorder(tax, Abundance, .fun = mean))) +
  geom_col(width = 0.8) +
  scale_fill_jco(name = "Taxonomy") +
  scale_fill_manual(values=rep(col_fil,5),name="Genus")+
  theme_classic() +
  theme(axis.line=element_line(size=0.5),
        #panel.border = element_blank(),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        axis.text.x=element_text(angle = -45, vjust = 0.8, hjust = 0),
        strip.text.x = element_text(angle = 0,  face = "bold"),
        legend.key.height = unit(0.5, 'cm'),
        legend.text = element_text(face = "italic"),
        text = element_text(family = 'serif', size = 12)
        #legend.position = "none")
  ) +
  ylab("Relative abundance")+
  facet_wrap("Sample.type",scales="free_x")


barplot.vag.rek.replicates

ggsave("barplot.replicates.vag-rec.pdf",width=50,height=4,limitsize = FALSE)
```

## Select highest read-duplicates

```{r, eval = T, include = T, echo = F, massage = F}
PSB.all.deduped <- ps_dedupe(PSB.all, vars = "id.type", method = "readcount", n = 1)
```

# SessionInfo

```{r}
sessionInfo()
```
