---
title: "2.1_Milk_and_faecal_characterization"
author: "Rasmus Riemer Jakobsen"
date: "`r Sys.time()`"
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
link-citations: yes
---

```{r setup, include=F, message=F, warning=F}
subdir_name <- "2.1_Milk_and_faecal_characterization"

#Load packages
source("Supplementary_scripts/package_load_install.R")
#Load utility functions
source("Supplementary_scripts/utils.R")

dir.create("results", showWarnings = FALSE, recursive = TRUE)

#source("Supplementary_scripts/create_stability.R")

#source("Phyloseq_import_CSS_16S_laca.R")
load("0.3_deduplicated_and_decontaminated.RData")


theme_set(theme_classic())

knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.fullwidth = TRUE,
                      root.dir = setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) #Set working directory to script directory 
)

#Coulour pallettes
col_fil <- brewer.pal(8, "Dark2") #pal_jco("default")(10)

col_scale <- scale_fill_distiller(palette = "Dark2")
```

```{r}
source(here::here("Supplementary_scripts","package_load_install.R")) 
source(here::here("Supplementary_scripts","utils.R")) # please see the src-folder for some custom functions used in 
#source(here::here("src", "run_clusterprofiler.R"))
```

```{r knit, echo=F, eval=FALSE}
rmarkdown::render(
  input = here(str_c(subdir_name, ".Rmd")),
  output_dir = here("results", "reports")
)
```


```{r, message = F,eval=F}
# Collapse all to Genus level

PSB <- tax_glom(PSB,taxrank="Genus")
PSB.all <- tax_glom(PSB.all,taxrank="Genus")

PSB.CSS <- tax_glom(PSB.CSS,taxrank="Genus")
PSB.CSS.all <- tax_glom(PSB.CSS.all,taxrank="Genus")

PSB.rar <- tax_glom(PSB.rar,taxrank="Genus")
PSB.rar.notrim <- tax_glom(PSB.rar.notrim,taxrank="Genus")
```

# Barplots

## Sample types

```{r, message = F}

#Collapse to genus level
bac.phyl <- PSB.genus

ps1 <- transform_sample_counts(bac.phyl, function(x) x / sum(x))

ps1<- merge_samples(ps1, "Sample.type")

ps1 <- transform_sample_counts(ps1, function(x) x / sum(x))

#Create melted dataframe
df <- psmelt(ps1)

#Select last non-empty taxonomic rank
df[df==""] <- NA
df$tax <- apply(df, 1, function(x) tail(na.omit(x), 1))

#Arrange samples by mean abundance
top <- df %>%
  group_by(tax) %>%
  dplyr::summarize(Mean = mean(Abundance)) %>%
  arrange(-Mean)
#Show top
#top

top10 <- top$tax[1:20]
df0 <- df %>%
  mutate(tax = forcats::fct_other(tax, c(as.matrix(top10))))

df0 <- df0[order(df0$Sample, decreasing = TRUE), ]

#Set order for samples
#df0$Treatment <- factor(df0$Treatment, levels = c("Control", "Amp", "FOP", "LM", "LM+Amp", "LM+FOP"))

#Set Hours as factor
#$Hours <- factor(df0$Hours, levels = c("3h", "6h", "24h"))

barplot.type.phylum.assigned <- ggplot(df0, aes(Sample, Abundance, fill = forcats::fct_reorder(tax, Abundance, .fun = mean))) +
  geom_col(width = 0.8) +
  scale_fill_manual(values=rep(col_fil,5),name="Genus")+
  theme_classic() +
  theme(        #panel.border = element_blank(),
        strip.background = element_rect(colour = "white", fill = "white"),
        axis.text.x=element_text(angle = -45, vjust = 0.8, hjust = 0),
        strip.text.x = element_text(angle = 0,  face = "bold"),
        legend.key.height = unit(0.5, 'cm'),
        legend.text = element_text(face = "italic"),
        #legend.position = "none")
  ) +
  ylab("Relative abundance")

barplot.type.phylum.assigned
```

## Milk samples

### Genus

```{r, message = F}

#Collapse to genus level
bac.phyl <- subset_samples(PSB.genus, Sample.type == "Milk")

ps1 <- transform_sample_counts(bac.phyl, function(x) x / sum(x))

ps1<- merge_samples(ps1, "Days")

ps1 <- transform_sample_counts(ps1, function(x) x / sum(x))

#Create melted dataframe
df <- psmelt(ps1)

#Select last non-empty taxonomic rank
df[df==""] <- NA
df$tax <- apply(df, 1, function(x) tail(na.omit(x), 1))

#Arrange samples by mean abundance
top <- df %>%
  group_by(tax) %>%
  dplyr::summarize(Mean = mean(Abundance)) %>%
  arrange(-Mean)
#Show top
#top

top10 <- top$tax[1:10]
df0 <- df %>%
  mutate(tax = forcats::fct_other(tax, c(as.matrix(top10))))

df0 <- df0[order(df0$Sample, decreasing = TRUE), ]

#Set order for samples
#df0$Treatment <- factor(df0$Treatment, levels = c("Control", "Amp", "FOP", "LM", "LM+Amp", "LM+FOP"))

#Set Hours as factor
#$Hours <- factor(df0$Hours, levels = c("3h", "6h", "24h"))

df0$tax <- forcats::fct_reorder(df0$tax, df0$Abundance, .fun = mean)
labels = levels(df0$tax)
labels[labels != "Other"] <- paste0("*", labels[labels != "Other"], "*")
bar.fill <- c("grey",rep(col_fil,5)[1:length(levels(df0$tax))-1]) # Set grey as last color = Other

barplot.milk.days.genus <- ggplot(df0, aes(Sample, Abundance, fill = forcats::fct_reorder(tax, Abundance, .fun = mean))) +
  geom_col(width = 0.8) +
  scale_fill_manual(values=bar.fill,name="Genus",labels = labels)+
  theme_classic() +
  theme(strip.background = element_rect(colour = "white", fill = "white"),
        strip.text.x = element_text(angle = 0,  face = "bold"),
        legend.key.height = unit(0.5, 'cm'),
        legend.text = element_markdown(),
        plot.title = element_text(hjust = 0.5)
  ) +
  ylab("Relative abundance")+
  xlab("Days after birth") +
  ggtitle("Milk")

sl(barplot.milk.days.genus,barplot.milk.days.genus,overwrite=T)


barplot.milk.days.genus

```

### Family

```{r, message = F}

#Collapse to genus level
bac.phyl <- subset_samples(PSB.family, Sample.type == "Milk")

ps1 <- transform_sample_counts(bac.phyl, function(x) x / sum(x))

ps1<- merge_samples(ps1, "Days")

ps1 <- transform_sample_counts(ps1, function(x) x / sum(x))

#Create melted dataframe
df <- psmelt(ps1)

#Select last non-empty taxonomic rank
df[df==""] <- NA
df$tax <- apply(df, 1, function(x) tail(na.omit(x), 1))

#Arrange samples by mean abundance
top <- df %>%
  group_by(tax) %>%
  dplyr::summarize(Mean = mean(Abundance)) %>%
  arrange(-Mean)
#Show top
#top

top10 <- top$tax[1:10]
df0 <- df %>%
  mutate(tax = forcats::fct_other(tax, c(as.matrix(top10))))

df0 <- df0[order(df0$Sample, decreasing = TRUE), ]

#Set order for samples
#df0$Treatment <- factor(df0$Treatment, levels = c("Control", "Amp", "FOP", "LM", "LM+Amp", "LM+FOP"))

#Set Hours as factor
#$Hours <- factor(df0$Hours, levels = c("3h", "6h", "24h"))

barplot.milk.days.family <- ggplot(df0, aes(Sample, Abundance, fill = forcats::fct_reorder(tax, Abundance, .fun = mean))) +
  geom_col(width = 0.8) +
  scale_fill_jco(name = "Taxonomy") +
  scale_fill_manual(values=rep(col_fil,5),name="Family")+
  theme_classic() +
  theme(        #panel.border = element_blank(),
        strip.background = element_rect(colour = "white", fill = "white"),
        #axis.text.x=element_text(angle = -45, vjust = 0.8, hjust = 0),
        strip.text.x = element_text(angle = 0,  face = "bold"),
        legend.key.height = unit(0.5, 'cm'),
        legend.text = element_text(face = "italic"),
        #legend.position = "none")
  ) +
  ylab("Relative abundance")+
  xlab("Day")

barplot.milk.days.family
```

## Infant faecal samples

### Genus

```{r, message = F}

#Collapse to genus level
bac.phyl <- subset_samples(PSB.genus, Sample.type == "Faecal sample")

ps1 <- transform_sample_counts(bac.phyl, function(x) x / sum(x))

ps1<- merge_samples(ps1, "Days")

ps1 <- transform_sample_counts(ps1, function(x) x / sum(x))

#Create melted dataframe
df <- psmelt(ps1)

#Select last non-empty taxonomic rank
df[df==""] <- NA
df$tax <- apply(df, 1, function(x) tail(na.omit(x), 1))

#Arrange samples by mean abundance
top <- df %>%
  group_by(tax) %>%
  dplyr::summarize(Mean = mean(Abundance)) %>%
  arrange(-Mean)
#Show top
#top

top10 <- top$tax[1:10]
df0 <- df %>%
  mutate(tax = forcats::fct_other(tax, c(as.matrix(top10))))

df0 <- df0[order(df0$Sample, decreasing = TRUE), ]

#Set order for samples
#df0$Treatment <- factor(df0$Treatment, levels = c("Control", "Amp", "FOP", "LM", "LM+Amp", "LM+FOP"))

#Set Hours as factor
#$Hours <- factor(df0$Hours, levels = c("3h", "6h", "24h"))

df0$tax <- forcats::fct_reorder(df0$tax, df0$Abundance, .fun = mean)
labels = levels(df0$tax)
labels[labels != "Other"] <- paste0("*", labels[labels != "Other"], "*")
bar.fill <- c("grey",rep(col_fil,5)[1:length(levels(df0$tax))-1]) # Set grey as last color = Other

barplot.faecal.days.genus <- ggplot(df0, aes(Sample, Abundance, fill = tax)) +
  geom_col(width = 0.8) +
  scale_fill_manual(values=bar.fill,name="Genus",labels = labels)+
  theme_classic() +
  theme(strip.background = element_rect(colour = "white", fill = "white"),
        strip.text.x = element_text(angle = 0,  face = "bold"),
        legend.key.height = unit(0.5, 'cm'),
        legend.text = element_markdown(),
        plot.title = element_text(hjust = 0.5)
  ) +
  ylab("Relative abundance")+
  xlab("Days after birth") +
  ggtitle("Infant faeces")

sl(barplot.faecal.days.genus,barplot.faecal.days.genus,overwrite = T)

barplot.faecal.days.genus
```



### Family

```{r, message = F}

#Collapse to genus level
bac.phyl <- subset_samples(PSB.family, Sample.type == "Milk")

ps1 <- transform_sample_counts(bac.phyl, function(x) x / sum(x))

ps1<- merge_samples(ps1, "Days")

ps1 <- transform_sample_counts(ps1, function(x) x / sum(x))

#Create melted dataframe
df <- psmelt(ps1)

#Select last non-empty taxonomic rank
df[df==""] <- NA
df$tax <- apply(df, 1, function(x) tail(na.omit(x), 1))

#Arrange samples by mean abundance
top <- df %>%
  group_by(tax) %>%
  dplyr::summarize(Mean = mean(Abundance)) %>%
  arrange(-Mean)
#Show top
#top

top10 <- top$tax[1:10]
df0 <- df %>%
  mutate(tax = forcats::fct_other(tax, c(as.matrix(top10))))

df0 <- df0[order(df0$Sample, decreasing = TRUE), ]

#Set order for samples
#df0$Treatment <- factor(df0$Treatment, levels = c("Control", "Amp", "FOP", "LM", "LM+Amp", "LM+FOP"))

#Set Hours as factor
#$Hours <- factor(df0$Hours, levels = c("3h", "6h", "24h"))

barplot.milk.days.family <- ggplot(df0, aes(Sample, Abundance, fill = forcats::fct_reorder(tax, Abundance, .fun = mean))) +
  geom_col(width = 0.8) +
  scale_fill_manual(values=rep(col_fil,5),name="Family")+
  theme_classic() +
  theme(        #panel.border = element_blank(),
        strip.background = element_rect(colour = "white", fill = "white"),
        #axis.text.x=element_text(angle = -45, vjust = 0.8, hjust = 0),
        strip.text.x = element_text(angle = 0,  face = "bold"),
        legend.key.height = unit(0.5, 'cm'),
        legend.text = element_text(face = "italic"),
        plot.title = element_text(hjust = 0.5)
        #legend.position = "none")
  ) +
  ylab("Relative abundance")+
  xlab("Day") + 
  ggtitle("Milk")

barplot.milk.days.family
```

### Species

```{r, message = F}

#Collapse to Species level
bac.phyl <- subset_samples(PSB.species, Sample.type == "Faecal sample")

ps1 <- transform_sample_counts(bac.phyl, function(x) x / sum(x))

ps1<- merge_samples(ps1, "Days")

ps1 <- transform_sample_counts(ps1, function(x) x / sum(x))

#Create melted dataframe
df <- psmelt(ps1)

#Select last non-empty taxonomic rank
df[df==""] <- NA
df$tax <- apply(df, 1, function(x) tail(na.omit(x), 1))

#Arrange samples by mean abundance
top <- df %>%
  group_by(tax) %>%
  dplyr::summarize(Mean = mean(Abundance)) %>%
  arrange(-Mean)
#Show top
#top

top10 <- top$tax[1:10]
df0 <- df %>%
  mutate(tax = forcats::fct_other(tax, c(as.matrix(top10))))

df0 <- df0[order(df0$Sample, decreasing = TRUE), ]

#Set order for samples
#df0$Treatment <- factor(df0$Treatment, levels = c("Control", "Amp", "FOP", "LM", "LM+Amp", "LM+FOP"))

#Set Hours as factor
#$Hours <- factor(df0$Hours, levels = c("3h", "6h", "24h"))

barplot.faecal.days.species <- ggplot(df0, aes(Sample, Abundance, fill = forcats::fct_reorder(tax, Abundance, .fun = mean))) +
  geom_col(width = 0.8) +
  scale_fill_manual(values=rep(col_fil,5),name="Species")+
  theme_classic() +
  theme(        #panel.border = element_blank(),
        strip.background = element_rect(colour = "white", fill = "white"),
        #axis.text.x=element_text(angle = -45, vjust = 0.8, hjust = 0),
        strip.text.x = element_text(angle = 0,  face = "bold"),
        legend.key.height = unit(0.5, 'cm'),
        legend.text = element_text(face = "italic"),
        plot.title = element_text(hjust = 0.5)
        #legend.position = "none")
  ) +
  ylab("Relative abundance")+
  xlab("Day") +
  ggtitle("Infant faeces")

barplot.faecal.days.species

sl(barplot.faecal.days.species,barplot.faecal.days.species,overwrite = T)
```

### BMI groups

```{r, message = F}

#Collapse to Species level
bac.phyl <- subset_samples(PSB.genus, Sample.type == "Faecal sample")

ps1 <- transform_sample_counts(bac.phyl, function(x) x / sum(x))

ps1<- merge_samples(ps1, "BMI.group")

ps1 <- transform_sample_counts(ps1, function(x) x / sum(x))

#Create melted dataframe
df <- psmelt(ps1)

#Select last non-empty taxonomic rank
df[df==""] <- NA
df$tax <- apply(df, 1, function(x) tail(na.omit(x), 1))

#Arrange samples by mean abundance
top <- df %>%
  group_by(tax) %>%
  dplyr::summarize(Mean = mean(Abundance)) %>%
  arrange(-Mean)
#Show top
#top

top10 <- top$tax[1:10]
df0 <- df %>%
  mutate(tax = forcats::fct_other(tax, c(as.matrix(top10))))

df0 <- df0[order(df0$Sample, decreasing = TRUE), ]

#Set order for samples
#df0$Treatment <- factor(df0$Treatment, levels = c("Control", "Amp", "FOP", "LM", "LM+Amp", "LM+FOP"))

#Set Hours as factor
#$Hours <- factor(df0$Hours, levels = c("3h", "6h", "24h"))

barplot.faecal.days.species <- ggplot(df0, aes(Sample, Abundance, fill = forcats::fct_reorder(tax, Abundance, .fun = mean))) +
  geom_col(width = 0.8) +
  scale_fill_manual(values=rep(col_fil,5),name="Genus")+
  theme_classic() +
  theme(        #panel.border = element_blank(),
        strip.background = element_rect(colour = "white", fill = "white"),
        #axis.text.x=element_text(angle = -45, vjust = 0.8, hjust = 0),
        strip.text.x = element_text(angle = 0,  face = "bold"),
        legend.key.height = unit(0.5, 'cm'),
        legend.text = element_text(face = "italic"),
        plot.title = element_text(hjust = 0.5)
        #legend.position = "none")
  ) +
  ylab("Relative abundance")+
  xlab("ppBMI") +
  ggtitle("Infant faeces")

barplot.faecal.days.species
```



# Rabuplot

## Time vs Sample type

### Facet wrapped -time

```{r, message = F}
#remotes::install_github("jstokholm/rabuplot")
library(rabuplot)

ps <- subset_samples(PSB, Days %in% c("3","30","60","90"))
ps@sam_data$Days <- factor(ps@sam_data$Days , levels = c("3","30","60","90"))
#ps@sam_data$Sample.type <- 

sl(rabu.faec.milk.days,
  rabuplot(
  ps,
  bar_chart_stacked =  TRUE,
  predictor = "Days",
  facet_wrap = "Sample.type",
  p_adjust = TRUE,
  N_taxa = 12,
  ylabs = "Relative abundance",
  main = "Days after birth",
  colors = c(RColorBrewer::brewer.pal(12, "Paired"),"#D3D3D3")
),
overwrite=T)

rabu.faec.milk.days
```

### Facet wrapped

#### BMI

```{r, message = F}
#remotes::install_github("jstokholm/rabuplot")
library(rabuplot)

ps <- subset_samples(PSB, Days %in% c("3","30","60","90"))
ps@sam_data$Days <- factor(ps@sam_data$Days , levels = c("3","30","60","90"))

sl(rabu.faec.milk.bmi,rabuplot(
  ps,
  bar_chart_stacked =  TRUE,
  Time = "Days",
  predictor = "BMI.group",
  facet_wrap = "Sample.type",
  p_adjust = TRUE,
  N_taxa = 12,
  text_angle_x = 45,
  ylabs = "Relative abundance",
  main = "Pre-pregnancy BMI",
  colors = c(RColorBrewer::brewer.pal(12, "Paired"),"#D3D3D3")
),
overwrite = T)

rabu.faec.milk.bmi
```

```{r}
ggarrange(rabu.faec.milk.bmi,rabu.faec.milk.days,nrow=2,common.legend = TRUE, legend = "right",heights = c(1.2,1))
```

### Plotted individually

#### Faecal samples

##### Days per BMI

```{r, message = F}
#remotes::install_github("jstokholm/rabuplot")
library(rabuplot)

ps <- subset_samples(PSB, Days %in% c("3","30","60","90") & Sample.type == "Faecal sample")
#ps@sam_data$Days <- factor(ps@sam_data$Days , levels = rev(c("3","30","60","90")))

rabu.faec.days.bmi <- rabuplot(
  ps,
  bar_chart_stacked =  TRUE,
  predictor = "Days",
  facet_wrap = "BMI.group",
  p_adjust = TRUE,
  N_taxa = 17
  #colors = rep(col_fil,4)[1:17]
)

rabu.faec.days.bmi
```

##### BMI per days

```{r, message = F}
#remotes::install_github("jstokholm/rabuplot")
library(rabuplot)

ps <- subset_samples(PSB, Days %in% c("3","30","60","90") & Sample.type == "Faecal sample")
#ps@sam_data$Days <- factor(ps@sam_data$Days , levels = rev(c("3","30","60","90")))

rabu.faec.bmi.days <- rabuplot(
  ps,
  bar_chart_stacked =  TRUE,
  predictor = "BMI.group",
  facet_wrap = "Days",
  p_adjust = TRUE,
  N_taxa = 17
  #colors = rep(col_fil,4)[1:17]
)

rabu.faec.bmi.days
```

```{r}
ggarrange(rabu.faec.days.bmi,rabu.faec.bmi.days,nrow=2,common.legend = TRUE, legend = "right")
```


#### Violin plots

```{r, message = F}
#remotes::install_github("jstokholm/rabuplot")
library(rabuplot)

ps <- subset_samples(PSB, Days %in% c("3","30","60","90"))
ps@sam_data$Days <- factor(ps@sam_data$Days , levels = rev(c("3","30","60","90")))

ps.milk <- subset_samples(ps,Sample.type=="Milk")

rabu.milk.days <- rabuplot(
  ps.milk,
  violin = TRUE,
  By_median = FALSE,
  predictor = "Days",
  legend_title = "Days after birth",
  italic_names = FALSE,
  p_adjust = TRUE,
  p_stars = TRUE,
  text_angle_x = 45,
  main = "Milk",
  colors = rev(col_fil[1:4])) # to share order with beta div plot colors

labels <- levels(factor(rabu.milk.days$data$variable))
labels[labels != "Other genus"] <- paste0("*", labels[labels != "Other genus"], "*")

rabu.milk.days <- rabu.milk.days + scale_x_discrete(labels = labels) + theme(axis.text.y = ggtext::element_markdown())
  

sl(rabu.milk.days,rabu.milk.days,overwrite = T)
#rabu.milk.days

ps.faec <- subset_samples(ps,Sample.type=="Faecal sample")

rabu.faec.days <- rabuplot(
  ps.faec,
  violin = TRUE,
  By_median = FALSE,
  predictor = "Days",
  legend_title = "Days after birth",
  italic_names = FALSE,
  p_adjust = TRUE,
  p_stars = TRUE,
  text_angle_x = 45,
  main = "Infant faeces",
  colors = rev(col_fil[2:4])) # to share order with beta div plot colors

labels <- levels(factor(rabu.faec.days$data$variable))
labels[labels != "Other genus"] <- paste0("*", labels[labels != "Other genus"], "*")

rabu.faec.days <- rabu.faec.days + scale_x_discrete(labels = labels) + theme(axis.text.y = ggtext::element_markdown())
  

sl(rabu.faec.days,rabu.faec.days,overwrite = T)

ggarrange(rabu.milk.days,rabu.faec.days,nrow=1,common.legend = TRUE,legend="right")
```

#### Bifidobacteria only

```{r, message = F}
#remotes::install_github("jstokholm/rabuplot")
library(rabuplot)

ps <- subset_samples(PSB.all, Days %in% c("3","30","60","90"))
ps@sam_data$Days <- factor(ps@sam_data$Days , levels = rev(c("3","30","60","90")))

ps.milk <- subset_samples(ps,Sample.type=="Milk")

rabu.milk.days <- rabuplot(
  ps.milk,
  violin = TRUE,
  predictor = "Days",
   type = "Species",
  select_taxa = "Bifidobacterium",
  #facet_wrap = "Sample.type",
  p_adjust = TRUE,
  p_stars = TRUE,
  text_angle_x = 45,
  main = "Milk",
  colors = rev(col_fil[1:4])) # to share order with beta div plot colors

#rabu.milk.days

ps.faec <- subset_samples(ps,Sample.type=="Faecal sample")

rabu.faec.days <- rabuplot(
  ps.faec,
  violin = TRUE,
  predictor = "Days",
  type = "Species",
  select_taxa = "Bifidobacterium",
  #facet_wrap = "Sample.type",
  p_adjust = TRUE,
  p_stars = TRUE,
  text_angle_x = 45,
  main = "Infant faeces",
  colors = rev(col_fil[2:4])) # to share order with beta div plot colors

sl(rabu.milk.faec.days.bifido,ggarrange(rabu.milk.days,rabu.faec.days,nrow=1,common.legend = TRUE,legend="right"),overwrite = T)

rabu.milk.faec.days.bifido
```

#### Boxplots

```{r, message = F}
#remotes::install_github("jstokholm/rabuplot")
library(rabuplot)

ps <- subset_samples(PSB.all, Days %in% c("3","30","60","90"))
ps@sam_data$Days <- factor(ps@sam_data$Days , levels = rev(c("3","30","60","90")))

ps.milk <- subset_samples(ps,Sample.type=="Milk")

rabu.box.milk.days <- rabuplot(
  ps.milk,
  violin = FALSE,
  predictor = "Days",
  legend_title = "Days after birth",
  italic_names = FALSE,
  p_adjust = TRUE,
  p_stars = TRUE,
  text_angle_x = 45,
  main = "Milk",
  colors = rev(col_fil[1:4])) # to share order with beta div plot colors

labels <- levels(factor(rabu.box.milk.days$data$variable))
labels[labels != "Other genus"] <- paste0("*", labels[labels != "Other genus"], "*")

rabu.box.milk.days <- rabu.box.milk.days + scale_x_discrete(labels = labels) + theme(axis.text.y = ggtext::element_markdown())

sl(rabu.box.milk.days, rabu.box.milk.days, overwrite = TRUE)

ps.faec <- subset_samples(ps,Sample.type=="Faecal sample")

rabu.box.faec.days <- rabuplot(
  ps.faec,
  violin = FALSE,
  predictor = "Days",
  legend_title = "Days after birth",
  italic_names = FALSE,
  p_adjust = TRUE,
  p_stars = TRUE,
  text_angle_x = 45,
  main = "Infant faeces",
  colors = rev(col_fil[2:4])) # to share order with beta div plot colors

labels <- levels(factor(rabu.box.faec.days$data$variable))
labels[labels != "Other genus"] <- paste0("*", labels[labels != "Other genus"], "*")

rabu.box.faec.days <- rabu.box.faec.days + scale_x_discrete(labels = labels) + theme(axis.text.y = ggtext::element_markdown())

sl(rabu.box.faec.days, rabu.box.faec.days, overwrite = TRUE)

ggarrange(rabu.box.milk.days,rabu.box.faec.days,nrow=1,common.legend = TRUE)
```

## BMI by sample type

### Milk

```{r, message = F}
#remotes::install_github("jstokholm/rabuplot")
library(rabuplot)

ps <- subset_samples(PSB, Days %in% c("3","30","60","90") & Sample.type=="Milk")
ps@sam_data$BMI.group <- factor(ps@sam_data$BMI.group , levels = rev(levels(ps@sam_data$BMI.group)))

rabu <- rabuplot(
  ps,
  violin = TRUE,
  By_median = FALSE,
  predictor = "BMI.group",
  legend_title = "Pre-pregnancy BMI",
  main = "Milk bacteria",
  italic_names = FALSE,
  text_angle_x = 45,
  p_stars = TRUE,
  colors = col_fil[3:1])

labels <- levels(factor(rabu$data$variable))
labels[labels != "Other genus"] <- paste0("*", labels[labels != "Other genus"], "*")

rabu <- rabu + scale_x_discrete(labels = labels) + theme(axis.text.y = ggtext::element_markdown())

sl(rabu.milk.bmi,rabu,overwrite = T)

rabu.milk.bmi
```

#### Secretor stratified

```{r, message = F}
#remotes::install_github("jstokholm/rabuplot")
library(rabuplot)

ps <- subset_samples(PSB, Days %in% c("3","30","60","90") & Sample.type=="Milk")
ps@sam_data$BMI.group <- factor(ps@sam_data$BMI.group , levels = rev(levels(ps@sam_data$BMI.group)))

rabu <- rabuplot(
  ps,
  violin = TRUE,
  By_median = FALSE,
  predictor = "BMI.group",
  legend_title = "Pre-pregnancy BMI",
  main = "Milk bacteria",
  facet_wrap = "Secretor",
  italic_names = FALSE,
  text_angle_x = 45,
  p_stars = TRUE,
  colors = col_fil[3:1])

labels <- levels(factor(rabu$data$variable))
labels[labels != "Other genus"] <- paste0("*", labels[labels != "Other genus"], "*")

rabu <- rabu + scale_x_discrete(labels = labels) + theme(axis.text.y = ggtext::element_markdown())

sl(rabu.milk.bmi.sec,rabu,overwrite = T)

rabu.milk.bmi.sec
```

### Faecal samples

```{r, message = F}
#remotes::install_github("jstokholm/rabuplot")
library(rabuplot)

ps <- subset_samples(PSB, Days %in% c("3","30","60","90") & Sample.type=="Faecal sample")
ps@sam_data$BMI.group <- factor(ps@sam_data$BMI.group , levels = rev(levels(ps@sam_data$BMI.group)))

rabu <- rabuplot(
  ps,
  violin = TRUE,
  By_median = FALSE,
  predictor = "BMI.group",
  legend_title = "Pre-pregnancy BMI",
  main = "Infant faecal bacteria",
  italic_names = FALSE,
  text_angle_x = 45,
  p_stars = TRUE,
  colors = col_fil[3:1])

labels <- levels(factor(rabu$data$variable))
labels[labels != "Other genus"] <- paste0("*", labels[labels != "Other genus"], "*")

rabu <- rabu + scale_x_discrete(labels = labels) + theme(axis.text.y = ggtext::element_markdown())

sl(rabu.faec.bmi,rabu,overwrite = T)

rabu.faec.bmi
```

#### Secretor stratified

```{r, message = F}
#remotes::install_github("jstokholm/rabuplot")
library(rabuplot)

ps <- subset_samples(PSB, Days %in% c("3","30","60","90") & Sample.type=="Faecal sample")
ps@sam_data$BMI.group <- factor(ps@sam_data$BMI.group , levels = rev(levels(ps@sam_data$BMI.group)))

rabu <- rabuplot(
  ps,
  violin = TRUE,
  By_median = FALSE,
  predictor = "BMI.group",
  legend_title = "Pre-pregnancy BMI",
  main = "Infant faecal bacteria",
  facet_wrap = "Secretor",
  italic_names = FALSE,
  text_angle_x = 45,
  p_stars = TRUE,
  colors = col_fil[3:1])

labels <- levels(factor(rabu$data$variable))
labels[labels != "Other genus"] <- paste0("*", labels[labels != "Other genus"], "*")

rabu <- rabu + scale_x_discrete(labels = labels) + theme(axis.text.y = ggtext::element_markdown())

sl(rabu.faec.bmi.sec,rabu,overwrite = T)

rabu.faec.bmi.sec
```

## Milk samples

### BMI over time

```{r, message = F}
#remotes::install_github("jstokholm/rabuplot")
library(rabuplot)

ps <- subset_samples(PSB, Days %in% c("3","30","60","90") & Sample.type == "Milk")
ps@sam_data$BMI.group <- factor(ps@sam_data$BMI.group , levels = rev(levels(ps@sam_data$BMI.group)))

ps@sam_data$RCID[duplicated(ps@sam_data$RCID)]

rabuplot(
  ps,
  violin = TRUE,
  predictor = "BMI.group",
  Time = "Days",
  facet_wrap = "Days",
  p_adjust = TRUE,
  colors = col_fil[1:3])
```

### Time by BMI

```{r, message = F}
#remotes::install_github("jstokholm/rabuplot")
library(rabuplot)

ps <- subset_samples(PSB, Days %in% c("3","30","60","90") & Sample.type == "Milk")
ps@sam_data$Days <- factor(ps@sam_data$Days , levels = rev(c("3","30","60","90")))


rabuplot(
  ps,
  violin = TRUE,
  predictor = "Days",
  #Time = "Days",
  facet_wrap = "BMI.group",
  p_adjust = TRUE,
  colors = col_fil[1:4])
```

## Faecal samples

### BMI over time

```{r, message = F}
#remotes::install_github("jstokholm/rabuplot")
library(rabuplot)

ps <- subset_samples(PSB, Days %in% c("3","30","60","90") & Sample.type == "Faecal sample")
#ps@sam_data$Days <- factor(ps@sam_data$Days , levels = rev(c("3","30","60","90")))

ps@sam_data$RCID[duplicated(ps@sam_data$RCID)]

rabuplot(
  ps,
  violin = TRUE,
  predictor = "BMI.group",
  Time = "Days",
  facet_wrap = "Days",
  p_adjust = FALSE,
  p_stars=TRUE,
  colors = col_fil[1:3])
```

#### Species

```{r, message = F}
#remotes::install_github("jstokholm/rabuplot")
library(rabuplot)

ps <- subset_samples(PSB, Days %in% c("3","30","60","90") & Sample.type == "Faecal sample")
#ps@sam_data$Days <- factor(ps@sam_data$Days , levels = rev(c("3","30","60","90")))

ps@sam_data$RCID[duplicated(ps@sam_data$RCID)]

rabuplot(
  ps,
  violin = TRUE,
  type = "species",
  predictor = "BMI.group",
  Time = "Days",
  facet_wrap = "Days",
  p_adjust = FALSE,
  colors = col_fil[1:3])
```


### Time per BMI

```{r, message = F}
#remotes::install_github("jstokholm/rabuplot")
library(rabuplot)

ps <- subset_samples(PSB, Days %in% c("3","30","60","90") & Sample.type == "Faecal sample")
#ps@sam_data$Days <- factor(ps@sam_data$Days , levels = rev(c("3","30","60","90")))

ps@sam_data$RCID[duplicated(ps@sam_data$RCID)]

rabuplot(
  ps,
  violin = TRUE,
  predictor = "BMI.group",
  facet_wrap = "Days",
  p_stars = TRUE,
  #p_adjust = TRUE,
  colors = col_fil[1:3])
```




# Alpha diversity

## Milk

### Days

```{r}

ps <- ps <- subset_samples(PSB, Days %in% c("3","30","60","90") & Sample.type == "Milk")
ps@sam_data$Days <- factor(ps@sam_data$Days , levels = c("3","30","60","90"))

#Normalize to mean read count
total = median(sample_sums(ps))
standf = function(x, t=total) round(t * (x / sum(x)))
ps.rel = transform_sample_counts(ps, standf)

#Set desired alpha diversity metrics

alpha_met <- c("Observed", "Shannon", "InvSimpson")

###########Boxplots

##Time

variable <- c("Days")

alpha.metrics.milk.days <- plot_richness(ps.rel, measures=alpha_met , x=variable, color=variable) + 
  geom_boxplot(alpha=0.1) +
  theme_classic() +
  scale_color_manual(values=col_fil) +
  theme(strip.background = element_rect(colour = "white", fill = "white"),
        axis.text.x=element_text(angle = 45, hjust = 1),
        strip.text.x = element_text(size = 8, angle = 0),
        legend.text = element_text(size = 8)
  ) +
  labs(x="Days",
       y="Alpha diversity") + 
  ggtitle("Milk") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))

comparisons <- levels(alpha.metrics.milk.days$data[[variable]])

combinations <- combn(levels(alpha.metrics.milk.days$data[[variable]]), 2, simplify = FALSE) %>%
  lapply(., function(x) c(x[1], x[2]))

ggboxplot(alpha.metrics.milk.days$data,x=variable,y="value",
          color="variable",
          facet.by = "variable",
          scales="free") +
  stat_compare_means(comparisons = combinations, label="p.format",method="wilcox.test",p.adjust.methods = "fdr") +
  theme_classic() +
  scale_color_manual(values=col_fil) 

sl(alpha.metrics.milk.days,alpha.metrics.milk.days,overwrite = T)

alpha.metrics.milk.days
```

```{r}
##Richness - all time points

richness = estimate_richness(ps.rel, measures = alpha_met)

#Time

rich <- cbind(richness, variable = sample_data(ps.rel)$Days)

TukeyHSD(aov(Shannon ~ variable, rich))

TukeyHSD(aov(Observed ~ variable, rich))
```

### BMI

```{r}

ps <- subset_samples(PSB, Sample.type == "Milk")
#ps@sam_data$BMI <- factor(ps@sam_data$BMI , levels = c("3","30","60","90"))

#Normalize to mean read count
total = median(sample_sums(ps))
standf = function(x, t=total) round(t * (x / sum(x)))
ps.rel = transform_sample_counts(ps, standf)

#Set desired alpha diversity metrics

alpha_met <- c("Observed", "Shannon", "InvSimpson")

###########Boxplots

##Time

variable <- c("BMI.group")

alpha.metrics.milk.bmi <- plot_richness(ps.rel, measures=alpha_met , x=variable, color=variable) + 
  geom_boxplot(alpha=0.1) +
  theme_classic() +
  scale_color_manual(values=col_fil) +
  theme(strip.background = element_rect(colour = "white", fill = "white"),
        axis.text.x=element_text(angle = 45, hjust = 1),
        strip.text.x = element_text(size = 8, angle = 0),
        legend.text = element_text(size = 8)
  ) +
  labs(x="BMI",
       y="Alpha diversity") + 
  ggtitle("Milk") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))

comparisons <- levels(alpha.metrics.milk.bmi$data[[variable]])

combinations <- combn(levels(alpha.metrics.milk.bmi$data[[variable]]), 2, simplify = FALSE) %>%
  lapply(., function(x) c(x[1], x[2]))


ggboxplot(alpha.metrics.milk.bmi$data,x=variable,y="value",
          color="variable",
          facet.by = "variable",
          scales="free") +
  stat_compare_means(comparisons = combinations, label="p.format",method="wilcox.test",p.adjust.methods = "fdr") +
  theme_classic() +
  scale_color_manual(values=col_fil) 

sl(alpha.metrics.milk.bmi,alpha.metrics.milk.bmi,overwrite = T)

alpha.metrics.milk.bmi
```

```{r}
##Richness - all time points

richness = estimate_richness(ps.rel, measures = alpha_met)

#Time

rich <- cbind(richness, variable = sample_data(ps.rel)$BMI.group)

TukeyHSD(aov(Shannon ~ variable, rich))

TukeyHSD(aov(Observed ~ variable, rich))
```

#### Only shannon

```{r}

ps <- subset_samples(PSB, Sample.type == "Milk")
#ps@sam_data$BMI <- factor(ps@sam_data$BMI , levels = c("3","30","60","90"))

#Normalize to mean read count
total = median(sample_sums(ps))
standf = function(x, t=total) round(t * (x / sum(x)))
ps.rel = transform_sample_counts(ps, standf)

#Set desired alpha diversity metrics

alpha_met <- c("Shannon")

###########Boxplots

##Time

variable <- c("BMI.group")

alpha.metrics.milk.bmi.shannon <- plot_richness(ps.rel, measures=alpha_met , x=variable, color=variable) + 
  geom_boxplot(alpha=0.1) +
  theme_classic() +
  scale_color_manual(values=col_fil) +
  theme(strip.background = element_rect(colour = "white", fill = "white"),
        axis.text.x=element_text(angle = 45, hjust = 1),
  ) +
  ggtitle("Milk") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))


combinations <- combn(levels(alpha.metrics.milk.bmi$data[[variable]]), 2, simplify = FALSE) %>%
  lapply(., function(x) c(x[1], x[2]))

sl(alpha.shannon.milk.bmi,
   ggboxplot(alpha.metrics.milk.bmi.shannon$data,x=variable,y="value",
          color=variable,
          scales="free") +
  stat_compare_means(comparisons = combinations, label="p.signif",method="wilcox.test",p.adjust.methods = "fdr") +
  theme(strip.background = element_rect(colour = "white", fill = "white"),
        axis.text.x=element_text(angle = 45, hjust = 1)) +
  labs(main="Milk",x="",y="Shannon diversity") +
  scale_color_manual(values=col_fil),
  overwrite = T)

alpha.shannon.milk.bmi
```

### Alfa diversity vs maternal age

```{r}
#Check whether alfa div and sample dataframes match
length(intersect(rownames(rich),rownames(ps@sam_data)))==nrow(rich)

df <- dplyr::select(rich,Shannon) %>%
  cbind(.,'Maternal age'=ps@sam_data$Maternal.age) %>%
  cbind(.,'BMI group'=ps@sam_data$BMI.group)

ggplot(df,aes(x=`Maternal age`,y=Shannon,color=`BMI group`)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_cor(method = "pearson") 
```

## Infant faeces

### Days

```{r}

ps <- ps <- subset_samples(PSB, Days %in% c("30","60","90") & Sample.type == "Faecal sample")
ps@sam_data$Days <- factor(ps@sam_data$Days , levels = c("30","60","90"))

#Normalize to mean read count
total = median(sample_sums(ps))
standf = function(x, t=total) round(t * (x / sum(x)))
ps.rel = transform_sample_counts(ps, standf)

#Set desired alpha diversity metrics

alpha_met <- c("Observed", "Shannon", "InvSimpson")

###########Boxplots

##Time

variable <- c("Days")

alpha.metrics.faec.days <- plot_richness(ps.rel, measures=alpha_met , x=variable, color=variable) + 
  geom_boxplot(alpha=0.1) +
  theme_classic() +
  scale_color_manual(values=col_fil) +
  theme(strip.background = element_rect(colour = "white", fill = "white"),
  ) +
  labs(x="Days",
       y="Alpha diversity") + 
  ggtitle("Infant faeces") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))

combinations <- combn(levels(alpha.metrics.faec.days$data[[variable]]), 2, simplify = FALSE) %>%
  lapply(., function(x) c(x[1], x[2]))

ggboxplot(alpha.metrics.faec.days$data,x=variable,y="value",
          color="variable",
          facet.by = "variable",
          scales="free") +
  stat_compare_means(comparisons = combinations, label="p.format",method="wilcox.test",p.adjust.methods = "fdr") +
  theme_classic() +
  scale_color_manual(values=col_fil) 

sl(alpha.metrics.faec.days,alpha.metrics.faec.days,overwrite = T)

alpha.metrics.faec.days
```

```{r}
##Richness - all time points

richness = estimate_richness(ps.rel, measures = alpha_met)

#Time

rich <- cbind(richness, variable = sample_data(ps.rel)$Days)

TukeyHSD(aov(Shannon ~ variable, rich))

TukeyHSD(aov(Observed ~ variable, rich))
```

#### Only shannon

```{r}

ps <- subset_samples(PSB, Sample.type == "Faecal sample")
#ps@sam_data$BMI <- factor(ps@sam_data$BMI , levels = c("3","30","60","90"))

#Normalize to mean read count
total = median(sample_sums(ps))
standf = function(x, t=total) round(t * (x / sum(x)))
ps.rel = transform_sample_counts(ps, standf)

#Set desired alpha diversity metrics

alpha_met <- c("Shannon")

###########Boxplots

##Time

variable <- c("Days")

alpha.metrics.faec.days.shannon <- plot_richness(ps.rel, measures=alpha_met , x=variable, color=variable) + 
  geom_boxplot(alpha=0.1) +
  theme_classic() +
  scale_color_manual(values=col_fil) +
  theme(strip.background = element_rect(colour = "white", fill = "white"),
        axis.text.x=element_text(angle = 45, hjust = 1),
  ) +
  ggtitle("Infant faeces") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))


combinations <- combn(levels(alpha.metrics.faec.days$data[[variable]]), 2, simplify = FALSE) %>%
  lapply(., function(x) c(x[1], x[2]))

sl(alpha.shannon.faec.days,
   ggboxplot(alpha.metrics.faec.days.shannon$data,x=variable,y="value",
          color=variable,
          scales="free") +
  stat_compare_means(comparisons = combinations, label="p.signif",method="wilcox.test",p.adjust.methods = "fdr") +
  theme(strip.background = element_rect(colour = "white", fill = "white"),
        axis.text.x=element_text(angle = 45, hjust = 1)) +
  labs(main="Infant faeces",x="",y="Shannon diversity") +
  #theme_classic() +
  scale_color_manual(values=col_fil),
  overwrite = T)

alpha.shannon.faec.days
```

### BMI

```{r}

ps <- subset_samples(PSB, Sample.type == "Faecal sample")
#ps@sam_data$BMI <- factor(ps@sam_data$BMI , levels = c("3","30","60","90"))

#Normalize to mean read count
total = median(sample_sums(ps))
standf = function(x, t=total) round(t * (x / sum(x)))
ps.rel = transform_sample_counts(ps, standf)

#Set desired alpha diversity metrics

alpha_met <- c("Observed", "Shannon", "InvSimpson")

###########Boxplots

##Time

variable <- c("BMI.group")

alpha.metrics.faec.bmi <- plot_richness(ps.rel, measures=alpha_met , x=variable, color=variable) + 
  geom_boxplot(alpha=0.1) +
  theme_classic() +
  scale_color_manual(values=col_fil) +
  theme(strip.background = element_rect(colour = "white", fill = "white")
  ) +
  labs(x="BMI",
       y="Alpha diversity") + 
  ggtitle("Infant faeces") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))


combinations <- combn(levels(alpha.metrics.faec.bmi$data[[variable]]), 2, simplify = FALSE) %>%
  lapply(., function(x) c(x[1], x[2]))

ggboxplot(alpha.metrics.faec.bmi$data,x=variable,y="value",
          color="variable",
          facet.by = "variable",
          scales="free") +
  stat_compare_means(comparisons = combinations, label="p.format",method="wilcox.test",p.adjust.methods = "fdr") +
  theme_classic() +
  scale_color_manual(values=col_fil) 

alpha.metrics.faec.bmi
```

```{r}
##Richness - all time points

richness = estimate_richness(ps.rel, measures = alpha_met)

#Time

rich <- cbind(richness, variable = sample_data(ps.rel)$BMI.group)

TukeyHSD(aov(Shannon ~ variable, rich))

TukeyHSD(aov(Observed ~ variable, rich))
```

#### Only shannon

```{r}

ps <- subset_samples(PSB, Sample.type == "Faecal sample")
#ps@sam_data$BMI <- factor(ps@sam_data$BMI , levels = c("3","30","60","90"))

#Normalize to mean read count
total = median(sample_sums(ps))
standf = function(x, t=total) round(t * (x / sum(x)))
ps.rel = transform_sample_counts(ps, standf)

#Set desired alpha diversity metrics

alpha_met <- c("Shannon")

###########Boxplots

##Time

variable <- c("BMI.group")

alpha.metrics.faec.bmi.shannon <- plot_richness(ps.rel, measures=alpha_met , x=variable, color=variable) + 
  geom_boxplot(alpha=0.1) +
  theme_classic() +
  scale_color_manual(values=col_fil) +
  theme(strip.background = element_rect(colour = "white", fill = "white"),
        axis.text.x=element_text(angle = 45, hjust = 1),
  ) +
  ggtitle("Infant faeces") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))


combinations <- combn(levels(alpha.metrics.faec.bmi$data[[variable]]), 2, simplify = FALSE) %>%
  lapply(., function(x) c(x[1], x[2]))

sl(alpha.shannon.faec.bmi,
   ggboxplot(alpha.metrics.faec.bmi.shannon$data,x=variable,y="value",
          color=variable,
          scales="free") +
  stat_compare_means(comparisons = combinations, label="p.signif",method="wilcox.test",p.adjust.methods = "fdr") +
  theme(strip.background = element_rect(colour = "white", fill = "white"),
        axis.text.x=element_text(angle = 45, hjust = 1)) +
  labs(main="Infant faeces",x="",y="Shannon diversity") +
  #theme_classic() +
  scale_color_manual(values=col_fil),
  overwrite = T)

alpha.shannon.faec.bmi
```

### Alfa diversity vs maternal age

```{r}
#Check whether alfa div and sample dataframes match
length(intersect(rownames(rich),rownames(ps@sam_data)))==nrow(rich)

df <- dplyr::select(rich,Shannon) %>%
  cbind(.,'Maternal age'=ps@sam_data$Maternal.age) %>%
  cbind(.,'BMI group'=ps@sam_data$BMI.group)

ggplot(df,aes(x=`Maternal age`,y=Shannon,color=`BMI group`)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_cor(method = "pearson") 
```

# Beta diversity

## Sample types

### Bray curtis

```{r, message = F}
ps <- PSB.CSS

#Do PCoA ordination based on Bray-curtis distance
GP.ord <- ordinate(ps, "PCoA", "bray")

#####################Run + Batch

#PSB.CSS@sam_data$Run <- as.factor(PSB.CSS@sam_data$Run)

#PSB.CSS.all@sam_data$Batch <- as.factor(ps@sam_data$Batch)

beta.type = plot_ordination(ps, GP.ord, color="Sample.type") + 
  stat_ellipse(geom = "polygon", level = 0.95, fill = NA, size = 1) +
  #scale_fill_manual(values= col_fil) +
  #scale_color_manual(values = col_fil) +
  theme_classic() + 
  theme(text = element_text(size = 12, family = 'serif'),
        axis.line=element_line(size=1),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        #axis.text.x=element_text(angle = 45, hjust = 1),
        strip.text.x = element_text(size = 12, angle = 0),
        legend.text = element_text(size = 12)
        #legend.position = "none"
  ) #+facet_wrap(~Sample.type + Batch,scales="free")

beta.type
```

### Weighted unifrac

```{r, message = F}
ps <- PSB.CSS

#Do PCoA ordination based on Bray-curtis distance
sl(GP.ord.wunifrac,ordinate(ps, "PCoA", "wunifrac"))

#####################Run + Batch

#PSB.CSS@sam_data$Run <- as.factor(PSB.CSS@sam_data$Run)

#PSB.CSS.all@sam_data$Batch <- as.factor(ps@sam_data$Batch)

beta.type = plot_ordination(ps, GP.ord.wunifrac, color="Sample.type") + 
  stat_ellipse(geom = "polygon", level = 0.95, fill = NA, size = 1) +
  #scale_fill_manual(values= col_fil) +
  #scale_color_manual(values = col_fil) +
  theme_classic() + 
  theme(text = element_text(size = 12, family = 'serif'),
        axis.line=element_line(size=1),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        #axis.text.x=element_text(angle = 45, hjust = 1),
        strip.text.x = element_text(size = 12, angle = 0),
        legend.text = element_text(size = 12)
        #legend.position = "none"
  ) #+facet_wrap(~Sample.type + Batch,scales="free")

beta.type
```

### NMDS

```{r, message = F}
load("0.3_deduplicated_and_decontaminated.RData")

ps <- PSB

#Do PCoA ordination based on Bray-curtis distance
#GP.ord <- ordinate(ps, "PCoA", "bray")
sl(GP.ord.nmds,ordinate(ps, "NMDS"))

#####################Run + Batch

#PSB.CSS@sam_data$Run <- as.factor(PSB.CSS@sam_data$Run)

#PSB.CSS.all@sam_data$Batch <- as.factor(ps@sam_data$Batch)

beta.type = plot_ordination(ps, GP.ord.nmds, color="Sample.type") + 
  stat_ellipse(geom = "polygon", level = 0.95, fill = NA, size = 1) +
  #scale_fill_manual(values= col_fil) +
  #scale_color_manual(values = col_fil) +
  theme_classic() + 
  theme(text = element_text(size = 12, family = 'serif'),
        axis.line=element_line(size=1),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        #axis.text.x=element_text(angle = 45, hjust = 1),
        strip.text.x = element_text(size = 12, angle = 0),
        legend.text = element_text(size = 12)
        #legend.position = "none"
  ) #+facet_wrap(~Sample.type + Batch,scales="free")

beta.type
```

## Milk

```{r}
ps <- subset_samples(PSB.CSS, Sample.type == "Milk" & Days %in% c("3","30","60","90"))
ps@sam_data$Days <- factor(ps@sam_data$Days, levels = c("3","30","60","90"))

#Do PCoA ordination based on Bray-curtis distance
GP.ord <- ordinate(ps, "PCoA", "bray")
```


### Days

```{r, message = F}

beta.milk.days = plot_ordination(ps, GP.ord, color="Days") + 
  stat_ellipse(geom = "polygon", level = 0.95, fill = NA, size = 1) +
  #scale_fill_manual(values= col_fil) +
  scale_color_manual(values = col_fil,name="Days after birth") +
  theme(strip.background = element_rect(colour = "white", fill = "white"),
        plot.title = element_text(hjust = 0.5)
        #legend.title = element_text("Days after birth")
  ) +
  ggtitle("Milk") +
  ggplot2::annotate("text", x = 0, y = 0.47, label = "All: p = 0.0002          Pairwise: p < 0.02 ")

sl(beta.milk.days,beta.milk.days,overwrite=T)

beta.milk.days
```

Permanova statistics

```{r, eval = T, include = T, echo = F, message = F}
PSB.CSS.no.na <- subset_samples(ps, !is.na(Days))

D_BC <- phyloseq::distance(PSB.CSS.no.na, "bray")

adonis <- adonis2(D_BC ~ Days, data = data.frame(sample_data(PSB.CSS.no.na)),permutations = 9999)
adonis
```

Pairwise PERMANOVA

```{r}
RVAideMemoire::pairwise.perm.manova(D_BC ,data.frame(sample_data(PSB.CSS.no.na))$Days,nperm=9999)
```

### BMI

```{r, message = F}
beta.milk.BMI = plot_ordination(ps, GP.ord, color="BMI.group") + 
  stat_ellipse(geom = "polygon", level = 0.95, fill = NA, size = 1) +
  #scale_fill_manual(values= col_fil) +
  scale_color_manual(values = col_fil,name="Pre-pregnancy BMI") +
  theme(strip.background = element_rect(colour = "white", fill = "white"),
        plot.title = element_text(hjust = 0.5)
        #legend.title = element_text("BMI after birth")
  ) +
  ggtitle("Milk") +
  ggplot2::annotate("text", x = 0, y = 0.47, label = "NW-OW:***NW-OB:***OW-OB:NS")

sl(beta.milk.BMI,beta.milk.BMI,overwrite=T)

beta.milk.BMI
```

Permanova statistics

```{r, eval = T, include = T, echo = F, message = F}
PSB.CSS.no.na <- subset_samples(ps, !is.na(BMI.group))

D_BC <- phyloseq::distance(PSB.CSS.no.na, "bray")

adonis <- adonis2(D_BC ~ BMI.group, data = data.frame(sample_data(PSB.CSS.no.na)),permutations = 9999)
adonis
```

Pairwise PERMANOVA

```{r}
RVAideMemoire::pairwise.perm.manova(D_BC ,data.frame(sample_data(PSB.CSS.no.na))$BMI.group,nperm=9999)
```

## Infant faeces

```{r}
ps <- subset_samples(PSB.CSS, Sample.type == "Faecal sample" & Days %in% c("3","30","60","90"))
ps@sam_data$Days <- factor(ps@sam_data$Days, levels = c("3","30","60","90"))

#Do PCoA ordination based on Bray-curtis distance
GP.ord <- ordinate(ps, "PCoA", "bray")
```


### Days

```{r, message = F}

beta.faec.days = plot_ordination(ps, GP.ord, color="Days") + 
  stat_ellipse(geom = "polygon", level = 0.95, fill = NA, size = 1) +
  #scale_fill_manual(values= col_fil) +
  scale_color_manual(values = col_fil[2:4],name="Days after birth") +
  theme(strip.background = element_rect(colour = "white", fill = "white"),
        plot.title = element_text(hjust = 0.5)
        #legend.title = element_text("Days after birth")
  ) +
  ggtitle("Infant faeces") +
  ggplot2::annotate("text", x = 0, y = 0.40, label = "All: p = 0.0003      Pairwise: p < 0.008")

sl(beta.faec.days,beta.faec.days,overwrite=T)

beta.faec.days
```

Permanova statistics

```{r, eval = T, include = T, echo = F, message = F}
PSB.CSS.no.na <- subset_samples(ps, !is.na(Days))

D_BC <- phyloseq::distance(PSB.CSS.no.na, "bray")

adonis <- adonis2(D_BC ~ Day, data = data.frame(sample_data(PSB.CSS.no.na)),permutations = 9999)
adonis
```

Pairwise PERMANOVA

```{r}
RVAideMemoire::pairwise.perm.manova(D_BC ,data.frame(sample_data(PSB.CSS.no.na))$Days,nperm=9999)
```

#### Bifidobacteria days

```{r}
ps <- subset_samples(PSB.CSS, Sample.type == "Faecal sample" & Days %in% c("3","30","60","90"))
ps@sam_data$Days <- factor(ps@sam_data$Days, levels = c("3","30","60","90"))
ps <- subset_taxa(ps,Genus=="Bifidobacterium")

#Do PCoA ordination based on Bray-curtis distance
GP.ord <- ordinate(ps, "PCoA", "bray")


beta.faec.bifi.days = plot_ordination(ps, GP.ord, color="Days") + 
  stat_ellipse(geom = "polygon", level = 0.95, fill = NA, size = 1) +
  #scale_fill_manual(values= col_fil) +
  scale_color_manual(values = col_fil,name="Days after birth") +
  theme(strip.background = element_rect(colour = "white", fill = "white"),
        plot.title = element_text(hjust = 0.5)
        #legend.title = element_text("Days after birth")
  ) +
  ggtitle("Infant faeces") +
  ggplot2::annotate("text", x = -0.35, y = 0.40, label = "p = 0.0008")

sl(beta.faec.bifi.days,beta.faec.days,overwrite=T)

beta.faec.bifi.days
```

Permanova statistics

```{r, eval = T, include = T, echo = F, message = F}

D_BC <- phyloseq::distance(ps, "bray")

adonis <- adonis2(D_BC ~ Day, data = data.frame(sample_data(ps)),permutations = 9999)
adonis
```

Pairwise PERMANOVA

```{r}
RVAideMemoire::pairwise.perm.manova(D_BC ,data.frame(sample_data(ps))$Day,nperm=9999)
```

### BMI

```{r, message = F}
beta.faec.BMI = plot_ordination(ps, GP.ord, color="BMI.group") + 
  stat_ellipse(geom = "polygon", level = 0.95, fill = NA, size = 1) +
  #scale_fill_manual(values= col_fil) +
  scale_color_manual(values = col_fil,name="Pre-pregnancy BMI") +
  theme(strip.background = element_rect(colour = "white", fill = "white"),
        plot.title = element_text(hjust = 0.5)
        #legend.title = element_text("BMI after birth")
  ) +
  ggtitle("Infant faeces") +
  ggplot2::annotate("text", x = 0.05, y = 0.50, label = "NW-OW:***NW-OB:***OW-OB:NS")

sl(beta.faec.BMI,beta.faec.BMI,overwrite=T)

beta.faec.BMI
```

Permanova statistics

```{r, eval = T, include = T, echo = F, message = F}
PSB.CSS.no.na <- subset_samples(ps, !is.na(BMI.group))

D_BC <- phyloseq::distance(PSB.CSS.no.na, "bray")

adonis <- adonis2(D_BC ~ BMI.group, data = data.frame(sample_data(PSB.CSS.no.na)),permutations = 9999)
adonis
```
Pairwise PERMANOVA

```{r}
RVAideMemoire::pairwise.perm.manova(D_BC ,data.frame(sample_data(PSB.CSS.no.na))$BMI.group,nperm=9999)
```

#### Bifidobacteria days

```{r}
ps <- subset_samples(PSB.CSS, Sample.type == "Faecal sample" & Days %in% c("3","30","60","90"))
ps@sam_data$Days <- factor(ps@sam_data$Days, levels = c("3","30","60","90"))
ps <- subset_taxa(ps,Genus=="Bifidobacterium")

#Do PCoA ordination based on Bray-curtis distance
GP.ord <- ordinate(ps, "PCoA", "bray")


beta.faec.bifi.bmi = plot_ordination(ps, GP.ord, color="BMI.group") + 
  stat_ellipse(geom = "polygon", level = 0.95, fill = NA, size = 1) +
  #scale_fill_manual(values= col_fil) +
  scale_color_manual(values = col_fil,name="Pre-pregnancy BMI") +
  theme(strip.background = element_rect(colour = "white", fill = "white"),
        plot.title = element_text(hjust = 0.5)
        #legend.title = element_text("Days after birth")
  ) +
  ggtitle("Infant faeces") +
  ggplot2::annotate("text", x = -0.40, y = 0.40, label = "p = 0.0001")

sl(beta.faec.bifi.bmi,beta.faec.bifi.bmi,overwrite=T)

beta.faec.bifi.bmi
```

Permanova statistics

```{r, eval = T, include = T, echo = F, message = F}

D_BC <- phyloseq::distance(ps, "bray")

adonis <- adonis2(D_BC ~ BMI.group, data = data.frame(sample_data(ps)),permutations = 9999)
adonis
```
Pairwise PERMANOVA

```{r}
RVAideMemoire::pairwise.perm.manova(D_BC ,data.frame(sample_data(ps))$BMI.group,nperm=9999)
```

## Time + Batch effects

```{r, message = F}
load("0.3_deduplicated_and_decontaminated.RData")

ps <- subset_samples(PSB.CSS, Sample.type %in% c("Milk","Faecal sample") & Days %in% c("3","30","60","90"))
ps@sam_data$Days <- as.factor(ps@sam_data$Days)


#Do PCoA ordination based on Bray-curtis distance
GP.ord <- ordinate(ps, "PCoA", "bray")

#####################Run + Batch

#PSB.CSS@sam_data$Run <- as.factor(PSB.CSS@sam_data$Run)

#PSB.CSS.all@sam_data$Batch <- as.factor(ps@sam_data$Batch)

beta.days = plot_ordination(ps, GP.ord, color="Days") + 
  stat_ellipse(geom = "polygon", level = 0.95, fill = NA, size = 1) +
  scale_color_manual(values= col_fil) +
  theme_classic() + 
  theme(text = element_text(size = 12, family = 'serif'),
        axis.line=element_line(size=1),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        #axis.text.x=element_text(angle = 45, hjust = 1),
        strip.text.x = element_text(size = 12, angle = 0),
        legend.text = element_text(size = 12)
        #legend.position = "none"
  ) +facet_wrap(~Sample.type ,scales="free")

beta.days

#batch

beta.batch = plot_ordination(ps, GP.ord, color="Batch") + 
  stat_ellipse(geom = "polygon", level = 0.95, fill = NA, size = 1) +
  scale_color_manual(values= col_fil) +
  theme_classic() + 
  theme(text = element_text(size = 12, family = 'serif'),
        axis.line=element_line(size=1),
        axis.text=element_text(size = 12, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        strip.background = element_rect(colour = "white", fill = "white"),
        #axis.text.x=element_text(angle = 45, hjust = 1),
        strip.text.x = element_text(size = 12, angle = 0),
        legend.text = element_text(size = 12)
        #legend.position = "none"
  ) +facet_wrap(~Sample.type ,scales="free")

beta.batch
```



# Related vs unrelated sample similarity

Make table of bray curtis distances between each sample with column specifying sample type and whether individuals are related or unrelated.

```{r, eval = T, include = T, echo = F, message = F, warning = F, fig.fullwidth = TRUE}
set.seed(1015)

#############Generate Bray curtis pairwise distance data frame:

#PSB.CSS.sub <- subset_samples(PSB.CSS.sub, Days %in% c("60",NA))

metadata <- meta.dat


################### Calculate the distance objects

## Weighted Unifrac

sl(wunifracDist,phyloseq::distance(PSB.CSS, "wunifrac"),overwrite = FALSE)

# Coerce distance object into matrix
wunifracDistMtx <- as.matrix(wunifracDist)

# Melt and remove values of 0 to eliminate distances of each sample to itself
wunifracDistMtx_melt <- subset(reshape2::melt(wunifracDistMtx, 
                                                value.name = "WUnifrac_Distance"))
                                 #WUnifrac_Distance != 0)

## Unifrac distance - binary

sl(unifracDist,phyloseq::distance(PSB, "unifrac"),overwrite = FALSE)

# Coerce distance object into matrix
unifracDistMtx <- as.matrix(unifracDist)

# Melt and remove values of 0 to eliminate distances of each sample to itself
unifracDistMtx_melt <- subset(reshape2::melt(unifracDistMtx, 
                                                value.name = "unifrac_Distance"))
                                 #unifrac_Distance != 0)

## Bray curtis

sl(brayDist,phyloseq::distance(PSB.CSS, "bray", binary = FALSE),overwrite = FALSE)

# Coerce distance object into matrix
brayDistMtx <- as.matrix(brayDist)

# Melt and remove values of 0 to eliminate distances of each sample to itself
brayDistMtx_melt <- subset(reshape2::melt(brayDistMtx, 
                                                value.name = "bray_Distance"))
                                 #bray_Distance != 0)

## Jaccard - binary

sl(jaccardDist,phyloseq::distance(PSB, "jaccard", binary = TRUE),overwrite = FALSE)

# Coerce distance object into matrix
jaccardDistMtx <- as.matrix(jaccardDist)

# Melt and remove values of 0 to eliminate distances of each sample to itself
jaccardDistMtx_melt <- subset(reshape2::melt(jaccardDistMtx, 
                                                value.name = "jaccard_Distance"))
                                 #jaccard_Distance != 0)


################# Combine the matrices 

CombDistMtx_melt <- tibble::add_column(wunifracDistMtx_melt,
                                       Unifrac_Distance = unifracDistMtx_melt$unifrac_Distance,
                                       BrayCurtis_Distance = brayDistMtx_melt$bray_Distance,
                                       Jaccard_Distance =  jaccardDistMtx_melt$jaccard_Distance,
                                       .after = "WUnifrac_Distance")

################ Relation distance function

get_relation_distances <- function(dist_matrix = "CombDistMtx_melt"){

#remove values of 0 to eliminate distances of each sample to itself

CombDistMtx_melt <- subset(CombDistMtx_melt,Unifrac_Distance != 0)

# Change Var1 and Var2 from factors to characters for easier use with dplyr
factorCols <- c("Var1", "Var2")
CombDistMtx_melt[factorCols] <- lapply(CombDistMtx_melt[factorCols], 
                                             as.character)

# Add columns for later subsetting when plotting.  This information will get
#   pulled from the metadata (metadataStool).  Within the *DistMtx_melt data 
#   frames, we need a column indicating whether Var1 and Var2 are infants,
#   mothers, and whether they're related.
CombDistDF <- CombDistMtx_melt

CombDistDF <- CombDistDF %>%
  mutate(Var1_id = metadata[CombDistDF$Var1, "ID"],
         Var2_id = metadata[CombDistDF$Var2, "ID"],
         Var1_sample = metadata[CombDistDF$Var1, "Sample.type"],
         Var2_sample = metadata[CombDistDF$Var2, "Sample.type"],
         Var1_days = metadata[CombDistDF$Var1, "Days"],
         Var2_days = metadata[CombDistDF$Var2, "Days"],
         Var1_batch = metadata[CombDistDF$Var1, "Batch"],
         Var2_batch = metadata[CombDistDF$Var2, "Batch"],
         Var1_run = metadata[CombDistDF$Var1, "Run"],
         Var2_run = metadata[CombDistDF$Var2, "Run"]
         )

CombDistDF <- CombDistDF %>% mutate(Relation = ifelse(Var1_id == Var2_id, "Related", "Unrelated")) %>%
  unite("sample_comparison",c(Var1_sample,Var2_sample),sep=" vs ",remove=FALSE) %>%
  unite("comparison_relation",c(sample_comparison,Relation), sep="_", remove=FALSE)

# Wihin each gorup select a subset of unrelated samples equal to the number of related samples

datalist = list()
comparisons = unique(CombDistDF$sample_comparison)

for (i in 1:length(comparisons)){
  compar = comparisons[i]
  print(compar)
  related.df <- subset(CombDistDF, sample_comparison == compar & Relation == "Related")
  unrelated.df <- subset(CombDistDF, sample_comparison == compar & Relation == "Unrelated")
  unrelated.df.sub <- dplyr::sample_n(unrelated.df, size = nrow(related.df))
  datalist[[i]] <- rbind(related.df,unrelated.df.sub)
}

CombDistDF = do.call(rbind, datalist)
return(CombDistDF)
}

sl(dist.relation,get_relation_distances(CombDistMtx_melt),overwrite = T)

dist.relation <- dist.relation %>%
  dplyr::rowwise() %>%
  dplyr::mutate(Timepoint = ifelse(Var1_days == Var2_days & !is.na(Var1_days), Var1_days, NA))

#Remove skin swabs due to too few samples
dist.relation.subset <- subset(dist.relation, Var1_sample != "Skin swab" & Var2_sample != "Skin swab")
```

## All samples

### Bray curtis distance boxplots

```{r, eval = T, include = T, echo = F, message = F}
Pairelation.all.bray <- ggplot(
  subset(dist.relation.subset,BrayCurtis_Distance!=0),  
    aes(Relation,BrayCurtis_Distance,fill = Relation)) +
  geom_boxplot(outlier.shape=NA, show.legend = FALSE)+
  #  geom_jitter(width = 0.2) +
  ylab("Bray-curtis distance") +
  #xlab("Infant faecal samples") +
  theme_classic() +
  scale_fill_aaas() +
  stat_compare_means(method = "anova" ,label.x.npc = 0, label.y.npc = 0.2)+
  facet_wrap("sample_comparison", scales = "free")

Pairelation.all.bray
```

### Weighted Unifrac distance boxplots

```{r, eval = T, include = T, echo = F, message = F}
Pairelation.all.wunifrac <- 
  ggplot(subset(dist.relation.subset,WUnifrac_Distance!=0),
         aes(Relation,WUnifrac_Distance,fill = Relation)) +
  geom_boxplot(outlier.shape=NA, show.legend = FALSE)+
  #  geom_jitter(width = 0.2) +
  ylab("Bray-curtis distance") +
  #xlab("Infant faecal samples") +
  theme_classic() +
  scale_fill_aaas() +
  stat_compare_means(method = "anova" ,label.x.npc = 0, label.y.npc = 0.2)+
  facet_wrap("sample_comparison", scales = "free")

Pairelation.all.wunifrac
```

### Jaccard (binary) distance boxplots

```{r, eval = T, include = T, echo = F, message = F}
Pairelation.all.jaccard <- 
  ggplot(subset(dist.relation.subset,Jaccard_Distance!=0),
         aes(Relation,Jaccard_Distance,fill = Relation)) +
  geom_boxplot(outlier.shape=NA, show.legend = FALSE)+
  #  geom_jitter(width = 0.2) +
  ylab("Bray-curtis distance") +
  #xlab("Infant faecal samples") +
  theme_classic() +
  scale_fill_aaas() +
  stat_compare_means(method = "anova" ,label.x.npc = 0, label.y.npc = 0.2)+
  facet_wrap("sample_comparison", scales = "free")

Pairelation.all.jaccard
```

### Unifrac (binary phylogenetic) distance boxplots

```{r, eval = T, include = T, echo = F, message = F}

Pairelation.all.unifrac <- 
  ggplot(subset(dist.relation.subset,Unifrac_Distance!=0),
         aes(Relation,Unifrac_Distance,fill = Relation)) +
  geom_boxplot(outlier.shape=NA, show.legend = FALSE)+
  #  geom_jitter(width = 0.2) +
  ylab("Bray-curtis distance") +
  #xlab("Infant faecal samples") +
  theme_classic() +
  scale_fill_aaas() +
  stat_compare_means(method = "anova" ,label.x.npc = 0, label.y.npc = 0.2)+
  facet_wrap("sample_comparison", scales = "free")

Pairelation.all.unifrac
```

## Over time

### Milk vs Faecal

#### Wunifrac

```{r, eval = T, include = T, echo = F, message = F}
CombDistDF.subset.milk <- subset(dist.relation.subset, sample_comparison %in% c("Milk vs Faecal sample")  & Timepoint != "NA" )

Pairelation.all.wunifrac <- 
  ggplot(CombDistDF.subset.milk,aes(Relation,WUnifrac_Distance,fill = Relation)) +
  geom_boxplot(outlier.shape=NA, show.legend = FALSE)+
  #  geom_jitter(width = 0.2) +½
  #ylab("Bray-curtis distance") +
  #xlab("Infant faecal samples") +
  theme_classic() +
  scale_fill_aaas() +
  stat_compare_means(method = "anova" ,label.x.npc = 0, label.y.npc = 0.4)+
  facet_wrap("Timepoint", scales = "free")

Pairelation.all.wunifrac
```
### Bray

```{r, eval = T, include = T, echo = F, message = F}
CombDistDF.subset.milk <- subset(dist.relation.subset, sample_comparison %in% c("Milk vs Faecal sample")  & Timepoint != "NA" )

Pairelation.all.bray <- 
  ggplot(CombDistDF.subset.milk,aes(Relation,BrayCurtis_Distance,fill = Relation)) +
  geom_boxplot(outlier.shape=NA, show.legend = FALSE)+
  #  geom_jitter(width = 0.2) +½
  #ylab("Bray-curtis distance") +
  #xlab("Infant faecal samples") +
  theme_classic() +
  scale_fill_aaas() +
  stat_compare_means(method = "anova" ,label.x.npc = 0, label.y.npc = 0.4)+
  facet_wrap("Timepoint", scales = "free")

Pairelation.all.bray
```

## Batch differences

### Milk

#### Batch

```{r, eval = T, include = T, echo = F, message = F}
CombDistDF.subset.milk <- subset(dist.relation.subset, sample_comparison %in% c("Milk vs Milk") )

Pairelation.milk.wunifrac.batch <- 
  ggplot(CombDistDF.subset.milk,aes(Relation,WUnifrac_Distance,fill = Relation)) +
  geom_boxplot(outlier.shape=NA, show.legend = FALSE)+
  #  geom_jitter(width = 0.2) +
  ylab("Bray-curtis distance") +
  #xlab("Infant faecal samples") +
  theme_classic() +
  scale_fill_aaas() +
  stat_compare_means(method = "anova" ,label.x.npc = 0, label.y.npc = 0.4)+
  facet_wrap("Var1_batch")

Pairelation.milk.wunifrac.batch

```

#### Run

```{r, eval = T, include = T, echo = F, message = F}
Pairelation.milk.wunifrac.run <- 
  ggplot(CombDistDF.subset.milk,aes(Relation,WUnifrac_Distance,fill = Relation)) +
  geom_boxplot(outlier.shape=NA, show.legend = FALSE)+
  #  geom_jitter(width = 0.2) +
  ylab("Bray-curtis distance") +
  #xlab("Infant faecal samples") +
  theme_classic() +
  scale_fill_aaas() +
  stat_compare_means(method = "anova" ,label.x.npc = 0, label.y.npc = 0.4)+
  facet_wrap(vars(Var1_batch,Var1_run))

Pairelation.milk.wunifrac.run
```

#### Check milk batch 2 info

```{r, eval = F, include = T, echo = F, message = F}
milk.b2 <- subset_samples(PSB,Sample.type == "Milk" & Batch == "2")

mb2.meta <- data.frame(sample_data(milk.b2))
```

### Faeces

```{r eval=T, echo=F, message=FALSE, include=T}
CombDistDF.subset.faeces <- subset(dist.relation.subset, sample_comparison %in% c("Faecal sample vs Faecal sample"))

Pairelation.all.wunifrac <- 
  ggplot(CombDistDF.subset.faeces,aes(Relation,WUnifrac_Distance,fill = Relation)) +
  geom_boxplot(outlier.shape=NA, show.legend = FALSE)+
  #  geom_jitter(width = 0.2) +
  ylab("Bray-curtis distance") +
  #xlab("Infant faecal samples") +
  theme_classic() +
  scale_fill_aaas() +
  stat_compare_means(method = "anova" ,label.x.npc = 0, label.y.npc = 0.2)

Pairelation.all.wunifrac
```

# Effect sizes

## Milk

First otu and metadata matrices are formated, and a suitable subset of variables to examine are selected

```{r, eval = T, include = F, echo = F, message = F}
######Importing-Data
# PSB.R <- microbiome::transform(
#   subset_samples(PSB, Sample.type == "Milk"),
#   "compositional")

PSB.R <- subset_samples(PSB.CSS, Sample.type == "Milk")

import.otu.table <- as.data.frame(otu_table(
  PSB.R
  )
  ) #%>%
  #mutate_all(as.numeric) #Convert to numeric where possible

cure.data = as.data.frame(as.matrix(PSB.R@sam_data))

#import.taxonomy = read.table('virome_taxonomy_LCA_known.txt', header = T)

rownames <- rownames(cure.data)

#Remove NA columns
cure.data <- cure.data[ , ! apply( cure.data , 2 , function(x) all(is.na(x)) ) ]

rownames(cure.data) <- rownames

#cure.data <- cure.data %>% mutate_if(is.character, as.numeric)

X8.nonrarefied = t(import.otu.table)                                                                  ###%%% non-rarefied


###rarefying|subsampling reads in every sample
#barplot(colSums(import.otu.table))
#min(colSums(import.otu.table))

#rare8 = rarefy(as.matrix(t(import.otu.table)), sample = 5000)
rare8 = t(import.otu.table)

X8.rarefied = as.data.frame(rare8)                                                                 ###%%% rarefied

###Subset metadata to match remaining samples after rarefaction

Y.all = cure.data[rownames(X8.rarefied),]

###Log tranformation
#X = log10(X8.rarefied+1)
X <- X8.rarefied

#Select relevant numeric column

Y <- dplyr::select(Y.all,all_of(effect.size.variables)) 
  
#Y <- Y.all %>% dplyr::select_if(is.factor)

#,"Weight_sept_2020", "Glucose_sept_GGT" , "Insulin_sept_GGT", "Cpeptide_sept_GGT", "HbA1c", "Glucagon_sept_GGT", "HOMA_IR"

colnames(Y) #Check remaining columns

###Select only numeric metadata culumns

Y <- Y %>%
  mutate(
    DNA.conc = as.numeric(DNA.conc),
    #Batch = as.numeric(Batch),
    #Days = as.numeric(Days),
    Maternal.age = as.numeric(Maternal.age),
    ppBMI = as.numeric(ppBMI),
    Gestational.age = as.numeric(Gestational.age),
    Birth.weight = as.numeric(Birth.weight),
    delta.whz = as.numeric(delta.whz),
    BMI.delta.3m = as.numeric(BMI.delta.3m),
    delta.waz = as.numeric(delta.waz),
    delta.haz = as.numeric(delta.waz)
  )

#Set numeric column class
Y <- Y %>% mutate_if(is.character,as.factor)

str(Y)

#colnames(Y) #Check remaining columns

##vifcor
##collinearity

#Select relevant columns for that are non-collinear

```

### Db anova


#### Bray curttis - constrained by batch + time

```{r, eval = T, include = T, echo = F, message = F, warning =  F, fig.width=10, fig.height=6, fig.fullwidth = TRUE}
###############Anova on capscale - all variables

anova.m <- matrix(nrow = 0, ncol =6) #Create matrix for results
colnames(anova.m) <- c("Df","SumOfSqs","F","Pr.F","Residual.Df","Residual.SumOfSqs")                
for (i in colnames(dplyr::select(Y,-c("Batch","Days")))){
  X.sub <- X[!is.na(Y[[i]]), ]
  Y.sub <- Y[!is.na(Y[[i]]), ]
  anova <- as.matrix(anova.cca(capscale(X.sub ~ Y.sub[,i] + Days + Condition(Batch), Y.sub, dist = "bray"),permutations = how(nperm=9999)))
  anova <- c(anova[1,],anova[2,1:2])
  anova.m <-rbind(anova.m,anova)
}

rownames(anova.m) <- colnames(dplyr::select(Y,-c("Batch","Days")))

#Prepare for plotting

anova.dat <- anova.m %>%
  as.data.frame() %>%
  tibble::add_column(Variable = rownames(anova.m),.name_repair = "minimal") %>%
  arrange(F)

#Add adjusted p-values
anova.dat$Pr.F.adj <- p.adjust(anova.dat$Pr.F, method = 'holm')

#Show table
#knitr::kable(anova.dat)

#Calculate sum of squares
anova.dat$R2 <- anova.dat$SumOfSqs / (anova.dat$SumOfSqs + anova.dat$Residual.SumOfSqs)

anova.dat <- anova.dat %>%
  arrange(R2)

##Fancy plot

anova.melt <- tidyr::pivot_longer(anova.dat,cols=c("R2"), names_to='measure', values_to="value") %>%
  add_significance(p.col = "Pr.F.adj", output.col = "PrF.sym",
                   cutpoints = c(0, 1e-04, 0.001, 0.01, 0.05, 0.1,1),
                   symbols = c("****", "***", "**", "*", "\U0387", "")
                   )

anova.melt.bac.milk.batch.time <- anova.melt

permanova.bac.milk.batch.time <- ggbarplot(anova.melt,
                                         x="Variable",
                                         y="value",
                                         #color = "Variable",
                                         orientation = c("horizontal"),
                                         color = NA,
                                         fill = "measure",
                                         position=position_dodge(0.7),
                                         label = anova.melt$PrF.sym,
                                         lab.size = 9,
                                         lab.vjust = 0.7,
                                         lab.hjust = -0.3
                                         
) +
  #stat_pvalue_manual(
  #  adonis.melt,  label = "PrF.sym", tip.length = 0.01
  #) +
  #geom_hline(yintercept = 0.05) +
  ggtitle("Capscale dbRDA effect size: all variables") +
  ylab(expression(R^2~"value")) +
  scale_fill_jco()

permanova.bac.milk.batch.time

```

#### Bray curttis -constrained by batch

```{r, eval = F, include = T, echo = F, message = F, warning =  F, fig.width=10, fig.height=6, fig.fullwidth = TRUE}
###############Anova on capscale - all variables

anova.m <- matrix(nrow = 0, ncol =6) #Create matrix for results
colnames(anova.m) <- c("Df","SumOfSqs","F","Pr.F","Residual.Df","Residual.SumOfSqs")                
for (i in colnames(dplyr::select(Y,-c("Batch")))){
  X.sub <- X[!is.na(Y[[i]]), ]
  Y.sub <- Y[!is.na(Y[[i]]), ]
  anova <- as.matrix(anova.cca(capscale(X.sub ~ Y.sub[,i] + Condition(Batch), Y.sub, dist = "bray"),permutations = how(nperm=9999)))
  anova <- c(anova[1,],anova[2,1:2])
  anova.m <-rbind(anova.m,anova)
}

rownames(anova.m) <- colnames(dplyr::select(Y,-c("Batch")))

#Prepare for plotting

anova.dat <- anova.m %>%
  as.data.frame() %>%
  tibble::add_column(Variable = rownames(anova.m),.name_repair = "minimal") %>%
  arrange(F)

#Add adjusted p-values
anova.dat$Pr.F.adj <- p.adjust(anova.dat$Pr.F, method = 'holm')

#Show table
#knitr::kable(anova.dat)

#Calculate sum of squares
anova.dat$R2 <- anova.dat$SumOfSqs / (anova.dat$SumOfSqs + anova.dat$Residual.SumOfSqs)

anova.dat <- anova.dat %>%
  arrange(R2)

##Fancy plot

anova.melt <- tidyr::pivot_longer(anova.dat,cols=c("R2"), names_to='measure', values_to="value") %>%
  add_significance(p.col = "Pr.F.adj", output.col = "PrF.sym",
                   cutpoints = c(0, 1e-04, 0.001, 0.01, 0.05, 0.1,1),
                   symbols = c("****", "***", "**", "*", "\U0387", "")
                   )

anova.melt.bac <- anova.melt

fancy.anova.bar.milk.bray.batch.day <- ggbarplot(anova.melt,
                                         x="Variable",
                                         y="value",
                                         #color = "Variable",
                                         orientation = c("horizontal"),
                                         color = NA,
                                         fill = "measure",
                                         position=position_dodge(0.7),
                                         label = anova.melt$PrF.sym,
                                         lab.size = 9,
                                         lab.vjust = 0.7,
                                         lab.hjust = -0.3
                                         
) +
  #stat_pvalue_manual(
  #  adonis.melt,  label = "PrF.sym", tip.length = 0.01
  #) +
  #geom_hline(yintercept = 0.05) +
  ggtitle("Capscale dbRDA effect size: all variables") +
  ylab(expression(R^2~"value")) +
  scale_fill_jco()

fancy.anova.bar.milk.bray.batch.day

```

#### Bray curttis - constrained by batch + time + Secretor

```{r, eval = T, include = T, echo = F, message = F, warning =  F, fig.width=10, fig.height=6, fig.fullwidth = TRUE}
###############Anova on capscale - all variables

anova.m <- matrix(nrow = 0, ncol =6) #Create matrix for results
colnames(anova.m) <- c("Df","SumOfSqs","F","Pr.F","Residual.Df","Residual.SumOfSqs")               

Y.sec <- Y[!is.na(Y[["Secretor"]]),]
X.sec <- X[!is.na(Y[["Secretor"]]),]              

for (i in colnames(dplyr::select(Y.sec,-c("Batch","Days","DNA.conc","Secretor","Lewis")))){
  X.sub <- X.sec[!is.na(Y.sec[[i]]), ]
  Y.sub <- Y.sec[!is.na(Y.sec[[i]]), ]
  anova <- as.matrix(anova.cca(capscale(X.sub ~ Y.sub[,i] + Days + Condition(Batch+Secretor+Lewis), Y.sub, dist = "bray"),permutations = how(nperm=9999)))
  anova <- c(anova[1,],anova[2,1:2])
  anova.m <-rbind(anova.m,anova)
}

rownames(anova.m) <- colnames(dplyr::select(Y,-c("Batch","Days","DNA.conc","Secretor","Lewis")))

#Prepare for plotting

anova.dat <- anova.m %>%
  as.data.frame() %>%
  tibble::add_column(Variable = rownames(anova.m),.name_repair = "minimal") %>%
  arrange(F)

#Add adjusted p-values
anova.dat$Pr.F.adj <- p.adjust(anova.dat$Pr.F, method = 'holm')

#Show table
#knitr::kable(anova.dat)

#Calculate sum of squares
anova.dat$R2 <- anova.dat$SumOfSqs / (anova.dat$SumOfSqs + anova.dat$Residual.SumOfSqs)

anova.dat <- anova.dat %>%
  arrange(R2)

##Fancy plot

anova.melt <- tidyr::pivot_longer(anova.dat,cols=c("R2"), names_to='measure', values_to="value") %>%
  add_significance(p.col = "Pr.F.adj", output.col = "PrF.sym",
                   cutpoints = c(0, 1e-04, 0.001, 0.01, 0.05, 0.1,1),
                   symbols = c("****", "***", "**", "*", "\U0387", "")
                   )

anova.melt.bac.milk.batch.time.sec <- anova.melt

permanova.bac.milk.batch.time.sec <- ggbarplot(anova.melt,
                                         x="Variable",
                                         y="value",
                                         #color = "Variable",
                                         orientation = c("horizontal"),
                                         color = NA,
                                         fill = "measure",
                                         position=position_dodge(0.7),
                                         label = anova.melt$PrF.sym,
                                         lab.size = 9,
                                         lab.vjust = 0.7,
                                         lab.hjust = -0.3
                                         
) +
  #stat_pvalue_manual(
  #  adonis.melt,  label = "PrF.sym", tip.length = 0.01
  #) +
  #geom_hline(yintercept = 0.05) +
  ggtitle("Capscale dbRDA effect size: all variables") +
  ylab(expression(R^2~"value")) +
  scale_fill_jco()

permanova.bac.milk.batch.time.sec

```


#### Bray curttis - constrained by batch + time + BMI

```{r, eval = T, include = T, echo = F, message = F, warning =  F, fig.width=10, fig.height=6, fig.fullwidth = TRUE}
###############Anova on capscale - all variables

anova.m <- matrix(nrow = 0, ncol =6) #Create matrix for results
colnames(anova.m) <- c("Df","SumOfSqs","F","Pr.F","Residual.Df","Residual.SumOfSqs")                
for (i in colnames(dplyr::select(Y,-c("Batch","Days","BMI")))){
  X.sub <- X[!is.na(Y[[i]]), ]
  Y.sub <- Y[!is.na(Y[[i]]), ]
  anova <- as.matrix(anova.cca(capscale(X.sub ~ Y.sub[,i] + Days + Condition(Batch+BMI), Y.sub, dist = "bray"),permutations = how(nperm=999)))
  anova <- c(anova[1,],anova[2,1:2])
  anova.m <-rbind(anova.m,anova)
}

rownames(anova.m) <- colnames(dplyr::select(Y,-c("Batch","Days","BMI")))

#Prepare for plotting

anova.dat <- anova.m %>%
  as.data.frame() %>%
  tibble::add_column(Variable = rownames(anova.m),.name_repair = "minimal") %>%
  arrange(F)

#Add adjusted p-values
anova.dat$Pr.F.adj <- p.adjust(anova.dat$Pr.F, method = 'holm')

#Show table
#knitr::kable(anova.dat)

#Calculate sum of squares
anova.dat$R2 <- anova.dat$SumOfSqs / (anova.dat$SumOfSqs + anova.dat$Residual.SumOfSqs)

anova.dat <- anova.dat %>%
  arrange(R2)

##Fancy plot

anova.melt <- tidyr::pivot_longer(anova.dat,cols=c("R2"), names_to='measure', values_to="value") %>%
  add_significance(p.col = "Pr.F.adj", output.col = "PrF.sym",
                   cutpoints = c(0, 1e-04, 0.001, 0.01, 0.05, 0.1,1),
                   symbols = c("****", "***", "**", "*", "\U0387", "")
                   )

anova.melt.bac.milk.batch.time.bmi <- anova.melt

permanova.bac.milk.batch.time.bmi <- ggbarplot(anova.melt,
                                         x="Variable",
                                         y="value",
                                         #color = "Variable",
                                         orientation = c("horizontal"),
                                         color = NA,
                                         fill = "measure",
                                         position=position_dodge(0.7),
                                         label = anova.melt$PrF.sym,
                                         lab.size = 9,
                                         lab.vjust = 0.7,
                                         lab.hjust = -0.3
                                         
) +
  #stat_pvalue_manual(
  #  adonis.melt,  label = "PrF.sym", tip.length = 0.01
  #) +
  #geom_hline(yintercept = 0.05) +
  ggtitle("Capscale dbRDA effect size: all variables") +
  ylab(expression(R^2~"value")) +
  scale_fill_jco()

permanova.bac.milk.batch.time.bmi

```





### RDA and dbRDA(Bray) variable selection - milk (for comparison)

```{r, eval = T, include = T, echo = F, message = F, warning =  F, fig.width=10, fig.height=4, fig.fullwidth = TRUE}

#RDA
Y.na <- na.omit(Y) #%>% select(rev(names(.)))
X.na <- X[rownames(Y.na),]

mod0 <- rda(X.na ~ 1 + Condition(Batch+Days), Y.na)  # Model with intercept only
mod1 <- rda(X.na ~ . + Condition(Batch+Days), Y.na)  # Model with all eX.naplanatory variables

## With scope present, the default direction is "both"
sl(ord.scope.both.milk,ordistep(mod0, scope = formula(mod1), direction="both", permutations = how(nperm = 999)),overwrite=FALSE)

ord.scope.both.milk$anova
ord.scope.both.milk$call

#Write our new model
spe.rda.signif <- rda(formula = X.na ~ Condition(Batch + Days) + Induction, data = Y.na)

RsquareAdj(spe.rda.signif)

anova.cca(spe.rda.signif, step = 1000)

anova.cca(spe.rda.signif, step = 1000, by = "term")

#dbRDA

mod0 <- dbrda(X.na ~ 1 + Condition(Batch+Days), Y.na, distance = "bray")  # Model with intercept only
mod1 <- dbrda(X.na ~ . + Condition(Batch+Days), Y.na, distance = "bray")  # Model with all eX.naplanatory variables

## With scope present, the default direction is "both"
sl(ord.scope.both.bray.milk,ordistep(mod0, scope = formula(mod1), direction="both",permutations = how(nperm = 999)),overwrite=FALSE)

ord.scope.both.bray.milk$anova
ord.scope.both.bray.milk$call

#Write our new model
spe.rda.signif <- dbrda(formula = X.na ~ Condition(Batch + Days) + BMI + Breastfeeding.issues + 
    Maternal.age, data = Y.na, distance = "bray")

RsquareAdj(spe.rda.signif)

anova.cca(spe.rda.signif, step = 1000)

anova.cca(spe.rda.signif, step = 1000, by = "term")

```

## Faeces

```{r, eval = T, include = F, echo = F, message = F}

######Importing-Data

# PSB.R <- microbiome::transform(
#   subset_samples(PSB, Sample.type == "Faecal sample"),
#   "compositional")

PSB.R <- subset_samples(PSB.CSS, Sample.type == "Faecal sample")


import.otu.table <- as.data.frame(otu_table(
  PSB.R
  )
  ) #%>%
  #mutate_all(as.numeric) #Convert to numeric where possible

cure.data = as.data.frame(as.matrix(PSB.R@sam_data))

#import.taxonomy = read.table('virome_taxonomy_LCA_known.txt', header = T)

rownames <- rownames(cure.data)

#Remove NA columns
cure.data <- cure.data[ , ! apply( cure.data , 2 , function(x) all(is.na(x)) ) ]

rownames(cure.data) <- rownames

#cure.data <- cure.data %>% mutate_if(is.character, as.numeric)

X8.nonrarefied = t(import.otu.table)                                                                  ###%%% non-rarefied


###rarefying|subsampling reads in every sample
#barplot(colSums(import.otu.table))
#min(colSums(import.otu.table))

#rare8 = rarefy(as.matrix(t(import.otu.table)), sample = 5000)
rare8 = t(import.otu.table)

X8.rarefied = as.data.frame(rare8)                                                                 ###%%% rarefied

###Subset metadata to match remaining samples after rarefaction

Y.all = cure.data[rownames(X8.rarefied),]

###Log tranformation
#X = log10(X8.rarefied+1)
X <- X8.rarefied

#Select relevant numeric column

Y <- dplyr::select(Y.all,all_of(effect.size.variables)) 
  
#Y <- Y.all %>% dplyr::select_if(is.factor)

#,"Weight_sept_2020", "Glucose_sept_GGT" , "Insulin_sept_GGT", "Cpeptide_sept_GGT", "HbA1c", "Glucagon_sept_GGT", "HOMA_IR"

colnames(Y) #Check remaining columns

###Select only numeric metadata culumns

Y <- Y %>%
  mutate(
    DNA.conc = as.numeric(DNA.conc),
    # Batch = as.numeric(Batch),
    # Days = as.numeric(Days),
    Maternal.age = as.numeric(Maternal.age),
    ppBMI = as.numeric(ppBMI),
    Gestational.age = as.numeric(Gestational.age),
    Birth.weight = as.numeric(Birth.weight),
    delta.whz = as.numeric(delta.whz),
    BMI.delta.3m = as.numeric(BMI.delta.3m),
    delta.waz = as.numeric(delta.waz),
    delta.haz = as.numeric(delta.waz)
  )

#Set numeric column class
Y <- Y %>% mutate_if(is.character,as.factor)

str(Y)

#colnames(Y) #Check remaining columns

##vifcor
##collinearity

#Select relevant columns for that are non-collinear

```

### Db anova


#### Bray-curtis - contrined by time + batch

```{r, eval = T, include = T, echo = F, message = F, warning =  F, fig.width=10, fig.height=6, fig.fullwidth = TRUE}
###############Anova on capscale - all variables

anova.m <- matrix(nrow = 0, ncol =6) #Create matrix for results
colnames(anova.m) <- c("Df","SumOfSqs","F","Pr.F","Residual.Df","Residual.SumOfSqs")                
for (i in colnames(dplyr::select(Y,-c("Batch","Days","DNA.conc")))){
   X.sub <- X[!is.na(Y[[i]]), ]
  Y.sub <- Y[!is.na(Y[[i]]), ]
  anova <- as.matrix(anova.cca(capscale(X.sub ~ na.omit(Y.sub[,i]) + Days + Condition(Batch), Y.sub, dist = "bray"),permutations = how(nperm=999)))
  anova <- c(anova[1,],anova[2,1:2])
  anova.m <-rbind(anova.m,anova)
}

rownames(anova.m) <- colnames(dplyr::select(Y,-c("Batch","Days","DNA.conc")))

#Prepare for plotting

anova.dat <- anova.m %>%
  as.data.frame() %>%
  tibble::add_column(Variable = rownames(anova.m),.name_repair = "minimal") %>%
  arrange(F)

#Add adjusted p-values
anova.dat$Pr.F.adj <- p.adjust(anova.dat$Pr.F, method = 'holm')

#Show table
#knitr::kable(anova.dat)

#Calculate sum of squares
anova.dat$R2 <- anova.dat$SumOfSqs / (anova.dat$SumOfSqs + anova.dat$Residual.SumOfSqs)

anova.dat <- anova.dat %>%
  arrange(R2)

##Fancy plot

anova.melt <- tidyr::pivot_longer(anova.dat,cols=c("R2"), names_to='measure', values_to="value") %>%
  add_significance(p.col = "Pr.F.adj", output.col = "PrF.sym",
                   cutpoints = c(0, 1e-04, 0.001, 0.01, 0.05, 0.1,1),
                   symbols = c("****", "***", "**", "*", "\U0387", "")
                   )

anova.melt.bac.faeces.batch.time <- anova.melt

permanova.bac.faeces.batch.time <- ggbarplot(anova.melt,
                                         x="Variable",
                                         y="value",
                                         #color = "Variable",
                                         orientation = c("horizontal"),
                                         color = NA,
                                         fill = "measure",
                                         position=position_dodge(0.7),
                                         label = anova.melt$PrF.sym,
                                         lab.size = 9,
                                         lab.vjust = 0.7,
                                         lab.hjust = -0.3
                                         
) +
  #stat_pvalue_manual(
  #  adonis.melt,  label = "PrF.sym", tip.length = 0.01
  #) +
  #geom_hline(yintercept = 0.05) +
  ggtitle("Capscale dbRDA effect size: all variables") +
  ylab(expression(R^2~"value")) +
  scale_fill_jco()

permanova.bac.faeces.batch.time

```

#### Bray curttis - constrained by batch + time + BMI

```{r, eval = T, include = T, echo = F, message = F, warning =  F, fig.width=10, fig.height=6, fig.fullwidth = TRUE}
###############Anova on capscale - all variables

anova.m <- matrix(nrow = 0, ncol =6) #Create matrix for results
colnames(anova.m) <- c("Df","SumOfSqs","F","Pr.F","Residual.Df","Residual.SumOfSqs")                
for (i in colnames(dplyr::select(Y,-c("Batch","Days","BMI")))){
  X.sub <- X[!is.na(Y[[i]]), ]
  Y.sub <- Y[!is.na(Y[[i]]), ]
  anova <- as.matrix(anova.cca(capscale(X.sub ~ Y.sub[,i] + Days + Condition(Batch+BMI), Y.sub, dist = "bray"),permutations = how(nperm=999)))
  anova <- c(anova[1,],anova[2,1:2])
  anova.m <-rbind(anova.m,anova)
}

rownames(anova.m) <- colnames(dplyr::select(Y,-c("Batch","Days","BMI")))

#Prepare for plotting

anova.dat <- anova.m %>%
  as.data.frame() %>%
  tibble::add_column(Variable = rownames(anova.m),.name_repair = "minimal") %>%
  arrange(F)

#Add adjusted p-values
anova.dat$Pr.F.adj <- p.adjust(anova.dat$Pr.F, method = 'holm')

#Show table
#knitr::kable(anova.dat)

#Calculate sum of squares
anova.dat$R2 <- anova.dat$SumOfSqs / (anova.dat$SumOfSqs + anova.dat$Residual.SumOfSqs)

anova.dat <- anova.dat %>%
  arrange(R2)

##Fancy plot

anova.melt <- tidyr::pivot_longer(anova.dat,cols=c("R2"), names_to='measure', values_to="value") %>%
  add_significance(p.col = "Pr.F.adj", output.col = "PrF.sym",
                   cutpoints = c(0, 1e-04, 0.001, 0.01, 0.05, 0.1,1),
                   symbols = c("****", "***", "**", "*", "\U0387", "")
                   )

anova.melt.bac.faec.batch.time.bmi <- anova.melt

permanova.bac.faec.batch.time.bmi <- ggbarplot(anova.melt,
                                         x="Variable",
                                         y="value",
                                         #color = "Variable",
                                         orientation = c("horizontal"),
                                         color = NA,
                                         fill = "measure",
                                         position=position_dodge(0.7),
                                         label = anova.melt$PrF.sym,
                                         lab.size = 9,
                                         lab.vjust = 0.7,
                                         lab.hjust = -0.3
                                         
) +
  #stat_pvalue_manual(
  #  adonis.melt,  label = "PrF.sym", tip.length = 0.01
  #) +
  #geom_hline(yintercept = 0.05) +
  ggtitle("Capscale dbRDA effect size: all variables") +
  ylab(expression(R^2~"value")) +
  scale_fill_jco()

permanova.bac.faec.batch.time.bmi

```

#### Bray curttis - constrained by batch + time + Secretor

```{r, eval = T, include = T, echo = F, message = F, warning =  F, fig.width=10, fig.height=6, fig.fullwidth = TRUE}
###############Anova on capscale - all variables

anova.m <- matrix(nrow = 0, ncol =6) #Create matrix for results
colnames(anova.m) <- c("Df","SumOfSqs","F","Pr.F","Residual.Df","Residual.SumOfSqs")               

Y.sec <- Y[!is.na(Y[["Secretor"]]),]
X.sec <- X[!is.na(Y[["Secretor"]]),]              

for (i in colnames(dplyr::select(Y.sec,-c("Batch","Days","DNA.conc","Secretor","Lewis")))){
  X.sub <- X.sec[!is.na(Y.sec[[i]]), ]
  Y.sub <- Y.sec[!is.na(Y.sec[[i]]), ]
  anova <- as.matrix(anova.cca(capscale(X.sub ~ Y.sub[,i] + Days + Condition(Batch+Secretor+Lewis), Y.sub, dist = "bray"),permutations = how(nperm=999)))
  anova <- c(anova[1,],anova[2,1:2])
  anova.m <-rbind(anova.m,anova)
}

rownames(anova.m) <- colnames(dplyr::select(Y,-c("Batch","Days","DNA.conc","Secretor","Lewis")))

#Prepare for plotting

anova.dat <- anova.m %>%
  as.data.frame() %>%
  tibble::add_column(Variable = rownames(anova.m),.name_repair = "minimal") %>%
  arrange(F)

#Add adjusted p-values
anova.dat$Pr.F.adj <- p.adjust(anova.dat$Pr.F, method = 'holm')

#Show table
#knitr::kable(anova.dat)

#Calculate sum of squares
anova.dat$R2 <- anova.dat$SumOfSqs / (anova.dat$SumOfSqs + anova.dat$Residual.SumOfSqs)

anova.dat <- anova.dat %>%
  arrange(R2)

##Fancy plot

anova.melt <- tidyr::pivot_longer(anova.dat,cols=c("R2"), names_to='measure', values_to="value") %>%
  add_significance(p.col = "Pr.F.adj", output.col = "PrF.sym",
                   cutpoints = c(0, 1e-04, 0.001, 0.01, 0.05, 0.1,1),
                   symbols = c("****", "***", "**", "*", "\U0387", "")
                   )

anova.melt.bac.faec.batch.time.sec <- anova.melt

permanova.bac.faec.batch.time.sec <- ggbarplot(anova.melt,
                                         x="Variable",
                                         y="value",
                                         #color = "Variable",
                                         orientation = c("horizontal"),
                                         color = NA,
                                         fill = "measure",
                                         position=position_dodge(0.7),
                                         label = anova.melt$PrF.sym,
                                         lab.size = 9,
                                         lab.vjust = 0.7,
                                         lab.hjust = -0.3
                                         
) +
  #stat_pvalue_manual(
  #  adonis.melt,  label = "PrF.sym", tip.length = 0.01
  #) +
  #geom_hline(yintercept = 0.05) +
  ggtitle("Capscale dbRDA effect size: all variables") +
  ylab(expression(R^2~"value")) +
  scale_fill_jco()

permanova.bac.faec.batch.time.sec

```

#### Bray-curtis - constrained by batch

```{r, eval = F, include = T, echo = F, message = F, warning =  F, fig.width=10, fig.height=6, fig.fullwidth = TRUE}
###############Anova on capscale - all variables

anova.m <- matrix(nrow = 0, ncol =6) #Create matrix for results
colnames(anova.m) <- c("Df","SumOfSqs","F","Pr.F","Residual.Df","Residual.SumOfSqs")                
for (i in colnames(dplyr::select(Y,-c("Batch")))){
  X.sub <- X[!is.na(Y[[i]]), ]
  Y.sub <- Y[!is.na(Y[[i]]), ]
  anova <- as.matrix(anova.cca(capscale(X.sub ~ na.omit(Y.sub[,i]) + Condition(Batch), Y.sub, dist = "bray"),permutations = how(nperm=999)))
  anova <- c(anova[1,],anova[2,1:2])
  anova.m <-rbind(anova.m,anova)
}

# 
# i <- "Gestational.age"
# 
# X.sub <- X[!is.na(Y[[i]]), ]
# Y.sub <- Y[!is.na(Y[[i]]), ]
#   
# anova <- as.matrix(anova.cca(capscale(X.sub ~ na.omit(Y[,i]) + Condition(Batch), Y.sub, dist = "bray"),permutations = how(nperm=99999)))
# anova <- c(anova[1,],anova[2,1:2])
# anova.m <-rbind(anova.m,anova)


rownames(anova.m) <- colnames(dplyr::select(Y,-c("Batch")))

#Prepare for plotting

anova.dat <- anova.m %>%
  as.data.frame() %>%
  tibble::add_column(Variable = rownames(anova.m),.name_repair = "minimal") %>%
  arrange(F)

#Add adjusted p-values
anova.dat$Pr.F.adj <- p.adjust(anova.dat$Pr.F, method = 'holm')

#Show table
#knitr::kable(anova.dat)

#Calculate sum of squares
anova.dat$R2 <- anova.dat$SumOfSqs / (anova.dat$SumOfSqs + anova.dat$Residual.SumOfSqs)

anova.dat <- anova.dat %>%
  arrange(R2)

##Fancy plot

anova.melt <- tidyr::pivot_longer(anova.dat,cols=c("R2"), names_to='measure', values_to="value") %>%
  add_significance(p.col = "Pr.F.adj", output.col = "PrF.sym",
                   cutpoints = c(0, 1e-04, 0.001, 0.01, 0.05, 0.1,1),
                   symbols = c("****", "***", "**", "*", "\U0387", "")
                   )

anova.melt.bac <- anova.melt

fancy.anova.bar.bac <- ggbarplot(anova.melt,
                                         x="Variable",
                                         y="value",
                                         #color = "Variable",
                                         orientation = c("horizontal"),
                                         color = NA,
                                         fill = "measure",
                                         position=position_dodge(0.7),
                                         label = anova.melt$PrF.sym,
                                         lab.size = 9,
                                         lab.vjust = 0.7,
                                         lab.hjust = -0.3
                                         
) +
  #stat_pvalue_manual(
  #  adonis.melt,  label = "PrF.sym", tip.length = 0.01
  #) +
  #geom_hline(yintercept = 0.05) +
  ggtitle("Capscale dbRDA effect size: all variables") +
  ylab(expression(R^2~"value")) +
  scale_fill_jco()

fancy.anova.bar.bac

```



## Combined effect size

### Time+Batch

```{r, eval = T, include = T, echo = F, message = F, warning =  F, fig.width=10, fig.height=4, fig.fullwidth = TRUE}
anova.melt.bac.milk.batch.time$Component <- "Milk bacteria"
anova.melt.bac.faeces.batch.time$Component <- "Faecal bacteria"

# Rename variables

anova.melt.comb <- rbind(anova.melt.bac.faeces.batch.time,anova.melt.bac.milk.batch.time) # bar chart order follows top df

anova.melt.comb$Variable.clean <- gsub("_"," ",anova.melt.comb$Variable)


esize.bac.feaces.milk <- ggbarplot(anova.melt.comb,
                                         x="Variable",
                                         y="value",
                                         ylab = FALSE,
                                         fill = "Component",
                                         orientation = c("horizontal"),
                                         #fill = "measure",
                                         position=position_dodge(0.7),
                                         label = anova.melt.comb$PrF.sym,
                                         lab.size = 9,
                                         lab.vjust = 0.8,
                                         lab.hjust = -0.3
                                         
) +
  #stat_pvalue_manual(
  #  adonis.melt,  label = "PrF.sym", tip.length = 0.01
  #) +
  #geom_hline(yintercept = 0.05) +
  #ggtitle("Capscale dbRDA effect size: all variables") +
  ylab(expression(R^2~"value")) +
  scale_fill_jco()

esize.bac.feaces.milk

save(anova.melt.comb,file="results/effect_size_bac_milk_faeces.RData")
```

### Time+Batch+Secretor

```{r, eval = T, include = T, echo = F, message = F, warning =  F, fig.width=10, fig.height=4, fig.fullwidth = TRUE}
anova.melt.bac.milk.batch.time.sec$Component <- "Milk bacteria"
anova.melt.bac.faec.batch.time.sec$Component <- "Faecal bacteria"

# Rename variables

anova.melt.comb <- rbind(anova.melt.bac.milk.batch.time.sec,anova.melt.bac.faec.batch.time.sec) # bar chart order follows top df

anova.melt.comb$Variable.clean <- gsub("_"," ",anova.melt.comb$Variable)


esize.bac.feaces.milk.sec <- ggbarplot(anova.melt.comb,
                                         x="Variable",
                                         y="value",
                                         ylab = FALSE,
                                         fill = "Component",
                                         orientation = c("horizontal"),
                                         #fill = "measure",
                                         position=position_dodge(0.7),
                                         label = anova.melt.comb$PrF.sym,
                                         lab.size = 9,
                                         lab.vjust = 0.8,
                                         lab.hjust = -0.3
                                         
) +
  #stat_pvalue_manual(
  #  adonis.melt,  label = "PrF.sym", tip.length = 0.01
  #) +
  #geom_hline(yintercept = 0.05) +
  #ggtitle("Capscale dbRDA effect size: all variables") +
  ylab(expression(R^2~"value")) +
  scale_fill_jco()

esize.bac.feaces.milk.sec

save(esize.bac.feaces.milk.sec,file="results/effect_size_bac_milk_faeces_secretor.RData")
```




# RDA plots (using microeco)

```{r}
meco_dat <- phyloseq2meco(PSB.CSS
  #transform_sample_counts(PSB, function(x) x / sum(x))
  )

#Tidy up taxonomy
meco_dat$tax_table <- tidy_taxonomy(meco_dat$tax_table)

meco_dat$tax_table

# Remove "[" and "]" from the Genus column
meco_dat$tax_table$Genus <- gsub("\\[|\\]", "", meco_dat$tax_table$Genus)

# Remove everything after the first "_" (including the "_")
meco_dat$tax_table$Genus <- gsub("(([^_]*_){2}[^_]*)_*.*", "\\1", meco_dat$tax_table$Genus)

meco_dat$tax_table$Genus

#Calculate relative abundences
meco_dat$cal_abund()

#Set days as factor
meco_dat$sample_table$Days <- as.factor(meco_dat$sample_table$Days)

#Calculta beta diversity metrics

meco_dat$cal_betadiv(unifrac = TRUE)

#Subset milk samples
meco_milk <- clone(meco_dat)

meco_milk$sample_table <- subset(meco_dat$sample_table, Sample.type == "Milk")

meco_milk$tidy_dataset()

#Subset faecal samples
meco_faec <- clone(meco_dat)

meco_faec$sample_table <- subset(meco_dat$sample_table, Sample.type == "Faecal sample")

meco_faec$tidy_dataset()

```

## Milk

### RDA plot

```{r, warning = F}
#Creating trans_env object 
# env <- select(meco_milk$sample_table,all_of(
#   c(effect.size.variables[-c(1:2)],"Secretor","Lewis")
#   )) %>%
#   mutate(Days = as.numeric(as.character(Days)),
#          Secretor = as.numeric(as.character(Secretor)),
#          Lewis = as.numeric(as.character(Lewis)),) %>%
#   select(-c("AB.infant")) %>%
#   select(where(is.numeric)) %>%
#   drop_na(Gestational.age) %>%
#   drop_na() %>%
#   select(where(~ !any(is.na(.))))

load("results/effect_size_bac_milk_faeces.RData")

#Select most significant variables
top.milk.vars <- anova.melt.comb %>%
 dplyr::filter(Component == "Milk bacteria") %>%
  arrange(-value) %>%
  slice_head(n = 10)



env <- select(meco_milk$sample_table,all_of(c(top.milk.vars$Variable,"Days"))) %>%
  mutate(Days = as.numeric(as.character(Days))
         ) %>%
         #Secretor = as.numeric(as.character(Secretor)),
         #Lewis = as.numeric(as.character(Lewis)),) %>%
  #select(-c("AB.infant")) %>%
  select(where(is.numeric)) %>%
  #drop_na(Gestational.age) %>%
  drop_na() %>%
  select(where(~ !any(is.na(.))))

colnames(env) <- str_replace_all(colnames(env), c("\\." = " ",
                                                              `BMI delta`="delta BMI",
                                                              delta="Δ",
                                                              `3m`="3 months",
                                                              waz = "WAZ",
                                                              whz = "WHZ",
                                                              haz = "HAZ",
                                                              AB="Antibiotics",
                                                              GDM="Gestational diabetes",
                                                              Secretor="Secretor status",
                                                              Lewis="Lewis status")
)
  
t1 <- trans_env$new(dataset = meco_milk,
                    add_data = env)
```

```{r, warning = F}
# use bray-curtis distance for dbRDA
t1$cal_ordination(method ="RDA", taxa_level = "Genus",scale_species = TRUE)
# t1$res_rda is the result list stored in the object
t1$trans_ordination(adjust_arrow_length = TRUE,
                    min_perc_env = 0.2,
                    max_perc_env = 1,
                    min_perc_tax = 2,
                    max_perc_tax = 4.5)
# t1$res_rda_trans is the transformed result for plotting
sl(rda.plot.milk,
   t1$plot_ordination(plot_color = "Days",
                   plot_type = c("point"),
                   env_text_size = 4,
                   taxa_text_size = 4,
                   taxa_text_color="darkred",
                   taxa_arrow_color="darkred",
                   point_alpha=0.2,
                   taxa_nudge_x = c(0, # Staph
                                    -5, # Lactobac
                                    -3, # Gemella
                                    0, # Strept
                                    0, # Bifido
                                    0, # Bacteroides
                                    0, #Pseudo
                                    4, # Blautia
                                    0, # Faecali
                                    0), #Acineto
                   taxa_nudge_y = c(-2, # Staph
                                    1, # Lactobac
                                    -1, # Gemella
                                    0.01, # Strept
                                    -5, # Bifido
                                    2, # Bacteroides
                                    1, #Pseudo
                                    0, # Blautia
                                    1, # Faecali
                                    0) #Acineto
                   )+
  ggrepel::geom_text_repel(force=100),
  overwrite = TRUE
)

rda.plot.milk

# plot(rda.plot.milk) +
#   ggrepel::geom_text_repel(force=100)

```

```{r,eval=FALSE}
t1$cal_ordination_anova()

head(t1$res_ordination_terms)
head(t1$res_ordination_axis)
```

```{r,eval=FALSE}
t1$cal_ordination_envfit()

head(t1$object$res_ordination_envfit)
```

## Faecal samples

Creating trans_env object 

```{r, warning = F}
# env <- select(meco_faec$sample_table,all_of(
#   c(effect.size.variables[-c(1:2)],"Secretor","Lewis")
#   )) %>%
#   mutate(Days = as.numeric(as.character(Days)),
#          Secretor = as.numeric(as.character(Secretor)),
#          Lewis = as.numeric(as.character(Lewis)),) %>%
#   select(-c("AB.infant")) %>%
#   select(where(is.numeric)) %>%
#   drop_na(Gestational.age) %>%
#   drop_na() %>%
#   select(where(~ !any(is.na(.))))

load("results/effect_size_bac_milk_faeces.RData")

#Select most significant variables
top.faec.vars <- anova.melt.comb %>%
 dplyr::filter(Component == "Faecal bacteria") %>%
  arrange(-value) %>%
  slice_head(n = 10)



env <- select(meco_faec$sample_table,all_of(c(top.faec.vars$Variable,"Days"))) %>%
  mutate(Days = as.numeric(as.character(Days)),
         #Secretor = as.numeric(as.character(Secretor))
         #Lewis = as.numeric(as.character(Lewis)),
         ) %>%
  #select(-c("AB.infant")) %>%
  select(where(is.numeric)) %>%
  #drop_na(Gestational.age) %>%
  drop_na() %>%
  select(where(~ !any(is.na(.))))

colnames(env) <- str_replace_all(colnames(env), c("\\." = " ",
                                                              `BMI delta`="delta BMI",
                                                              delta="Δ",
                                                              `3m`="3 months",
                                                              waz = "WAZ",
                                                              whz = "WHZ",
                                                              haz = "HAZ",
                                                              AB="Antibiotics",
                                                              GDM="Gestational diabetes",
                                                              Secretor="Secretor status",
                                                              Lewis="Lewis status")
)
  
t1 <- trans_env$new(dataset = meco_faec,
                    add_data = env)
```

### RDA plot

```{r, warning = F}
# use bray-curtis distance for dbRDA
t1$cal_ordination(method ="RDA",scale_species = TRUE)
# t1$res_rda is the result list stored in the object
t1$trans_ordination(adjust_arrow_length = TRUE,
                    min_perc_env = 0.2,
                    max_perc_env = 1.5,
                    min_perc_tax = 2,
                    max_perc_tax = 3.5)
# t1$res_rda_trans is the transformed result for plotting
sl(rda.plot.faec,
   t1$plot_ordination(plot_color = "Days",
                   plot_type = c("point"),
                   env_text_size = 4,
                   taxa_text_size = 4,
                   taxa_text_color="darkred",
                   taxa_arrow_color="darkred",
                   point_alpha=0.2,
                   taxa_nudge_x = c(0,#Bifido
                                    0, #Bacteroides
                                    0, #Parabac
                                    0, #Strepto
                                    0, #E. col
                                    0, #Clostri
                                    0, #Veillo
                                    -5, #Ruminococc
                                    0, #Klebsiella
                                    0), # Collinsella
                   taxa_nudge_y = c(-2,#Bifido
                                    5, #Bacteroides
                                    -3, #Parabac
                                    0, #Strepto
                                    -7, #E. col
                                    -1, #Clostri
                                    3, #Veillo
                                    -0.001, #Ruminococc
                                    2, #Klebsiella
                                    -1) # Collinsella
                   )+
  ggrepel::geom_label_repel(force=500),
  overwrite = TRUE
)

rda.plot.faec
```

# Upset plots

## Milk

### Over time

```{r}
load("0.3_deduplicated_and_decontaminated.RData")


#devtools::install_github("hms-dbmi/UpSetR")
library("MicrobiotaProcess")


ps <- subset_samples(PSB.rar, Sample.type == "Milk", )
ps@sam_data$Days <- as.factor(ps@sam_data$Days)

# vennlist <- get_vennlist(obj=ps, factorNames="Days")

upsetda <- get_upset(obj=ps, factorNames="Days")

# vennp <- venn.diagram(vennlist,
#                       height=5,
#                       width=5, 
#                       filename=NULL, 
#                       #fill=col_fil,
#                       #cat.col=col_fil,
#                       alpha = 0.85, 
#                       fontfamily = "serif",
#                       fontface = "bold",
#                       cex = 1.2,
#                       cat.cex = 1.3,
#                       cat.default.pos = "outer",
#                       cat.dist=0.1,
#                       margin = 0.1, 
#                       lwd = 3,
#                       lty ='dotted',
#                       imagetype = "svg")
# grid::grid.draw(vennp)

sl(upset.milk.time,
   upset(upsetda, sets=unique(as.vector(sample_data(ps)$Group)), 
      sets.bar.color = "#56B4E9",
      order.by = "freq", 
      empty.intersections = "on")
   ,overwrite = T)

upset.milk.time
```

## Faeces

### Over time

```{r}
load("0.3_deduplicated_and_decontaminated.RData")

library("MicrobiotaProcess")

ps <- subset_samples(PSB.rar, Sample.type == "Faecal sample", )
ps <- subset_samples(ps, !is.na(Days))
ps@sam_data$Days <- factor(ps@sam_data$Days,levels=c("30","60","90"))

# vennlist <- get_vennlist(obj=ps, factorNames="Days")
upsetda <- get_upset(obj=ps, factorNames="Days")

# vennp <- venn.diagram(vennlist,
#                       height=5,
#                       width=5, 
#                       filename=NULL, 
#                       #fill=col_fil,
#                       #cat.col=col_fil,
#                       alpha = 0.85, 
#                       fontfamily = "serif",
#                       fontface = "bold",
#                       cex = 1.2,
#                       cat.cex = 1.3,
#                       cat.default.pos = "outer",
#                       cat.dist=0.1,
#                       margin = 0.1, 
#                       lwd = 3,
#                       lty ='dotted',
#                       imagetype = "svg")
# grid::grid.draw(vennp)

sl(upset.faeces.time,
   upset(upsetda, sets=unique(as.vector(sample_data(ps)$Group)), 
      sets.bar.color = "#56B4E9",
      order.by = "freq", 
      empty.intersections = "on")
   ,overwrite = T)

upset.faeces.time

library(ggplotify)

cowplot::plot_grid(as.ggplot(upset.milk.time),as.ggplot(upset.faeces.time))
```

Unload MicrobiotaProcess to avoid phyloseq conflicts in other sections

```{r}
detach(package:MicrobiotaProcess, unload=TRUE)
```


# Session info

```{r}
sessionInfo()
```
