---
title: "3.2_Transfer_by_ASV"
author: "Rasmus Riemer Jakobsen"
date: "`r Sys.time()`"
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
link-citations: yes
---

# Load data

```{r setup, include=F, message=F, warning=F}
#Set working directory to script directory
#setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

subdir_name <- "3.2_Transfer_by_ASV"

#Load packages
source("Supplementary_scripts/package_load_install.R")
#Load utility functions
source("Supplementary_scripts/utils.R")

dir.create("results", showWarnings = FALSE, recursive = TRUE)

#source("Supplementary_scripts/create_stability.R")

#source("Phyloseq_import_CSS_16S_laca.R")
load("0.3_deduplicated_and_decontaminated.RData")


theme_set(theme_classic())

knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.fullwidth = TRUE,
                      root.dir = setwd(dirname(rstudioapi::getActiveDocumentContext()$path))) #Set working directory to script directory 

```

```{r}
source(here::here("Supplementary_scripts","package_load_install.R")) 
source(here::here("Supplementary_scripts","utils.R")) # please see the src-folder for some custom functions used in 
#source(here::here("src", "run_clusterprofiler.R"))
```

```{r}
ps <- PSB
```

```{r knit, echo=F, eval=FALSE}
rmarkdown::render(
  input = here(str_c(subdir_name, ".Rmd")),
  output_dir = here("results", "reports")
)
```




# Temporal persistence 

## Frequency over time by child id- Eval when data is updated

```{r, eval = F, message = F}
#Cannot be made into function due to multiple uses of "subset"

#Milk

PSB.sub <- PSB %>% subset_samples(Sample.type == "Milk")
Remove low prevalance AND abundance ASVs
# prev <- 2/nrow(PSB.sub@sam_data)
# PSB.sub = phyloseq_filter_prevalence(PSB.sub, prev.trh = prev, abund.trh = 1000, abund.type = "total", threshold_condition = "AND")

ps.sub <- subset_samples(PSB.sub, ID %in% PSB.sub@sam_data$ID[duplicated(PSB.sub@sam_data$ID)])

datalist = vector("list", length = nrow(ps.sub@otu_table)) #Create empty datalist

for (id in sort(unique(ps.sub@sam_data$ID))){
    print(id)
    ps <- subset_samples(ps.sub,ID == id) %>% ps_arrange(Days) #Arrange ps object by days  
    otu <- as.data.frame(otu_table(ps))
    otu <- otu %>%
    rowwise() %>%
    mutate(`freq` = sum(c_across(everything()) != 0)/ncol(otu))

    datalist[[paste0("freq_ID_",id,sep="")]] <- otu$freq
}

otu.freq <- dplyr::bind_cols(datalist) #Combine datalist entries to dataframe

otu.freq.mean <- otu.freq %>% 
          rowwise() %>% 
          mutate(freq.mean = mean(c_across(where(is.numeric))))

otu.freq.mean.nz <- otu.freq %>%
   mutate(across(where(is.numeric), ~na_if(., 0))) %>%
          rowwise() %>%
          mutate(freq.mean.nz = mean(c_across(where(is.numeric)),na.rm=TRUE))
  
otu.id <- rownames(ps@otu_table)

otu.freq.milk <- cbind(otu.id,otu.freq.mean,select(otu.freq.mean.nz,"freq.mean.nz"))
sl(otu.freq.milk, otu.freq.milk, overwrite = T) 

#Faeces

PSB.sub <- PSB %>% subset_samples(Sample.type == "Faecal sample")
ps.sub <- subset_samples(PSB.sub, ID %in% PSB.sub@sam_data$ID[duplicated(PSB.sub@sam_data$ID)])
Remove low prevalance AND abundance ASVs
# prev <- 2/nrow(PSB.sub@sam_data)
# PSB.sub = phyloseq_filter_prevalence(PSB.sub, prev.trh = prev, abund.trh = 1000, abund.type = "total", threshold_condition = "AND")

datalist = vector("list", length = nrow(ps.sub@otu_table)) #Create empty datalist

for (id in sort(unique(ps.sub@sam_data$ID))){
    print(id)
    ps <- subset_samples(ps.sub,ID == id) %>% ps_arrange(Days) #Arrange ps object by days  
    otu <- as.data.frame(otu_table(ps))
    otu <- otu %>%
    rowwise() %>%
    mutate(`freq` = sum(c_across(everything()) != 0)/ncol(otu))

    datalist[[paste0("freq_ID_",id,sep="")]] <- otu$freq
}

otu.freq <- dplyr::bind_cols(datalist) #Combine datalist entries to dataframe

otu.freq.mean <- otu.freq %>% 
          rowwise() %>% 
          mutate(freq.mean = mean(c_across(where(is.numeric))))

otu.freq.mean.nz <- otu.freq %>%
   mutate(across(where(is.numeric), ~na_if(., 0))) %>%
          rowwise() %>%
          mutate(freq.mean.nz = mean(c_across(where(is.numeric)),na.rm=TRUE))
  
otu.id <- rownames(ps@otu_table)

otu.freq.faeces <- cbind(otu.id,otu.freq.mean,select(otu.freq.mean.nz,"freq.mean.nz"))

sl(otu.freq.faeces, otu.freq.faeces, overwrite = T) 
```

## Freq mean distribution

```{r}
#Load frequency tables
sl(otu.freq.milk, otu.freq.milk, overwrite = FALSE) 
sl(otu.freq.faeces, otu.freq.faeces, overwrite = FALSE) 


frq.milk.hist <- ggplot(otu.freq.milk, aes(x=freq.mean.nz))+
  geom_histogram(bins=1000)+
  geom_density(aes(y = ..density..)) +
  ggtitle("Milk")+
  xlab("Mean intra-individual fequency")

frq.milk.hist

frq.faeces.hist <- ggplot(otu.freq.faeces, aes(x=freq.mean.nz))+
  geom_histogram(bins=1000)+
  geom_density(aes(y = ..density..)) +
  ggtitle("Faeces") 

ggarrange(frq.milk.hist,frq.faeces.hist)
```

## Freq vs mean abundance

```{r, warning = F}
#Load frequency tables
sl(otu.freq.milk, otu.freq.milk, overwrite = FALSE) 
sl(otu.freq.faeces, otu.freq.faeces, overwrite = FALSE) 

sl(PSB.abund.type,metagMisc::phyloseq_average(PSB,group = "Niche", avg_type = "acomp"), overwrite = FALSE)

otu.avg <- data.frame(otu_table(PSB.abund.type))
otu.avg <- rownames_to_column(otu.avg, var = "otu.id")

#Milk
freq.abund.milk <- left_join(otu.freq.milk,otu.avg, by = "otu.id")

frq.abund.plot.milk <- ggplot(freq.abund.milk,aes(x=freq.mean.nz,y=Milk))+
  geom_point() +
  geom_smooth()+
  ggtitle("Milk")+
  xlab("Mean intra-individual fequency")+
  ylab("Average relative abundance")

#Faeces
freq.abund.faeces <- left_join(otu.freq.faeces,otu.avg, by = "otu.id")

frq.abund.plot.faeces <- ggplot(freq.abund.faeces,aes(x=freq.mean.nz,y=Infant.faeces))+
  geom_point() +
  geom_smooth()+
  ggtitle("Infant faeces")+
  xlab("Mean intra-individual fequency")+
  ylab("Average relative abundance")
  
ggarrange(frq.abund.plot.milk,frq.abund.plot.faeces)
```

### Color by genus - very slow! only run for new data (output saved)

```{r, warning = F}
ps <- subset_samples(PSB,Sample.type == "Milk")

# prev <- 2/nrow(ps@sam_data)
# ps = phyloseq_filter_prevalence(ps, prev.trh = prev, abund.trh = 1000, abund.type = "total", threshold_condition = "AND")

otu.id <- rownames(ps@otu_table)

sl(tax_best_milk,
   get_tax_rank_best(ps,
                     n_show = 7),
  overwrite = FALSE)

tax_best_milk <- tax_best_milk %>% dplyr::rename(otu.id = "OTU")

freq.abund.milk.tax <- left_join(freq.abund.milk,
                                 distinct(tax_best_milk,otu.id, .keep_all = TRUE),
                                 ,by = "otu.id")

frq.abund.plot.milk.genus <- ggplot(freq.abund.milk.tax,aes(x=freq.mean.nz,y=Milk,color=best_genus_other))+
  geom_point() +
  scale_color_manual(values=rep(col_fil,5),name="Genus")+
  ggtitle("Milk")+
  xlab("Mean intra-individual fequency")+
  ylab("Mean relative abundance") 

#frq.abund.plot.milk.genus

frq.abund.plot.milk.family <- ggplot(freq.abund.milk.tax,aes(x=freq.mean.nz,y=Milk,color=best_family_other))+
  geom_point() +
  scale_color_manual(values=rep(col_fil,5),name="Family")+
  ggtitle("Milk")+
  xlab("Mean intra-individual fequency")+
  ylab("Mean relative abundance") 

#frq.abund.plot.milk.family

####################################Faeces
ps <- subset_samples(PSB,Sample.type == "Faecal sample")

# prev <- 2/nrow(ps@sam_data)
# ps = phyloseq_filter_prevalence(ps, prev.trh = prev, abund.trh = 1000, abund.type = "total", threshold_condition = "AND")

otu.id <- rownames(ps@otu_table)

sl(tax_best_faec,
   get_tax_rank_best(ps,
                     n_show = 7),
  overwrite = FALSE)

tax_best_faec <- tax_best_faec %>% dplyr::rename(otu.id = "OTU")

freq.abund.faeces.tax <- left_join(freq.abund.faeces,
                                   distinct(tax_best_faec,otu.id, .keep_all = TRUE),
                                   ,by = "otu.id",keep = )

frq.abund.plot.faeces.genus <- ggplot(freq.abund.faeces.tax,aes(x=freq.mean.nz,y=Infant.faeces,color=best_genus_other))+
  geom_point()+
  scale_color_manual(values=rep(col_fil,5),name="Genus")+
  ggtitle("Infant faeces")+
  xlab("Mean intra-individual fequency")+
  ylab("Mean relative abundance") 

#frq.abund.plot.faeces.genus

frq.abund.plot.faeces.family <- ggplot(freq.abund.faeces.tax,aes(x=freq.mean.nz,y=Infant.faeces,color=best_family_other))+
  geom_point() +
  scale_color_manual(values=rep(col_fil,5),name="Family")+
  ggtitle("Infant faeces")+
  xlab("Mean intra-individual fequency")+
  ylab("Mean relative abundance") 

#frq.abund.plot.faeces.family
  
(sl(freq.abun.tax.plots,
   ggarrange(frq.abund.plot.milk.genus,
             frq.abund.plot.milk.family,
             frq.abund.plot.faeces.genus,
             frq.abund.plot.faeces.family,
             nrow = 2),
   overwrite = T))

ggsave("results/freq.abun.tax.plots.pdf",width=20,height=20,limitsize = FALSE, units = "cm", dpi = 400)

#freq.abun.tax.plots
```

## Mean distribution

### Milk

#### By Genus

```{r, warning = F}
#Get order of means per group

bymean <- with(freq.abund.milk.tax, reorder(best_family_other, freq.mean.nz, sum))
#bymean

freq.abund.milk.tax %>%
  group_by(best_genus_other) %>%
  dplyr::summarise(mean = count(freq.mean.nz))


freq.abund.milk.tax$best_genus_other <- fct_reorder(.f = freq.abund.milk.tax$best_genus_other,
             .x = freq.abund.milk.tax$freq.mean.nz,
             .fun = mean,
             .desc = TRUE
             )


mean.freq.plot.milk.genus <- ggboxplot(freq.abund.milk.tax,
                                 x="best_genus_other",
                                 y="freq.mean.nz",
                                 fill = "best_genus_other",
                                 error.plot = "pointrange",
                                 add = "mean_sd",
                                 #legend = "none",
                                 palette = col_fil) +
  #stat_compare_means(method = "anova") +      # Add global p-value
  stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "Other") +
  theme_classic() +
  theme(legend.position = "none",
        strip.background = element_rect(colour = "white", fill = "white"),
        axis.text.x=element_text(angle = -45, vjust = 0.8, hjust = 0)
  ) +
  xlab("") +
  ylab("ASV persistence") +
  scale_y_continuous(expand = expansion(mult = 0.1))

sl(mean.freq.plot.milk.genus,mean.freq.plot.milk.genus,overwrite = T)

```

### Faeces

#### By genus

```{r, warning = F}
#Get order of means per group

bymean <- with(freq.abund.faeces.tax, reorder(best_genus_other, freq.mean.nz, sum))
#bymean

freq.abund.faeces.tax %>%
  group_by(best_genus_other) %>%
  dplyr::summarise(mean = count(freq.mean.nz))


freq.abund.faeces.tax$best_genus_other <- fct_reorder(.f = freq.abund.faeces.tax$best_genus_other,
             .x = freq.abund.faeces.tax$freq.mean.nz,
             .fun = mean,
             .desc = TRUE
             )


mean.freq.plot.faeces.genus <- ggboxplot(ungroup(freq.abund.faeces.tax),
                                 x="best_genus_other",
                                 y="freq.mean.nz",
                                 fill = "best_genus_other",
                                 error.plot = "pointrange",
                                 add = c("mean_sd"),
                                 #legend = "none",
                                 palette = col_fil) +
  #stat_compare_means(method = "anova") +      # Add global p-value
  stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "Other") +
  theme_classic() +
  theme(legend.position = "none",
        strip.background = element_rect(colour = "white", fill = "white"),
        axis.text.x=element_text(angle = -45, vjust = 0.8, hjust = 0)
  ) +
  ylim(0,1.05) +
  xlab("") +
  ylab("ASV persistence") +
  scale_y_continuous(expand = expansion(mult = 0.1))

# mean.freq.plot.faeces.genus

sl(mean.freq.plot.faeces.genus,mean.freq.plot.faeces.genus,overwrite = T)

ggarrange(mean.freq.plot.milk.genus,mean.freq.plot.faeces.genus,nrow=2)
```


# Shao

## Functions 

```{r}
ps_to_pa <- function(ps, suffix = "_inf", ...) {
  ps %>%
    ps_to_df() %>%
    pivot_longer(all_of(taxa_names(ps)), names_to = "ASV", values_to = "RA") %>%
    mutate(pa = if_else(RA >= 0.001, 1, 0)) %>%
    select(ID, sample_id, ASV, pa, ...) %>%
    rename_with(~str_c(., suffix), !c(ASV, ID))
}

#pa_test <- ps_to_pa(ps)

get_total_transfer <- function(ps_obj) {

  pa_mother <- ps_obj %>%
    subset_samples(Origin == "Mother" & !Niche == "Skin") %>%
    #subset_samples(!only_mother) %>%
    ps_to_pa(., "_mother", Niche) %>%
    filter(pa_mother == "1") # select all ASVs present in mothers (i.e. at least once)
  
  pa_inf <- ps_obj %>%
    subset_samples(Origin == "Infant"& Niche != "Skin") %>%
    ps_to_pa(., "_inf", Niche, Birth.mode) %>% #, Days
    rename_with(~str_remove(., "_inf"), starts_with("Birth.mode"))

  total_transfer <- inner_join(pa_mother, pa_inf, by = c("ID", "ASV")) %>%
    unite(pa_mother, pa_inf, col = "pa_mi", sep = "_", remove = F) %>%
    mutate(trans_yn = if_else(pa_mi == "1_1", "yes", "no")) %>%
    unite(col = "niche_from_to", Niche_mother, Niche_inf, sep = "-", remove = F)
  
  return(total_transfer)
}


#total_transfer_test <- get_total_transfer(ps)


get_trans_rate <- function(total_transfer, ...) {
  total_transfer %>% 
    dplyr::count(ASV, ID, Niche_mother, Niche_inf, niche_from_to, trans_yn, ...) %>%
    # number of transfers (yes/no) for each ASV/subject
    pivot_wider(names_from = trans_yn, values_from = n, values_fill = 0, names_prefix = "trans_") %>%
    mutate(any_trans_yn = if_else(trans_yes == 0 & trans_no > 0, "no", "yes")) %>%
    dplyr::count(ASV, Niche_mother, Niche_inf, niche_from_to, any_trans_yn, ..., name = "n_any") %>% #, ...
    # number of [any] transfers for each ASV/subject
    pivot_wider(names_from = any_trans_yn, values_from = n_any, values_fill = 0, names_prefix = "any_trans_") %>%
    mutate(mat_presence = any_trans_yes + any_trans_no,
           trans_rate = any_trans_yes / (mat_presence))
}


#trans_rate_test <- get_trans_rate(total_transfer_test)

format_fisher_p <- function(fisher_p) {
  
  get_trans_rate <- function(data, bm = "C-section") {
    data %>%
      filter(Birth.mode == bm) %>%
      pull(trans_rate)
  }
   
  fisher_p %>% 
    group_by(niche_from_to) %>% 
    mutate(col = map_chr(data, ~ if_else(get_trans_rate(.x, "C-section") > get_trans_rate(.x, "Vaginal"), "sc", "vag")),
           q.value = p.adjust(p.value, method = "BH"),
           label = glue::glue("_p_ = {scales::pvalue_format()(p.value)}<br> _q_ = {scales::pvalue_format()(q.value)}")) %>%
    filter(p.value < 0.1) 
}

```

## Plot - Mother to infant

### Milk vs infant faeces days 30-90

```{r, ow = '100%', w = 10.5, h = 7}
ps.sub <- subset_samples(PSB, Sample.type %in% c("Milk","Faecal sample") & Days %in% c(30,60,90))

df <- data.frame(ps.sub@tax_table)
df[df==""] <- NA
df$tax <- apply(df, 1, function(x) tail(na.omit(x), 1))
taxa_names(ps.sub )<- paste0(
  df$tax,
  str_remove(taxa_names(ps.sub),"OTU")
)
 
#Overwrite ID with RC_ID(time sensitive ID)
ps.sub@sam_data$ID <- ps.sub@sam_data$RCID

count <- data.frame(ps@sam_data) %>% dplyr::count(ID, sort = TRUE) %>% filter(n>1)

PSB.RA <- transform_sample_counts(
  ps.sub,
  function(x) x / sum(x))

ps <- PSB.RA #subset_samples(PSB.RA, ID %in% count$ID)

sl(total_transfer_milk_faec, get_total_transfer(ps), overwrite = T) 

total_transfer <- total_transfer_milk_faec

str(subset(total_transfer, trans_yn == "yes"))

#col_bm <- cols_mb$birth_mode

#names(col_bm) <- NULL
levels(as.factor(total_transfer$niche_from_to))
niche_sel = c("Milk-Infant faeces")

trans_rate <- get_trans_rate(total_transfer, Birth.mode)

ASV_sel_df <- get_trans_rate(total_transfer) %>% 
  filter(niche_from_to %in% niche_sel) %>% 
  group_by(niche_from_to) %>% 
  mutate(max_mat_presence = max(mat_presence)) %>% 
  filter(mat_presence > (0.1 * max_mat_presence)) %>% # Shao: > 10 pairs
  slice_max(trans_rate, n = 15, with_ties = F) # ties for vag-NP

total_trans <- left_join(
  ASV_sel_df %>% 
    select(niche_from_to, ASV),
  trans_rate, by = c("niche_from_to", "ASV")
  ) %>% 
  add_count(ASV, wt = mat_presence, name = "total_mat_presence") %>%
  ungroup() %>% 
  mutate(ASV = format_OTU(ASV), #%>% fct_reorder(., mat_presence))
         niche_from_to = fct_relevel(niche_from_to, niche_sel)) %>% #"vag-skin", 
  drop_na(Birth.mode)
 
total_trans_reordered <- total_trans %>% 
  mutate(ASV = tidytext::reorder_within(ASV, mat_presence, niche_from_to))

shao_fisher_p <- function(total_trans) {

  total_trans %>%
    group_by(ASV, niche_from_to) %>%
    nest() %>% 
    mutate(fit = map(data, ~fisher.test(.x %>% select(any_trans_no, any_trans_yes))),
           tidied = map(fit, broom::tidy)) %>%
    unnest(tidied)
}

shao_facet <- total_trans_reordered %>% 
  ggplot(aes(x = mat_presence, y = ASV, color = Birth.mode)) +
    geom_col(position = position_dodge(), fill = "white", width = 0.8) +
    scale_color_manual(name = "Birth mode", labels = list(sc = "C-section", vag = "Vaginal birth"), values = col_fil) +
    scale_x_continuous(expand = expansion(mult = c(0.02, 0.5))) +
    geom_col(data = total_trans_reordered, 
             aes(x = any_trans_yes, y = ASV, fill = Birth.mode), 
             position = position_dodge(), inherit.aes = F, width = 0.8) +
    scale_fill_manual(name = "Birth mode", labels = list(sc = "C-section", vag = "Vaginal birth"), values = col_fil) +
    facet_wrap(~ niche_from_to, scales = "free", nrow = 3) +
    tidytext::scale_y_reordered(expand = expansion(mult = c(0.08, 0.08))) +
    labs(x = "Number of mother-infant pairs") +
    theme(axis.text.y = element_markdown())

shao_facet

# add p-values

# p_data <- shao_fisher_p(total_trans_reordered) %>% ungroup()
# 
# shao_facet_p <- shao_facet + 
#   geom_richtext(data = p_data %>% format_fisher_p, aes(x = Inf, y = ASV, label = label, color = col), inherit.aes = F, hjust = "right", fill = NA, label.color = NA, size = 3, lineheight = .8)
#   
# legend_cols <- total_trans %>%
#     ggplot(aes(x = ASV, y = mat_presence, fill = Birth.mode)) + #as.numeric(interaction(birth_mode, ASV))
#       geom_col(color = "black") +
#       scale_fill_manual(name = "Transmission", labels = list(`C-section` = "Transmission", `Vaginal` = "No transmission"), values = c("black", "white"))
# 
# shao_comb <- cowplot::plot_grid(
#   shao_facet_p + theme(legend.position = "none"),
#   cowplot::plot_grid(
#     NULL,
#     cowplot::get_legend(shao_facet_p),
#     cowplot::get_legend(legend_cols),
#     NULL,
#     ncol = 1, align = "v",
#     rel_heights  = c(0.3, 0.2, 0.2, 0.3)),
#   nrow = 1, rel_widths = c(0.85, 0.15))
# 
# #sl(shao_comb, shao_comb, overwrite = T)
# shao_comb

```

### Milk cs infant faeces - day3 milk vs day 30 infant faeces

```{r, ow = '100%', w = 10.5, h = 7}
ps.sub <- subset_samples(PSB, Sample.type %in% c("Milk","Faecal sample") & Days %in% c(30,60,90))

df <- data.frame(ps.sub@tax_table)
df[df==""] <- NA
df$tax <- apply(df, 1, function(x) tail(na.omit(x), 1))
taxa_names(ps.sub )<- paste0(
  df$tax,
  str_remove(taxa_names(ps.sub),"OTU")
)

#Overwrite ID with RC_ID(time sensitive ID)
#ps.sub@sam_data$ID <- ps.sub@sam_data$RCID

count <- data.frame(ps.sub @sam_data) %>% dplyr::count(ID, sort = TRUE) %>% filter(n>1)

PSB.RA <- transform_sample_counts(
  ps.sub,
  function(x) x / sum(x))

ps <- subset_samples(PSB.RA, ID %in% count$ID)

sl(total_transfer_milk_faec, get_total_transfer(ps), overwrite = T) 

total_transfer <- total_transfer_milk_faec

str(subset(total_transfer, trans_yn == "yes"))

#col_bm <- cols_mb$birth_mode

#names(col_bm) <- NULL
levels(as.factor(total_transfer$niche_from_to))
niche_sel = c("Milk-Infant faeces")

trans_rate <- get_trans_rate(total_transfer, Birth.mode)

ASV_sel_df <- get_trans_rate(total_transfer) %>% 
  filter(niche_from_to %in% niche_sel) %>% 
  group_by(niche_from_to) %>% 
  mutate(max_mat_presence = max(mat_presence)) %>% 
  filter(mat_presence > (0.1 * max_mat_presence)) %>% # Shao: > 10 pairs
  slice_max(trans_rate, n = 15, with_ties = F) # ties for vag-NP

total_trans <- left_join(
  ASV_sel_df %>% 
    select(niche_from_to, ASV),
  trans_rate, by = c("niche_from_to", "ASV")
  ) %>% 
  add_count(ASV, wt = mat_presence, name = "total_mat_presence") %>%
  ungroup() %>% 
  mutate(ASV = format_OTU(ASV), #%>% fct_reorder(., mat_presence))
         niche_from_to = fct_relevel(niche_from_to, niche_sel)) %>% #"vag-skin", 
  drop_na(Birth.mode)
 
total_trans_reordered <- total_trans %>% 
  mutate(ASV = tidytext::reorder_within(ASV, mat_presence, niche_from_to))

shao_fisher_p <- function(total_trans) {

  total_trans %>%
    group_by(ASV, niche_from_to) %>%
    nest() %>% 
    mutate(fit = map(data, ~fisher.test(.x %>% select(any_trans_no, any_trans_yes))),
           tidied = map(fit, broom::tidy)) %>%
    unnest(tidied)
}

shao_facet <- total_trans_reordered %>% 
  ggplot(aes(x = mat_presence, y = ASV, color = Birth.mode)) +
    geom_col(position = position_dodge(), fill = "white", width = 0.8) +
    scale_color_manual(name = "Birth mode", labels = list(sc = "C-section", vag = "Vaginal birth"), values = col_fil) +
    scale_x_continuous(expand = expansion(mult = c(0.02, 0.5))) +
    geom_col(data = total_trans_reordered, 
             aes(x = any_trans_yes, y = ASV, fill = Birth.mode), 
             position = position_dodge(), inherit.aes = F, width = 0.8) +
    scale_fill_manual(name = "Birth mode", labels = list(sc = "C-section", vag = "Vaginal birth"), values = col_fil) +
    facet_wrap(~ niche_from_to, scales = "free", nrow = 3) +
    tidytext::scale_y_reordered(expand = expansion(mult = c(0.08, 0.08))) +
    labs(x = "Number of mother-infant pairs") +
    theme(axis.text.y = element_markdown())

#shao_facet

# add p-values
p_data <- shao_fisher_p(total_trans_reordered) %>% ungroup()

shao_facet_p <- shao_facet + 
  geom_richtext(data = p_data %>% format_fisher_p, aes(x = Inf, y = ASV, label = label, color = col), inherit.aes = F, hjust = "right", fill = NA, label.color = NA, size = 3, lineheight = .8)
  
legend_cols <- total_trans %>%
    ggplot(aes(x = ASV, y = mat_presence, fill = Birth.mode)) + #as.numeric(interaction(birth_mode, ASV))
      geom_col(color = "black") +
      scale_fill_manual(name = "Transmission", labels = list(`C-section` = "Transmission", `Vaginal` = "No transmission"), values = c("black", "white"))

shao_comb <- cowplot::plot_grid(
  shao_facet_p + theme(legend.position = "none"),
  cowplot::plot_grid(
    NULL,
    cowplot::get_legend(shao_facet_p),
    cowplot::get_legend(legend_cols),
    NULL,
    ncol = 1, align = "v",
    rel_heights  = c(0.3, 0.2, 0.2, 0.3)),
  nrow = 1, rel_widths = c(0.85, 0.15))

#sl(shao_comb, shao_comb, overwrite = T)
shao_comb

```

### All samples with day 3 milk and day 30 faeces

```{r, ow = '100%', w = 10.5, h = 7}
ps.sub <- subset_samples(PSB, !Sample.type %in% c("Faecal sample") & !Days %in% c(30,60,90) | Sample.type %in% c("Faecal sample") & Days %in% 30)

df <- data.frame(ps.sub@tax_table)
df[df==""] <- NA
df$tax <- apply(df, 1, function(x) tail(na.omit(x), 1))
taxa_names(ps.sub)<- paste0(
  df$tax,
  str_remove(taxa_names(ps.sub),"OTU")
)

count <- meta.dat %>% dplyr::count(ID, sort = TRUE) %>% filter(n>1)
PSB.RA <- transform_sample_counts(
  ps.sub,
  function(x) x / sum(x))

ps <- subset_samples(PSB.RA, ID %in% count$ID)

sl(total_transfer_early, get_total_transfer(ps), overwrite = T) 

str(subset(total_transfer_early, trans_yn == "yes"))

total_transfer <- total_transfer_early

str(subset(total_transfer, trans_yn == "yes"))

#col_bm <- cols_mb$birth_mode

#names(col_bm) <- NULL
levels(as.factor(total_transfer$niche_from_to))

trans_rate <- get_trans_rate(total_transfer, Birth.mode)

ASV_sel_df <- get_trans_rate(total_transfer) %>% 
  #filter(niche_from_to %in% niche_sel) %>% 
  group_by(niche_from_to) %>% 
  mutate(max_mat_presence = max(mat_presence)) %>% 
  filter(mat_presence > (0.1 * max_mat_presence)) %>% # Shao: > 10 pairs
  slice_max(trans_rate, n = 15, with_ties = F) # ties for vag-NP

total_trans <- left_join(
  ASV_sel_df %>% 
    select(niche_from_to, ASV),
  trans_rate, by = c("niche_from_to", "ASV")
  ) %>% 
  add_count(ASV, wt = mat_presence, name = "total_mat_presence") %>%
  ungroup() %>% 
  mutate(ASV = format_OTU(ASV), #%>% fct_reorder(., mat_presence))
         niche_from_to = fct_relevel(niche_from_to, niche_sel)) %>% #"vag-skin", 
  drop_na(Birth.mode)
 
total_trans_reordered <- total_trans %>% 
  mutate(ASV = tidytext::reorder_within(ASV, mat_presence, niche_from_to))

shao_fisher_p <- function(total_trans) {

  total_trans %>%
    group_by(ASV, niche_from_to) %>%
    nest() %>% 
    mutate(fit = map(data, ~fisher.test(.x %>% select(any_trans_no, any_trans_yes))),
           tidied = map(fit, broom::tidy)) %>%
    unnest(tidied)
}

shao_facet <- total_trans_reordered %>% 
  ggplot(aes(x = mat_presence, y = ASV, color = Birth.mode)) +
    geom_col(position = position_dodge(), fill = "white", width = 0.8) +
    scale_color_manual(name = "Birth mode", labels = list(sc = "C-section", vag = "Vaginal birth"), values = col_fil) +
    scale_x_continuous(expand = expansion(mult = c(0.02, 0.5))) +
    geom_col(data = total_trans_reordered, 
             aes(x = any_trans_yes, y = ASV, fill = Birth.mode), 
             position = position_dodge(), inherit.aes = F, width = 0.8) +
    scale_fill_manual(name = "Birth mode", labels = list(sc = "C-section", vag = "Vaginal birth"), values = col_fil) +
    facet_wrap(~ niche_from_to, scales = "free", nrow = 3) +
    tidytext::scale_y_reordered(expand = expansion(mult = c(0.08, 0.08))) +
    labs(x = "Number of mother-infant pairs") +
    theme(axis.text.y = element_markdown())

shao_facet

# add p-values

# p_data <- shao_fisher_p(total_trans_reordered) %>% ungroup()
# 
# shao_facet_p <- shao_facet + 
#   geom_richtext(data = p_data %>% format_fisher_p, aes(x = Inf, y = ASV, label = label, color = col), inherit.aes = F, hjust = "right", fill = NA, label.color = NA, size = 3, lineheight = .8)
#   
# legend_cols <- total_trans %>%
#     ggplot(aes(x = ASV, y = mat_presence, fill = Birth.mode)) + #as.numeric(interaction(birth_mode, ASV))
#       geom_col(color = "black") +
#       scale_fill_manual(name = "Transmission", labels = list(`C-section` = "Transmission", `Vaginal` = "No transmission"), values = c("black", "white"))
# 
# shao_comb <- cowplot::plot_grid(
#   shao_facet_p + theme(legend.position = "none"),
#   cowplot::plot_grid(
#     NULL,
#     #cowplot::get_legend(shao_facet_p),
#     cowplot::get_legend(legend_cols),
#     NULL,
#     ncol = 1, align = "v",
#     rel_heights  = c(0.3, 0.2, 0.2, 0.3)),
#   nrow = 1, rel_widths = c(0.85, 0.15))
# 
# sl(shao_comb, shao_comb, overwrite = T)
# shao_comb
```

## Plot - Infant to mother

### Milk vs infant faeces days 30-90

```{r, ow = '100%', w = 10.5, h = 7}
ps.sub <- subset_samples(PSB, Sample.type %in% c("Milk","Faecal sample") & Days %in% c(30,60,90))
# Reverse infant-mother encoding to calculate mother-infant transmission instead
ps.sub@sam_data$Origin <- dplyr::recode(ps.sub@sam_data$Origin, Infant = "Mother", Mother = "Infant")

df <- data.frame(ps.sub@tax_table)
df[df==""] <- NA
df$tax <- apply(df, 1, function(x) tail(na.omit(x), 1))
taxa_names(ps.sub )<- paste0(
  df$tax,
  str_remove(taxa_names(ps.sub),"OTU")
)
 
#Overwrite ID with RC_ID(time sensitive ID)
ps.sub@sam_data$ID <- ps.sub@sam_data$RCID

count <- data.frame(ps.sub@sam_data) %>% dplyr::count(ID, sort = TRUE) %>% filter(n>1)

PSB.RA <- transform_sample_counts(
  ps.sub,
  function(x) x / sum(x))

ps <- PSB.RA #subset_samples(PSB.RA, ID %in% count$ID)

sl(total_transfer_faec_milk, get_total_transfer(ps), overwrite = T) 

total_transfer <- total_transfer_faec_milk

str(subset(total_transfer, trans_yn == "yes"))

#col_bm <- cols_mb$birth_mode

#names(col_bm) <- NULL
levels(as.factor(total_transfer$niche_from_to))
niche_sel = c("Milk-Infant faeces")

trans_rate <- get_trans_rate(total_transfer, Birth.mode)

ASV_sel_df <- get_trans_rate(total_transfer) %>% 
  #filter(niche_from_to %in% niche_sel) %>% 
  group_by(niche_from_to) %>% 
  mutate(max_mat_presence = max(mat_presence)) %>% 
  filter(mat_presence > (0.1 * max_mat_presence)) %>% # Shao: > 10 pairs
  slice_max(trans_rate, n = 15, with_ties = F) # ties for vag-NP

total_trans <- left_join(
  ASV_sel_df %>% 
    select(niche_from_to, ASV),
  trans_rate, by = c("niche_from_to", "ASV")
  ) %>% 
  add_count(ASV, wt = mat_presence, name = "total_mat_presence") %>%
  ungroup() %>% 
  mutate(ASV = format_OTU(ASV), #%>% fct_reorder(., mat_presence))
         niche_from_to = fct_relevel(niche_from_to, niche_sel)) %>% #"vag-skin", 
  drop_na(Birth.mode)
 
total_trans_reordered <- total_trans %>% 
  mutate(ASV = tidytext::reorder_within(ASV, mat_presence, niche_from_to))

shao_fisher_p <- function(total_trans) {

  total_trans %>%
    group_by(ASV, niche_from_to) %>%
    nest() %>% 
    mutate(fit = map(data, ~fisher.test(.x %>% select(any_trans_no, any_trans_yes))),
           tidied = map(fit, broom::tidy)) %>%
    unnest(tidied)
}

shao_facet <- total_trans_reordered %>% 
  ggplot(aes(x = mat_presence, y = ASV, color = Birth.mode)) +
    geom_col(position = position_dodge(), fill = "white", width = 0.8) +
    scale_color_manual(name = "Birth mode", labels = list(sc = "C-section", vag = "Vaginal birth"), values = col_fil) +
    scale_x_continuous(expand = expansion(mult = c(0.02, 0.5))) +
    geom_col(data = total_trans_reordered, 
             aes(x = any_trans_yes, y = ASV, fill = Birth.mode), 
             position = position_dodge(), inherit.aes = F, width = 0.8) +
    scale_fill_manual(name = "Birth mode", labels = list(sc = "C-section", vag = "Vaginal birth"), values = col_fil) +
    facet_wrap(~ niche_from_to, scales = "free", nrow = 3) +
    tidytext::scale_y_reordered(expand = expansion(mult = c(0.08, 0.08))) +
    labs(x = "Number of mother-infant pairs") +
    theme(axis.text.y = element_markdown())

shao_facet

# add p-values

p_data <- shao_fisher_p(total_trans_reordered) %>% ungroup()

shao_facet_p <- shao_facet + 
  geom_richtext(data = p_data %>% format_fisher_p, aes(x = Inf, y = ASV, label = label, color = col), inherit.aes = F, hjust = "right", fill = NA, label.color = NA, size = 3, lineheight = .8)
  
legend_cols <- total_trans %>%
    ggplot(aes(x = ASV, y = mat_presence, fill = Birth.mode)) + #as.numeric(interaction(birth_mode, ASV))
      geom_col(color = "black") +
      scale_fill_manual(name = "Transmission", labels = list(`C-section` = "Transmission", `Vaginal` = "No transmission"), values = c("black", "white"))

shao_comb <- cowplot::plot_grid(
  shao_facet_p + theme(legend.position = "none"),
  cowplot::plot_grid(
    NULL,
    cowplot::get_legend(shao_facet_p),
    cowplot::get_legend(legend_cols),
    NULL,
    ncol = 1, align = "v",
    rel_heights  = c(0.3, 0.2, 0.2, 0.3)),
  nrow = 1, rel_widths = c(0.85, 0.15))

#sl(shao_comb, shao_comb, overwrite = T)
shao_comb
```

### Milk cs infant faeces - day3 milk vs day 30 infant faeces

```{r, ow = '100%', w = 10.5, h = 7}
ps.sub <- subset_samples(PSB, Sample.type %in% c("Milk","Faecal sample") & Days %in% c(30,60,90))
# Reverse infant-mother encoding to calculate mother-infant transmission instead
ps.sub@sam_data$Origin <- dplyr::recode(ps.sub@sam_data$Origin, Infant = "Mother", Mother = "Infant")

df <- data.frame(ps.sub@tax_table)
df[df==""] <- NA
df$tax <- apply(df, 1, function(x) tail(na.omit(x), 1))
taxa_names(ps.sub )<- paste0(
  df$tax,
  str_remove(taxa_names(ps.sub),"OTU")
)

#Overwrite ID with RC_ID(time sensitive ID)
#ps.sub@sam_data$ID <- ps.sub@sam_data$RCID

count <- data.frame(ps.sub @sam_data) %>% dplyr::count(ID, sort = TRUE) %>% filter(n>1)

PSB.RA <- transform_sample_counts(
  ps.sub,
  function(x) x / sum(x))

ps <- subset_samples(PSB.RA, ID %in% count$ID)

sl(total_transfer_milk_faec_early, get_total_transfer(ps), overwrite = T) 

total_transfer <- total_transfer_milk_faec_early

str(subset(total_transfer, trans_yn == "yes"))

#col_bm <- cols_mb$birth_mode

#names(col_bm) <- NULL
levels(as.factor(total_transfer$niche_from_to))
niche_sel = c("Milk-Infant faeces")

trans_rate <- get_trans_rate(total_transfer, Birth.mode)

ASV_sel_df <- get_trans_rate(total_transfer) %>% 
  #filter(niche_from_to %in% niche_sel) %>% 
  group_by(niche_from_to) %>% 
  mutate(max_mat_presence = max(mat_presence)) %>% 
  filter(mat_presence > (0.1 * max_mat_presence)) %>% # Shao: > 10 pairs
  slice_max(trans_rate, n = 15, with_ties = F) # ties for vag-NP

total_trans <- left_join(
  ASV_sel_df %>% 
    select(niche_from_to, ASV),
  trans_rate, by = c("niche_from_to", "ASV")
  ) %>% 
  add_count(ASV, wt = mat_presence, name = "total_mat_presence") %>%
  ungroup() %>% 
  mutate(ASV = format_OTU(ASV), #%>% fct_reorder(., mat_presence))
         niche_from_to = fct_relevel(niche_from_to, niche_sel)) %>% #"vag-skin", 
  drop_na(Birth.mode)
 
total_trans_reordered <- total_trans %>% 
  mutate(ASV = tidytext::reorder_within(ASV, mat_presence, niche_from_to))

shao_fisher_p <- function(total_trans) {

  total_trans %>%
    group_by(ASV, niche_from_to) %>%
    nest() %>% 
    mutate(fit = map(data, ~fisher.test(.x %>% select(any_trans_no, any_trans_yes))),
           tidied = map(fit, broom::tidy)) %>%
    unnest(tidied)
}

shao_facet <- total_trans_reordered %>% 
  ggplot(aes(x = mat_presence, y = ASV, color = Birth.mode)) +
    geom_col(position = position_dodge(), fill = "white", width = 0.8) +
    scale_color_manual(name = "Birth mode", labels = list(sc = "C-section", vag = "Vaginal birth"), values = col_fil) +
    scale_x_continuous(expand = expansion(mult = c(0.02, 0.5))) +
    geom_col(data = total_trans_reordered, 
             aes(x = any_trans_yes, y = ASV, fill = Birth.mode), 
             position = position_dodge(), inherit.aes = F, width = 0.8) +
    scale_fill_manual(name = "Birth mode", labels = list(sc = "C-section", vag = "Vaginal birth"), values = col_fil) +
    facet_wrap(~ niche_from_to, scales = "free", nrow = 3) +
    tidytext::scale_y_reordered(expand = expansion(mult = c(0.08, 0.08))) +
    labs(x = "Number of mother-infant pairs") +
    theme(axis.text.y = element_markdown())

shao_facet

# add p-values

p_data <- shao_fisher_p(total_trans_reordered) %>% ungroup()

shao_facet_p <- shao_facet + 
  geom_richtext(data = p_data %>% format_fisher_p, aes(x = Inf, y = ASV, label = label, color = col), inherit.aes = F, hjust = "right", fill = NA, label.color = NA, size = 3, lineheight = .8)
  
legend_cols <- total_trans %>%
    ggplot(aes(x = ASV, y = mat_presence, fill = Birth.mode)) + #as.numeric(interaction(birth_mode, ASV))
      geom_col(color = "black") +
      scale_fill_manual(name = "Transmission", labels = list(`C-section` = "Transmission", `Vaginal` = "No transmission"), values = c("black", "white"))

shao_comb <- cowplot::plot_grid(
  shao_facet_p + theme(legend.position = "none"),
  cowplot::plot_grid(
    NULL,
    cowplot::get_legend(shao_facet_p),
    cowplot::get_legend(legend_cols),
    NULL,
    ncol = 1, align = "v",
    rel_heights  = c(0.3, 0.2, 0.2, 0.3)),
  nrow = 1, rel_widths = c(0.85, 0.15))

#sl(shao_comb, shao_comb, overwrite = T)
shao_comb

```

### All samples with day 3 milk and day 30 faeces

```{r, ow = '100%', w = 10.5, h = 7}
ps.sub <- subset_samples(PSB, !Sample.type %in% c("Faecal sample") & !Days %in% c(30,60,90) | Sample.type %in% c("Faecal sample") & Days %in% 30)
# Reverse infant-mother encoding to calculate mother-infant transmission instead
ps.sub@sam_data$Origin <- dplyr::recode(ps.sub@sam_data$Origin, Infant = "Mother", Mother = "Infant")

df <- data.frame(ps.sub@tax_table)
df[df==""] <- NA
df$tax <- apply(df, 1, function(x) tail(na.omit(x), 1))
taxa_names(ps.sub)<- paste0(
  df$tax,
  str_remove(taxa_names(ps.sub),"OTU")
)

count <- meta.dat %>% dplyr::count(ID, sort = TRUE) %>% filter(n>1)
PSB.RA <- transform_sample_counts(
  ps.sub,
  function(x) x / sum(x))

ps <- subset_samples(PSB.RA, ID %in% count$ID)

sl(total_transfer_all_inf_to_mother, get_total_transfer(ps), overwrite = T) 

str(subset(total_transfer_early, trans_yn == "yes"))

total_transfer <- total_transfer_all_inf_to_mother

#col_bm <- cols_mb$birth_mode

#names(col_bm) <- NULL
levels(as.factor(total_transfer$niche_from_to))

trans_rate <- get_trans_rate(total_transfer, Birth.mode)

ASV_sel_df <- get_trans_rate(total_transfer) %>% 
  #filter(niche_from_to %in% niche_sel) %>% 
  group_by(niche_from_to) %>% 
  mutate(max_mat_presence = max(mat_presence)) %>% 
  filter(mat_presence > (0.1 * max_mat_presence)) %>% # Shao: > 10 pairs
  slice_max(trans_rate, n = 15, with_ties = F) # ties for vag-NP

total_trans <- left_join(
  ASV_sel_df %>% 
    select(niche_from_to, ASV),
  trans_rate, by = c("niche_from_to", "ASV")
  ) %>% 
  add_count(ASV, wt = mat_presence, name = "total_mat_presence") %>%
  ungroup() %>% 
  mutate(ASV = format_OTU(ASV), #%>% fct_reorder(., mat_presence))
         niche_from_to = fct_relevel(niche_from_to, niche_sel)) %>% #"vag-skin", 
  drop_na(Birth.mode)
 
total_trans_reordered <- total_trans %>% 
  mutate(ASV = tidytext::reorder_within(ASV, mat_presence, niche_from_to))

shao_fisher_p <- function(total_trans) {

  total_trans %>%
    group_by(ASV, niche_from_to) %>%
    nest() %>% 
    mutate(fit = map(data, ~fisher.test(.x %>% select(any_trans_no, any_trans_yes))),
           tidied = map(fit, broom::tidy)) %>%
    unnest(tidied)
}

shao_facet <- total_trans_reordered %>% 
  ggplot(aes(x = mat_presence, y = ASV, color = Birth.mode)) +
    geom_col(position = position_dodge(), fill = "white", width = 0.8) +
    scale_color_manual(name = "Birth mode", labels = list(sc = "C-section", vag = "Vaginal birth"), values = col_fil) +
    scale_x_continuous(expand = expansion(mult = c(0.02, 0.5))) +
    geom_col(data = total_trans_reordered, 
             aes(x = any_trans_yes, y = ASV, fill = Birth.mode), 
             position = position_dodge(), inherit.aes = F, width = 0.8) +
    scale_fill_manual(name = "Birth mode", labels = list(sc = "C-section", vag = "Vaginal birth"), values = col_fil) +
    facet_wrap(~ niche_from_to, scales = "free", nrow = 3) +
    tidytext::scale_y_reordered(expand = expansion(mult = c(0.08, 0.08))) +
    labs(x = "Number of mother-infant pairs") +
    theme(axis.text.y = element_markdown())

shao_facet

# add p-values
p_data <- shao_fisher_p(total_trans_reordered) %>% ungroup()

shao_facet_p <- shao_facet + 
  geom_richtext(data = p_data %>% format_fisher_p, aes(x = Inf, y = ASV, label = label, color = col), inherit.aes = F, hjust = "right", fill = NA, label.color = NA, size = 3, lineheight = .8)
  
legend_cols <- total_trans %>%
    ggplot(aes(x = ASV, y = mat_presence, fill = Birth.mode)) + #as.numeric(interaction(birth_mode, ASV))
      geom_col(color = "black") +
      scale_fill_manual(name = "Transmission", labels = list(`C-section` = "Transmission", `Vaginal` = "No transmission"), values = c("black", "white"))

shao_comb <- cowplot::plot_grid(
  shao_facet_p + theme(legend.position = "none"),
  cowplot::plot_grid(
    NULL,
    #cowplot::get_legend(shao_facet_p),
    cowplot::get_legend(legend_cols),
    NULL,
    ncol = 1, align = "v",
    rel_heights  = c(0.3, 0.2, 0.2, 0.3)),
  nrow = 1, rel_widths = c(0.85, 0.15))

sl(shao_comb, shao_comb, overwrite = T)
shao_comb
```

### Between all sample types - using sum abundance for feacal + milk samples

# Total transmissions

PM: align code below with code above.

```{r}
sl(cum_transfer, {
  pa_mother <- ps$RA %>%
    subset_samples((mother_inf == "mother" & !niche == "bm") | (niche == "bm" & first_bm)) %>%
    pres_abund_filter(pres = 1, abund = 0.001) %>%
    ps_to_pa(., "_mother", niche)
  
  pa_inf <- ps$RA %>%
    subset_samples(mother_inf == "infant") %>%
    #pres_abund_filter(pres = 1, abund = 0.001) %>%
    ps_to_pa(., "_inf", niche, birth_mode)
  
  cum_transfer <- inner_join(pa_mother, pa_inf, by = c("subject", "ASV")) %>%
      filter(pa_mother == "1") %>%  # select all ASVs present in mothers (i.e. at least once)
      unite(pa_mother, pa_inf, col = "pa_mi", sep = "_", remove = F)
    
  n_small <- cum_transfer %>% 
    distinct(subject, birth_mode_inf) %>% 
    count(birth_mode_inf) %>% 
    slice_min(n) %>% 
    pull(n) # 46
  
  downsampled <- cum_transfer %>%
    distinct(subject, birth_mode_inf) %>% 
    group_by(birth_mode_inf) %>%
    slice_sample(n = n_small)
    
  cum_transfer <- cum_transfer %>%
    filter(subject %in% downsampled$subject) %>%
    count(pa_mi, niche_mother, niche_inf, birth_mode_inf) %>%
    pivot_wider(names_from = "pa_mi", values_from = "n", values_fill = 0) %>% # transfer/no transfer of present ASVs
    mutate(rate = `1_1` / (`1_0` + `1_1`))
  
  cum_transfer
}, overwrite = FALSE)

```

We used a downsampled dataset to not obscure our transmission detection rate with a higher chance of co-occurrence of ASVs in a larger number of subjects (e.g. vaginal > sc). Here we used a set consisting of 46 sc and 46 random vaginally born infants.

```{r, ow = '70%', w = 7, h = 7}
cum_transfer_pval <- cum_transfer %>%
  group_by(niche_mother, niche_inf) %>%
  nest() %>%
  mutate(fit = map(data, ~ .x %>% select(`1_0`, `1_1`) %>% fisher.test),
         tidied = map(fit, broom::tidy)) %>%
  unnest(tidied) %>%
  group_by(niche_mother) %>%
  mutate(q.value = p.adjust(p.value, method = "BH"),
         q.value_sci = gsub("e", " × 10^", scales::scientific_format(digits = 2)(q.value)) %>% str_replace_all(., "10\\^-0", "10^-"),
         qlab1 = if_else(q.value < 0.001, q.value_sci, scales::pvalue_format()(q.value)),
         qlab2 = if_else(q.value < 0.05, glue::glue("_q_ = **{qlab1}**"), glue::glue("_q_ = {qlab1}")))

cum_transfer %>%
  #unite(c(niche_mother, niche_inf), col = "from_to", sep = " - ", remove = F) %>%
  ggplot(aes(x = rate, y = niche_inf, color = birth_mode_inf)) +
    geom_point() + 
    facet_grid(niche_mother ~ ., scales = "free") +
    geom_richtext(data = cum_transfer_pval, aes(y = niche_inf, label = qlab2), x = 0.4, hjust = "left", fill = NA, label.color = NA, size = 3, inherit.aes = F)
```

# By relative abundance

# Personal ASV transmission 

## functions

```{r}
ps_to_pa <- function(ps, suffix = "_inf", ...) {
  ps %>%
    ps_to_df() %>%
    pivot_longer(all_of(taxa_names(ps)), names_to = "ASV", values_to = "RA") %>%
    mutate(pa = if_else(RA >= 0.001, 1, 0)) %>%
    select(subject, sample_id, ASV, pa, ...) %>%
    rename_with(~str_c(., suffix), !c(ASV, subject))
}

ps_to_pa()

get_rates <- function(mother_pa, inf_pa, ...) {
  rates <- inner_join(mother_pa, inf_pa, by = c("subject", "ASV")) %>%
    unite(pa_mother, pa_inf, col = "pa_mi", sep = "_", remove = F) %>%
    filter(pa_mother == "1") %>% # select all ASVs present in mothers (i.e. at least once)
    count(ASV, pa_mi, ...) %>%
    pivot_wider(names_from = "pa_mi", values_from = "n", values_fill = 0) %>% # transfer/no transfer of present ASVs
    mutate(
      maternal_presence = (`1_0` + `1_1`),
      rate = `1_1` / maternal_presence)# %>%
    #filter(maternal_presence > 5)
  return(rates)
}

rand_mother_pa <- function(mother_pa, seed = 666) {
  set.seed(seed)
  
  mother_pa %>%
    group_by(subject) %>%
    nest() %>%
    ungroup() %>%
    mutate(subject = sample(subject, replace = F)) %>%
    unnest(data)
}

rand_pa_wrapper <- function(mother_niche, infant_niche, infant_time, perm = 1000, seed = 22332) {
  
  ps_mother <- ps$RA %>%
    prune_samples(sample_data(.)$mother_inf == "mother" & sample_data(.)$niche == mother_niche, .) %>%
    pres_abund_filter(pres = 1, abund = 0.001)
  
  ps_inf <- ps$RA %>%
    prune_samples(sample_data(.)$mother_inf == "infant" & sample_data(.)$niche == infant_niche & sample_data(.)$time == infant_time, .) %>%
    pres_abund_filter(pres = 1, abund = 0.001)
  
  mother_pa <- ps_to_pa(ps_mother, "_mother")
  inf_pa <- ps_to_pa(ps_inf, "_inf")
  
  true_rates <- get_rates(mother_pa, inf_pa)
  
  set.seed(seed)
  rand_rates <- tibble(seed = sample.int(n = 1000, size = perm, replace = F)) %>% # 1000 pseudorandom seeds based on set.seed() above
    mutate(mother_pa = map(seed, ~rand_mother_pa(mother_pa, .x)),
           rates = map(mother_pa, ~get_rates(.x, inf_pa))) %>%
    unnest(rates) %>%
    select(-mother_pa)
  return(list(true = true_rates, rand = rand_rates, mother_pa = mother_pa, inf_pa = inf_pa, n_mat = nsamples(ps_mother)))
  
}

get_rates_Z <- function(rand_rates, true_rates, n_mat, mat_pres_rate = 0.2, mti_rate = 0.1) { 
  # maternal presence rate = minimum 20% of maternal samples
  # true mother-to-infant rate = minimum 10%
  
  rand_rates_df <- rand_rates %>%
    group_by(ASV) %>%
    summarise(mean_rate = mean(rate),
              median_rate = median(rate),
              min_rate = min(rate),
              max_rate = max(rate),
              sd_rate = sd(rate)) %>%
    left_join(true_rates, by = "ASV") %>%
    mutate(
       diff = (rate - mean_rate),
       Z_diff = diff / sd_rate) %>%
    arrange(-diff) %>%
    filter(maternal_presence >= floor(n_mat * mat_pres_rate) &
           rate >= mti_rate)
  
  print(glue::glue("A total of {n_mat} maternal samples were included, therefore the minimum maternal presence for an ASV to be considered in this analysis is {floor(n_mat * mat_pres_rate)}."))
  
  return(rand_rates_df)
}

format_data <- function(data, ASVs) {
  data %>%
    filter(ASV %in% ASVs) %>%
    mutate(ASV = fct_relevel(ASV, rev(ASVs))) %>%
    arrange(ASV) %>%
    mutate(ASV = format_OTU(ASV) %>% fct_inorder())
}  
  
plot_rates <- function(rand_rates, true_rates, rates_Z, n = 20) {
  
  rand_rates %>%
    format_data(rates_Z$ASV[1:n]) %>%
    ggplot(aes(x = rate, y = ASV)) +
      ggdist::stat_halfeye() +
      geom_point(data = true_rates %>% format_data(rates_Z$ASV[1:n]), color = "black", fill = "grey80", shape = 21, size = 2.5) +
    scale_x_continuous(expand = c(0.025, 0.025), limits = c(0, 1), labels = scales::percent, breaks = scales::pretty_breaks(5)) +
    labs(x = "Transmission rate") +
    theme(axis.text.y = element_markdown(), panel.grid.major.y = element_blank(), panel.grid.minor.y = element_blank())
} 

create_bg_list <- function(ps_mother) { 
  mother_tax <- ps_mother %>%
    tax_table() %>%
    as.data.frame() %>%
    add_count(family) %>%
    filter(!is.na(family) & n >= 5) %>%
    select(family, ASV)
  
  mother_tax_list <- split(mother_tax, mother_tax[,1], drop = F) %>%
    map(~.x %>% pull(ASV))
  return(mother_tax_list)
}

```

## skin - skin (w1)

```{r, ow = '80%', w = 8, h = 6}
sl(skin_skin_w1_rates, rand_pa_wrapper("skin", "skin", "w1", perm = 1000, seed = 22331), overwrite = FALSE)

skin_skin_w1_Z <- get_rates_Z(skin_skin_w1_rates$rand, skin_skin_w1_rates$true, skin_skin_w1_rates$n_mat) 

sl(rand_skin_skin_w1_p, plot_rates(skin_skin_w1_rates$rand, skin_skin_w1_rates$true, skin_skin_w1_Z, n = 15), overwrite = T)
rand_skin_skin_w1_p

# skin_mother_tax_list <- ps$RA %>% 
#  subset_samples(mother_inf == "mother" & niche == "skin") %>% 
#  pres_abund_filter(pres = 1, abund = 0.001) %>%
#  create_bg_list()
# run_clusterProfiler(skin_skin_w1_Z %>% filter(Z_diff > 2) %>% pull(ASV), db_list = skin_mother_tax_list, p.adjustCutoff = 1, xlsx_name = "skin_skin_w1")
```

## skin - skin (w2)

```{r, ow = '80%', w = 8, h = 6}
sl(skin_skin_w2_rates, rand_pa_wrapper("skin", "skin", "w2", perm = 1000, seed = 22332), overwrite = FALSE)

skin_skin_w2_Z <- get_rates_Z(skin_skin_w2_rates$rand, skin_skin_w2_rates$true, skin_skin_w2_rates$n_mat) 

sl(rand_skin_skin_w2_p, plot_rates(skin_skin_w2_rates$rand, skin_skin_w2_rates$true, skin_skin_w2_Z, n = 15))
rand_skin_skin_w2_p

# run_clusterProfiler(skin_skin_w2_Z %>% filter(Z_diff > 2) %>% pull(ASV), db_list = skin_mother_tax_list, p.adjustCutoff = 1, xlsx_name = "skin_skin_w2")
```

## skin - skin (m1)

```{r, ow = '80%', w = 8, h = 6}
sl(skin_skin_m1_rates, rand_pa_wrapper("skin", "skin", "m1", perm = 1000, seed = 22333), overwrite = FALSE)

skin_skin_m1_Z <- get_rates_Z(skin_skin_m1_rates$rand, skin_skin_m1_rates$true, skin_skin_m1_rates$n_mat) 

sl(rand_skin_skin_m1_p, plot_rates(skin_skin_m1_rates$rand, skin_skin_m1_rates$true, skin_skin_m1_Z, n = 15))
rand_skin_skin_m1_p

# run_clusterProfiler(skin_skin_m1_Z %>% filter(Z_diff > 2) %>% pull(ASV), db_list = skin_mother_tax_list, p.adjustCutoff = 1, xlsx_name = "skin_skin_m1")
```

A higher number of 'maternal presence' can occur since we may find new combinations of maternal presence/infant presence, in case of infants w/o a mother with presence of a certain ASV they possess.

Rates did not perform well when stratifying for birth mode/breastfeeding etc. Numbers seem to become too low in that case.

```{r}
check_topn <- function(rates_Z, n = 15) {
   rates_Z$ASV[1:n]
}

list(w1 = skin_skin_w1_Z, w2 = skin_skin_w2_Z, m1 = skin_skin_m1_Z) %>%
  map(check_topn) %>% 
  enframe(name = "ASV") %>%
  unnest(value) %>%
  count(value, sort = T)
```

Overlap likely caused by mother samples which were kept stable.

## fae - fae (w2)

```{r, ow = '70%', w = 7, h = 4.5}
sl(fae_fae_w2_rates, rand_pa_wrapper("fae", "fae", "w2", perm = 1000, seed = 22332), overwrite = FALSE)

fae_fae_w2_Z <- get_rates_Z(fae_fae_w2_rates$rand, fae_fae_w2_rates$true, fae_fae_w2_rates$n_mat) 

sl(rand_fae_fae_w2_p, plot_rates(fae_fae_w2_rates$rand, fae_fae_w2_rates$true, fae_fae_w2_Z, n = 15))
rand_fae_fae_w2_p

fae_mother_tax_list <- ps$RA %>% 
  subset_samples(mother_inf == "mother" & niche == "fae") %>% 
  pres_abund_filter(pres = 1, abund = 0.001) %>%
  create_bg_list()

#run_clusterProfiler(fae_fae_w2_Z %>% filter(Z_diff > 2) %>% pull(ASV), db_list = fae_mother_tax_list, p.adjustCutoff = 1, xlsx_name = "fae_fae_w2")
```

It seems this idea performs best in skin given the high level of transmissions (this was much less for fae-fae).
We also tested bm-fae, but since we only had 30 breastmilk samples, these analyses did not perform well.

Note: hashed 'out' the run_clusterProfiler-section so it does not interfere with 'phyloseq'.


```{r, message = F}

```

```{r, message = F}

```

```{r, message = F}

```

```{r, message = F}

```

```{r, message = F}

```

```{r, message = F}

```

```{r, message = F}

```

```{r, message = F}

```

```{r, message = F}

```

```{r, message = F}

```

```{r, message = F}

```

```{r, message = F}

```

```{r, message = F}

```