---
title: "4.1_Metabolome_milk"
author: "Rasmus Riemer Jakobsen"
date: "`r Sys.time()`"
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
link-citations: yes
---

```{r setup, include=F, message=F, warning=F}
subdir_name <- "4.1_Metabolome_milk"

#Load packages
source("Supplementary_scripts/package_load_install.R")
#Load utility functions
source("Supplementary_scripts/utils.R")

dir.create("results", showWarnings = FALSE, recursive = TRUE)

# Set theme and colort

theme_set(theme_classic())

knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.fullwidth = TRUE)

effect.size.variables <- append(effect.size.variables[-c(1:2)],"RCID") #"DNA conc" and "batch" only for sequencing data

#Coulour pallettes
col_fil <- brewer.pal(8, "Dark2")#pal_jco("default")(10)

col_scale <- scale_fill_distiller(palette = "Dark2")
```

```{r knit, echo=F, eval=FALSE}
#Run to generate
rmarkdown::render(
  input = here::here(str_c(subdir_name, ".Rmd")),
  output_dir = here::here("results", "reports")
)
```

# Load data

```{r}
df.id <- readr::read_tsv("Data/metabolome_milk.txt")
meta <- readr::read_tsv("Data/metadata_metabome_milk.txt")
metab.class <- readr::read_tsv("Data/metabolite_categories.txt")

#Remove sample 383_2 - possible error?
df.id <- df.id %>%
  filter(RCID != "383_2")
meta <- meta %>%
  filter(RCID != "383_2")

  
#Check if dataframes match
setdiff(df.id$RCID,meta$RCID)


df <- df.id %>% dplyr::select(-RCID) 

# Set columns to factor
meta <- meta %>%
  dplyr::mutate(across(c("Days","Siblings", "GDM", "C.section", "AB.mother", "AB.infant", "Breast.infection", "Breastfeeding.issues", "Secretor", "Lewis"),
    as.factor))

meta$BMI.group <- factor(meta$BMI.group, levels = c("Normal","Overweight","Obese"))

str(meta$Lewis)
```

# Sumammary table

## Ordered by class concentration (mMol/mL)

```{r}
library(vtable)
library(kableExtra)

 

# Sort by mean within each class
metab.class <- metab.class[match(colnames(df), metab.class$Metabolite), ]
table(metab.class$Class)

means <- colMeans(df)

metab.class$Means <- unname(means)

metab.class.sorted <- metab.class %>%
  group_by(Class) %>%
  arrange(-Means) %>%
  arrange(Class)
#summarise(df.met_sorted)

# Add days and apply within-class ordering
df.met <- df[metab.class.sorted$Metabolite]
df.met <- cbind(df.met,"Days"=c(meta$Days))

# Days to sortin mat

metab.class.na <- rbind(NA,ungroup(metab.class.sorted))

#table(c("Timepoint",metab.class.sorted$Class))

Sys.setenv(OPENSSL_CONF="/dev/null")

#Grouped by category

st(dplyr::select(df.met,-c("Days")),
   out = "return",
   #group = 'Days',
   #group.test = TRUE,
   summ = c('mean(x)','sd(x)'),
   ) %>%
   knitr::kable("html") %>%
    kable_styling(htmltable_class="lightable-classic") %>%
    pack_rows(index = table(c(metab.class.sorted$Class)),
label_row_css = "background-color: #666; color: #fff;") %>%
    #row_spec(1, bold = TRUE) %>%
    save_kable("results/metabolites.html", density = 400)
webshot::webshot("results/metabolites.html", "results/metabolites.pdf")
webshot::webshot("results/metabolites.html", "results/metabolites.png")
```

## Ordered by class concentration (mMol/mL) - Days

```{r}
library(vtable)
library(kableExtra)

 

# Sort by mean within each class
metab.class <- metab.class[match(colnames(df), metab.class$Metabolite), ]
table(metab.class$Class)

means <- colMeans(df)

metab.class$Means <- unname(means)

metab.class.sorted <- metab.class %>%
  group_by(Class) %>%
  arrange(-Means) %>%
  arrange(Class)

#summarise(df.met_sorted)

# Add days and apply within-class ordering
df.met <- df[metab.class.sorted$Metabolite]
df.met <- cbind(df.met,"Days"=c(meta$Days))

# Days to sortin mat

metab.class.na <- rbind(NA,ungroup(metab.class.sorted))

#table(c("Timepoint",metab.class.sorted$Class))

Sys.setenv(OPENSSL_CONF="/dev/null")

#Grouped by category

sum_table_mt_group <- st(df.met,
   out = "return",
   group = 'Days',
   group.test = TRUE,
   summ = c('mean(x)','sd(x)'),
   )



sum_table_mt_group %>%
   knitr::kable("html") %>%
    kable_styling(htmltable_class="lightable-classic") %>%
    pack_rows(index = table(c(" ",metab.class.sorted$Class)),
label_row_css = "background-color: #666; color: #fff;") %>%
    #row_spec(1, bold = TRUE) %>%
    save_kable("results/metabolites_grouped.html", density = 400)
webshot::webshot("results/metabolites_grouped.html", "results/metabolites_grouped.pdf")
webshot::webshot("results/metabolites_grouped.html", "results/metabolites_grouped.png")

```

#### Export xlsx

```{r}
library(expss)
library(openxlsx)

wb = createWorkbook()
sh = addWorksheet(wb, "Tables")

#Export - we should specify workbook and worksheet.

xl_write(sum_table_mt_group, wb, sh)

#And, finally, we save workbook with table to the xlsx file.


saveWorkbook(wb, "results/metabolites_grouped.xlsx", overwrite = TRUE)

```

# PCA plot

## Normalisation comparison

```{r}
library(factoextra)

pca_res <- prcomp(df, scale. = FALSE, center = TRUE) #Apply UV-scaling

#Square root log
df.sq <- df %>%
  dplyr::mutate(across(where(is.numeric), ~ifelse(. < 0, NA_real_, sqrt(.))))
pca_res_sqrt <- prcomp(df.sq, scale. = FALSE, center = TRUE) #Apply UV-scaling

#Log10 norm
df.log <- log10(df+1)
pca_res_log <- prcomp(df.log, scale. = FALSE, center = TRUE) #Apply UV-scaling

ggarrange(
fviz_eig(pca_res) + ggtitle("Only centering"),
fviz_eig(pca_res_sqrt) + ggtitle("Square root normalisation"),
fviz_eig(pca_res_log) + ggtitle("log10 normalisation"),
nrow=1
)

```

##Non normalised

```{r}
library(ggfortify)

# Non-normalised

autoplot(pca_res, data = meta, colour = "Lewis", frame = TRUE, frame.type = 'norm')
autoplot(pca_res, data = meta, colour = "Secretor", frame = TRUE, frame.type = 'norm')
autoplot(pca_res, data = meta, colour = "Days", frame = TRUE, frame.type = 'norm')

```

## Square root norm

```{r}
library(ggfortify)

# Non-normalised

autoplot(pca_res, data = meta, colour = "Lewis", frame = TRUE, frame.type = 'norm')
autoplot(pca_res, data = meta, colour = "Secretor", frame = TRUE, frame.type = 'norm')
autoplot(pca_res, data = meta, colour = "Days", frame = TRUE, frame.type = 'norm')

# Sqrt normalised

autoplot(pca_res_sqrt, data = meta, colour = "Lewis", fill=NA,frame = TRUE, frame.type = 'norm') +
  scale_color_manual(values=col_fil) +
  scale_fill_manual(values=col_fil)

sl(metab.pca.milk.secretor,
autoplot(pca_res_sqrt, data = meta, colour = "Secretor") +
  scale_color_manual(values=col_fil) +
  stat_ellipse(aes(color=Secretor),geom = "polygon", level = 0.95, fill = NA, size = 1) +
  ggplot2::annotate("text", x = -0.08, y = 0.12, label = "p < 0.0001"),
  overwrite = T
  )

metab.pca.milk.secretor

sl(metab.pca.milk.days,
  autoplot(pca_res_sqrt, data = meta, colour = "Days") +
  scale_color_manual(values=col_fil[4:1],) +
  stat_ellipse(aes(color=Days),geom = "polygon", level = 0.95, fill = NA, size = 1) +
  ggplot2::annotate("text", x = -0.02, y = 0.12, label = "All: p < 0.0001      Pairwise: p < 0.004"),
  overwrite = T
  )+
  ggtitle("Milk metabolites")

metab.pca.milk.days

autoplot(pca_res_sqrt, data = meta, colour = "delta.whz", frame = TRUE, frame.type = 'norm')
autoplot(pca_res_sqrt, data = meta, colour = "BMI.delta.3m", frame = TRUE, frame.type = 'norm')


pca_res_sqrt.bmi <- prcomp(df.sq[!is.na(meta$BMI.group),], scale. = FALSE, center = TRUE) #Apply UV-scaling

sl(metab.pca.milk.bmi,
  autoplot(pca_res_sqrt.bmi, data = meta[!is.na(meta$BMI.group),], colour = "BMI.group") +
  scale_color_manual(values=col_fil[1:3],name="BMI group") +
  stat_ellipse(aes(color=BMI.group),geom = "polygon", level = 0.95, fill = NA, size = 1) +
  ggplot2::annotate("text", x = 0, y = 0.12, label = "NW-OW:*NW-OB:*OW-OB:NS"),
  overwrite = T
  ) +
  ggtitle("Milk metabolites")

metab.pca.milk.bmi

autoplot(pca_res_sqrt, data = meta, colour = "Gestational.age", frame = TRUE, frame.type = 'norm')

autoplot(pca_res_sqrt, data = meta, colour = "Days", shape="Secretor",frame = TRUE, frame.type = 'norm')

######################GGbiplot

#install_github("vqv/ggbiplot")
library(ggbiplot)

ggbiplot::ggbiplot(pca_res_sqrt)

ggbiplot(pca_res_sqrt,ellipse=TRUE,choices=c(3,4),labels=rownames(meta), groups=as.factor(meta$Lewis), var.axes=TRUE) #Loadings + points

#ggbiplot(pca_res,ellipse=TRUE,choices=c(3,4),labels=rownames(data), groups=data$Age, var.axes=FALSE) #No loadings

################Styling

biplot <- ggbiplot(pca_res_sqrt,ellipse=TRUE,choices=c(3,4),labels=rownames(meta), groups=meta$BMI.group, var.axes=TRUE)

biplot +
  scale_colour_jco(name = "BMI") +
  #scale_fill_manual(values=rep(col_fil,5),name="Genus")+
  theme_classic() +
  theme(
    #text = element_text(linewidth = 10, colour = "Black"),
        axis.line=element_line(size=0.5),
        #panel.border = element_blank(),
        axis.text=element_text(size = 10, colour = "Black"),
        axis.ticks=element_line(size=1, colour = "Black"),
        #strip.background = element_rect(colour = "white", fill = "white"),
        #axis.text.x=element_text(angle = 45, hjust = 1),
        strip.text.x = element_text(angle = 0, size=12, face = "bold")
        #legend.position = "none"
  )

        
```

Pairwise PERMANOVA

```{r}
dist <- df %>%
  dplyr::mutate(across(where(is.numeric), ~ifelse(. < 0, NA_real_, sqrt(.)))) %>%
  dist(.,"euclidean")

```

Days
```{r}
RVAideMemoire::pairwise.perm.manova(dist ,meta$Days,nperm=9999)
```

BMI
```{r}
RVAideMemoire::pairwise.perm.manova(dist ,meta$BMI.group,nperm=9999)
```

## Loadings plot and variable contributions

```{r}
library(factoextra)

res.pca <- pca_res_sqrt

#fviz_pca_biplot(res.pca, repel = TRUE)



sl(metab.pca.var,fviz_pca_var(res.pca, col.var="contrib",
             geom ="text",
             gradient.cols = c("darkgrey","black"),
             repel = TRUE # Avoid text overlapping
             ), overwrite = T)

metab.pca.var

sl(metab.res.pca1,fviz_contrib(res.pca, choice = "var", axes = 1, top = 10), overwrite = T)
metab.res.pca1

sl(metab.res.pca2,fviz_contrib(res.pca, choice = "var", axes = 2, top = 10), overwrite = T)
metab.res.pca2

```

# Effect size

## Sqrt

```{r, eval = T, include = T, echo = F, message = F}
#import.taxonomy = read.table('virome_taxonomy_LCA_known.txt', header = T)

cure.data <- meta
rownames <- rownames(cure.data)

#Remove NA columns
cure.data <- cure.data[ , ! apply( cure.data , 2 , function(x) all(is.na(x)) ) ]

#Set numeric column class
cure.data <- cure.data %>% plyr::mutate(across(where(is.character)),as.factor)

#Sqrt normalise and transpose
X <- df %>%
  dplyr::mutate(across(where(is.numeric), ~ifelse(. < 0, NA_real_, sqrt(.))))

names <- colnames(X)
rownames <- rownames(X)
X <- X %>% as.matrix() %>% t() %>% as.data.frame()
rownames(X) <- names
colnames(X) <- rownames

Y.all = cure.data

#Select relevant numeric column

Y <- dplyr::select(Y.all,all_of(
  append(effect.size.variables,"RCID")) #Omit batch+dna conc variables as these are unique to sequencing data
  ) %>% 
  tibble::column_to_rownames("RCID")

#Y <- Y.all %>% dplyr::select_if(is.factor)

#,"Weight_sept_2020", "Glucose_sept_GGT" , "Insulin_sept_GGT", "Cpeptide_sept_GGT", "HbA1c", "Glucagon_sept_GGT", "HOMA_IR"

colnames(Y) #Check remaining columns

###Select only numeric metadata culumns

Y  <- Y %>%dplyr::mutate(across(everything(), as.numeric))

colnames(Y) #Check remaining columns

##vifcor
##collinearity

#Select relevant columns for that are non-collinear
# 
# 
# Y.ad = Y %>% as.data.frame()
# #Y.ad <- dplyr::select(Y,-c("HMO.group"))
# 
# # Remove samples with missing metadata
# Y.ad <- Y.ad %>% drop_na() %>% as.data.frame()
# ## Remove corresponding samples from metadata
# colnames(X) <- df.id$RCID
# X.ad <- X %>% dplyr::select(rownames(Y.ad))

```

#### Bray curttis - constrained by time

```{r, eval = T, include = T, echo = F, message = F, warning =  F, fig.width=10, fig.height=6, fig.fullwidth = TRUE}
###############Anova on capscale - all variables

anova.m <- matrix(nrow = 0, ncol =6) #Create matrix for results
colnames(anova.m) <- c("Df","SumOfSqs","F","Pr.F","Residual.Df","Residual.SumOfSqs")                
for (i in colnames(dplyr::select(Y,-c("Days")))){
  X.sub <- X[,!is.na(Y[[i]])]
  Y.sub <- Y[!is.na(Y[[i]]), ]
  anova <- as.matrix(anova.cca(capscale(t(X.sub) ~ Y.sub[,i] + Days, Y.sub, dist = "bray"),permutations = how(nperm=9999)))
  anova <- c(anova[1,],anova[2,1:2])
  anova.m <-rbind(anova.m,anova)
}

rownames(anova.m) <- colnames(dplyr::select(Y,-c("Days")))

#Prepare for plotting

anova.dat <- anova.m %>%
  as.data.frame() %>%
  tibble::add_column(Variable = rownames(anova.m),.name_repair = "minimal") %>%
  arrange(F)

#Add adjusted p-values
anova.dat$Pr.F.adj <- p.adjust(anova.dat$Pr.F, method = 'holm')


#Calculate sum of squares
anova.dat$R2 <- anova.dat$SumOfSqs / (anova.dat$SumOfSqs + anova.dat$Residual.SumOfSqs)

anova.dat <- anova.dat %>%
  arrange(R2)

##Fancy plot

anova.melt <- tidyr::pivot_longer(anova.dat,cols=c("R2"), names_to='measure', values_to="value") %>%
  add_significance(p.col = "Pr.F.adj", output.col = "PrF.sym",
                   cutpoints = c(0, 1e-04, 0.001, 0.01, 0.05, 0.1,1),
                   symbols = c("****", "***", "**", "*", "\U0387", "")
                   )

anova.melt.metab.milk.time <- anova.melt

permanova.metab.milk.day <- ggbarplot(anova.melt,
                                         x="Variable",
                                         y="value",
                                         #color = "Variable",
                                         orientation = c("horizontal"),
                                         color = NA,
                                         fill = "measure",
                                         position=position_dodge(0.7),
                                         label = anova.melt$PrF.sym,
                                         lab.size = 9,
                                         lab.vjust = 0.7,
                                         lab.hjust = -0.3
                                         
) +
  #stat_pvalue_manual(
  #  adonis.melt,  label = "PrF.sym", tip.length = 0.01
  #) +
  #geom_hline(yintercept = 0.05) +
  ggtitle("Constrained for time") +
  ylab(expression(R^2~"value")) +
  #scale_y_log10()+
  scale_fill_jco()

permanova.metab.milk.day

save(anova.melt.metab.milk.time,file="results/effect_size_milk_metab.RData")
```

#### Bray curttis - constrained by time + lewis +secretor

```{r, eval = T, include = T, echo = F, message = F, warning =  F, fig.width=10, fig.height=6, fig.fullwidth = TRUE}
###############Anova on capscale - all variables

anova.m <- matrix(nrow = 0, ncol =6) #Create matrix for results
colnames(anova.m) <- c("Df","SumOfSqs","F","Pr.F","Residual.Df","Residual.SumOfSqs")                
for (i in colnames(dplyr::select(Y,-c("Days","Lewis","Secretor")))){
  X.sub <- X[,!is.na(Y[[i]])]
  Y.sub <- Y[!is.na(Y[[i]]), ]
  anova <- as.matrix(anova.cca(capscale(t(X.sub) ~ Y.sub[,i] + Days+Secretor+Lewis, Y.sub, dist = "bray"),permutations = how(nperm=9999)))
  anova <- c(anova[1,],anova[2,1:2])
  anova.m <-rbind(anova.m,anova)
}

rownames(anova.m) <- colnames(dplyr::select(Y,-c("Days","Lewis","Secretor")))

#Prepare for plotting

anova.dat <- anova.m %>%
  as.data.frame() %>%
  tibble::add_column(Variable = rownames(anova.m),.name_repair = "minimal") %>%
  arrange(F)

#Add adjusted p-values
anova.dat$Pr.F.adj <- p.adjust(anova.dat$Pr.F, method = 'holm')


#Calculate sum of squares
anova.dat$R2 <- anova.dat$SumOfSqs / (anova.dat$SumOfSqs + anova.dat$Residual.SumOfSqs)

anova.dat <- anova.dat %>%
  arrange(R2)

##Fancy plot

anova.melt <- tidyr::pivot_longer(anova.dat,cols=c("R2"), names_to='measure', values_to="value") %>%
  add_significance(p.col = "Pr.F.adj", output.col = "PrF.sym",
                   cutpoints = c(0, 1e-04, 0.001, 0.01, 0.05, 0.1,1),
                   symbols = c("****", "***", "**", "*", "\U0387", "")
                   )

anova.melt.metab.milk.day.lewis.secretor <- anova.melt

permanova.metab.milk.day.lewis.secretor <- ggbarplot(anova.melt,
                                         x="Variable",
                                         y="value",
                                         #color = "Variable",
                                         orientation = c("horizontal"),
                                         color = NA,
                                         fill = "measure",
                                         position=position_dodge(0.7),
                                         label = anova.melt$PrF.sym,
                                         lab.size = 9,
                                         lab.vjust = 0.7,
                                         lab.hjust = -0.3
                                         
) +
  #stat_pvalue_manual(
  #  adonis.melt,  label = "PrF.sym", tip.length = 0.01
  #) +
  #geom_hline(yintercept = 0.05) +
  ggtitle("Constrained for time+Lewis+Secretor") +
  ylab(expression(R^2~"value")) +
  #scale_y_log10()+
  scale_fill_jco()

permanova.metab.milk.day.lewis.secretor

save(permanova.metab.milk.day.lewis.secretor,file="results/effect_size_milk_metab_sec.RData")
```

#### Secretor + lewis constrained comparison

```{r, eval = T, include = T, echo = F, message = F, warning =  F, fig.width=10, fig.height=6, fig.fullwidth = TRUE}
ggarrange(permanova.metab.milk.day,permanova.metab.milk.day.lewis.secretor)
```

# Rabuplot for metabolites

```{r}


samples <- sample_data(meta)
rownames(samples) <- samples$RCID


tax <- phyloseq::tax_table(as.data.frame(colnames(df)))

otu <- otu_table(log10(df+1)
                 ,taxa_are_rows=FALSE)

colnames(otu) <- rownames(tax)
rownames(otu) <- row.names(samples)

sl(psm,phyloseq(otu,tax,samples),overwrite=TRUE)
```

```{r}
#remotes::install_github("jstokholm/rabuplot")
library(rabuplot)

#ps <- subset_samples(PSB.all, Days %in% c("3","30","60","90"))
psm@sam_data$Days <- factor(psm@sam_data$Days , levels = rev(c("3","30","60","90")))

psm@sam_data$BMI.group <- factor(psm@sam_data$BMI.group , levels = rev(levels(psm@sam_data$BMI.group)))

psm@sam_data$Secretor <- ifelse(psm@sam_data$Secretor == 1, "Secretor", "Non-secretor")

tax <- as.data.frame(phyloseq::tax_table(psm)) %>%
 dplyr::mutate(`metabolites` = ta1)
psm@tax_table <- phyloseq::tax_table(as.matrix(tax))
```

## Metabolites vs days

```{r}
sl(rabu.metab.milk.days,
   rabuplot(
  psm,
  violin = FALSE,
  log = TRUE,
  #log_max = 200,
  type = "metabolites",
  relative_abun = FALSE,
  N_taxa = 15,
  predictor = "Days", 
  stats = "non-parametric",
  italic_names = FALSE,
  p_val = TRUE,
  p_stars =TRUE,
  text_angle_x = 45,
  colors = col_fil[1:4],
  main = "Milk metabolites - days",
  xlabs = "mmol/ml",
  legend_title = "Days after birth")+
  # scale_y_continuous(limits=c(0,2.5),labels = scales::label_number(accuracy = 1,scale = 100)), 
  scale_y_log10(breaks=c(.001,.01,0.3,1,1.7,3,9),labels=c("0","1","2","5","50","200", "P-value")),
    overwrite=T)


rabu.metab.milk.days
```

```{r,eval=F,inculde=F}
sl(rabu.metab.milk.days,
   rabuplot(
  psm,
  violin = FALSE,
  log = TRUE,
  #log_max = 200,
  type = "metabolites",
  relative_abun = FALSE,
  N_taxa = 15,
  predictor = "Days", 
  stats = "non-parametric",
  italic_names = FALSE,
  p_val = TRUE,
  p_stars =TRUE,
  text_angle_x = 45,
  colors = col_fil[1:4],
  main = "",
  xlabs = "mmol/ml",
  legend_title = "Days after birth")+
  # scale_y_continuous(limits=c(0,2.5),labels = scales::label_number(accuracy = 1,scale = 100)), 
  scale_y_log10(breaks=c(.001,.01,0.3,1,1.7,3,9),labels=c("0","1","2","5","50","200", "P-value")),
    overwrite=T)

rabu.metab.milk.days
```

### Days by BMI group

```{r,eval=T,inculde=T}
sl(rabu.metab.milk.bmi,
   rabuplot(
  psm,
  violin = FALSE,
  log = TRUE,
  #log_max = 200,
  type = "metabolites",
  relative_abun = FALSE,
  N_taxa = 15,
  predictor = "BMI.group", 
  stats = "non-parametric",
  italic_names = FALSE,
  p_val = TRUE,
  p_stars =TRUE,
  text_angle_x = 45,
  colors =  col_fil[3:1],
  main = "",
  xlabs = "mmol/ml",
  legend_title = "Pre-pregnancy BMI")+
  # scale_y_continuous(limits=c(0,2.5),labels = scales::label_number(accuracy = 1,scale = 100)), 
  scale_y_log10(breaks=c(.001,.01,0.3,1,1.7,3,9),labels=c("0","1","2","5","50","200", "P-value")),
    overwrite=T)

rabu.metab.milk.bmi
```

```{r,eval=F,inculde=F}
sl(rabu.metab.milk.bmi.roel,
   rabuplot(
  psm,
  violin = FALSE,
  log = TRUE,
  #log_max = 200,
  type = "metabolites",
  relative_abun = FALSE,
  N_taxa = 15,
  predictor = "BMI.group", 
  stats = "non-parametric",
  italic_names = FALSE,
  p_val = TRUE,
  p_stars =TRUE,
  text_angle_x = 45,
  colors = c("darkgreen","darkgoldenrod","darkred"),
  main = "",
  xlabs = "mmol/ml",
  legend_title = "Pre-pregnancy BMI")+
  # scale_y_continuous(limits=c(0,2.5),labels = scales::label_number(accuracy = 1,scale = 100)), 
  scale_y_log10(breaks=c(.001,.01,0.3,1,1.7,3,9),labels=c("0","1","2","5","50","200", "P-value")),
    overwrite=T)

rabu.metab.milk.bmi.roel
```

## Metabolites vs secretor status

```{r}
psm.pseu <- psm
psm.pseu@otu_table <- otu_table(psm.pseu) + 0.0001 # Addd small pseudocount to allow significance detection for metabolites with concentration 0 in one secretor group

sl(rabu.metab.milk.sec,
  rabuplot(
  psm,
  violin = FALSE,
  log = TRUE,
  #log_max = 200,
  type = "metabolites",
  relative_abun = FALSE,
  N_taxa = 15,
  predictor = "Secretor", 
  stats = "non-parametric",
  italic_names = FALSE,
  p_val = TRUE,
  p_stars =TRUE,
  p_adjust = TRUE,
  text_angle_x = 45,
  colors = col_fil[1:2],
  main = "",
  xlabs = "mmol/ml",
  legend_title = "Secretor status")+
  # scale_y_continuous(limits=c(0,2.5),labels = scales::label_number(accuracy = 1,scale = 100)), 
  scale_y_log10(breaks=c(.001,.01,0.3,1,1.7,3,9),labels=c("0","1","2","5","50","200", "P-value")),
    overwrite=T)

rabu.metab.milk.sec
```

### Arrange

```{r}
ggarrange(rabu.metab.milk.days,rabu.metab.milk.sec,nrow=1)
```



### Top 40

```{r}
psm.bmi <- subset_samples(psm, !is.na(BMI.group))
psm.bmi@sam_data$BMI.group <- factor(psm.bmi@sam_data$BMI.group , levels = rev(levels(psm.bmi@sam_data$BMI.group)))

sl(rabu.metab.milk.bmi.40,rabuplot(
  psm.bmi,
  violin = FALSE,
  log = TRUE,
  type = "metabolites",
  relative_abun = FALSE,
  N_taxa = 75,
  predictor = "BMI.group", 
  italic_names = FALSE,
  stats = "parametric",
  p_stars =TRUE,
  colors = col_fil[3:1],
  text_angle_x = 45,
  legend_title = "BMI group",
  main = "Milk metabolites",
  xlabs = "mmol/ml")+
  # scale_y_continuous(limits=c(0,2.5),labels = scales::label_number(accuracy = 1,scale = 100)), 
  scale_y_log10(breaks=c(.001,.01,0.3,1,1.7,3,9),labels=c("0","1","2","5","50","200", "P-value")),
    overwrite=T)

rabu.metab.milk.bmi.40
```

### By days

```{r}
psm.rev <- psm.bmi
psm.rev@sam_data$Days <- factor(psm.rev@sam_data$Days, levels = rev(levels(psm.rev@sam_data$Days)))

rabu.metab.milk.bmi.days <- rabuplot(
  psm.rev,
  violin = FALSE,
  log = TRUE,
  type = "metabolites",
  relative_abun = FALSE,
  N_taxa = 20,
  predictor = "BMI.group", 
  italic_names = FALSE,
  stats = "parametric",
  p_stars =TRUE,
  colors = col_fil[3:1],
  main = "Milk metabolites",
  xlabs = "mmol/ml",
  facet_wrap = "Days")+
  # scale_y_continuous(limits=c(0,2.5),labels = scales::label_number(accuracy = 1,scale = 100)), 
  scale_y_log10(breaks=c(.001,.01,0.3,1,1.7,3,9),labels=c("0","1","2","5","50","200", "P-value"))

rabu.metab.milk.bmi.days
```

# RDA plot

```{r}
sl(psm,phyloseq(otu,tax,samples),overwrite=FALSE)

#psm <- phyloseq(otu,tax,samples)

meco_dat <- phyloseq2meco(
  psm)

meco_dat$tidy_dataset()

#Tidy up taxonomy
meco_dat$tax_table <- tidy_taxonomy(meco_dat$tax_table)

#Calculate relative abundences
meco_dat$cal_abund()

#Set days as factor
meco_dat$sample_table$Days <- factor(meco_dat$sample_table$Days,levels=c("3","30","60","90"))

#Calculta beta diversity metrics

meco_dat$cal_betadiv(unifrac = FALSE)
```
# Explainable class

Creating trans_env object 

```{r, warning = F}
#Select most significant variables
top.metab.vars <- anova.melt.metab.milk.time %>%
  arrange(-value) %>%
  slice_head(n = 10)



env <- select(meco_dat$sample_table,all_of(c(top.metab.vars$Variable,"Days"))) %>%
  mutate(Days = as.numeric(as.character(Days))
         )  %>%
 dplyr::mutate(Days = as.numeric(as.character(Days)),
         Secretor = as.numeric(as.character(Secretor)),
         Lewis = as.numeric(as.character(Lewis)),) %>%
  #select(-c("AB.infant")) %>%
  dplyr::mutate_all(as.numeric) %>% #select(where(is.numeric)) %>%
  #drop_na(Gestational.age) %>%
  drop_na() %>%
  select(where(~ !any(is.na(.))))


colnames(env) <- str_replace_all(colnames(env), c("\\." = " ",
                                                              `BMI delta`="delta BMI",
                                                              delta="Δ",
                                                              `3m`="3 months",
                                                              waz = "WAZ",
                                                              whz = "WHZ",
                                                              haz = "HAZ",
                                                              AB="Antibiotics",
                                                              GDM="Gestational diabetes",
                                                              Secretor="Secretor status",
                                                              Lewis="Lewis status")
                    )
  
t1 <- trans_env$new(dataset = meco_dat,
                    add_data = env)
```

## PCA RDA

```{r, warning = F}
# use bray-curtis distance for dbRDA
t1$cal_ordination(method ="RDA",scale_species = TRUE, taxa_level = "ta1")
# t1$res_rda is the result list stored in the object

t1$trans_ordination(adjust_arrow_length = TRUE,
                    min_perc_env = 0.2,
                    max_perc_env = 1.5,
                    min_perc_tax = 2,
                    max_perc_tax = 4)
# t1$res_rda_trans is the transformed result for plotting
sl(rda.plot.metab,
   t1$plot_ordination(plot_color = "Days",
                   plot_type = c("point"),
                   env_text_size = 4,
                   taxa_text_size = 4,
                   taxa_text_color="darkred",
                   taxa_arrow_color="darkred",
                   point_alpha=0.2,
                   taxa_nudge_x = c(-0.05,0.1,-0.1,-0.3,-0.3,0,-0.05,0,0,0.1),
                   taxa_nudge_y = c(-0.1, #2-fl
                                    -0.03, #3-FL
                                    0, #HMO 5.10
                                    0.07, #LNDFH I
                                    0.02, #HMO 5.29
                                    0, #LNT
                                    0, #HMO 1.70
                                    0.2, #LNDFH II
                                    -0.2, #LDFT
                                    0) # Glutamine
                   )+
  ggrepel::geom_text_repel(force=30),
  overwrite = TRUE
)

rda.plot.metab
```

## PCOA RDA

```{r, warning = F}
# use bray-curtis distance for dbRDA
t1$cal_ordination(method = "dbRDA", use_measure = "bray")
# t1$res_rda is the result list stored in the object
t1$trans_ordination(adjust_arrow_length = TRUE, max_perc_env = 1.5)
# t1$res_rda_trans is the transformed result for plotting
t1$plot_ordination(plot_color = "Days", plot_type = c("point"))
```


## Metabolites vs biomarkers

```{r, warning = F}
env <- select(meco_dat$sample_table,matches("bio_")) %>%
  drop_na()

t1 <- trans_env$new(dataset = meco_dat,
                    add_data = env)
```

```{r, warning = F}
# 'p_adjust_type = "Env"' means p adjustment is performed for each environmental variable separately.
t1$cal_cor(use_data = "ta1", p_adjust_method = "fdr", p_adjust_type = "Env")
```

Then, we can plot the correlation results using plot_cor function.

```{r, warning = F}
# default ggplot2 method with clustering
#t1$plot_cor()
```

There are too many genera. We can use the filter_feature parameter in plot_cor to filter some taxa that do not have any significance < 0.001.

```{r, warning = F}
# filter genera that donot have at least one ** or ***
t1$plot_cor(filter_feature = c("", "*"), cluster_ggplot = "both")
```



# Session info

```{r}
sessionInfo()
```
