---
title: "5.1_Metabolome_microbiome_relations"
author: "Rasmus Riemer Jakobsen"
date: "`r Sys.time()`"
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
link-citations: yes
---

```{r setup, include=F, message=F, warning=F}
subdir_name <- "5.1_Metabolome_microbiome_relations"

#Load packages
source("Supplementary_scripts/package_load_install.R")
#Load utility functions
source("Supplementary_scripts/utils.R")

dir.create("results", showWarnings = FALSE, recursive = TRUE)

# Set theme and colort

theme_set(theme_classic())

knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.fullwidth = TRUE)


```

```{r}
source(here::here("Supplementary_scripts","package_load_install.R")) 
source(here::here("Supplementary_scripts","utils.R")) # please see the src-folder for some custom functions used in 
#source(here::here("src", "run_clusterprofiler.R"))
```

```{r knit, echo=F, eval=FALSE}
#Run to generate
rmarkdown::render(
  input = here::here(str_c(subdir_name, ".Rmd")),
  output_dir = here::here("results", "reports")
)
```

# Load data

```{r}
#Microbiome
load("0.3_deduplicated_and_decontaminated.RData")

#Metabolome
metab.id <- readr::read_tsv("Data/metabolome_milk.txt")
meta <- readr::read_tsv("Data/metadata_metabome_milk.txt")

## Check if dataframes match
setdiff(metab.id$RCID,meta$RCID)

metab <- metab.id %>% dplyr::select(-RCID) 

log.metab <- 
  log10(metab+1) 

# Set key metadata columns to factor
meta <- meta %>%
  mutate_at(
    c("Days","Siblings", "GDM", "Induction", "AB.mother", "AB.infant", "Breast.infection", "Breastfeeding.issues", "Secretor", "Lewis"),
    as.factor)
```


# Combined effect size plots

## Import permanova results

```{r}
load(file="results/effect_size_milk_metab.RData")

load(file="results/effect_size_bac_milk_faeces.RData")
```

## Combine matrices

```{r}
anova.melt.metab.milk.time$Component <- "Milk metabolites"
anova.melt.comb$Variable.clean <- gsub("\\."," ",anova.melt.comb$Variable)
anova.melt.metab.milk.time$Variable.clean <- gsub("\\."," ",anova.melt.metab.milk.time$Variable)

anova.melt.comb <- rbind(anova.melt.metab.milk.time,anova.melt.comb)

anova.melt.comb <- anova.melt.comb %>% 
  filter(!str_detect(Variable, "DNA.conc")) #Remove DNA conc as it only pertains to 16S data
```


## Combined plot

### Non-scaled

```{r}
library(ggbreak) 

anova.ordered <- anova.melt.comb %>% 
    mutate(Variable.clean.comp = tidytext::reorder_within(Variable.clean, value, within = Component)) %>%
    mutate(Variable.clean = str_remove(forcats::fct_inorder(Variable.clean.comp), "__.*")) %>%
    mutate(Variable.clean = str_replace_all(Variable.clean, c(`BMI delta`="delta BMI",
                                                              ppBMI = "Pre-pregnancy BMI",
                                                              delta="Δ",
                                                              `3m`="3 months",
                                                              waz = "WAZ",
                                                              whz = "WHZ",
                                                              haz = "HAZ",
                                                              AB="Antibiotics",
                                                              GDM="Gestational diabetes",
                                                              Secretor="Secretor status",
                                                              Lewis="Lewis status",
                                                              `C section`="C-section"))
           )



#anova.ordered$Variable.clean

anova.ordered <- anova.ordered %>%
  mutate(Category = case_when(
    Variable.clean == "Antibiotics infant" ~ "Antibiotics",
    Variable.clean == "Gestational diabetes" ~ "Maternal",
    Variable.clean == "Infant sex" ~ "Infant",
    Variable.clean == "Breast infection" ~ "Breastfeeding",
    Variable.clean == "Breastfeeding issues" ~ "Breastfeeding",
    Variable.clean == "Maternal age" ~ "Maternal",
    Variable.clean == "Siblings" ~ "Maternal",
    Variable.clean == "Antibiotics mother" ~ "Antibiotics",
    Variable.clean == "Birth weight" ~ "Infant",
    Variable.clean == "Induction" ~ "Maternal",
    Variable.clean == "Pre-pregnancy BMI" ~ "Maternal",
    Variable.clean == "Δ BMI 3 months" ~ "Maternal",
    Variable.clean == "Δ WHZ" ~ "Infant",
    Variable.clean == "Gestational age" ~ "Infant",
    Variable.clean == "Δ WAZ" ~ "Infant",
    Variable.clean == "Δ HAZ" ~ "Infant",
    Variable.clean == "Lewis status" ~ "Genetic",
    Variable.clean == "Secretor status" ~ "Genetic",
    Variable.clean == "C-section" ~ "Birth mode",
    TRUE ~ NA_character_  # For any values not specified above
  ))

# Add categories



anova.milk.metab <- subset(anova.ordered,Component=="Milk metabolites")
anova.milk.bac <- subset(anova.ordered,Component=="Milk bacteria")
anova.faec.bac <- subset(anova.ordered,Component=="Faecal bacteria")

y.lab <- expression(R^2~value)

p.milk.metab <- ggbarplot(anova.milk.metab,
                                         x="Variable.clean",
                                         y="value",
                                         ylab = FALSE,
                                         fill = "Category",
                                         orientation = c("horizontal"),
                                         #fill = "measure",
                                         position=position_dodge(0.7),
                                         label = anova.milk.metab$PrF.sym,
                                         lab.size = 9,
                                         lab.vjust = 0.8,
                                         lab.hjust = -0.3
) +
  scale_fill_manual(values=col_fil)  +
  theme(legend.position = "none",
        axis.text.x.top = element_blank(),
        axis.line.x.top = element_blank(),
        axis.ticks.x.top = element_blank(),
        axis.text.y = element_text(angle = 35, vjust = 0.5, hjust=1)
        ) +
  #scale_y_break(c(0.1,0.43),expand=FALSE) +
  #scale_y_continuous(limits=c(0,0.45),breaks=c(0.025,0.05,0.075,0.44))
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  ggtitle("Milk metabolites")+
  labs(y=y.lab)

p.milk.metab                    

p.milk.bac <- ggbarplot(anova.milk.bac,
                                         x="Variable.clean",
                                         y="value",
                                         ylab = FALSE,
                                         fill = "Category",
                                         orientation = c("horizontal"),
                                         #fill = "measure",
                                         position=position_dodge(0.7),
                                         label = anova.milk.bac$PrF.sym,
                                         lab.size = 9,
                                         lab.vjust = 0.8,
                                         lab.hjust = -0.3
) +
  scale_fill_manual(values=col_fil) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.275))) +
  theme(legend.position = "none",
        axis.text.y = element_text(angle = 35, vjust = 0.5, hjust=1)) +
  ggtitle("Milk bacteria")+
  labs(y=y.lab)



p.faec.bac <- ggbarplot(anova.faec.bac,
                                         x="Variable.clean",
                                         y="value",
                                         ylab = FALSE,
                                         fill = "Category",
                                         orientation = c("horizontal"),
                                         #fill = "measure",
                                         position=position_dodge(0.7),
                                         label = anova.milk.bac$PrF.sym,
                                         lab.size = 9,
                                         lab.vjust = 0.8,
                                         lab.hjust = -0.3
) +
  scale_fill_manual(values=col_fil) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.35))) +
  theme(legend.position = "none",
        axis.text.y = element_text(angle = 35, vjust = 0.5, hjust=1)) +
  ggtitle("Infant faecal bacteria")+
  labs(y=y.lab)


sl(esize.bac.metab,
   list(p.milk.metab+theme(plot.margin = unit(c(0,0,0.5,0.2), "cm")),
        p.milk.bac+theme(plot.margin = unit(c(0,0,0.5,0.1), "cm")),
        p.faec.bac+theme(plot.margin = unit(c(0,0,0.5,0.1), "cm"))
        ),
   overwrite = TRUE)

esize.bac.metab

```


### With Breaks

```{r}
library(ggbreak) 


esize.bac.metab.breaks <- ggbarplot(anova.melt.comb,
                                         x="Variable.clean",
                                         y="value",
                                        ylab = FALSE,
                                         fill = "Component",
                                         orientation = c("horizontal"),
                                         #fill = "measure",
                                         position=position_dodge(0.7),
                                         label = anova.melt.comb$PrF.sym,
                                         lab.size = 9,
                                         lab.vjust = 0.8,
                                         lab.hjust = -0.3
                                         
) +
  scale_y_break(c(0.022,0.063), ticklabels=c(0.065,0.07)) +
  scale_y_break(c(0.075,0.43), ticklabels=c(0.43,0.44)) +
  #scale_y_continuous(trans = log_trans()) +
  ylab(expression(R^2~"value")) +
  scale_fill_jco()

esize.bac.metab.breaks
```

# PROCrustes comparison matrix correlation quantifications

## Subset samples found in both matrices

```{r}
# test <- data.frame(sample_data(PSB.CSS)) %>%
#   as.data.frame() %>% dplyr::arrange(Sample.type,RCID)
# 
# test$RCID
# meta$RCID

#Bacteria
PSB.CSS.milk <- subset_samples(PSB.CSS, Sample.type == "Milk") 
PSB.CSS.faeces <- subset_samples(PSB.CSS, Sample.type == "Faecal sample")


colnames(PSB.CSS.milk@otu_table)



PSB.CSS.milk.both.bac <- subset_samples(PSB.CSS.milk, RCID %in% PSB.CSS.faeces@sam_data$RCID)
PSB.CSS.milk.both.bac
PSB.CSS.milk.both.metab <- subset_samples(PSB.CSS.milk, RCID %in% meta$RCID)
PSB.CSS.milk.both.metab

PSB.CSS.faeces.both.bac <- subset_samples(PSB.CSS.faeces, RCID %in% PSB.CSS.milk@sam_data$RCID)
PSB.CSS.faeces.both.bac
PSB.CSS.faeces.both.metab <- subset_samples(PSB.CSS.faeces, RCID %in% meta$RCID)
PSB.CSS.faeces.both.metab

##Samples present in all three matrices
PSB.CSS.milk.all <- subset_samples(PSB.CSS.milk.both.bac, RCID %in% meta$RCID)
PSB.CSS.milk.all
PSB.CSS.faeces.all <- subset_samples(PSB.CSS.faeces.both.bac, RCID %in% meta$RCID)
PSB.CSS.faeces.all

#Metabolome
df.milk <- metab[which(meta$RCID %in% PSB.CSS.milk.both.metab@sam_data$RCID),]
df.faeces <- metab[which(meta$RCID %in% PSB.CSS.faeces.both.metab@sam_data$RCID),]
df.both <- metab[which(meta$RCID %in% PSB.CSS.milk.all@sam_data$RCID),]
str(df.both)


## Check if dataframes match
PSB.CSS.milk@sam_data$RCID
meta$RCID
setdiff(df.both$RCID,PSB.CSS.milk.all@sam_data$RCID)
```


## PCoAs

```{r}
get_pcoa <- function(ps="PSB.CSS"){
otu <- data.frame(ps@otu_table) %>% t()
bray <- vegan::vegdist(otu, method = "bray")
add <-  !(ade4::is.euclid(bray))
pcoa <- cmdscale(bray, k = nrow(otu)-1, eig = TRUE, add = add)
return(pcoa)
}

get_pca <- function(df="PSB.CSS"){
otu <- log10(df+1)
bray <- vegan::vegdist(otu, method = "euclidian")
#add <-  !(ade4::is.euclid(bray))
pca <- cmdscale(bray, k = nrow(otu)-1, eig = TRUE
                #, add = add
                )
return(pca)
}



pcoa.bac <- get_pcoa(PSB.CSS)
vegan::ordiplot(pcoa.bac, type = "text", main = "PCoA for Bacteria Data, Bray Distances")


pca.metab <- get_pca(df.both)
vegan::ordiplot(pca.metab, type = "text", main = "PCA for milk metabolome")
```

## Procrustes

### All samples

```{r, warning = F}
#All time points
pcoa.milk <- PSB.CSS.milk.both.bac %>% get_pcoa()
pcoa.faeces <- PSB.CSS.faeces.both.bac %>% get_pcoa()

sl(procrustes.test.bac,vegan::protest(X = pcoa.milk, Y = pcoa.faeces, scores = "sites", symmetric = T, permutations = 999), overwrite = FALSE)

# procrustes.test.bac

#Only samples in all datasets

pcoa.milk.all <- PSB.CSS.milk.all %>% get_pcoa()
pcoa.faeces.all <- PSB.CSS.faeces.all %>% get_pcoa()

sl(procrustes.test.bac.sub,vegan::protest(X = pcoa.milk.all, Y = pcoa.faeces.all, scores = "sites", symmetric = T, permutations = 999), overwrite = FALSE)

# procrustes.test.bac.sub

# Milk vs milk metabolites

## All samples found in bac and metab data

pcoa.milk <- PSB.CSS.milk.both.metab %>% get_pcoa()
pca.milk <- df.milk %>% get_pca()

sl(procrustes.test.milk,vegan::protest(X = pcoa.milk, Y = pca.milk, scores = "sites", symmetric = T, permutations = 999), overwrite = FALSE)

# procrustes.test.milk

## Samples with all 3 data types

pca.all <- df.both %>% get_pca()

sl(procrustes.test.milk.sub,vegan::protest(X = pcoa.milk.all, Y = pca.all, scores = "sites", symmetric = T, permutations = 999), overwrite = FALSE)

# procrustes.test.milk.sub

# Faeces vs milk metabolites

pcoa.faeces <- PSB.CSS.faeces.both.metab %>% get_pcoa()
pca.faeces <- df.faeces %>% get_pca()

sl(procrustes.test.faeces,vegan::protest(X = pcoa.faeces, Y = pca.faeces, scores = "sites", symmetric = T, permutations = 999), overwrite = FALSE)

# procrustes.test.faeces

## Samples with all 3 data types

sl(procrustes.test.faeces.sub,vegan::protest(X = pcoa.faeces.all, Y = pca.all, scores = "sites", symmetric = T, permutations = 999), overwrite = FALSE)

# procrustes.test.faeces.sub
```

```{r}
get_proc <- function(test="Vegan::protest output"){
res <- c(test$ss,test$scale,test$signif)
return(res)
}

bac_milk_bac_faeces <- get_proc(procrustes.test.bac.sub)
bac_milk_metab_milk <- get_proc(procrustes.test.milk.sub)
bac_faeces_metab_milk <- get_proc(procrustes.test.faeces.sub)

m <- matrix(nrow = 0, ncol =3)
colnames(m) <- c("Sum of Squares", "Procrustes Correlation", "Pr.F")

m <- rbind(m,bac_milk_bac_faeces,bac_milk_metab_milk,bac_faeces_metab_milk)
m

rownames(m) <- c("Milk bac vs feacal bac", "Milk bac vs milk metab", "Fecal bac vs milk metab")

#Prepare for plotting

dat <- m %>%
  as.data.frame() %>%
  tibble::add_column(Variable = rownames(m),.name_repair = "minimal") 

#Add adjusted p-values
dat$Pr.F.adj <- p.adjust(dat$Pr.F, method = 'holm')

#Show table
#knitr::kable(anova.dat)

##Fancy plot

melt <- tidyr::pivot_longer(dat,cols=c("Procrustes Correlation"), names_to='measure', values_to="value") %>%
  add_significance(p.col = "Pr.F", output.col = "PrF.sym",
                   cutpoints = c(0, 1e-04, 0.001, 0.01, 0.05, 0.1,1),
                   symbols = c("****", "***", "**", "*", "\u02D9", "")
                   ) #%>%
  #arrange(value)

procrustes.bar <- ggbarplot(melt,
                                         x="Variable",
                                         y="value",
                                         #color = "Variable",
                                         orientation = c("horizontal"),
                                         color = NA,
                                         fill = "measure",
                                         position=position_dodge(0.7),
                                         label = melt$PrF.sym,
                                         lab.size = 9,
                                         lab.vjust = 0.7,
                                         lab.hjust = -0.3
                                         
) +
  #stat_pvalue_manual(
  #  adonis.melt,  label = "PrF.sym", tip.length = 0.01
  #) +
  #geom_hline(yintercept = 0.05) +
  ggtitle("") +
  xlab("") +
  ylab("Procrustes test correlation") +
  scale_fill_jco() +
  scale_x_discrete(limits = rev)

procrustes.bar
```

### Per time point

#### Subset matrices

```{r}
#Samples present in all three matrices

## Day 3
PSB.CSS.milk.all.3 <- subset_samples(PSB.CSS.milk.both.metab, Days == 3)
#PSB.CSS.faeces.all.3 <- subset_samples(PSB.CSS.faeces.all, Days == 3)
PSB.CSS.milk.all.3
df.both.3 <- metab[which(meta$RCID %in% PSB.CSS.milk.all.3@sam_data$RCID),]

## Day 30
PSB.CSS.milk.all.30 <- subset_samples(PSB.CSS.milk.all, Days == 30)
PSB.CSS.faeces.all.30 <- subset_samples(PSB.CSS.faeces.all, Days == 30)
PSB.CSS.milk.all.30
df.both.30 <- metab[which(meta$RCID %in% PSB.CSS.milk.all.30@sam_data$RCID),]

## Day 60
PSB.CSS.milk.all.60 <- subset_samples(PSB.CSS.milk.all, Days == 60)
PSB.CSS.faeces.all.60 <- subset_samples(PSB.CSS.faeces.all, Days == 60)
PSB.CSS.milk.all.60
df.both.60 <- metab[which(meta$RCID %in% PSB.CSS.milk.all.60@sam_data$RCID),]

## Day 90
PSB.CSS.milk.all.90 <- subset_samples(PSB.CSS.milk.all, Days == 90)
PSB.CSS.faeces.all.90 <- subset_samples(PSB.CSS.faeces.all, Days == 90)
PSB.CSS.milk.all.90
df.both.90 <- metab[which(meta$RCID %in% PSB.CSS.milk.all.90@sam_data$RCID),]
```

#### Plot  

```{r}
sl(proctest.milkmet.3,vegan::protest(X = get_pcoa(PSB.CSS.milk.all.3), Y = get_pca(df.both.3), scores = "sites", symmetric = T, permutations = 999), overwrite = FALSE)
# sl(proctest.feacesmet.0,vegan::protest(X = get_pcoa(PSB.CSS.faeces.all.3), Y = get_pca(df.both.3), scores = "sites", symmetric = T, permutations = 999), overwrite = FALSE) # No day 3 faecal samples
# sl(proctest.milkfaec.0,vegan::protest(X = get_pcoa(PSB.CSS.milk.all.3), Y = get_pcoa(PSB.CSS.faeces.all.3), scores = "sites", symmetric = T, permutations = 999), overwrite = FALSE)

sl(proctest.milkmet.30,vegan::protest(X = get_pcoa(PSB.CSS.milk.all.30), Y = get_pca(df.both.30), scores = "sites", symmetric = T, permutations = 999), overwrite = FALSE)
sl(proctest.feacesmet.30,vegan::protest(X = get_pcoa(PSB.CSS.faeces.all.30), Y = get_pca(df.both.30), scores = "sites", symmetric = T, permutations = 999), overwrite = FALSE) # No day 30 faecal samples
sl(proctest.milkfaec.30,vegan::protest(X = get_pcoa(PSB.CSS.milk.all.30), Y = get_pcoa(PSB.CSS.faeces.all.30), scores = "sites", symmetric = T, permutations = 999), overwrite = FALSE)

sl(proctest.milkmet.60,vegan::protest(X = get_pcoa(PSB.CSS.milk.all.60), Y = get_pca(df.both.60), scores = "sites", symmetric = T, permutations = 999), overwrite = FALSE)
sl(proctest.feacesmet.60,vegan::protest(X = get_pcoa(PSB.CSS.faeces.all.60), Y = get_pca(df.both.60), scores = "sites", symmetric = T, permutations = 999), overwrite = FALSE) # No day 60 faecal samples
sl(proctest.milkfaec.60,vegan::protest(X = get_pcoa(PSB.CSS.milk.all.60), Y = get_pcoa(PSB.CSS.faeces.all.60), scores = "sites", symmetric = T, permutations = 999), overwrite = FALSE)

sl(proctest.milkmet.90,vegan::protest(X = get_pcoa(PSB.CSS.milk.all.90), Y = get_pca(df.both.90), scores = "sites", symmetric = T, permutations = 999), overwrite = FALSE)
sl(proctest.feacesmet.90,vegan::protest(X = get_pcoa(PSB.CSS.faeces.all.90), Y = get_pca(df.both.90), scores = "sites", symmetric = T, permutations = 999), overwrite = FALSE) # No day 90 faecal samples
sl(proctest.milkfaec.90,vegan::protest(X = get_pcoa(PSB.CSS.milk.all.90), Y = get_pcoa(PSB.CSS.faeces.all.90), scores = "sites", symmetric = T, permutations = 999), overwrite = FALSE)

# Combine outputs

m <- matrix(nrow = 0, ncol =3)
colnames(m) <- c("Sum of Squares", "Procrustes Correlation", "Pr.F")

m <- rbind(m,
           get_proc(proctest.milkmet.3),
           get_proc(proctest.milkmet.30),
           get_proc(proctest.milkmet.60),
           get_proc(proctest.milkmet.90))

m <- as.data.frame(m)
m$Days <- c("Day 3", "Day 30", "Day 60", "Day 90")
m$Comparison <- "Milk bacteria vs milk metabolome"

m.1 <- m


m <- matrix(nrow = 0, ncol =3)
colnames(m) <- c("Sum of Squares", "Procrustes Correlation", "Pr.F")

m <- rbind(m,
           get_proc(proctest.feacesmet.30),
           get_proc(proctest.feacesmet.60),
           get_proc(proctest.feacesmet.90))

m <- as.data.frame(m)
m$Days <- c("Day 30", "Day 60", "Day 90")
m$Comparison <- "Faecal bacteria vs milk metabolome"

m.2 <- m



m <- matrix(nrow = 0, ncol =3)
colnames(m) <- c("Sum of Squares", "Procrustes Correlation", "Pr.F")

m <- rbind(m,
           get_proc(proctest.milkfaec.30),
           get_proc(proctest.milkfaec.60),
           get_proc(proctest.milkfaec.90))

m <- as.data.frame(m)
m$Days <- c("Day 30", "Day 60", "Day 90")
m$Comparison <- "Milk bacteria vs faecal bacteria"

m.3 <- m

#Prepare for plotting

dat <- rbind(m.1,m.2,m.3)

# dat <- m %>%
#   as.data.frame() %>%
#   tibble::add_column(Variable = rownames(m),.name_repair = "minimal") 

#Add adjusted p-values
dat$Pr.F.adj <- p.adjust(dat$Pr.F, method = 'holm')

#Show table
#knitr::kable(anova.dat)

##Fancy plot

melt <- tidyr::pivot_longer(dat,cols=c("Procrustes Correlation"), names_to='measure', values_to="Procrustes Correlation") %>%
  add_significance(p.col = "Pr.F", output.col = "PrF.sym",
                   cutpoints = c(0, 1e-04, 0.001, 0.01, 0.05, 0.1,1),
                   symbols = c("****", "***", "**", "*", "\u02D9", "")
                   ) #%>%
  #arrange(value)

#Set order
melt$Days <- factor(melt$Days, levels = c("Day 3", "Day 30", "Day 60", "Day 90"))

procrustes.bar <- ggbarplot(melt,
                                         x="Comparison",
                                         y="Procrustes Correlation",
                                         #color = "Variable",
                                         orientation = c("vertical"),
                                         color = NA,
                                         fill = "Days",
                                         legend = "right",
                                         position=position_dodge(0.7),
                                         label = melt$PrF.sym,
                                         lab.size = 9,
                                         lab.vjust = 0.5,
                                         lab.hjust = 0.5
                                         
) +
  #stat_pvalue_manual(
  #  adonis.melt,  label = "PrF.sym", tip.length = 0.01
  #) +
  #geom_hline(yintercept = 0.05) +
  ggtitle("") +
  xlab("") +
  #ylab("Procrustes test significance") +
  scale_fill_jco() +
  scale_x_discrete(limits = rev)

procrustes.bar

#Set order
melt$Days <- factor(melt$Days, levels = rev(c("Day 3", "Day 30", "Day 60", "Day 90")))

sl(procrustes.bar.comb,ggbarplot(melt,
                                         x="Days",
                                         y="Procrustes Correlation",
                                         #color = "Variable",
                                         orientation = c("vertical"),
                                         color = NA,
                                         fill = "Comparison",
                                         legend = "none",
                                         position=position_dodge(0.7),
                                         label = melt$PrF.sym,
                                         lab.size = 9,
                                         lab.vjust = 0.5,
                                         lab.hjust = 0.5,
                                         facet.by = "Comparison",
                                         scales = "free_x",
                                         
                                         
) +
  #stat_pvalue_manual(
  #  adonis.melt,  label = "PrF.sym", tip.length = 0.01
  #) +
  #geom_hline(yintercept = 0.05) +
  ggtitle("") +
  xlab("") +
  #ylab("Procrustes test significance") +
  scale_fill_manual(values=col_fil) +
  theme(strip.background = element_rect(colour = "white", fill = "white"),
        #axis.text.x=element_text(angle = 45, hjust = 1),
        strip.text.x = element_text(size = 12, angle = 0),
        legend.text = element_text(size = 12)
        #panel.border = element_rect(fill = NA, color = NA)
  ),
overwrite = TRUE)

procrustes.bar.comb
```

# Correlation heatmaps

## Correlation test function

```{r}

# Correlation heatmap function

cor_pheatmap <- function(x="numeric matrix with RCID",y="numeric matrix with RCID", p_cutoff = 0.05, cor_cutoff = 0.1){

# Remove zero-sum rows and columns
y <- y %>%
  select(-RCID) %>%
  filter(rowSums(.,na.rm=TRUE) != 0) %>%
  dplyr::select(where(~ sum(., na.rm = TRUE) != 0)) %>%
  bind_cols(y['RCID'], .)

x <- x %>%
  select(-RCID) %>%
  filter(rowSums(.,na.rm=TRUE) != 0) %>%
  dplyr::select(where(~ sum(., na.rm = TRUE) != 0)) %>%
  bind_cols(x['RCID'], .)
  

# Align dataframes
y <- y[which(y$RCID %in% x$RCID),] %>% arrange(RCID)
x <- x[which(y$RCID %in% y$RCID),] %>% arrange(RCID)



y <- y %>% select(-RCID)
x <- x %>% select(-RCID)

# initialize empty matrices
mat_cor <- mat_pval <- mat_signif <- matrix(NA,
                                            nrow = ncol(x),
                                            ncol = ncol(y),
                                            dimnames = list(colnames(x),
                                                            colnames(y)))

# for each pair, compute one cell of the matrice
for(metab in colnames(x)){
  for(param in colnames(y)){
    
    # correlation coefficient
    mat_cor[metab, param] <- cor(x[[metab]],
                                 y[[param]],
                                 method = "pearson",
                                 use="pairwise.complete.obs" #Removal of NA values for each pairwise correlations
                                 #use="complete.obs" # All cases with missing values removed
                                 )
    
    # pvalue
    pval <- cor.test(x[[metab]],
                     y[[param]],
                     method = "pearson",
                     use="pairwise.complete.obs"
                     #use="complete.obs"
                     )[["p.value"]]
    
    # decide nb of stars
    if(pval >= 0.05){
      stars <- ""
    } else if(pval >= 0.01){
      stars <- "⁎"
    } else if(pval >= 1e-3){
      stars <- "⁎⁎"
    } else{
      stars <- "⁎⁎⁎"
    }
    
    # save pvalue and stars in the corresponding matrices
    mat_pval[metab, param] <- pval
    mat_signif[metab, param] <- stars
  }
}

# Select only significance values below threshold
mat_pval_sig <- mat_pval %>%
  as.data.frame() %>%
  dplyr::select(where(~ any(. < p_cutoff))) %>%
  dplyr::filter(rowSums(. < p_cutoff) > 0) 

#Subset matrices
mat_cor_sig <- mat_cor %>%
  as.data.frame() %>%
  select(colnames(mat_pval_sig)) %>%
  filter(row.names(mat_cor) %in% rownames(mat_pval_sig))

mat_signif_sig <- mat_signif %>%
  as.data.frame() %>%
  select(colnames(mat_pval_sig)) %>%
  filter(row.names(mat_signif) %in% rownames(mat_pval_sig)) 

#Subset correlations by threshold
mat_cor_sig_cor <- mat_cor_sig %>%
  as.data.frame() %>%
  dplyr::select(where(~ any(. > cor_cutoff | . < -cor_cutoff))) %>%
  dplyr::filter(rowSums(. < cor_cutoff | . > -cor_cutoff) > 0) 

mat_pval_sig_cor <- mat_pval_sig %>%
  as.data.frame() %>%
  select(colnames(mat_cor_sig_cor)) %>%
  filter(row.names(mat_signif_sig) %in% rownames(mat_cor_sig_cor)) %>%
  dplyr::select(where(~ any(. < p_cutoff))) %>%
  dplyr::filter(rowSums(. < p_cutoff) > 0) 

mat_cor_sig_cor <- mat_cor_sig_cor %>%
  as.data.frame() %>%
  select(colnames(mat_pval_sig_cor)) %>%
  filter(row.names(mat_cor_sig_cor) %in% rownames(mat_pval_sig_cor))

## Repeat removal of non-significant or correlating rows/columns
mat_cor_sig_cor <- mat_cor_sig_cor %>%
  as.data.frame() %>%
  dplyr::select(where(~ any(. > cor_cutoff | . < -cor_cutoff))) %>%
  dplyr::filter(rowSums(. < cor_cutoff | . > -cor_cutoff) > 0) 

mat_pval_sig_cor <- mat_pval_sig_cor %>%
  as.data.frame() %>%
  select(colnames(mat_cor_sig_cor)) %>%
  filter(row.names(mat_pval_sig_cor) %in% rownames(mat_cor_sig_cor)) %>%
  dplyr::select(where(~ any(. < p_cutoff))) %>%
  dplyr::filter(rowSums(. < p_cutoff) > 0) 

mat_cor_sig_cor <- mat_cor_sig_cor %>%
  as.data.frame() %>%
  select(colnames(mat_pval_sig_cor)) %>%
  filter(row.names(mat_cor_sig_cor) %in% rownames(mat_pval_sig_cor))

mat_signif_sig_cor <- mat_signif_sig %>%
  as.data.frame() %>%
  select(colnames(mat_cor_sig_cor)) %>%
  filter(row.names(mat_signif_sig) %in% rownames(mat_cor_sig_cor))

  
# Save objects
cor_list <- list()
cor_list$mat_cor <- mat_cor
cor_list$mat_cor_sig <- mat_cor_sig
cor_list$mat_pval <- mat_pval
cor_list$mat_pval_sig <- mat_pval_sig
cor_list$mat_signif <- mat_signif
cor_list$mat_signif_sig <- mat_signif_sig
cor_list$mat_cor_sig_cor <- mat_cor_sig_cor
cor_list$mat_signif_sig_cor <- mat_signif_sig_cor

return(cor_list)
}
```


## Maternal blood samples vs milk metabolites

```{r}
library(pheatmap)
set.seed(123)

# Select dataframes

## Blood markers

x <- meta %>%
  select(c(bio_kreak:bio_urat)) %>%
  dplyr::mutate(across(everything(), ~str_replace_all(., ">", ""))) %>%
  dplyr::mutate(across(everything(), ~str_replace_all(., ",", "."))) %>% 
  dplyr::mutate(across(everything(), as.numeric))

x <- x %>% select(-bio_crp) # Too many missing values
x$RCID <- meta$RCID # Add id's again
x <- x %>% drop_na() # alternately filer only rows with all NAs filter(rowSums(is.na(.)) < ncol(.))

## Milk metabolites
y <- metab.id



# cor <- cor_pheatmap(x,y,p_cutoff=0.01)
cor <- cor_pheatmap(x,y,p_cutoff=0.01,cor_cutoff = 0.15)

pheatmap(cor$mat_cor_sig,
         display_numbers = cor$mat_signif_sig,
         fontsize_number = 20,
         angle_col = 45)

pheatmap(cor$mat_cor_sig_cor,
         display_numbers = cor$mat_signif_sig_cor,
         fontsize_number = 20,
         angle_col = 45)
```

## Milk bacteria vs milk metabolites

```{r}
# Select dataframes

## Milk bacteria

ps <- subset_samples(PSB.genus,Sample.type=="Milk") %>%
  transform_sample_counts(., function(x) x / sum(x))
  

tax <- tax_best(ps)

ps.df <- as.data.frame(otu_table(ps)) %>%
  t() %>%
  `colnames<-`(tax$best) %>%
  as.data.frame()

col_sums <- colSums(ps.df) # Select 30 highest abundance ASVs
top_cols <- names(sort(col_sums, decreasing = TRUE)[1:30])
x <- ps.df %>%
  select(all_of(top_cols))

x <- x %>%
  mutate_all(~na_if(., 0))

x$RCID <- sample_data(ps)$RCID

## Metabolites

y <- metab.id

y <- y %>%
  select(-RCID) %>%
  select(-Acetoin) %>%
  mutate_all(~na_if(., 0)) %>%
  bind_cols(y['RCID'], .)

## Person correlations

cor <- cor_pheatmap(x,y,p_cutoff=0.01,cor_cutoff = 0.2)

pheatmap(cor$mat_cor_sig,
         display_numbers = cor$mat_signif_sig,
         fontsize_number = 20,
         angle_col = 45)

pheatmap(cor$mat_cor_sig_cor,
         display_numbers = cor$mat_signif_sig_cor,
         fontsize_number = 20,
         angle_col = 45)
```

## Faecal bacteria vs milk Metabolites

### Genus

```{r}
# Select dataframes

## Faecal bacteria bacteria

ps <- subset_samples(PSB.genus,Sample.type=="Faecal sample") %>%
  transform_sample_counts(., function(x) x / sum(x))

## Select high abundance+prevalance ASVs

# prev <- 0.03 #10/nrow(PSB.all@sam_data)
# ps = phyloseq_filter_prevalence(ps, prev.trh = prev, abund.trh = 0.005, abund.type = "mean", threshold_condition = "AND")


tax <- tax_best(ps)

ps.df <- as.data.frame(otu_table(ps)) %>%
  t() %>%
  `colnames<-`(tax$best) %>%
  as.data.frame()

x <- ps.df

# x <- ps.df %>%
#   mutate_all(~na_if(., 0))
# 


x$RCID <- sample_data(ps)$RCID

## Metabolites

y <- metab.id

# y <- y %>%
#   select(-RCID) %>%
#   select(-Acetoin) %>%
#   mutate_all(~na_if(., 0)) %>%
#   bind_cols(y['RCID'], .)

## Person correlations

cor <- cor_pheatmap(x,y,p_cutoff=0.05,cor_cutoff = 0.5)

pheatmap(cor$mat_cor_sig,
         display_numbers = cor$mat_signif_sig,
         fontsize_number = 20,
         angle_col = 45)

pheatmap(cor$mat_cor_sig_cor,
         display_numbers = cor$mat_signif_sig_cor,
         fontsize_number = 20,
         angle_col = 45)
```

### Species

```{r}
# Select dataframes

## Faecal bacteria bacteria

ps <- subset_samples(PSB.species,Sample.type=="Faecal sample") %>%
  transform_sample_counts(., function(x) x / sum(x))
  

tax <- tax_best(ps)

ps.df <- as.data.frame(otu_table(ps)) %>%
  t() %>%
  `colnames<-`(tax$best) %>%
  as.data.frame()

col_sums <- colSums(ps.df) # Select 30 highest abundance ASVs
top_cols <- names(sort(col_sums, decreasing = TRUE)[1:30])
x <- ps.df %>%
  select(all_of(top_cols))

x$RCID <- sample_data(ps)$RCID

## Metabolites

y <- metab.id

## Person correlations

cor <- cor_pheatmap(x,y,p_cutoff=0.01,cor_cutoff = 0.15)

pheatmap(cor$mat_cor_sig,
         display_numbers = cor$mat_signif_sig,
         fontsize_number = 20,
         angle_col = 45)

pheatmap(cor$mat_cor_sig_cor,
         display_numbers = cor$mat_signif_sig_cor,
         fontsize_number = 20,
         angle_col = 45)
```

## Milk vs Faecal bacteria?

```{r}
# Select dataframes

## Milk bacteria

ps <- subset_samples(PSB.genus,Sample.type=="Milk") %>%
  transform_sample_counts(., function(x) x / sum(x))
  

tax <- tax_best(ps)

ps.df <- as.data.frame(otu_table(ps)) %>%
  t() %>%
  #`colnames<-`(paste(tax$best," HM")) %>%
  `colnames<-`(tax$best) %>%
  as.data.frame()

col_sums <- colSums(ps.df) # Select 30 highest abundance ASVs
top_cols <- names(sort(col_sums, decreasing = TRUE)[1:60])
x <- ps.df %>%
  select(all_of(top_cols))

x$RCID <- sample_data(ps)$RCID

## Faecal bacteria

ps <- subset_samples(PSB.genus,Sample.type=="Faecal sample") %>%
  transform_sample_counts(., function(x) x / sum(x))
  

tax <- tax_best(ps)

ps.df <- as.data.frame(otu_table(ps)) %>%
  t() %>%
  #`colnames<-`(paste(tax$best," FS")) %>%
  `colnames<-`(tax$best) %>%
  as.data.frame()

col_sums <- colSums(ps.df) # Select 30 highest abundance ASVs
top_cols <- names(sort(col_sums, decreasing = TRUE)[1:60])
y <- ps.df %>%
  select(all_of(top_cols))

y$RCID <- sample_data(ps)$RCID


# Select only co-ocurring bacteria

y <- y[,intersect(colnames(y),colnames(x))]
x <- x[,intersect(colnames(y),colnames(x))]

## Person correlations

cor <- cor_pheatmap(x,y,p_cutoff=0.05,cor_cutoff = 0.2)

pheatmap(cor$mat_cor_sig,
         display_numbers = cor$mat_signif_sig,
         fontsize_number = 20,
         angle_col = 45)

pheatmap(cor$mat_cor_sig_cor,
         display_numbers = cor$mat_signif_sig_cor,
         fontsize_number = 20,
         angle_col = 45)

```

# Bifidobacteria correlations

## Milk bacteria vs milk metabolites

```{r}
# Select dataframes

## Milk bacteria

ps <- subset_samples(PSB.species,Sample.type=="Milk") %>%
  transform_sample_counts(., function(x) x / sum(x)) %>%
  subset_taxa(Genus == "Bifidobacterium") 
  

tax <- tax_best(ps)

ps.df <- as.data.frame(otu_table(ps)) %>%
  t() %>%
  `colnames<-`(tax$best) %>%
  as.data.frame()

# col_sums <- colSums(ps.df) # Select 30 highest abundance ASVs
#top_cols <- names(sort(col_sums, decreasing = TRUE)[1:50])
# x <- ps.df %>%
#   select(all_of(top_cols))

x <- ps.df
x$RCID <- sample_data(ps)$RCID
x <- x %>% filter(rowSums(select_if(., is.numeric)) != 0)


## Metabolites

y <- metab.id

## Person correlations

cor <- cor_pheatmap(x,y,p_cutoff=0.05,cor_cutoff = 0.1)

pheatmap(cor$mat_cor_sig,
         display_numbers = cor$mat_signif_sig,
         fontsize_number = 20,
         angle_col = 45)

pheatmap(cor$mat_cor_sig_cor,
         display_numbers = cor$mat_signif_sig_cor,
         fontsize_number = 20,
         angle_col = 45)
```

## Faecal bacteria vs milk Metabolites

```{r}
# Select dataframes

## Faecal bacteria bacteria

ps <- subset_samples(PSB.species,Sample.type=="Faecal sample") %>%
  transform_sample_counts(., function(x) x / sum(x))  %>%
  subset_taxa(Genus == "Bifidobacterium") 
  

tax <- tax_best(ps)

ps.df <- as.data.frame(otu_table(ps)) %>%
  t() %>%
  `colnames<-`(tax$best) %>%
  as.data.frame()

# col_sums <- colSums(ps.df) # Select 30 highest abundance ASVs
# top_cols <- names(sort(col_sums, decreasing = TRUE)[1:30])
# x <- ps.df %>%
#   select(all_of(top_cols))

x <- ps.df
x$RCID <- sample_data(ps)$RCID
x <- x %>% filter(rowSums(select_if(., is.numeric)) != 0)

## Metabolites

y <- metab.id

## Person correlations

cor <- cor_pheatmap(x,y,p_cutoff=0.05,cor_cutoff = 0.1)

pheatmap(cor$mat_cor_sig,
         display_numbers = cor$mat_signif_sig,
         fontsize_number = 20,
         angle_col = 45)

pheatmap(cor$mat_cor_sig_cor,
         display_numbers = cor$mat_signif_sig_cor,
         fontsize_number = 20,
         angle_col = 45)
```

## Milk vs Faecal bacteria?

```{r}
# Select dataframes

## Milk bacteria

ps <- subset_samples(PSB.species,Sample.type=="Milk") %>%
  transform_sample_counts(., function(x) x / sum(x)) %>%
  subset_taxa(Genus == "Bifidobacterium") 
  

tax <- tax_best(ps)

ps.df <- as.data.frame(otu_table(ps)) %>%
  t() %>%
  #`colnames<-`(paste(tax$best," HM")) %>%
  `colnames<-`(tax$best) %>%
  as.data.frame()

x <- ps.df
x$RCID <- sample_data(ps)$RCID
x <- x %>% filter(rowSums(select_if(., is.numeric)) != 0)

## Faecal bacteria

ps <- subset_samples(PSB.species,Sample.type=="Faecal sample") %>%
  transform_sample_counts(., function(x) x / sum(x)) %>%
  subset_taxa(Genus == "Bifidobacterium") 
  

tax <- tax_best(ps)

ps.df <- as.data.frame(otu_table(ps)) %>%
  t() %>%
  #`colnames<-`(paste(tax$best," FS")) %>%
  `colnames<-`(tax$best) %>%
  as.data.frame()

y <- ps.df
y$RCID <- sample_data(ps)$RCID
y <- y %>% filter(rowSums(select_if(., is.numeric)) != 0)


# Select only co-ocurring bacteria

# y <- y[,intersect(colnames(y),colnames(x))]
# x <- x[,intersect(colnames(y),colnames(x))]

## Person correlations

cor <- cor_pheatmap(x,y,p_cutoff=0.05,cor_cutoff = 0.1)

pheatmap(cor$mat_cor_sig,
         display_numbers = cor$mat_signif_sig,
         fontsize_number = 20,
         angle_col = 45)

pheatmap(cor$mat_cor_sig_cor,
         display_numbers = cor$mat_signif_sig_cor,
         fontsize_number = 20,
         angle_col = 45)
```
  
# Age propensity scoring

```{r}
library(tidyverse)
library(caret)
library(MLeval)

# Set x2 to be a dataframe with n samples and then the features as columns. ONLY include the predictors and the endpoint. Name the endpoint age or change below. 

# Prep data

# df.days <- cbind(
#   df,
#   Days = as.numeric(meta$Days)
#   )

repCV10 <- trainControl(method = "repeatedcv", number = 5,
                        repeats = 10,
                        returnResamp = "all",
                        savePredictions = "all",
                        allowParallel = T,
                        verboseIter = F
                        )
```

## Model selection and score calculations

### Milk samples

```{r}
ps <- subset_samples(PSB,Sample.type == "Milk") %>%
  transform_sample_counts(function(x) {x/sum(x)})
prev <- 4/nrow(ps@sam_data) #Remove ASVs not found in at least 4 samples 
ps = phyloseq_filter_prevalence(ps, prev.trh = prev, abund.trh = 0.001, abund.type = "mean", threshold_condition = "AND")
taxa <- rownames(ps@tax_table)
sum(sample_sums(ps))/nrow(ps@sam_data)

df <- as.data.frame(ps@otu_table) %>% t() %>% as.data.frame()
df$Days <- ps@sam_data$Days

# Run the model to set 
sl(prop_milk,
  train(
  Days ~., data = df, 
  method = 'spls',
  trControl = repCV10,
  preProcess = c("scale","center"),
  tuneGrid = expand.grid(K = 1:5, # K sets number of components
                         eta = seq(0,0.9,0.1), # eta sets sparsity kappa = 0.5),
                         kappa = 0.5)
  )
  #,overwrite=TRUE
)


#plot(prop_milk)


# modelres <- evalm(model)
# 
# # Select a model based on the plots and you may dig out the predictions by:   
# prd <- predict(model,newdata = x2[,-1] , type = "prob")

# ...Or by looking into the CV predictions stored in the model object. 
# Where you filter on the model you want to use.

# CSS model 
ps <- subset_samples(PSB.CSS,Sample.type == "Milk") 
#Keep taxa based on relative abundance filtering
ps <- prune_taxa(taxa,ps)
#sum(sample_sums(ps))/nrow(ps@sam_data)

df.css <- as.data.frame(ps@otu_table) %>% t() %>% as.data.frame()
df.css$Days <- ps@sam_data$Days

sl(prop_milk_css,
train(
  Days ~., data = df.css, 
  method = 'spls',
  trControl = repCV10,
  preProcess = c("scale","center"),
  tuneGrid = expand.grid(K = 1:5, # K sets number of components
                         eta = seq(0,0.9,0.1), # eta sets sparsity kappa = 0.5),
                         kappa = 0.5)
)
#,overwrite=TRUE
)

ggarrange(plot(prop_milk),plot(prop_milk_css))

propCV.milk <- prop_milk$pred %>% 
filter(eta=="0.2" & K == 2) %>% 
group_by(rowIndex) %>% 
dplyr::summarize(predCV  = mean(pred))

propCV.milk.css <- prop_milk_css$pred %>% 
filter(eta=="0.2" & K == 2) %>% 
group_by(rowIndex) %>% 
dplyr::summarize(predCV  = mean(pred))

propCV.milk$RCID <- ps@sam_data$RCID
propCV.milk$propCV.milk.css <- propCV.milk.css$predCV

ggplot(propCV.milk, aes(x=predCV, y=propCV.milk.css)) + 
  geom_point()+
  geom_smooth(method=lm)

propCV.milk.sel <- propCV.milk %>%
  filter(predCV >= 0, predCV <= 100, propCV.milk.css >= 0, propCV.milk.css <= 100)

ggplot(propCV.milk.sel, aes(x=predCV, y=propCV.milk.css)) + 
  geom_point()+
  geom_smooth(method=lm)

# Subset cv scores depending on whether normalised or not were best
prop.milk.bac <- propCV.milk.sel %>% dplyr::select(RCID,propCV.milk.css)
```

### Faecal samples

```{r}
ps <- subset_samples(PSB,Sample.type == "Faecal sample") %>%
  transform_sample_counts(function(x) {x/sum(x)})
prev <- 4/nrow(ps@sam_data) #Remove ASVs not found in at least 4 samples
ps = phyloseq_filter_prevalence(ps, prev.trh = prev, abund.trh = 0.001, abund.type = "mean", threshold_condition = "AND")
taxa <- rownames(ps@tax_table)
sum(sample_sums(ps))/nrow(ps@sam_data)

df <- as.data.frame(ps@otu_table) %>% t() %>% as.data.frame()
df$Days <- ps@sam_data$Days

# Run the model to set 
sl(prop_faec,
   train(
  Days ~., data = df, 
  method = 'spls',
  trControl = repCV10,
  preProcess = c("scale","center"),
  tuneGrid = expand.grid(K = 1:5, # K sets number of components
                         eta = seq(0,0.9,0.1), # eta sets sparsity kappa = 0.5),
                         kappa = 0.5)
  )
)


#plot(prop_faec)

# modelres <- evalm(model)
# 
# # Select a model based on the plots and you may dig out the predictions by:   
# prd <- predict(model,newdata = x2[,-1] , type = "prob")

# ...Or by looking into the CV predictions stored in the model object. 
# Where you filter on the model you want to use.

# CSS model 
ps <- subset_samples(PSB.CSS,Sample.type == "Faecal sample") 
#Keep taxa based on relative abundance filtering
ps <- prune_taxa(taxa,ps)
#sum(sample_sums(ps))/nrow(ps@sam_data)

df.css <- as.data.frame(ps@otu_table) %>% t() %>% as.data.frame()
df.css$Days <- ps@sam_data$Days

sl(prop_faec_css,
  train(
  Days ~., data = df.css, 
  method = 'spls',
  trControl = repCV10,
  preProcess = c("scale","center"),
  tuneGrid = expand.grid(K = 1:5, # K sets number of components
                         eta = seq(0,0.9,0.1), # eta sets sparsity kappa = 0.5),
                         kappa = 0.5)
  )
)

ggarrange(plot(prop_faec),plot(prop_faec_css))

propCV.faec <- prop_faec$pred %>% 
filter(eta=="0.1" & K == 2) %>% 
group_by(rowIndex) %>% 
dplyr::summarize(predCV  = mean(pred))

propCV.faec.css <- prop_faec_css$pred %>% 
filter(eta=="0.1" & K == 2) %>% 
group_by(rowIndex) %>% 
dplyr::summarize(predCV  = mean(pred))

propCV.faec$RCID <- ps@sam_data$RCID
propCV.faec$propCV.faec.css <- propCV.faec.css$predCV

ggplot(propCV.faec, aes(x=predCV, y=propCV.faec.css)) + 
  geom_point()+
  geom_smooth(method=lm)

propCV.faec.sel <- propCV.faec %>% filter(predCV >= 0, predCV <= 100, propCV.faec.css >= 0, propCV.faec.css <= 100)

ggplot(propCV.faec.sel, aes(x=predCV, y=propCV.faec.css)) + 
  geom_point()+
  geom_smooth(method=lm)

# Subset cv scores depending on whether normalised or not were best
prop.faec.bac <- propCV.faec.sel %>% dplyr::select(RCID,propCV.faec.css)
```

### Metabolome

```{r}

# Set x2 to be a dataframe with n samples and then the features as columns. ONLY include the predictors and the endpoint. Name the endpoint age or change below. 

# Prep data

metab.days <- cbind(
  metab,
  Days = as.numeric(as.character(meta$Days))
  )



# Run the model to set 
sl(prop_metab_milk,
   train(
  Days ~., data = metab.days, 
  method = 'spls',
  trControl = repCV10,
  preProcess = c("scale","center"),
  tuneGrid = expand.grid(K = 1:5, # K sets number of components
                         eta = seq(0,0.9,0.1), # eta sets sparsity kappa = 0.5),
                         kappa = 0.5)
  )
)

## Run model on log normalised metabolome

sl(prop_metab_milk_log,
   train(
  Days ~., data = metab.days, 
  method = 'spls',
  trControl = repCV10,
  preProcess = c("scale","center"),
  tuneGrid = expand.grid(K = 1:5, # K sets number of components
                         eta = seq(0,0.9,0.1), # eta sets sparsity kappa = 0.5),
                         kappa = 0.5)
  )
)

ggarrange(plot(prop_metab_milk),plot(prop_metab_milk_log))

# Select model from based

propCV_milk_metab_sel <- prop_metab_milk$pred %>% 
filter(eta=="0.8" & K == 2) %>% 
group_by(rowIndex) %>% 
dplyr::summarize(predCV  = mean(pred))

propCV_milk_metab_log_sel <- prop_metab_milk_log$pred %>% 
filter(eta=="0.9" & K ==4) %>% 
group_by(rowIndex) %>% 
dplyr::summarize(predCV  = mean(pred))

propCV_milk_metab_sel$RCID <- meta$RCID
propCV_milk_metab_sel$propCV.milk.metab.log <- propCV_milk_metab_log_sel$predCV

ggplot(propCV_milk_metab_sel, aes(x=predCV, y=propCV.milk.metab.log)) + 
  geom_point()+
  geom_smooth(method=lm)

#Save scores from selected model and normalisation

prop.milk.metab <- propCV_milk_metab_sel %>% dplyr::select(c(RCID,predCV)) %>% dplyr::rename( propCV.milk.metab = "predCV")

```

### Propensity-score loadings

```{r}
m <- prop_metab_milk_log
m$finalModel$projection %>% 
  data.frame() %>% 
  rownames_to_column('mtb') %>% 
  gather(cmp,loads,-mtb) %>% 
  ggplot(data = ., aes(mtb,loads, fill = cmp)) + 
  geom_bar(stat = 'identity', position = position_dodge()) + 
  coord_flip()

#Og for reggressionsvektoren: 
b <- m$finalModel$betahat %>% 
  data.frame() %>% 
  rownames_to_column('mtb') 
colnames(b)[2] <- 'b'
b %>% 
  ggplot(data = ., aes(mtb,b)) + 
  geom_bar(stat = 'identity', position = position_dodge()) + 
  coord_flip()

# 
# coeffs <- coef(prop_milk$finalModel,unlist(prop_milk$bestTune$lambda))
# 
# prop_milk$finalModel$projection
# 
# models <- as.data.table(coeffs$coefficients)
# winnermodelscoeffs <- models[which(coeffs$fraction==cvresult$bestTune$fraction)]

# Metabolome

# coeffs <- coef(prop_metab_milk_log$finalModel,unlist(prop_metab_milk_log$bestTune))

```

#### Combine and export

```{r,eval=T}
# Combine 
# propensityCV.merged <- dplyr::full_join(prop.milk.bac,prop.faec.bac,by="RCID") %>%
#   dplyr::full_join(.,prop.milk.metab,by="RCID")

sl(propensityCV.merged,dplyr::full_join(prop.milk.metab,prop.milk.bac,by="RCID") %>%
  dplyr::full_join(.,prop.faec.bac,by="RCID"))

# 
# # Add to phyloseq object
# df.faec <- propCV.faec.filt %>%
#   select(c(RCID.uniq,predCV,propCV.faec.css)) %>%
#   
#   mutate(Sample.type="Faecal sample")
# 
# df.milk <- propCV.milk.filt %>%
#   select(c(RCID.uniq,predCV,propCV.milk.css)) %>%
#   mutate(Sample.type="Milk")
# 
# propCV.bac <- 
# 
# df <- as.data.frame(sample_data(PSB)) -> left_join()
# PSB@sam_data <- 


#Save scores only

```

## Linear comparisons

### All time points

```{r,warning=F}
prop.dat <- propensityCV.merged %>%
  dplyr::mutate(Days = str_extract(RCID, "(?<=_).*")) %>%
  dplyr::mutate(Days = case_when(
    Days == "3" ~ "30",
    Days == "4" ~ "60",
    Days == "5" ~ "90",
    Days == "2" ~ "3",
    TRUE ~ Days  # Keeps all other values as they are
  )) %>%
  dplyr::mutate(ID = str_extract(RCID, "^[^_]*")) %>%
  dplyr::mutate_at(dplyr::vars(Days ,RCID, ID), factor)

#Format
prop.dat$Days <- factor(prop.dat$Days, levels=c("3","30","60","90"))

prop.dat$`Milk bacterial maturity` <- prop.dat$propCV.milk.css
prop.dat$`Faecal bacterial maturity` <- prop.dat$propCV.faec.css
prop.dat$`Milk metabolome maturity` <- prop.dat$propCV.milk.metab

prop.plot.milk.bac.faec.bac <- ggplot(prop.dat, aes(x=`Milk bacterial maturity`, y=`Faecal bacterial maturity`)) + 
  geom_smooth(method=lm) +
  geom_point(aes(color=Days))+
  scale_color_manual(values=col_fil,name="Days of life",breaks=c(3,30,60,90))+
  #ggtitle("Milk bacteria vs faecal bacteria") +
  stat_cor(method = "pearson", label.x = 10, label.y = 20)

prop.plot.milk.bac.milk.metab <- ggplot(prop.dat, aes(x=`Milk bacterial maturity`, y=`Milk metabolome maturity`)) + 
  geom_point(aes(color=Days))+
  stat_smooth(method=lm) +
  scale_color_manual(values=col_fil,name="Days of life",breaks=c(3,30,60,90))+
  #ggtitle("Milk bacteria vs milk metabolites") +
  stat_cor(method = "pearson", label.x = 10, label.y = -25)

prop.plot.faec.bac.milk.metab <- ggplot(prop.dat, aes(x=`Faecal bacterial maturity`, y=`Milk metabolome maturity`)) + 
  geom_point(aes(color=Days))+
  stat_smooth(method=lm) +
  scale_color_manual(values=col_fil,name="Days of life",breaks=c(3,30,60,90))+
  #ggtitle("Faecal bacteria vs milk metabolites") +
  stat_cor(method = "pearson", label.x = 10, label.y = -25)

sl(age.propensity.cor,ggarrange(prop.plot.milk.bac.milk.metab,
                                prop.plot.faec.bac.milk.metab,
                                prop.plot.milk.bac.faec.bac,
          nrow=1,
          labels = c("A","B","C"),
          font.label = list(size = 30),
          hjust = 0.3,
          vjust = 1,
          common.legend = TRUE,
          legend="right"),
   overwrite = TRUE
   )

age.propensity.cor
```

### By time point

```{r,warning=F}
prop.plot.milk.bac.faec.bac.days <- ggplot(prop.dat, aes(x=`Milk bacterial maturity`, y=`Faecal bacterial maturity`,color=Days))+
  geom_smooth(method=lm,show.legend=FALSE) +
  geom_point()+
  scale_color_manual(values=col_fil,name="Days of life",breaks=c(3,30,60,90))+
  #ggtitle("Milk bacteria vs faecal bacteria") +
  stat_cor(method = "pearson",show.legend = FALSE)+
  scale_y_continuous(limits=c(-10,150))

prop.plot.milk.bac.milk.metab.days <- ggplot(prop.dat, aes(x=`Milk bacterial maturity`, y=`Milk metabolome maturity`,color=Days)) + 
  geom_point()+
  scale_color_manual(values=col_fil,name="Days of life",breaks=c(3,30,60,90))+
  stat_smooth(method=lm,show.legend=FALSE) +
  #ggtitle("Milk bacteria vs milk metabolites") +
  stat_cor(method = "pearson",show.legend = FALSE)+
  scale_y_continuous(limits=c(-50,170))

prop.plot.faec.bac.milk.metab.days <- ggplot(prop.dat, aes(x=`Faecal bacterial maturity`, y=`Milk metabolome maturity`,color=Days)) + 
  geom_point()+
  scale_color_manual(values=col_fil,name="Days of life",breaks=c(3,30,60,90))+
  stat_smooth(method=lm,show.legend=FALSE) +
  #ggtitle("Faecal bacteria vs milk metabolites") +
  stat_cor(method = "pearson",show.legend = FALSE) +
  scale_y_continuous(limits=c(-10,150))

sl(age.propensity.cor.days,ggarrange(prop.plot.milk.bac.milk.metab.days,
                                     prop.plot.faec.bac.milk.metab.days,
                                     prop.plot.milk.bac.faec.bac.days,
          nrow=1,
          labels = c("D","E","F"),
          font.label = list(size = 30),
          hjust = 0.3,
          vjust = 1,
          common.legend = TRUE,
          legend="right"),
   overwrite = TRUE
   )

age.propensity.cor.days
```

## Linear models 

See section 5.2 for complex models - not run here due to conflicts of the "lme4" package with "phyloseq"

## Metadata correlations

### Pure microbial maturity scores


```{r}
meta.prop.merged <- left_join(meta,propensityCV.merged,by="RCID")

x <- meta.prop.merged %>% dplyr::select(effect.size.variables[! effect.size.variables %in% c('Batch', 'DNA.conc','Days')]) %>% mutate_all(as.numeric)
x$RCID <- meta.prop.merged$RCID

y <- meta.prop.merged %>% dplyr::select(c(RCID,propCV.milk.css, propCV.faec.css, propCV.milk.metab))

cor <- cor_pheatmap(x,y,p_cutoff=0.05,cor_cutoff = 0.1)

pheatmap(cor$mat_cor,
         display_numbers = cor$mat_signif,
         fontsize_number = 20,
         angle_col = 45)

pheatmap(cor$mat_cor_sig_cor,
         display_numbers = cor$mat_signif_sig_cor,
         fontsize_number = 20,
         angle_col = 45)
```

### MAZ scores

```{r}
prop.dat<-readRDS(here::here("results","RData","5.2_Age_propensity_glm_models","prop.dat.glm.RDS"))

prop.dat$Days <- as.factor(prop.dat$Days)

meta.prop.merged <- left_join(meta,prop.dat,by=c("RCID","Days"),keep = FALSE)

x <- meta.prop.merged %>%
  dplyr::select(effect.size.variables[! effect.size.variables %in% c('Batch', 'DNA.conc','Days')]) %>% mutate_all(as.numeric)
x$RCID <- meta.prop.merged$RCID

y <- meta.prop.merged %>% dplyr::select(c(RCID,
                                          milk.bac.maturity=propCV.milk.css,
                                          faecal.bac.maturity=propCV.faec.css,
                                          milk.metab.maturity=propCV.milk.metab,
                                          milk.bac.z,
                                          faec.bac.z,
                                          milk.metab.z))

cor <- cor_pheatmap(x,y,p_cutoff=0.05,cor_cutoff = 0.05)

pheatmap(cor$mat_cor,
         display_numbers = cor$mat_signif,
         fontsize_number = 20,
         angle_col = 45)

#Clean column names
cor$mat_cor_sig_cor <- cor$mat_cor_sig_cor %>%
  dplyr::rename(
    `Milk bacterial maturity` = milk.bac.maturity,
    `Faecal bacterial maturity` = faecal.bac.maturity,
    `Milk metabolome maturity` = milk.metab.maturity,
    `Milk bacterial maturity-for-age` = milk.bac.z,
    `Faecal bacterial maturity-for-age` = faec.bac.z,
    `Milk metabolome maturity-for-age` = milk.metab.z
  )

rownames(cor$mat_cor_sig_cor) <- rownames(cor$mat_cor_sig_cor) %>% str_replace_all(., c(`BMI delta`="delta BMI",
                                                              ppBMI = "Pre-pregnancy BMI",
                                                              delta.="Δ",
                                                              `3m`="3 months",
                                                              waz = "WAZ",
                                                              whz = "WHZ",
                                                              haz = "HAZ",
                                                              AB="Antibiotics",
                                                              GDM="Gestational diabetes",
                                                              Secretor="Secretor status",
                                                              Lewis="Lewis status",
                                                              `C section`="C-section",
                                                              "\\." =" "))


sl(age.prop.z,pheatmap(cor$mat_cor_sig_cor,
         display_numbers = cor$mat_signif_sig_cor,
         fontsize_number = 20,
         angle_col = 45,
         ),
   overwrite=TRUE
)

age.prop.z

```

### Mean MAZ scores

```{r}
# Calculate averages for each ID
averages <- prop.dat %>%
  group_by(ID) %>%
  summarize(
    avg_milk_bac_z = mean(milk.bac.z, na.rm = TRUE),
    avg_faec_bac_z = mean(faec.bac.z, na.rm = TRUE),
    avg_milk_metab_z = mean(milk.metab.z, na.rm = TRUE)
  )


meta$ID <- as.factor(meta$ID)
meta.id <- meta %>%
  group_by(ID) %>%
  dplyr::slice(1) %>%
  ungroup()

meta.prop.merged <- left_join(averages,meta.id,by="ID",keep = FALSE)

x <- meta.prop.merged %>%
  dplyr::select(effect.size.variables[! effect.size.variables %in% c('Batch', 'DNA.conc','AB.infant','Days')]) %>% mutate_all(as.numeric) 
x$RCID <- as.factor(meta.prop.merged$ID)

y <- meta.prop.merged %>% dplyr::select(c(avg_milk_bac_z,
                                          avg_faec_bac_z,
                                          avg_milk_metab_z,
                                          RCID = ID))

#Remove samples with missing RCID (dropout subjects)
y <- y[!is.na(meta.prop.merged$RCID),]
x <- x[!is.na(meta.prop.merged$RCID),]

# x <- x[,c(1:11,19)]

cor <- cor_pheatmap(x,y,p_cutoff=0.05,cor_cutoff = 0.05)

pheatmap(cor$mat_cor,
         display_numbers = cor$mat_signif,
         fontsize_number = 20,
         angle_col = 45)

pheatmap(cor$mat_cor_sig_cor,
         display_numbers = cor$mat_signif_sig_cor,
         fontsize_number = 20,
         angle_col = 45)

```

#### Mean vs mean scatterplots

```{r}
prop.plot.milk.bac.faec.bac.days <- ggplot(meta.prop.merged, aes(x=avg_milk_bac_z, y=avg_faec_bac_z)) + 
  geom_smooth(method=lm) +
  geom_point()+
  ggtitle("Milk bacteria vs faecal bacteria") +
  stat_cor(method = "pearson")

prop.plot.milk.bac.milk.metab.days <- ggplot(meta.prop.merged, aes(x=avg_milk_bac_z, y=avg_milk_metab_z)) + 
  geom_point()+
  stat_smooth(method=lm) +
  ggtitle("Milk bacteria vs milk metabolites") +
  stat_cor(method = "pearson")

prop.plot.faec.bac.milk.metab.days <- ggplot(meta.prop.merged, aes(x=avg_faec_bac_z, y=avg_milk_metab_z)) + 
  geom_point()+
  stat_smooth(method=lm) +
  ggtitle("Faecal bacteria vs milk metabolites") +
  stat_cor(method = "pearson")

ggarrange(prop.plot.milk.bac.faec.bac.days,
          prop.plot.milk.bac.milk.metab.days,
          prop.plot.faec.bac.milk.metab.days,
          nrow=1,
          common.legend = TRUE,legend="bottom")
```

### Spread of MAZ scores

```{r}
# Calculate averages for each ID
spread <- prop.dat %>%
  group_by(ID) %>%
  filter(n() >= 2) %>%
  summarise(
    spread_milk_bac_z = ifelse(sum(!is.na(milk.bac.z)) >= 2, max(milk.bac.z, na.rm = TRUE) - min(milk.bac.z, na.rm = TRUE), NA_real_),
    spread_faec_bac_z = ifelse(sum(!is.na(faec.bac.z)) >= 2, max(faec.bac.z, na.rm = TRUE) - min(faec.bac.z, na.rm = TRUE), NA_real_),
    spread_milk_metab_z = ifelse(sum(!is.na(milk.metab.z)) >= 2, max(milk.metab.z, na.rm = TRUE) - min(milk.metab.z, na.rm = TRUE), NA_real_)
  )

  # summarise(spread_milk_bac_z = max(milk.bac.z, na.rm = TRUE) - min(milk.bac.z, na.rm = TRUE),
  #   spread_faec_bac_z = max(faec.bac.z, na.rm = TRUE) - min(faec.bac.z, na.rm = TRUE),
  #   spread_milk_metab_z = max(milk.metab.z, na.rm = TRUE) - min(milk.metab.z, na.rm = TRUE)
  # )

# View the results
#print(spread)

meta$ID <- as.factor(meta$ID)
meta.id <- meta %>%
  group_by(ID) %>%
  dplyr::slice(1) %>%
  ungroup()

meta.prop.merged <- left_join(spread,meta.id,by="ID",keep = FALSE)

x <- meta.prop.merged %>%
  dplyr::select(effect.size.variables[! effect.size.variables %in% c('Batch', 'DNA.conc','AB.infant','Days')]) %>% mutate_all(as.numeric) 
x$RCID <- as.factor(meta.prop.merged$ID)

y <- meta.prop.merged %>% dplyr::select(c(
                                          # avg_milk_bac_z,
                                          # avg_faec_bac_z,
                                          # avg_milk_metab_z,
                                          spread_milk_bac_z,
                                          spread_faec_bac_z,
                                          spread_milk_metab_z,
                                          RCID = ID))

#Remove samples with missing RCID (dropout subjects)
y <- y[!is.na(meta.prop.merged$RCID),]
x <- x[!is.na(meta.prop.merged$RCID),]

# x <- x[,c(1:11,19)]

cor <- cor_pheatmap(x,y,p_cutoff=0.2,cor_cutoff = 0.1)

pheatmap(cor$mat_cor,
         display_numbers = cor$mat_signif,
         fontsize_number = 20,
         angle_col = 45)

pheatmap(cor$mat_cor_sig_cor,
         display_numbers = cor$mat_signif_sig_cor,
         fontsize_number = 20,
         angle_col = 45)

```

# Session info


```{r}
sessionInfo()
```
