---
title: "5.3_Metabolome_microbiome_correlations"
author: "Rasmus Riemer Jakobsen"
date: "`r Sys.time()`"
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
link-citations: yes
---

```{r setup, include=F, message=F, warning=F}
subdir_name <- "5.3_Metabolome_microbiome_correlations"

#Load packages
source("Supplementary_scripts/package_load_install.R")
#Load utility functions
source("Supplementary_scripts/utils.R")

dir.create("results", showWarnings = FALSE, recursive = TRUE)

# Set theme and colort

theme_set(theme_classic())

knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.fullwidth = TRUE)

# Load data

#Microbiome
load("0.3_deduplicated_and_decontaminated.RData")

#Metabolome
metab.id <- readr::read_tsv("Data/metabolome_milk.txt")
metab.meta <- readr::read_tsv("Data/metadata_metabome_milk.txt")
metab.class <- readr::read_tsv("Data/metabolite_categories.txt") %>%
  mutate(Class = ifelse(Class == "Amino acids and derivatives", "Amino acids", Class)) # Shorten for better plot legend
  

## Check if dataframes match
setdiff(metab.id$RCID,metab.meta$RCID)

metab <- metab.id %>% dplyr::select(-RCID) 

log.metab <- 
  log10(metab+1) 

# Set key metadata columns to factor
metab.meta <- metab.meta %>%
  mutate_at(
    c("Days","Siblings", "GDM", "Induction", "AB.mother", "AB.infant", "Breast.infection", "Breastfeeding.issues", "Secretor", "Lewis"),
    as.factor)
```

```{r}
source(here::here("Supplementary_scripts","package_load_install.R")) 
source(here::here("Supplementary_scripts","utils.R")) # please see the src-folder for some custom functions used in 
#source(here::here("src", "run_clusterprofiler.R"))
```

```{r knit, echo=F, eval=FALSE}
#Run to generate
rmarkdown::render(
  input = here::here(str_c(subdir_name, ".Rmd")),
  output_dir = here::here("results", "reports")
)
```


```{r}

```

# RCCA

## Milk metab vs milk bac

### Conversion and screeplot

```{r, eval = T, include = T}
#Subset samples present in both
ps <- subset_samples(PSB.CSS, Sample.type == "Milk" & RCID %in% metab.id$RCID)

#rcid.shared <- intersect(metab.meta$RCID,ps.milk@sam_data$RCID)
x <- subset(metab.id, RCID %in% ps@sam_data$RCID) %>% dplyr::select(-RCID) %>%
  dplyr::mutate(across(where(is.numeric), ~ifelse(. < 0, NA_real_, sqrt(.)))) 
#colnames(x) <- make.names(colnames(x))
tax <- data.frame(ps@tax_table) 

#Extract ASV data
y <- data.frame(ps@otu_table)
##Select last non-empty taxonomic rank
tax.bac <- as.data.frame(ps@tax_table) %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured" taxonomies (uninformative) 
  mutate(across(.cols = where(is.character), ~str_replace_all(., "_sensu_stricto_1", ""))) %>% # remove all "sens strictu" (uninformative)
  apply(.,1, function(x) tail(stats::na.omit(x), 1))
#Combine otu_ids with taxonomies for unique identiiers
rownames(y) <- paste(rownames(y),tax.bac, sep = "-")


meta <- data.frame(ps@sam_data)


y <- t(y)


    ##Converting and removing near-zero variance values
library(mixOmics)

nzv <- nearZeroVar(x)
test <- x[, -nzv$Position]
           
           
#vir <- mvirome[,-c(1)]
nzv <- nearZeroVar(y) # If any
y <- y[, -nzv$Position]


  #PCA
library(vegan)
mod.x <- rda(x, scale = TRUE)
plot(mod.x, scaling = 3)
screeplot(mod.x)
mod.y <- rda(y, scale = TRUE)
plot(mod.y, scaling = 3)
screeplot(mod.y)

scaling = "ridge"
  #Apply scaling 
result_1comp <- rcc(x, y, method = scaling, ncomp = 1)
result_2comp <- rcc(x, y, method = scaling, ncomp = 2)
result_3comp <- rcc(x, y, method = scaling, ncomp = 3)
result_4comp <- rcc(x, y, method = scaling, ncomp = 4)
result_5comp <- rcc(x, y, method = scaling, ncomp = 5)
result_6comp <- rcc(x, y, method = scaling, ncomp = 6)

par(mfrow = c(2, 3))
plot(result_1comp,main="1 Component")
plot(result_2comp,main="2 Components")
plot(result_3comp,main="3 Components")
plot(result_4comp,main="4 Components")
plot(result_5comp,main="5 Components")
plot(result_6comp,main="6 Components")

# produce a heat map of the cross correlation matrix
imgCor(x, y, sideColors = c("purple", "green")) 
```

### Model tuning

```{r, eval = T, include = T}
# set grid search values for each regularisation parameter
grid1 <- seq(0.001, 0.2, length = 10) 
grid2 <- seq(0.001, 0.2, length = 10)

# optimise the regularisation parameter values
sl(cv.tune.rcc.milk.metab.milk.bac, tune.rcc(x, y, grid1 = grid1, grid2 = grid2, 
                                   validation = "loo"),overwrite = FALSE)

# examine the results of CV tuning
plot(cv.tune.rcc.milk.metab.milk.bac)
cv.tune.rcc.milk.metab.milk.bac

#manually set lambda scores
#opt.l1 <- 1.7
#opt.l2 <- 0.003
opt.l1 <- cv.tune.rcc.milk.metab.milk.bac$opt.lambda1
opt.l2 <- cv.tune.rcc.milk.metab.milk.bac$opt.lambda2
```

### Model and component selection

```{r, eval = T, include = T}
shrink.rcc <- rcc(x, y, method = "shrinkage")

CV.rcc <- rcc(x, y, method = "ridge", 
                         lambda1 = opt.l1,
                         lambda2 = opt.l2) 


par(mfrow = c(1, 2))
# barplot of cross validation method rCCA canonical correlations
plot(CV.rcc, type = "barplot", main = "Cross Validation") 

# barplot of shrinkage method rCCA canonical correlations
plot(shrink.rcc, type = "barplot", main = "Shrinkage")
```

### Run model

```{r, eval = T, include = T}
#Perfrom optimised CV rCCA

result.CV <- rcc(x, y, method = "ridge",
                         lambda1 = opt.l1,
                         lambda2 = opt.l2,
              ncomp=3)

result.shr <- rcc(x, y, method = "shrinkage",ncomp=3)

# Select ridge or shrinkage method
result <- result.CV

#colnames(result$x) <- gsub(".*\\-", "", colnames(result$x)) %>% make.unique(sep="_")
#colnames(result$Y) <- gsub(".*\\-", "", colnames(result$Y)) %>% make.unique(sep="_")

# rownames(result$loadings$X) <- gsub(".*\\-", "", rownames(result$loadings$X))
# rownames(result$loadings$Y) <- gsub(".*\\-", "", rownames(result$loadings$Y))


```

### Visualization

#### Heatmaps

```{r, eval = T, include = T, echo = T, message = F, warning =  F, fig.fullwidth = TRUE}
#Defne pallette to only add colour to only high
##Generate new pallette
palette <- hcl.colors(200, palette = "Zissou 1", alpha = NULL, rev = FALSE, fixup = TRUE) #Seclect palette https://developer.r-project.org/Blog/public/2019/04/01/hcl-based-color-palettes-in-grdevices/
palette[61:140] <- "white" #Select which part of pallette to replace with white

# high4_3c <- cim(result, comp = 1:3, ylab = "Bacteria", xlab = "Virome host prediction", margins = c(15,20),cutoff = 0.4, color = palette)
high.25_3c <- cim(result, comp = 1:3, ylab = "Bacteria", xlab = "Virome host prediction", margins = c(13,10),cutoff = 0.25, color = palette)
high.2_3c <- cim(result, comp = 1:3, ylab = "Bacteria", xlab = "Virome host prediction", margins = c(13,10),cutoff = 0.2, color = palette)

high.15_3c <- cim(result, comp = 1:3, ylab = "Bacteria", xlab = "Virome host prediction", margins = c(13,10),cutoff = 0.15, color = palette
                  #,cut.tree = c(0,0.5)
                  )


```

#### Scatterplots

```{r}
mat.25_3c <- high.25_3c$mat
mat.2_3c <- high.2_3c$mat

#Make linear plots of all bac and vir
x.25_3c <- x[,rownames(mat.25_3c)]
y.25_3c <- y[,colnames(mat.25_3c)]
xy.25_3c <- cbind(x.25_3c,y.25_3c)
x.2_3c <- x[,rownames(mat.2_3c)]
y.2_3c <- y[,colnames(mat.2_3c)]
xy.2_3c <- cbind(x.2_3c,y.2_3c)

#ggpairs
library(GGally)

#.25 matrix ggpairs
pairs.25 <- ggpairs(as.data.frame(xy.25_3c),
        diag = list(continuous = "densityDiag"),
        upper = list(continuous = wrap("cor", size = 4)),
        axisLabels = "show")
file_path <- "results/pairs.milk.bac.milk.metab.rcca.25.pdf"
if (!file.exists(file_path)) {
   ggsave(file_path,pairs.25,height=30,width=25,units="cm")
} else {
    cat("File already exists:", file_path, "\n")
}

#.2 matrix ggpairs
pairs.2 <-ggpairs(as.data.frame(xy.2_3c),
        diag = list(continuous = "densityDiag"),
        upper = list(continuous = wrap("cor", size = 4)),
        axisLabels = "show")
file_path <- "results/pairs.milk.bac.milk.metab.rcca.2.pdf"
if (!file.exists(file_path)) {
   ggsave(file_path,pairs.25,height=30,width=25,units="cm")
} else {
    cat("File already exists:", file_path, "\n")
}
```

#### Heatmaps - Pheatmap

##### 0.2 cutoff

```{r, eval = T, include = T, echo = T, message = F, warning =  F, fig.fullwidth = TRUE}
###### Select RCCA output(with cutoff)
rcca <- data.frame(high.2_3c$mat)


library(pheatmap)

palette <- hcl.colors(200, palette = "Zissou 1", alpha = NULL, rev = FALSE, fixup = TRUE) #Seclect palette https://developer.r-project.org/Blog/public/2019/04/01/hcl-based-color-palettes-in-grdevices/
palette[46:155] <- "white" #Select which part of pallette to replace with white

#Remove OTU IDs to only show taxonomy
tax <- gsub(".*\\.", "", colnames(rcca)) %>% make.unique(sep="")
tax
otu <- gsub("\\..*", "", colnames(rcca)) %>% make.unique(sep="")
otu

colnames(rcca) <- tax

######################## Annotations

###########  x-axis annotation

######## Phylogeny 1

rank <-"Family"

tax_levels <- ps@tax_table[otu,] %>%
    data.frame() %>%
  dplyr::select(1:rank) 

##Select last non-empty taxonomic rank
tax.best <- tax_levels %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured"
  apply(.,1, function(x) tail(stats::na.omit(x), 1)) %>%
  sub("_.*", "",.)
  

tax_levels$best <- tax.best 
sample_col <- tax_levels %>% dplyr::select(best) %>%
  dplyr::rename(!!rank := best)

######## Phylogeny 1

rank <-"Class"

tax_levels <- ps@tax_table[otu,] %>%
    data.frame() %>%
  dplyr::select(1:rank) 

##Select last non-empty taxonomic rank
tax.best <- tax_levels %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured"
  apply(.,1, function(x) tail(stats::na.omit(x), 1)) %>%
  sub("_.*", "",.)
  

tax_levels$best <- tax.best 
sample_col_2 <- tax_levels %>% dplyr::select(best) %>%
  dplyr::rename(!!rank := best)

############ Y-axis annotation

## Matbolite class

rownames(rcca) 

metab.class.sel <- metab.class[match(rownames(rcca), metab.class$Metabolite), ]

# Convert the "Metabolite" column to row names
rownames <- metab.class.sel$Metabolite

# Remove the "Metabolite" column from the dataframe

metab.class.sel <- metab.class.sel %>% as.data.frame() %>% dplyr::select(Class) 
rownames(metab.class.sel) <- rownames
metab.class.sel <- rename(metab.class.sel, Class = "Metabolite category")

######################### Clean species names 

##Select last non-empty taxonomic rank
rank <-"Species"

# x-axis annotation
## Phylogeny
tax_levels <- ps@tax_table[otu,] %>%
    data.frame() %>%
  dplyr::select(1:rank) 

##Select last non-empty taxonomic rank
tax.best <- tax_levels %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured"
  dplyr::mutate(Genus = str_remove(Genus, "_.*")) %>% #Remove odd genus annotations
  mutate(across(.cols = where(is.character), ~str_replace_all(., "_NK3B31_group", ""))) %>% # Fix wierd database match
  apply(.,1, function(x) tail(stats::na.omit(x), 1))
  

tax_levels$best <- tax.best 

tax_levels <- tax_levels %>%
  dplyr::mutate_all(~na_if(., '')) %>%
  dplyr::mutate(best = ifelse(is.na(.data[[rank]]), paste(best, "unclassified"), best)) 

colnames(rcca) <- make.unique(tax_levels$best) %>% gsub("_", " ",.) %>% gsub("_", " ",.)

# Align color vecors with clean tax names
row.names(sample_col) <- colnames(rcca)
#Add second annotation row
sample_col <- cbind(sample_col,sample_col_2)

#############

#Set colors
ann_colors <- list(Class = rep(col_fil,3)[1:n_distinct(sample_col[,1])])

newnames <- lapply(
  colnames(rcca),
  function(x) bquote(italic(.(x))))

sl(pheat.milkmet.milkbac.2,pheatmap(rcca,
         color=palette,
         #border_color = NA,
         #angle_col = 45,
         annotation_col = sample_col,
         annotation_row =  metab.class.sel,
         cutree_rows = 3,
         cutree_cols = 3,
         labels_col = as.expression(newnames),
         main = "A - Milk bacteria vs Milk metabolites"
         #annotation_colors= ann_colors[1]#colorRampPalette(col_fil)
         ),
   overwrite = TRUE
)


gridExtra::grid.arrange(pheat.milkmet.milkbac.2$gtable, vp=viewport(width=0.95, height=1))

```

#### Network plots

##### Shrinkage

```{r, eval = T, include = T, echo = T, message = F, warning =  F, fig.width=10, fig.height=6, fig.fullwidth = TRUE}
# Reformat for gephi

## Save blocks as polygon to determine shape in network

result$names$Polygon <- result$names$data
result$names$Polygon <- replace(result$names$data, c("bac","vir"),c(3,4)) %>% as.integer()

## Remove ubique markers to allow shared coloring in Gephi
result$names_simple$colnames$X <- gsub("*.\\_", "", result$names$colnames$X)
result$names_simple$colnames$Y <- gsub("*.\\_", "", result$names$colnames$Y)

#Show network


#Export network result
network.res <- network(result, comp = 1:3, interactive = FALSE,
        lwd.edge = 2,
        cutoff = 0.20,
        color.node = col_fil[1:2])

network.res <- network(result, comp = 1:3, interactive = FALSE,
        lwd.edge = 2,
        cutoff = 0.15,
        color.node = col_fil[1:2])

library(igraph)

#network.res 

write.graph(network.res$gR, file = "network.shrinkage0.3.gml", format = "gml")
```

##### CV

```{r, eval = T, include = T, echo = T, message = F, warning =  F, fig.width=10, fig.height=6, fig.fullwidth = TRUE}
result.CV$names_simple$colnames$X <- gsub("*.\\_", "", result.CV$names$colnames$X)
result.CV$names_simple$colnames$Y <- gsub("*.\\_", "", result.CV$names$colnames$Y)

#Show network
network(result.CV, comp = 1:3, interactive = FALSE,
        lwd.edge = 2,
        cutoff = 0.2,
        color.node = col_fil[1:2]) 

network(result.CV, comp = 1:3, interactive = FALSE,
        lwd.edge = 2,
        cutoff = 0.15,
        color.node = col_fil[1:2]) 

# network(result.CV, comp = 1:3, interactive = FALSE,
#         lwd.edge = 2,
#         cutoff = 0.5,
#         color.node = col_fil[1:2]) 


library(igraph)
network.res <- network(result, comp = 1:3, interactive = FALSE,
        lwd.edge = 2,
        cutoff = 0.2,
        color.node = col_fil[1:2])

write.graph(network.res$gR, file = "network.CV0.2.gml", format = "gml")
```

## Milk metab vs infant faecal bac

### Conversion and screeplot

```{r, eval = T, include = T}
#Subset samples present in both
ps <- subset_samples(PSB.CSS, Sample.type == "Faecal sample" & RCID %in% metab.id$RCID)
#rcid.shared <- intersect(metab.meta$RCID,ps.milk@sam_data$RCID)
x <- subset(metab.id, RCID %in% ps@sam_data$RCID) %>% dplyr::select(-RCID) %>%
  dplyr::mutate(across(where(is.numeric), ~ifelse(. < 0, NA_real_, sqrt(.))))
#colnames(x) <- make.names(colnames(x))
tax <- data.frame(ps@tax_table)

#Extract ASV data
y <- data.frame(ps@otu_table)
##Select last non-empty taxonomic rank
tax.bac <- as.data.frame(ps@tax_table) %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured" taxonomies (uninformative)
  apply(.,1, function(x) tail(stats::na.omit(x), 1))
#Combine otu_ids with taxonomies for unique identiiers
rownames(y) <- paste(rownames(y),tax.bac, sep = "-")


meta <- data.frame(ps@sam_data)


y <- t(y)


    ##Converting and removing near-zero variance values
library(mixOmics)

nzv <- nearZeroVar(x)
x <- x[, -nzv$Position]
           
           
#vir <- mvirome[,-c(1)]
nzv <- nearZeroVar(y) # If any
y <- y[, -nzv$Position]


  #PCA
library(vegan)
mod.x <- rda(x, scale = FALSE, center = FALSE)
plot(mod.x, scaling = 3)
screeplot(mod.x)
mod.y <- rda(y, scale = FALSE, center = FALSE)
plot(mod.y, scaling = 3)
screeplot(mod.y)

scaling = "ridge"
  #Apply scaling 
result_1comp <- rcc(x, y, method = scaling, ncomp = 1)
result_2comp <- rcc(x, y, method = scaling, ncomp = 2)
result_3comp <- rcc(x, y, method = scaling, ncomp = 3)
result_4comp <- rcc(x, y, method = scaling, ncomp = 4)
result_5comp <- rcc(x, y, method = scaling, ncomp = 5)
result_6comp <- rcc(x, y, method = scaling, ncomp = 6)

par(mfrow = c(2, 3))
plot(result_1comp,main="1 Component")
plot(result_2comp,main="2 Components")
plot(result_3comp,main="3 Components")
plot(result_4comp,main="4 Components")
plot(result_5comp,main="5 Components")
plot(result_6comp,main="6 Components")

# produce a heat map of the cross correlation matrix
imgCor(x, y, sideColors = c("purple", "green")) 
```

### Model tuning

```{r, eval = T, include = T}
# set grid search values for each regularisation parameter
grid1 <- seq(0.001, 0.2, length = 20) 
grid2 <- seq(0.001, 0.2, length = 20)

# optimise the regularisation parameter values
sl(cv.tune.rcc.milk.metab.faecal.bac, tune.rcc(x, y, grid1 = grid1, grid2 = grid2, 
                                   validation = "loo"),overwrite = FALSE)

# examine the results of CV tuning
plot(cv.tune.rcc.milk.metab.faecal.bac)
cv.tune.rcc.milk.metab.faecal.bac

#manually set lambda scores
#opt.l1 <- 1.7
#opt.l2 <- 0.003
opt.l1 <- cv.tune.rcc.milk.metab.faecal.bac$opt.lambda1
opt.l2 <- cv.tune.rcc.milk.metab.faecal.bac$opt.lambda2
```

### Model and component selection

```{r, eval = T, include = T}
shrink.rcc <- rcc(x, y, method = "shrinkage")

CV.rcc <- rcc(x, y, method = "ridge", 
                         lambda1 = opt.l1,
                         lambda2 = opt.l2) 

# barplot of cross validation method rCCA canonical correlations
plot(CV.rcc, type = "barplot", main = "Cross Validation") 

# barplot of shrinkage method rCCA canonical correlations
plot(shrink.rcc, type = "barplot", main = "Shrinkage")
```

### Run model

```{r, eval = T, include = T}
#Perfrom optimised CV rCCA

result.CV <- rcc(x, y, method = "ridge",
                         lambda1 = opt.l1,
                         lambda2 = opt.l2,
              ncomp=2)

result.raw <- rcc(x, y, method = "shrinkage",ncomp=2)

#Remove OTU IDs to only show taxonomy
result <- result.raw

result$names$colnames$X <- gsub(".*\\-", "", result$names$colnames$X) %>% make.unique(sep="_")
result$names$colnames$Y <- gsub(".*\\-", "", result$names$colnames$Y) %>% make.unique(sep="_")


# colnames(result$X) <- gsub(".*\\-", "", colnames(result$X))
# colnames(result$Y) <- gsub(".*\\-", "", colnames(result$Y))

# rownames(result$loadings$X) <- gsub(".*\\-", "", rownames(result$loadings$X))
# rownames(result$loadings$Y) <- gsub(".*\\-", "", rownames(result$loadings$Y))


```

### Visualization

#### Heatmaps 

```{r, eval = T, include = T, echo = T, message = F, warning =  F, fig.fullwidth = TRUE}
#Defne pallette to only add colour to only high
##Generate new pallette
palette <- hcl.colors(200, palette = "Zissou 1", alpha = NULL, rev = FALSE, fixup = TRUE) #Seclect palette https://developer.r-project.org/Blog/public/2019/04/01/hcl-based-color-palettes-in-grdevices/
palette[61:140] <- "white" #Select which part of pallette to replace with white

# high4_2c <- cim(result, comp = 1:2, ylab = "Bacteria", xlab = "Virome", margins = c(12,18),cutoff = 0.4, color = palette)
# high45_2c <- cim(result, comp = 1:2, ylab = "Bacteria", xlab = "Virome", margins = c(12,18),cutoff = 0.45, color = palette)
#high5_3c <- cim(result, comp = 1:2, ylab = "Bacteria", xlab = "Virome", margins = c(12,18),cutoff = 0.5, color = palette)

#pdf(filename="high5_3c.pdf", res = 600, width = 4000, height = 4000)
#high5_3c
#dev.off()

######


high.25_2c <- cim(result, comp = 1:2, ylab = "Bacteria", xlab = "Virome host prediction", margins = c(15,20),cutoff = 0.25, color = palette)

high.225_2c <- cim(result, comp = 1:2, ylab = "Bacteria", xlab = "Virome host prediction", margins = c(15,20),cutoff = 0.225, color = palette)

high.21_2c <- cim(result, comp = 1:2, ylab = "Bacteria", xlab = "Virome host prediction", margins = c(15,20),cutoff = 0.21, color = palette)

high.2_2c <- cim(result, comp = 1:2, ylab = "Bacteria", xlab = "Virome host prediction", margins = c(15,20),cutoff = 0.2, color = palette)

high.175_2c <- cim(result, comp = 1:2, ylab = "Bacteria", xlab = "Virome host prediction", margins = c(15,20),cutoff = 0.175, color = palette)

high.15_2c <- cim(result, comp = 1:2, ylab = "Bacteria", xlab = "Virome host prediction", margins = c(15,20),cutoff = 0.15, color = palette)

mat.2_2c <- high.2_2c$mat

 #Output Matrix
#write.table(mat.2_3c, "Matrix_high_3-Component_0.45.txt", sep="\t")


```

#### Scatterplots

```{r}
mat.25_2c <- high.25_2c$mat
mat.2_2c <- high.2_2c$mat

#Make linear plots of all bac and vir
x.25_2c <- x[,rownames(mat.25_2c)]
y.25_2c <- y[,colnames(mat.25_2c)]
xy.25_2c <- cbind(x.25_2c,y.25_2c)
x.2_2c <- x[,rownames(mat.2_2c)]
y.2_2c <- y[,colnames(mat.2_2c)]
xy.2_2c <- cbind(x.2_2c,y.2_2c)

#ggpairs
library(GGally)

#.25 matrix ggpairs
pairs.25 <- ggpairs(as.data.frame(xy.25_2c))
file_path <- "results/pairs.faec.bac.milk.metab.rcca.25.pdf"
if (!file.exists(file_path)) {
   ggsave(file_path,pairs.25,height=30,width=25,units="cm")
} else {
    cat("File already exists:", file_path, "\n")
}

#.2 matrix ggpairs
pairs.2 <-ggpairs(as.data.frame(xy.2_2c))
file_path <- "results/pairs.faec.bac.milk.metab.rcca.2.pdf"
if (!file.exists(file_path)) {
   ggsave(file_path,pairs.25,height=30,width=25,units="cm")
} else {
    cat("File already exists:", file_path, "\n")
}
```

#### Heatmaps - Pheatmap

##### 0.2 cutoff

```{r, eval = T, include = T, echo = T, message = F, warning =  F, fig.fullwidth = TRUE}
###### Select RCCA output(with cutoff)
rcca <- data.frame(high.2_2c$mat)


library(pheatmap)

palette <- hcl.colors(200, palette = "Zissou 1", alpha = NULL, rev = FALSE, fixup = TRUE) #Seclect palette https://developer.r-project.org/Blog/public/2019/04/01/hcl-based-color-palettes-in-grdevices/
palette[46:155] <- "white" #Select which part of pallette to replace with white

#Remove OTU IDs to only show taxonomy
tax <- gsub(".*\\.", "", colnames(rcca)) %>% make.unique(sep="")
tax
otu <- gsub("\\..*", "", colnames(rcca)) %>% make.unique(sep="")
otu

colnames(rcca) <- tax

######################## Annotations

###########  x-axis annotation

######## Phylogeny 1

rank <-"Family"

tax_levels <- ps@tax_table[otu,] %>%
    data.frame() %>%
  dplyr::select(1:rank) 

##Select last non-empty taxonomic rank
tax.best <- tax_levels %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured"
  apply(.,1, function(x) tail(stats::na.omit(x), 1)) %>%
  sub("_.*", "",.)
  

tax_levels$best <- tax.best 
sample_col <- tax_levels %>% dplyr::select(best) %>%
  dplyr::rename(!!rank := best)

######## Phylogeny 2

rank <-"Class"

tax_levels <- ps@tax_table[otu,] %>%
    data.frame() %>%
  dplyr::select(1:rank) 

##Select last non-empty taxonomic rank
tax.best <- tax_levels %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured"
  apply(.,1, function(x) tail(stats::na.omit(x), 1)) %>%
  sub("_.*", "",.)
  

tax_levels$best <- tax.best 
sample_col_2 <- tax_levels %>% dplyr::select(best) %>%
  dplyr::rename(!!rank := best)

############ Y-axis annotation

## Matbolite class

rownames(rcca) 

metab.class.sel <- metab.class[match(rownames(rcca), metab.class$Metabolite), ]

# Convert the "Metabolite" column to row names
rownames <- metab.class.sel$Metabolite

# Remove the "Metabolite" column from the dataframe

metab.class.sel <- metab.class.sel %>% as.data.frame() %>% dplyr::select(Class) 
rownames(metab.class.sel) <- rownames
metab.class.sel <- rename(metab.class.sel, Class = "Metabolite category")

######################### Clean species names 

##Select last non-empty taxonomic rank
rank <-"Species"

# x-axis annotation
## Phylogeny
tax_levels <- ps@tax_table[otu,] %>%
    data.frame() %>%
  dplyr::select(1:rank) 

##Select last non-empty taxonomic rank
tax.best <- tax_levels %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured"
  dplyr::mutate(Genus = str_remove(Genus, "_.*")) %>% #Remove odd genus annotations
  mutate(across(.cols = where(is.character), ~str_replace_all(., "_NK3B31_group", ""))) %>% # Fix wierd database match
  apply(.,1, function(x) tail(stats::na.omit(x), 1))
  

tax_levels$best <- tax.best 

tax_levels <- tax_levels %>%
  dplyr::mutate_all(~na_if(., '')) %>%
  dplyr::mutate(best = ifelse(is.na(.data[[rank]]), paste(best, "unclassified"), best)) 

colnames(rcca) <- make.unique(tax_levels$best) %>% gsub("_", " ",.)

# Align color vecors with clean tax names
row.names(sample_col) <- colnames(rcca)
#Add second annotation row
sample_col <- cbind(sample_col,sample_col_2)

#############

#Set colors
ann_colors <- list(Class = rep(col_fil,3)[1:n_distinct(sample_col[,1])])

sl(pheat.milkmet.faecbac.2,pheatmap(rcca,
         color=palette,
         #border_color = NA,
         angle_col = 45,
         annotation_col = sample_col,
         annotation_row =  metab.class.sel,
         cutree_rows = 3,
         cutree_cols = 3,
         main = "B - Infant faecal bacteria vs Milk metabolites"
         ),
   overwrite = TRUE
)

gridExtra::grid.arrange(pheat.milkmet.faecbac.2$gtable, vp=viewport(width=0.95, height=1))

```

##### 0.21 cutoff

```{r, eval = T, include = T, echo = T, message = F, warning =  F, fig.fullwidth = TRUE}
###### Select RCCA output(with cutoff)
rcca <- data.frame(high.21_2c$mat)


library(pheatmap)

palette <- hcl.colors(200, palette = "Zissou 1", alpha = NULL, rev = FALSE, fixup = TRUE) #Seclect palette https://developer.r-project.org/Blog/public/2019/04/01/hcl-based-color-palettes-in-grdevices/
palette[46:155] <- "white" #Select which part of pallette to replace with white

#Remove OTU IDs to only show taxonomy
tax <- gsub(".*\\.", "", colnames(rcca)) %>% make.unique(sep="")
tax
otu <- gsub("\\..*", "", colnames(rcca)) %>% make.unique(sep="")
otu

colnames(rcca) <- tax

######################## Annotations

###########  x-axis annotation

######## Phylogeny 1

rank <-"Family"

tax_levels <- ps@tax_table[otu,] %>%
    data.frame() %>%
  dplyr::select(1:rank) 

##Select last non-empty taxonomic rank
tax.best <- tax_levels %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured"
  apply(.,1, function(x) tail(stats::na.omit(x), 1)) %>%
  sub("_.*", "",.)
  

tax_levels$best <- tax.best 
sample_col <- tax_levels %>% dplyr::select(best) %>%
  dplyr::rename(!!rank := best)

######## Phylogeny 1

rank <-"Class"

tax_levels <- ps@tax_table[otu,] %>%
    data.frame() %>%
  dplyr::select(1:rank) 

##Select last non-empty taxonomic rank
tax.best <- tax_levels %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured"
  apply(.,1, function(x) tail(stats::na.omit(x), 1)) %>%
  sub("_.*", "",.)
  

tax_levels$best <- tax.best 
sample_col_2 <- tax_levels %>% dplyr::select(best) %>%
  dplyr::rename(!!rank := best)

############ Y-axis annotation

## Matbolite class

rownames(rcca) 

metab.class.sel <- metab.class[match(rownames(rcca), metab.class$Metabolite), ]

# Convert the "Metabolite" column to row names
rownames <- metab.class.sel$Metabolite

# Remove the "Metabolite" column from the dataframe

metab.class.sel <- metab.class.sel %>% as.data.frame() %>% dplyr::select(Class) 
rownames(metab.class.sel) <- rownames
metab.class.sel <- rename(metab.class.sel, Class = "Metabolite category")

######################### Clean species names 

##Select last non-empty taxonomic rank
rank <-"Species"

# x-axis annotation
## Phylogeny
tax_levels <- ps@tax_table[otu,] %>%
    data.frame() %>%
  dplyr::select(1:rank) 

##Select last non-empty taxonomic rank
tax.best <- tax_levels %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured"
  dplyr::mutate(Genus = str_remove(Genus, "_.*")) %>% #Remove odd genus annotations
  mutate(across(.cols = where(is.character), ~str_replace_all(., "_NK3B31_group", ""))) %>% # Fix wierd database match
  apply(.,1, function(x) tail(stats::na.omit(x), 1))
  

tax_levels$best <- tax.best 

tax_levels <- tax_levels %>%
  dplyr::mutate_all(~na_if(., '')) %>%
  dplyr::mutate(best = ifelse(is.na(.data[[rank]]), paste(best, "unclassified"), best)) 

colnames(rcca) <- make.unique(tax_levels$best) %>% gsub("_", " ",.)

# Align color vecors with clean tax names
row.names(sample_col) <- colnames(rcca)
#Add second annotation row
sample_col <- cbind(sample_col,sample_col_2)

#############

#Set colors
ann_colors <- list(Class = rep(col_fil,3)[1:n_distinct(sample_col[,1])])

sl(pheat.milkmet.faecbac.21,pheatmap(rcca,
         color=palette,
         #border_color = NA,
         angle_col = 45,
         annotation_col = sample_col,
         annotation_row =  metab.class.sel,
         cutree_rows = 3,
         cutree_cols = 3,
         main = "B - Infant faecal bacteria vs Milk metabolites"
         ),
   overwrite = TRUE
)

gridExtra::grid.arrange(pheat.milkmet.faecbac.21$gtable, vp=viewport(width=0.95, height=1))

```

##### 0.225 cutoff

```{r, eval = T, include = T, echo = T, message = F, warning =  F, fig.fullwidth = TRUE}
###### Select RCCA output(with cutoff)
rcca <- data.frame(high.225_2c$mat)


library(pheatmap)

palette <- hcl.colors(200, palette = "Zissou 1", alpha = NULL, rev = FALSE, fixup = TRUE) #Seclect palette https://developer.r-project.org/Blog/public/2019/04/01/hcl-based-color-palettes-in-grdevices/
palette[46:155] <- "white" #Select which part of pallette to replace with white

#Remove OTU IDs to only show taxonomy
tax <- gsub(".*\\.", "", colnames(rcca)) %>% make.unique(sep="")
tax
otu <- gsub("\\..*", "", colnames(rcca)) %>% make.unique(sep="")
otu

colnames(rcca) <- tax

######################## Annotations

###########  x-axis annotation

######## Phylogeny 1

rank <-"Family"

tax_levels <- ps@tax_table[otu,] %>%
    data.frame() %>%
  dplyr::select(1:rank) 

##Select last non-empty taxonomic rank
tax.best <- tax_levels %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured"
  apply(.,1, function(x) tail(stats::na.omit(x), 1)) %>%
  sub("_.*", "",.)
  

tax_levels$best <- tax.best 
sample_col <- tax_levels %>% dplyr::select(best) %>%
  dplyr::rename(!!rank := best)

######## Phylogeny 1

rank <-"Class"

tax_levels <- ps@tax_table[otu,] %>%
    data.frame() %>%
  dplyr::select(1:rank) 

##Select last non-empty taxonomic rank
tax.best <- tax_levels %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured"
  apply(.,1, function(x) tail(stats::na.omit(x), 1)) %>%
  sub("_.*", "",.)
  

tax_levels$best <- tax.best 
sample_col_2 <- tax_levels %>% dplyr::select(best) %>%
  dplyr::rename(!!rank := best)

############ Y-axis annotation

## Matbolite class

rownames(rcca) 

metab.class.sel <- metab.class[match(rownames(rcca), metab.class$Metabolite), ]

# Convert the "Metabolite" column to row names
rownames <- metab.class.sel$Metabolite

# Remove the "Metabolite" column from the dataframe

metab.class.sel <- metab.class.sel %>% as.data.frame() %>% dplyr::select(Class) 
rownames(metab.class.sel) <- rownames
metab.class.sel <- rename(metab.class.sel, Class = "Metabolite category")

######################### Clean species names 

##Select last non-empty taxonomic rank
rank <-"Species"

# x-axis annotation
## Phylogeny
tax_levels <- ps@tax_table[otu,] %>%
    data.frame() %>%
  dplyr::select(1:rank) 

##Select last non-empty taxonomic rank
tax.best <- tax_levels %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured"
  dplyr::mutate(Genus = str_remove(Genus, "_.*")) %>% #Remove odd genus annotations
  mutate(across(.cols = where(is.character), ~str_replace_all(., "_NK3B31_group", ""))) %>% # Fix wierd database match
  apply(.,1, function(x) tail(stats::na.omit(x), 1))
  

tax_levels$best <- tax.best 

tax_levels <- tax_levels %>%
  dplyr::mutate_all(~na_if(., '')) %>%
  dplyr::mutate(best = ifelse(is.na(.data[[rank]]), paste(best, "unclassified"), best)) 

colnames(rcca) <- make.unique(tax_levels$best) %>% gsub("_", " ",.)

# Align color vecors with clean tax names
row.names(sample_col) <- colnames(rcca)
#Add second annotation row
sample_col <- cbind(sample_col,sample_col_2)

#############

#Set colors
ann_colors <- list(Class = rep(col_fil,3)[1:n_distinct(sample_col[,1])])

newnames <- lapply(
  colnames(rcca),
  function(x) bquote(italic(.(x))))

sl(pheat.milkmet.faecbac.225,pheatmap(rcca,
         color=palette,
         #border_color = NA,
         #angle_col = 45,
         annotation_col = sample_col,
         annotation_row =  metab.class.sel,
         cutree_rows = 3,
         cutree_cols = 3,
         labels_col = as.expression(newnames),
         main = "B - Infant faecal bacteria vs Milk metabolites"
         ),
   overwrite = TRUE
)

gridExtra::grid.arrange(pheat.milkmet.faecbac.225$gtable, vp=viewport(width=0.95, height=1))

```

##### 0.25 cutoff

```{r, eval = T, include = T, echo = T, message = F, warning =  F, fig.fullwidth = TRUE}
###### Select RCCA output(with cutoff)
rcca <- data.frame(high.25_2c$mat)


library(pheatmap)

palette <- hcl.colors(200, palette = "Zissou 1", alpha = NULL, rev = FALSE, fixup = TRUE) #Seclect palette https://developer.r-project.org/Blog/public/2019/04/01/hcl-based-color-palettes-in-grdevices/
palette[46:155] <- "white" #Select which part of pallette to replace with white

#Remove OTU IDs to only show taxonomy
tax <- gsub(".*\\.", "", colnames(rcca)) %>% make.unique(sep="")
tax
otu <- gsub("\\..*", "", colnames(rcca)) %>% make.unique(sep="")
otu

colnames(rcca) <- tax

######################## Annotations

###########  x-axis annotation

######## Phylogeny 1

rank <-"Family"

tax_levels <- ps@tax_table[otu,] %>%
    data.frame() %>%
  dplyr::select(1:rank) 

##Select last non-empty taxonomic rank
tax.best <- tax_levels %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured"
  apply(.,1, function(x) tail(stats::na.omit(x), 1)) %>%
  sub("_.*", "",.)
  

tax_levels$best <- tax.best 
sample_col <- tax_levels %>% dplyr::select(best) %>%
  dplyr::rename(!!rank := best)

######## Phylogeny 1

rank <-"Class"

tax_levels <- ps@tax_table[otu,] %>%
    data.frame() %>%
  dplyr::select(1:rank) 

##Select last non-empty taxonomic rank
tax.best <- tax_levels %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured"
  apply(.,1, function(x) tail(stats::na.omit(x), 1)) %>%
  sub("_.*", "",.)
  

tax_levels$best <- tax.best 
sample_col_2 <- tax_levels %>% dplyr::select(best) %>%
  dplyr::rename(!!rank := best)

############ Y-axis annotation

## Matbolite class

rownames(rcca) 

metab.class.sel <- metab.class[match(rownames(rcca), metab.class$Metabolite), ]

# Convert the "Metabolite" column to row names
rownames <- metab.class.sel$Metabolite

# Remove the "Metabolite" column from the dataframe

metab.class.sel <- metab.class.sel %>% as.data.frame() %>% dplyr::select(Class) 
rownames(metab.class.sel) <- rownames
metab.class.sel <- rename(metab.class.sel, Class = "Metabolite category")

######################### Clean species names 

##Select last non-empty taxonomic rank
rank <-"Species"

# x-axis annotation
## Phylogeny
tax_levels <- ps@tax_table[otu,] %>%
    data.frame() %>%
  dplyr::select(1:rank) 

##Select last non-empty taxonomic rank
tax.best <- tax_levels %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured"
  dplyr::mutate(Genus = str_remove(Genus, "_.*")) %>% #Remove odd genus annotations
  mutate(across(.cols = where(is.character), ~str_replace_all(., "_NK3B31_group", ""))) %>% # Fix wierd database match
  apply(.,1, function(x) tail(stats::na.omit(x), 1))
  

tax_levels$best <- tax.best 

tax_levels <- tax_levels %>%
  dplyr::mutate_all(~na_if(., '')) %>%
  dplyr::mutate(best = ifelse(is.na(.data[[rank]]), paste(best, "unclassified"), best)) 

colnames(rcca) <- make.unique(tax_levels$best) %>% gsub("_", " ",.)

# Align color vecors with clean tax names
row.names(sample_col) <- colnames(rcca)
#Add second annotation row
sample_col <- cbind(sample_col,sample_col_2)

#############

#Set colors
ann_colors <- list(Class = rep(col_fil,3)[1:n_distinct(sample_col[,1])])

newnames <- lapply(
  colnames(rcca),
  function(x) bquote(italic(.(x))))

sl(pheat.milkmet.faecbac.25,pheatmap(rcca,
         color=palette,
         #border_color = NA,
         #angle_col = 45,
         annotation_col = sample_col,
         annotation_row =  metab.class.sel,
         cutree_rows = 3,
         cutree_cols = 3,
         labels_col = as.expression(newnames),
         main = "B - Infant faecal bacteria vs Milk metabolites"
         ),
   overwrite = TRUE
   )

gridExtra::grid.arrange(pheat.milkmet.faecbac.25$gtable, vp=viewport(width=0.97, height=1))

```

## Only Bifidobacteria vs milk metabolites

### Conversion and screeplot

```{r, eval = T, include = T}
#Subset samples present in both
ps <- subset_samples(PSB.CSS, Sample.type == "Faecal sample" & RCID %in% metab.id$RCID)
ps <- subset_taxa(ps, Family=="Bifidobacteriaceae")#subset_taxa(ps, Genus=="Bifidobacterium")
#rcid.shared <- intersect(metab.meta$RCID,ps.milk@sam_data$RCID)
x <- subset(metab.id, RCID %in% ps@sam_data$RCID) %>% dplyr::select(-RCID) %>%
  dplyr::mutate(across(where(is.numeric), ~ifelse(. < 0, NA_real_, sqrt(.))))
#colnames(x) <- make.names(colnames(x))
tax <- data.frame(ps@tax_table)

#Extract ASV data
y <- data.frame(ps@otu_table)
##Select last non-empty taxonomic rank
tax.bac <- as.data.frame(ps@tax_table) %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured" taxonomies (uninformative)
  apply(.,1, function(x) tail(stats::na.omit(x), 1))
#Combine otu_ids with taxonomies for unique identiiers
rownames(y) <- paste(rownames(y),tax.bac, sep = "-")


meta <- data.frame(ps@sam_data)


y <- t(y)


    ##Converting and removing near-zero variance values
library(mixOmics)

nzv <- nearZeroVar(x)
x <- x[, -nzv$Position]
           
           
#vir <- mvirome[,-c(1)]
nzv <- nearZeroVar(y) # If any
y <- y[, -nzv$Position]


  #PCA
library(vegan)
mod.x <- rda(x, scale = FALSE, center = FALSE)
plot(mod.x, scaling = 3)
screeplot(mod.x)
mod.y <- rda(y, scale = FALSE, center = FALSE)
plot(mod.y, scaling = 3)
screeplot(mod.y)

scaling = "ridge"
  #Apply scaling 
result_1comp <- rcc(x, y, method = scaling, ncomp = 1)
result_2comp <- rcc(x, y, method = scaling, ncomp = 2)
result_3comp <- rcc(x, y, method = scaling, ncomp = 3)
result_4comp <- rcc(x, y, method = scaling, ncomp = 4)
result_5comp <- rcc(x, y, method = scaling, ncomp = 5)
result_6comp <- rcc(x, y, method = scaling, ncomp = 6)

par(mfrow = c(2, 3))
plot(result_1comp,main="1 Component")
plot(result_2comp,main="2 Components")
plot(result_3comp,main="3 Components")
plot(result_4comp,main="4 Components")
plot(result_5comp,main="5 Components")
plot(result_6comp,main="6 Components")

# produce a heat map of the cross correlation matrix
imgCor(x, y, sideColors = c("purple", "green")) 
```

### Model tuning

```{r, eval = T, include = T}
# set grid search values for each regularisation parameter
grid1 <- seq(0.001, 0.2, length = 20) 
grid2 <- seq(0.001, 0.2, length = 20)

# optimise the regularisation parameter values
sl(cv.tune.rcc.milk.metab.faecal.bifido, tune.rcc(x, y, grid1 = grid1, grid2 = grid2, 
                                   validation = "loo"),overwrite = FALSE)

# examine the results of CV tuning
plot(cv.tune.rcc.milk.metab.faecal.bifido)
cv.tune.rcc.milk.metab.faecal.bifido

#manually set lambda scores
#opt.l1 <- 1.7
#opt.l2 <- 0.003
opt.l1 <- cv.tune.rcc.milk.metab.faecal.bifido$opt.lambda1
opt.l2 <- cv.tune.rcc.milk.metab.faecal.bifido$opt.lambda2
```

### Model and component selection

```{r, eval = T, include = T}
shrink.rcc <- rcc(x, y, method = "shrinkage")

CV.rcc <- rcc(x, y, method = "ridge", 
                         lambda1 = opt.l1,
                         lambda2 = opt.l2) 

# barplot of cross validation method rCCA canonical correlations
plot(CV.rcc, type = "barplot", main = "Cross Validation") 

# barplot of shrinkage method rCCA canonical correlations
plot(shrink.rcc, type = "barplot", main = "Shrinkage")
```

### Run model

```{r, eval = T, include = T}
#Perfrom optimised CV rCCA

result.CV <- rcc(x, y, method = "ridge",
                         lambda1 = opt.l1,
                         lambda2 = opt.l2,
              ncomp=2)

result.raw <- rcc(x, y, method = "shrinkage",ncomp=2)

#Remove OTU IDs to only show taxonomy
result <- result.raw

result$names$colnames$X <- gsub(".*\\-", "", result$names$colnames$X) %>% make.unique(sep="_")
result$names$colnames$Y <- gsub(".*\\-", "", result$names$colnames$Y) %>% make.unique(sep="_")


# colnames(result$X) <- gsub(".*\\-", "", colnames(result$X))
# colnames(result$Y) <- gsub(".*\\-", "", colnames(result$Y))

# rownames(result$loadings$X) <- gsub(".*\\-", "", rownames(result$loadings$X))
# rownames(result$loadings$Y) <- gsub(".*\\-", "", rownames(result$loadings$Y))


```

### Visualization

#### Heatmaps 

```{r, eval = T, include = T, echo = T, message = F, warning =  F, fig.fullwidth = TRUE}
#Defne pallette to only add colour to only high
##Generate new pallette
palette <- hcl.colors(200, palette = "Zissou 1", alpha = NULL, rev = FALSE, fixup = TRUE) #Seclect palette https://developer.r-project.org/Blog/public/2019/04/01/hcl-based-color-palettes-in-grdevices/
palette[61:140] <- "white" #Select which part of pallette to replace with white

# high4_2c <- cim(result, comp = 1:2, ylab = "Bacteria", xlab = "Virome", margins = c(12,18),cutoff = 0.4, color = palette)
# high45_2c <- cim(result, comp = 1:2, ylab = "Bacteria", xlab = "Virome", margins = c(12,18),cutoff = 0.45, color = palette)
#high5_3c <- cim(result, comp = 1:2, ylab = "Bacteria", xlab = "Virome", margins = c(12,18),cutoff = 0.5, color = palette)

#pdf(filename="high5_3c.pdf", res = 600, width = 4000, height = 4000)
#high5_3c
#dev.off()

######

# high6_3c <- cim(result, comp = 1:3, ylab = "Bacteria", xlab = "Virome host prediction", margins = c(7,9),cutoff = 0.6, color = palette)
# high5_3c <- cim(result, comp = 1:3, ylab = "Bacteria", xlab = "Virome host prediction", margins = c(7,9),cutoff = 0.5, color = palette)
# high45_3c<- cim(result, comp = 1:3, ylab = "Bacteria", xlab = "Virome host prediction", margins = c(15,20),cutoff = 0.45, color = palette)
high.25_2c <- cim(result, comp = 1:2, ylab = "Bacteria", xlab = "Virome host prediction", margins = c(15,20),cutoff = 0.25, color = palette)

high.2_2c <- cim(result, comp = 1:2, ylab = "Bacteria", xlab = "Virome host prediction", margins = c(15,20),cutoff = 0.2, color = palette)

high.21_2c <- cim(result, comp = 1:2, ylab = "Bacteria", xlab = "Virome host prediction", margins = c(15,20),cutoff = 0.21, color = palette)

high.175_2c <- cim(result, comp = 1:2, ylab = "Bacteria", xlab = "Virome host prediction", margins = c(15,20),cutoff = 0.175, color = palette)

high.15_2c <- cim(result, comp = 1:2, ylab = "Bacteria", xlab = "Virome host prediction", margins = c(15,20),cutoff = 0.15, color = palette)

mat.2_2c <- high.2_2c$mat

 #Output Matrix
#write.table(mat.2_3c, "Matrix_high_3-Component_0.45.txt", sep="\t")


```

#### Scatterplots

```{r}
mat.25_2c <- high.25_2c$mat
mat.2_2c <- high.2_2c$mat

#Make linear plots of all bac and vir
x.25_2c <- x[,rownames(mat.25_2c)]
y.25_2c <- y[,colnames(mat.25_2c)]
xy.25_2c <- cbind(x.25_2c,y.25_2c)
x.2_2c <- x[,rownames(mat.2_2c)]
y.2_2c <- y[,colnames(mat.2_2c)]
xy.2_2c <- cbind(x.2_2c,y.2_2c)

#ggpairs
library(GGally)

#.25 matrix ggpairs
pairs.25 <- ggpairs(as.data.frame(xy.25_2c))
file_path <- "results/pairs.faec.bifido.milk.metab.rcca.25.pdf"
if (!file.exists(file_path)) {
   ggsave(file_path,pairs.25,height=30,width=25,units="cm")
} else {
    cat("File already exists:", file_path, "\n")
}

#.2 matrix ggpairs
pairs.2 <-ggpairs(as.data.frame(xy.2_2c))
file_path <- "results/pairs.faec.bifido.milk.metab.rcca.2.pdf"
if (!file.exists(file_path)) {
   ggsave(file_path,pairs.25,height=30,width=25,units="cm")
} else {
    cat("File already exists:", file_path, "\n")
}
```

#### Heatmaps - Pheatmap

##### 0.21 cutoff

```{r, eval = T, include = T, echo = T, message = F, warning =  F, fig.fullwidth = TRUE}
###### Select RCCA output(with cutoff)
rcca <- data.frame(high.21_2c$mat)


library(pheatmap)

palette <- hcl.colors(200, palette = "Zissou 1", alpha = NULL, rev = FALSE, fixup = TRUE) #Seclect palette https://developer.r-project.org/Blog/public/2019/04/01/hcl-based-color-palettes-in-grdevices/
palette[46:155] <- "white" #Select which part of pallette to replace with white

#Remove OTU IDs to only show taxonomy
tax <- gsub(".*\\.", "", colnames(rcca)) %>% make.unique(sep="")
tax
otu <- gsub("\\..*", "", colnames(rcca)) %>% make.unique(sep="")
otu

colnames(rcca) <- tax

######################## Annotations

###########  x-axis annotation

######## Phylogeny 1

rank <-"Family"

tax_levels <- ps@tax_table[otu,] %>%
    data.frame() %>%
  dplyr::select(1:rank) 

##Select last non-empty taxonomic rank
tax.best <- tax_levels %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured"
  apply(.,1, function(x) tail(stats::na.omit(x), 1)) %>%
  sub("_.*", "",.)
  

tax_levels$best <- tax.best 
sample_col <- tax_levels %>% dplyr::select(best) %>%
  dplyr::rename(!!rank := best)

######## Phylogeny 1

rank <-"Class"

tax_levels <- ps@tax_table[otu,] %>%
    data.frame() %>%
  dplyr::select(1:rank) 

##Select last non-empty taxonomic rank
tax.best <- tax_levels %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured"
  apply(.,1, function(x) tail(stats::na.omit(x), 1)) %>%
  sub("_.*", "",.)
  

tax_levels$best <- tax.best 
sample_col_2 <- tax_levels %>% dplyr::select(best) %>%
  dplyr::rename(!!rank := best)

############ Y-axis annotation

## Matbolite class

rownames(rcca) 

metab.class.sel <- metab.class[match(rownames(rcca), metab.class$Metabolite), ]

# Convert the "Metabolite" column to row names
rownames <- metab.class.sel$Metabolite

# Remove the "Metabolite" column from the dataframe

metab.class.sel <- metab.class.sel %>% as.data.frame() %>% dplyr::select(Class) 
rownames(metab.class.sel) <- rownames
metab.class.sel <- rename(metab.class.sel, Class = "Metabolite category")

######################### Clean species names 

##Select last non-empty taxonomic rank
rank <-"Species"

# x-axis annotation
## Phylogeny
tax_levels <- ps@tax_table[otu,] %>%
    data.frame() %>%
  dplyr::select(1:rank) 

##Select last non-empty taxonomic rank
tax.best <- tax_levels %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured"
  dplyr::mutate(Genus = str_remove(Genus, "_.*")) %>% #Remove odd genus annotations
  mutate(across(.cols = where(is.character), ~str_replace_all(., "_NK3B31_group", ""))) %>% # Fix wierd database match
  apply(.,1, function(x) tail(stats::na.omit(x), 1))
  

tax_levels$best <- tax.best 

tax_levels <- tax_levels %>%
  dplyr::mutate_all(~na_if(., '')) %>%
  dplyr::mutate(best = ifelse(is.na(.data[[rank]]), paste(best, "unclassified"), best)) 

colnames(rcca) <- make.unique(tax_levels$best) %>% gsub("_", " ",.)

# Align color vecors with clean tax names
row.names(sample_col) <- colnames(rcca)
#Add second annotation row
sample_col <- cbind(sample_col,sample_col_2)

#############

#Set colors
ann_colors <- list(Class = rep(col_fil,3)[1:n_distinct(sample_col[,1])])

sl(pheat.milkmet.faecbifido.21,pheatmap(rcca,
         color=palette,
         #border_color = NA,
         angle_col = 45,
         annotation_col = sample_col,
         annotation_row =  metab.class.sel,
         cutree_rows = 3,
         cutree_cols = 3,
         main = "C - Infant faecal bifidobacteria vs Milk metabolites"
         ),
   overwrite = TRUE
)

gridExtra::grid.arrange(pheat.milkmet.faecbifido.21$gtable, vp=viewport(width=0.95, height=1))

```

##### 0.2 cutoff

```{r, eval = T, include = T, echo = T, message = F, warning =  F, fig.fullwidth = TRUE}
###### Select RCCA output(with cutoff)
rcca <- data.frame(high.2_2c$mat)


library(pheatmap)

palette <- hcl.colors(200, palette = "Zissou 1", alpha = NULL, rev = FALSE, fixup = TRUE) #Seclect palette https://developer.r-project.org/Blog/public/2019/04/01/hcl-based-color-palettes-in-grdevices/
palette[46:155] <- "white" #Select which part of pallette to replace with white

#Remove OTU IDs to only show taxonomy
tax <- gsub(".*\\.", "", colnames(rcca)) %>% make.unique(sep="")
tax
otu <- gsub("\\..*", "", colnames(rcca)) %>% make.unique(sep="")
otu

colnames(rcca) <- tax

######################## Annotations

###########  x-axis annotation

######## Phylogeny 1

rank <-"Species"

tax_levels <- ps@tax_table[otu,] %>%
    data.frame() %>%
  dplyr::select(1:rank) 

##Select last non-empty taxonomic rank
tax.best <- tax_levels %>% dplyr::mutate_all(~na_if(., '')) %>%
  dplyr::mutate(Species = replace(Species, is.na(Species), "unclassified")) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured"
  apply(.,1, function(x) tail(stats::na.omit(x), 1)) #%>%
  #sub("_.*", "",.)
  

tax_levels$best <-  sapply(strsplit(gsub("_", " ", tax.best), " "), function(x) tail(x, 1))

sample_col <- tax_levels %>% dplyr::select(best) %>%
  dplyr::rename(!!rank := best)

######## Phylogeny 2

rank <-"Class"

tax_levels <- ps@tax_table[otu,] %>%
    data.frame() %>%
  dplyr::select(1:rank) 

##Select last non-empty taxonomic rank
tax.best <- tax_levels %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured"
  apply(.,1, function(x) tail(stats::na.omit(x), 1)) %>%
  sub("_.*", "",.)
  

tax_levels$best <- tax.best 
sample_col_2 <- tax_levels %>% dplyr::select(best) %>%
  dplyr::rename(!!rank := best)

############ Y-axis annotation

## Matbolite class

rownames(rcca) 

metab.class.sel <- metab.class[match(rownames(rcca), metab.class$Metabolite), ]

# Convert the "Metabolite" column to row names
rownames <- metab.class.sel$Metabolite

# Remove the "Metabolite" column from the dataframe

metab.class.sel <- metab.class.sel %>% as.data.frame() %>% dplyr::select(Class) 
rownames(metab.class.sel) <- rownames
metab.class.sel <- rename(metab.class.sel, Class = "Metabolite category")

######################### Clean species names 

##Select last non-empty taxonomic rank
rank <-"Species"

# x-axis annotation
## Phylogeny
tax_levels <- ps@tax_table[otu,] %>%
    data.frame() %>%
  dplyr::select(1:rank) 

##Select last non-empty taxonomic rank
tax.best <- tax_levels %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured"
  dplyr::mutate(Genus = str_remove(Genus, "_.*")) %>% #Remove odd genus annotations
  mutate(across(.cols = where(is.character), ~str_replace_all(., "_NK3B31_group", ""))) %>% # Fix wierd database match
  apply(.,1, function(x) tail(stats::na.omit(x), 1))
  

tax_levels$best <- tax.best 

tax_levels <- tax_levels %>%
  dplyr::mutate_all(~na_if(., '')) %>%
  dplyr::mutate(best = ifelse(is.na(.data[[rank]]), paste(best, "unclassified"), best))

colnames(rcca) <- make.unique(tax_levels$best) %>% gsub("_", " ",.)

# Align color vecors with clean tax names
row.names(sample_col) <- colnames(rcca)
#Add second annotation row
#sample_col <- cbind(sample_col,sample_col_2)

#############

#Set colors
ann_colors <- list(Class = rep(col_fil,3)[1:n_distinct(sample_col[,1])])

sl(pheat.milkmet.faecbifido.2,pheatmap(rcca,
         color=palette,
         #border_color = NA,
         angle_col = 45,
         annotation_col = sample_col,
         annotation_row =  metab.class.sel,
         cutree_rows = 3,
         cutree_cols = 3,
         main = "C - Infant faecal bifidobacteria vs Milk metabolites"
         ),
   overwrite = TRUE
)

gridExtra::grid.arrange(pheat.milkmet.faecbifido.2$gtable, vp=viewport(width=0.95, height=1))

```

##### 0.175 cutoff

```{r, eval = T, include = T, echo = T, message = F, warning =  F, fig.fullwidth = TRUE}
###### Select RCCA output(with cutoff)
rcca <- data.frame(high.175_2c$mat)


library(pheatmap)

palette <- hcl.colors(200, palette = "Zissou 1", alpha = NULL, rev = FALSE, fixup = TRUE) #Seclect palette https://developer.r-project.org/Blog/public/2019/04/01/hcl-based-color-palettes-in-grdevices/
palette[46:155] <- "white" #Select which part of pallette to replace with white

#Remove OTU IDs to only show taxonomy
tax <- gsub(".*\\.", "", colnames(rcca)) %>% make.unique(sep="")
tax
otu <- gsub("\\..*", "", colnames(rcca)) %>% make.unique(sep="")
otu

colnames(rcca) <- tax

######################## Annotations

###########  x-axis annotation

######## Phylogeny 1

rank <-"Species"

tax_levels <- ps@tax_table[otu,] %>%
    data.frame() %>%
  dplyr::select(1:rank) 

##Select last non-empty taxonomic rank
tax.best <- tax_levels %>% dplyr::mutate_all(~na_if(., '')) %>%
  dplyr::mutate(Species = replace(Species, is.na(Species), "unclassified")) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured"
  apply(.,1, function(x) tail(stats::na.omit(x), 1)) #%>%
  #sub("_.*", "",.)
  

tax_levels$best <-  sapply(strsplit(gsub("_", " ", tax.best), " "), function(x) tail(x, 1))

sample_col <- tax_levels %>% dplyr::select(best) %>%
  dplyr::rename(!!rank := best)

######## Phylogeny 2 - not used 

rank <-"Class"

tax_levels <- ps@tax_table[otu,] %>%
    data.frame() %>%
  dplyr::select(1:rank) 

##Select last non-empty taxonomic rank
tax.best <- tax_levels %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured"
  apply(.,1, function(x) tail(stats::na.omit(x), 1)) %>%
  sub("_.*", "",.)
  

tax_levels$best <- tax.best 
sample_col_2 <- tax_levels %>% dplyr::select(best) %>%
  dplyr::rename(!!rank := best)

############ Y-axis annotation

## Matbolite class

rownames(rcca) 

metab.class.sel <- metab.class[match(rownames(rcca), metab.class$Metabolite), ]

# Convert the "Metabolite" column to row names
rownames <- metab.class.sel$Metabolite

# Remove the "Metabolite" column from the dataframe

metab.class.sel <- metab.class.sel %>% as.data.frame() %>% dplyr::select(Class) 
rownames(metab.class.sel) <- rownames
metab.class.sel <- rename(metab.class.sel, Class = "Metabolite category")

######################### Clean species names 

##Select last non-empty taxonomic rank
rank <-"Species"

# x-axis annotation
## Phylogeny
tax_levels <- ps@tax_table[otu,] %>%
    data.frame() %>%
  dplyr::select(1:rank) 

##Select last non-empty taxonomic rank
tax.best <- tax_levels %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured"
  dplyr::mutate(Genus = str_remove(Genus, "_.*")) %>% #Remove odd genus annotations
  mutate(across(.cols = where(is.character), ~str_replace_all(., "_NK3B31_group", ""))) %>% # Fix wierd database match
  apply(.,1, function(x) tail(stats::na.omit(x), 1))
  

tax_levels$best <- tax.best 

tax_levels <- tax_levels %>%
  dplyr::mutate_all(~na_if(., '')) %>%
  dplyr::mutate(best = ifelse(is.na(.data[[rank]]), paste(best, "unclassified"), best)) 

colnames(rcca) <- make.unique(tax_levels$best) %>% gsub("_", " ",.)

# Align color vecors with clean tax names
row.names(sample_col) <- colnames(rcca)
#Add second annotation row
#sample_col <- cbind(sample_col,sample_col_2)

#############

#Set colors
ann_colors <- list(Class = rep(col_fil,3)[1:n_distinct(sample_col[,1])])

newnames <- lapply(
  colnames(rcca),
  function(x) bquote(italic(.(x))))

sl(pheat.milkmet.faecbifido.175,pheatmap(rcca,
         color=palette,
         #border_color = NA,
         #angle_col = 45,
         annotation_col = sample_col,
         annotation_row =  metab.class.sel,
         cutree_rows = 4,
         cutree_cols = 4,
         labels_col = as.expression(newnames),
         main = "C - Infant faecal bifidobacteria vs Milk metabolites"
         ),
   overwrite = TRUE
)

gridExtra::grid.arrange(pheat.milkmet.faecbifido.175$gtable, vp=viewport(width=0.95, height=1))

```

#### Network plots

##### Shrinkage

```{r, eval = T, include = T, echo = T, message = F, warning =  F, fig.width=10, fig.height=6, fig.fullwidth = TRUE}
# Reformat for gephi

## Save blocks as polygon to determine shape in network

result$names$Polygon <- result$names$data
result$names$Polygon <- replace(result$names$data, c("x","vy"),c(3,4)) %>% as.integer()

## Remove ubique markers to allow shared coloring in Gephi
result$names_simple$colnames$X <- gsub("*.\\_", "", result$names$colnames$X)
result$names_simple$colnames$Y <- gsub("*.\\_", "", result$names$colnames$Y)

#Show network

network(result, comp = 1:2, interactive = FALSE,
        lwd.edge = 2,
        cutoff = 0.2,
        color.node = col_fil[1:2])

network(result, comp = 1:2, interactive = FALSE,
        lwd.edge = 2,
        cutoff = 0.15,
        color.node = col_fil[1:2]) 


library(igraph)

#network.res 

write.graph(network.res$gR, file = "network.shrinkage0.3.gml", format = "gml")
```

##### CV

```{r, eval = T, include = T, echo = T, message = F, warning =  F, fig.width=10, fig.height=6, fig.fullwidth = TRUE}
result.CV$names_simple$colnames$X <- gsub("*.\\_", "", result.CV$names$colnames$X)
result.CV$names_simple$colnames$Y <- gsub("*.\\_", "", result.CV$names$colnames$Y)

#Show network
network(result.CV, comp = 1:2, interactive = FALSE,
        lwd.edge = 2,
        cutoff = 0.25,
        color.node = col_fil[1:2]) 

network(result.CV, comp = 1:2, interactive = FALSE,
        lwd.edge = 2,
        cutoff = 0.2,
        color.node = col_fil[1:2]) 

# network(result.CV, comp = 1:3, interactive = FALSE,
#         lwd.edge = 2,
#         cutoff = 0.5,
#         color.node = col_fil[1:2]) 


library(igraph)
network.res <- network(result, comp = 1:2, interactive = FALSE,
        lwd.edge = 2,
        cutoff = 0.175,
        color.node = col_fil[1:2])

write.graph(network.res$gR, file = "network.CV0.3.gml", format = "gml")
```

## Milk bac vs infant faecal bac

### Conversion and screeplot

```{r, eval = F, include = T}
#Subset samples present in both
ps <- subset_samples(PSB.CSS, Sample.type == "Faecal sample" & RCID %in% PSB.CSS@sam_data$RCID)
ps2 <- subset_samples(PSB.CSS, Sample.type == "Milk" & RCID %in% metab.id$RCID)

ps <- subset_samples(ps, RCID %in% ps2@sam_data$RCID)
ps2 <- subset_samples(ps2, RCID %in% ps@sam_data$RCID)

ps <- prune_taxa(taxa_sums(ps) > 1, ps)
ps2 <- prune_taxa(taxa_sums(ps2) > 1, ps2)


#Extract ASV data
x <- data.frame(ps2@otu_table)
##Select last non-emptx taxonomic rank
tax.bac <- as.data.frame(ps2@tax_table) %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured" taxonomies (uninformative) 
  mutate(across(.cols = where(is.character), ~str_replace_all(., "_sensu_stricto_1", ""))) %>% # remove all "sens strictu" (uninformative)
  apply(.,1, function(x) tail(stats::na.omit(x), 1))
#Combine otu_ids with taxonomies for unique identiiers
rownames(x) <- paste(rownames(x),tax.bac, sep = "-")


meta2 <- data.frame(ps2@sam_data)

x<- t(x)

#Extract ASV data
y <- data.frame(ps@otu_table)
##Select last non-empty taxonomic rank
tax.bac <- as.data.frame(ps@tax_table) %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured" taxonomies (uninformative)
  apply(.,1, function(x) tail(stats::na.omit(x), 1))
#Combine otu_ids with taxonomies for unique identiiers
rownames(y) <- paste(rownames(y),tax.bac, sep = "-")


meta <- data.frame(ps@sam_data)


y <- t(y)


    ##Converting and removing near-zero variance values
library(mixOmics)

nzv <- nearZeroVar(x)
x <- x[, -nzv$Position]
       
           
#vir <- mvirome[,-c(1)]
nzv <- nearZeroVar(y) # If any
y <- y[, -nzv$Position]

  #PCA
library(vegan)
mod.x <- rda(x, scale = FALSE, center = FALSE)
plot(mod.x, scaling = 3)
screeplot(mod.x)
mod.y <- rda(y, scale = FALSE, center = FALSE)
plot(mod.y, scaling = 3)
screeplot(mod.y)

scaling = "ridge"
  #Apply scaling 
result_1comp <- rcc(x, y, method = scaling, ncomp = 1)
result_2comp <- rcc(x, y, method = scaling, ncomp = 2)
result_3comp <- rcc(x, y, method = scaling, ncomp = 3)
result_4comp <- rcc(x, y, method = scaling, ncomp = 4)
result_5comp <- rcc(x, y, method = scaling, ncomp = 5)
result_6comp <- rcc(x, y, method = scaling, ncomp = 6)

par(mfrow = c(2, 3))
plot(result_1comp,main="1 Component")
plot(result_2comp,main="2 Components")
plot(result_3comp,main="3 Components")
plot(result_4comp,main="4 Components")
plot(result_5comp,main="5 Components")
plot(result_6comp,main="6 Components")

# produce a heat map of the cross correlation matrix
imgCor(x, y, sideColors = c("purple", "green")) 
```

### Model tuning

```{r, eval = T, include = T}
# set grid search values for each regularisation parameter
grid1 <- seq(0.001, 0.2, length = 10) 
grid2 <- seq(0.001, 0.2, length = 10)

# optimise the regularisation parameter values
sl(cv.tune.rcc.milk.bac.faecal.bac, tune.rcc(x, y, grid1 = grid1, grid2 = grid2, 
                                   validation = "loo"),overwrite = FALSE)

# examine the results of CV tuning
plot(cv.tune.rcc.milk.bac.faecal.bac)
cv.tune.rcc.milk.bac.faecal.bac

#manually set lambda scores
#opt.l1 <- 1.7
#opt.l2 <- 0.003
opt.l1 <- cv.tune.rcc.milk.bac.faecal.bac$opt.lambda1
opt.l2 <- cv.tune.rcc.milk.bac.faecal.bac$opt.lambda2
```

### Model and component selection

```{r, eval = T, include = T}
shrink.rcc <- rcc(x, y, method = "shrinkage")

CV.rcc <- rcc(x, y, method = "ridge", 
                         lambda1 = opt.l1,
                         lambda2 = opt.l2) 

# barplot of cross validation method rCCA canonical correlations
plot(CV.rcc, type = "barplot", main = "Cross Validation") 

# barplot of shrinkage method rCCA canonical correlations
plot(shrink.rcc, type = "barplot", main = "Shrinkage")
```

### Run model

```{r, eval = T, include = T}
#Perfrom optimised CV rCCA

result.CV <- rcc(x, y, method = "ridge",
                         lambda1 = opt.l1,
                         lambda2 = opt.l2,
              ncomp=2)

result.raw <- rcc(x, y, method = "shrinkage",ncomp=3)

#Remove OTU IDs to only show taxonomy
result <- result.raw

result$names$colnames$X <- gsub(".*\\-", "", result$names$colnames$X) %>% make.unique(sep="_")
result$names$colnames$Y <- gsub(".*\\-", "", result$names$colnames$Y) %>% make.unique(sep="_")


# colnames(result$X) <- gsub(".*\\-", "", colnames(result$X))
# colnames(result$Y) <- gsub(".*\\-", "", colnames(result$Y))

# rownames(result$loadings$X) <- gsub(".*\\-", "", rownames(result$loadings$X))
# rownames(result$loadings$Y) <- gsub(".*\\-", "", rownames(result$loadings$Y))


```

### Visualization

#### Heatmaps 

```{r, eval = T, include = T, echo = T, message = F, warning =  F, fig.fullwidth = TRUE}
#Defne pallette to only add colour to only high
##Generate new pallette
palette <- hcl.colors(200, palette = "Zissou 1", alpha = NULL, rev = FALSE, fixup = TRUE) #Seclect palette https://developer.r-project.org/Blog/public/2019/04/01/hcl-based-color-palettes-in-grdevices/
palette[61:140] <- "white" #Select which part of pallette to replace with white

# high4_2c <- cim(result, comp = 1:2, ylab = "Bacteria", xlab = "Virome", margins = c(12,18),cutoff = 0.4, color = palette)
# high45_2c <- cim(result, comp = 1:2, ylab = "Bacteria", xlab = "Virome", margins = c(12,18),cutoff = 0.45, color = palette)
#high5_3c <- cim(result, comp = 1:2, ylab = "Bacteria", xlab = "Virome", margins = c(12,18),cutoff = 0.5, color = palette)

#pdf(filename="high5_3c.pdf", res = 600, width = 4000, height = 4000)
#high5_3c
#dev.off()

######

high.3_3c <- cim(result, comp = 1:3, ylab = "Faecal bacteria", xlab = "Milk bacteria", margins = c(15,20),cutoff = 0.3, color = palette)

high.275_3c <- cim(result, comp = 1:3, ylab = "Faecal bacteria", xlab = "Milk bacteria", margins = c(15,20),cutoff = 0.275, color = palette)

high.25_3c <- cim(result, comp = 1:3, ylab = "Faecal bacteria", xlab = "Milk bacteria", margins = c(15,20),cutoff = 0.25, color = palette)

high.225_3c <- cim(result, comp = 1:3, ylab = "Faecal bacteria", xlab = "Milk bacteria", margins = c(15,20),cutoff = 0.225, color = palette)

high.21_3c <- cim(result, comp = 1:3, ylab = "Faecal bacteria", xlab = "Milk bacteria", margins = c(15,20),cutoff = 0.21, color = palette)

high.2_3c <- cim(result, comp = 1:3, ylab = "Faecal bacteria", xlab = "Milk bacteria", margins = c(15,20),cutoff = 0.2, color = palette)

high.175_3c <- cim(result, comp = 1:3, ylab = "Faecal bacteria", xlab = "Milk bacteria", margins = c(15,20),cutoff = 0.175, color = palette)

high.15_3c <- cim(result, comp = 1:3, ylab = "Faecal bacteria", xlab = "Milk bacteria", margins = c(15,20),cutoff = 0.15, color = palette)

mat.2_3c <- high.2_3c$mat

 #Output Matrix
#write.table(mat.2_3c, "Matrix_high_3-Component_0.45.txt", sep="\t")


```

#### Scatterplots

```{r}
mat.25_2c <- high.25_2c$mat
mat.2_2c <- high.2_2c$mat

#Make linear plots of all bac and vir
x.25_2c <- x[,rownames(mat.25_2c)]
y.25_2c <- y[,colnames(mat.25_2c)]
xy.25_2c <- cbind(x.25_2c,y.25_2c)
x.2_2c <- x[,rownames(mat.2_2c)]
y.2_2c <- y[,colnames(mat.2_2c)]
xy.2_2c <- cbind(x.2_2c,y.2_2c)

#ggpairs
library(GGally)

#.25 matrix ggpairs
pairs.25 <- ggpairs(as.data.frame(xy.25_2c))
file_path <- "results/pairs.faec.bac.milk.metab.rcca.25.pdf"
if (!file.exists(file_path)) {
   ggsave(file_path,pairs.25,height=30,width=25,units="cm")
} else {
    cat("File already exists:", file_path, "\n")
}

#.2 matrix ggpairs
pairs.2 <-ggpairs(as.data.frame(xy.2_2c))
file_path <- "results/pairs.faec.bac.milk.metab.rcca.2.pdf"
if (!file.exists(file_path)) {
   ggsave(file_path,pairs.25,height=30,width=25,units="cm")
} else {
    cat("File already exists:", file_path, "\n")
}
```

#### Heatmaps - Pheatmap

##### 0.3 cutoff

```{r, eval = T, include = T, echo = T, message = F, warning =  F, fig.fullwidth = TRUE}
###### Select RCCA output(with cutoff)
rcca <- data.frame(high.3_3c$mat)


library(pheatmap)

palette <- hcl.colors(200, palette = "Zissou 1", alpha = NULL, rev = FALSE, fixup = TRUE) #Seclect palette https://developer.r-project.org/Blog/public/2019/04/01/hcl-based-color-palettes-in-grdevices/
palette[46:155] <- "white" #Select which part of pallette to replace with white

#Remove OTU IDs to only show taxonomy
tax <- gsub(".*\\.", "", colnames(rcca)) %>% make.unique(sep="")
otu <- gsub("\\..*", "", colnames(rcca)) %>% make.unique(sep="")

tax_y <- gsub(".*\\.", "", rownames(rcca)) %>% make.unique(sep="")
otu_y <- gsub("\\-.*", "", rownames(rcca)) %>% make.unique(sep="")


colnames(rcca) <- tax

######################## Annotations

###########  x-axis annotation

######## Phylogeny 1

rank <-"Family"

tax_levels <- ps@tax_table[otu,] %>%
    data.frame() %>%
  dplyr::select(1:rank) 

##Select last non-empty taxonomic rank
tax.best <- tax_levels %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured"
  apply(.,1, function(x) tail(stats::na.omit(x), 1)) %>%
  sub("_.*", "",.)
  

tax_levels$best <- tax.best 
sample_col <- tax_levels %>% dplyr::select(best) %>%
  dplyr::rename(!!rank := best)

######## Phylogeny 2

rank <-"Class"

tax_levels <- ps@tax_table[otu,] %>%
    data.frame() %>%
  dplyr::select(1:rank) 

##Select last non-empty taxonomic rank
tax.best <- tax_levels %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured"
  apply(.,1, function(x) tail(stats::na.omit(x), 1)) %>%
  sub("_.*", "",.)

tax_levels$best <- tax.best 
sample_col2 <- tax_levels %>% dplyr::select(best) %>%
  dplyr::rename(!!rank := best)

############ Y-axis annotation

######## Phylogeny 1

rank <-"Family"

tax_levels_2 <- ps2@tax_table[otu_y,] %>%
    data.frame() %>%
  dplyr::select(1:rank) 

##Select last non-empty taxonomic rank
tax.best <- tax_levels_2 %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured"
  apply(.,1, function(x) tail(stats::na.omit(x), 1)) %>%
  sub("_.*", "",.)
  

tax_levels_2$best <- tax.best 
sample_col_y <- tax_levels_2 %>% dplyr::select(best) %>%
  dplyr::rename(!!rank := best)

######## Phylogeny 2

rank <-"Class"

tax_levels_2 <- ps2@tax_table[otu_y,] %>%
    data.frame() %>%
  dplyr::select(1:rank) 

##Select last non-empty taxonomic rank
tax.best <- tax_levels_2 %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured"
  apply(.,1, function(x) tail(stats::na.omit(x), 1)) %>%
  sub("_.*", "",.)

tax_levels_2$best <- tax.best 
sample_col_y_2 <- tax_levels_2 %>% dplyr::select(best) %>%
  dplyr::rename(!!rank := best)

################################## Clean species names 

##Select last non-empty taxonomic rank
rank <-"Species"

################## x-axis annotation

## Phylogeny
tax_levels <- ps@tax_table[otu,] %>%
    data.frame() %>%
  dplyr::select(1:rank) 

##Select last non-empty taxonomic rank
tax.best <- tax_levels %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured"
  dplyr::mutate(Genus = str_remove(Genus, "_.*")) %>% #Remove odd genus annotations
  mutate(across(.cols = where(is.character), ~str_replace_all(., "_NK3B31_group", ""))) %>% # Fix wierd database match
  apply(.,1, function(x) tail(stats::na.omit(x), 1))
  

tax_levels$best <- tax.best 

tax_levels <- tax_levels %>%
  dplyr::mutate_all(~na_if(., '')) %>%
  dplyr::mutate(best = ifelse(is.na(.data[[rank]]), paste(best, "unclassified"), best)) 

colnames(rcca) <- make.unique(tax_levels$best) %>% gsub("_", " ",.)

# Align color vecors with clean tax names
row.names(sample_col) <- colnames(rcca)
#Add second annotation row
sample_col <- cbind(sample_col,sample_col_2)

################## y-axis annotation

## Phylogeny
tax_levels <- ps2@tax_table[otu_y,] %>%
    data.frame() %>%
  dplyr::select(1:rank) 

##Select last non-empty taxonomic rank
tax.best <- tax_levels %>% dplyr::mutate_all(~na_if(., '')) %>%
  mutate(across(.cols = where(is.character), ~if_else(str_detect(., "uncultured"), NA_character_, .))) %>% # remove all "_uncultured"
  dplyr::mutate(Genus = str_remove(Genus, "_.*")) %>% #Remove odd genus annotations
  mutate(across(.cols = where(is.character), ~str_replace_all(., "_NK3B31_group", ""))) %>% # Fix wierd database match
  apply(.,1, function(x) tail(stats::na.omit(x), 1))
  

tax_levels$best <- tax.best 

tax_levels <- tax_levels %>%
  dplyr::mutate_all(~na_if(., '')) %>%
  dplyr::mutate(best = ifelse(is.na(.data[[rank]]), paste(best, "unclassified"), best)) 
rownames(rcca) <- make.unique(tax_levels$best) %>% gsub("_", " ",.)

# Align color vecors with clean tax names
row.names(sample_col_y) <- rownames(rcca)
#Add second annotation row
sample_col_y <- cbind(sample_col_y,sample_col_y_2)


#############

sample_col <- sample_col %>%
  dplyr::rename("Class faec" = Class, "Family faec" = Family)

#Set colors
ann_colors <- list(Class = rep(col_fil,3)[1:n_distinct(sample_col[,1])])

sl(pheat.milkbac.faecbac.3,pheatmap(rcca,
         color=palette,
         #border_color = NA,
         angle_col = 45,
         annotation_col = sample_col,
         annotation_row =  sample_col_y,
         cutree_rows = 3,
         cutree_cols = 3,
         main = "Infant faecal bacteria vs Milk bacteria"
         ),
   overwrite = TRUE
)

gridExtra::grid.arrange(pheat.milkbac.faecbac.3$gtable, vp=viewport(width=0.95, height=1))

```



## Arrange heatmaps 

```{r}
library(ggplotify)
library(patchwork)

#overall

p <- gridExtra::grid.arrange(pheat.milkmet.milkbac.2$gtable,
                        pheat.milkmet.faecbac.21$gtable,
                        heights	= c(1,1),
                        vp=viewport(width=0.95, height=1))

p

ggsave("Figures/RCCA.tiff",p, width = 30, height = 30, units = "cm", dpi = 400) 

ggsave("Figures/RCCA.pdf",p,width = 30, height = 30, units = "cm", dpi = 400)

# Bifido plots

p2 <- gridExtra::grid.arrange(pheat.milkmet.faecbifido.2$gtable,
                        pheat.milkmet.faecbifido.175$gtable,
                        vp=viewport(width=0.95, height=1))


ggsave("Figures/RCCA_bifido.tiff",p2, width = 30, height = 30, units = "cm", dpi = 400) 

ggsave("Figures/RCCA_bifido.pdf",p2,width = 30, height = 30, units = "cm", dpi = 400)

# Combined

p3 <- gridExtra::grid.arrange(pheat.milkmet.milkbac.2$gtable,
                        pheat.milkmet.faecbac.225$gtable,
                        pheat.milkmet.faecbifido.175$gtable,
                        vp=viewport(width=0.95, height=1))


ggsave("Figures/RCCA_comb.tiff", p3, width = 30, height = 45, units = "cm", dpi = 400) 

ggsave("Figures/RCCA_comb.pdf", p3,width = 30, height = 45, units = "cm", dpi = 400)

```


# Multiblock spls

```{r}
library(mixOmics) # import the mixOmics library

set.seed(123) # for reproducibility, remove for normal use
```

```{r}
#Get shared id's
ps1 <- subset_samples(PSB.CSS, Sample.type=="Milk") 

ps2 <- subset_samples(PSB.CSS, Sample.type=="Faecal sample") 

ids.shared <- intersect(intersect(ps1@sam_data$RCID, ps2@sam_data$RCID), metab.id$RCID) %>% sort()


#Subset shared id and order by RCID 
ps1 <- subset_samples(ps1, RCID%in%ids.shared)
X1 <- data.frame(otu_table(ps1))[,order(ps1@sam_data$RCID)] %>%
  t()
rownames(X1) <- NULL

ps2 <- subset_samples(ps2, RCID%in%ids.shared)
X2 <- data.frame(otu_table(ps2))[,order(ps2@sam_data$RCID)] %>%
  t()
rownames(X2) <- NULL

 
X3 <- subset(metab.id, RCID %in% ids.shared) %>%
  arrange(RCID) %>%
  dplyr::select(-RCID) %>%
  dplyr::mutate(across(where(is.numeric), ~ifelse(. < 0, NA_real_, sqrt(.)))) %>% # Squere root normalization 
  scale(.,scale=FALSE,center=TRUE)

# compile these into a single X object
X <- list('Milk_bacteria' = X1, 'Faecal_bacteria' = X2, 'Milk_metabolites' = X3) 

lapply(X, dim) # check dimensions
```

##Initial Analysis

###Pairwise PLS Comparisons

Prior to tuning and fitting a MB-sPLS model, the datasets should be explored in a pairwise manner. While the focus is understanding how the two explantory datasets relate to the 'Milk bacteria' data, the relationship between the 'Faecal bacteria' and 'Milk metabolites' datasets is important to assess. These pairwise comparisons can be seen in Figure 1. Note that there were only 25 features from each dataframe selected and features with a correlation value lower than 0.5 were removed from the plot.

Observing Figure 1(a), it seems that the selected 'Milk bacteria' features are all positively associated with the first component. There are a few 'Faecal bacteria' features that are highly correlated with these, but most are negatively associated with the first component. The opposite of this is the case when looking at the relationships between the 'Milk bacteria' and 'Milk metabolites' features (Figure 1(b)). The methlyation and 'Milk metabolites' features are highly negatively correlated with on another as can be seen in Figure 1©.

```{r}
# select arbitrary values of features to keep
list.keepX = c(40, 40)
list.keepY = c(40, 40)

# generate three pairwise PLS models
pls1 <- spls(X[["Milk_bacteria"]], X[["Faecal_bacteria"]], 
             keepX = list.keepX, keepY = list.keepY)
pls2 <- spls(X[["Milk_bacteria"]], X[["Milk_metabolites"]], 
             keepX = list.keepX, keepY = list.keepY)
pls3 <- spls(X[["Faecal_bacteria"]], X[["Milk_metabolites"]], 
             keepX = list.keepX, keepY = list.keepY)

# plot features of first PLS
plotVar(pls1, cutoff = 0.3, title = "(a) 'Milk bacteria' vs 'Faecal bacteria'",
        legend = c("'Milk bacteria'", "'Faecal bacteria'"),
        var.names = FALSE, style = 'graphics',
        pch = c(16, 17), cex = c(2,2),
        col = c('darkorchid', 'lightgreen'))

# plot features of second PLS
plotVar(pls2, cutoff = 0.3, title = "(b) 'Milk bacteria' vs 'Milk metabolites'",
        legend = c("'Milk bacteria'", "'Milk metabolites'"),
        var.names = FALSE, style = 'graphics',
        pch = c(16, 17), cex = c(2,2),
        col = c('darkorchid', 'lightgreen'))

# plot features of third PLS
plotVar(pls3, cutoff = 0.3, title = "(c) 'Faecal bacteria' vs 'Milk metabolites'",
        legend = c("'Faecal bacteria'", "'Milk metabolites'"),
        var.names = FALSE, style = 'graphics',
        pch = c(16, 17), cex = c(2,2),
        col = c('darkorchid', 'lightgreen'))
```

Below the correlation between the X and Y components for each of the above sPLS models is shown.The RNA and methylation components are the most strongly correlated. The first components of the other two sPLS models showed correlation scores around 0.6.

```{r}
# calculate correlation of RNA and methylation
cor(pls1$variates$X, pls1$variates$Y)
```

```{r}
# calculate correlation of RNA and accessibility
cor(pls2$variates$X, pls2$variates$Y)
```

```{r}
# calculate correlation of methylation and accessibility
cor(pls3$variates$X, pls3$variates$Y) 
```

## Initial Multiblock sPLS Model

### Design

As discussed in the N-Integration Methods page, the “design” of the model needs to be created with the type of relationships to be explored in mind. Based on the correlation values seen above, a full design of value 0.5 will be used. In reality, there would be more exploration and tuning involved in forming this design matrix.

With a design in place, the initial MB-sPLS model can be generated. An arbitrarily high number of components (ncomp = 5) will be used.

```{r}
# for square matrix filled with 0.5s
design = matrix(0.3, ncol = length(X), nrow = length(X), 
                dimnames = list(names(X), names(X)))

diag(design) = 0 # set diagonal to 0s

basic.mbspls.model = block.spls(X, indY = 1, # generate basic model
                                ncomp = 5, 
                                design = design)
```

### Tuning

Unfortunately, there are not yet methods within mixOmics which allow for the tuning of block.spls objects. Hence, the number of of components (and features per component) to use cannot be determined experimentally as in all the other case studies. If tuning is necessary, manually undergoing repeated cross-validation, assessing the regression's Mean Squared Error (MSE) and R2

scores as components are added, is the way to achieve optimal parameter values.

For simplicity within this vignette, arbitrary values will be selected and can be seen below. 

```{r}
choice.ncomp <- 3 
choice.keepX <- list(Milk_bacteria = rep(100, 3), # 50 features per each component 
                     Faecal_bacteria = rep(100, 3), 
                     Milk_metabolites = rep(50, 3))
```

### Final model

Using these selected parameter values, the final MB-sPLS model can be constructed.

```{r}
# generate final model using "tuned" parameters
final.mbspls.model = block.spls(X, indY = 1,  
                                ncomp = choice.ncomp, 
                                keepX = choice.keepX,
                                design = design)
```
```{r}
plotIndiv(final.mbspls.model, ind.names = FALSE,
          group = as.factor(ps1@sam_data$BMI.group), 
          pch = as.factor(ps1@sam_data$Days),
          col.per.group = col_fil[1:3], 
          legend = TRUE, legend.title = 'Days', legend.title.pch = 'BMI group')

```


```{r}
set.seed(422)

#Sample random subset of samples
samples <- sample(1:dim(X1)[1], 30)

X1 <- X1[samples, ] 
X2 <- X2[samples, ] 
X3 <- X3[samples, ] 
X.arrow <- list(Milk_bacteria = X1, Faecal_bacteria = X2, Milk_metabolites = X3)

final.mbspls.model.arrow = block.spls(X.arrow, indY = 1, 
                                ncomp = choice.ncomp, 
                                keepX = choice.keepX,
                                design = design)

symbols <- list(Milk_bacteria = 1, Faecal_bacteria = 6, Milk_metabolites = 7)

plotArrow(final.mbspls.model, ind.names = FALSE,
          group = as.factor(ps1@sam_data$ID),
          pch = symbols, pch.size = 3)

plotArrow(final.mbspls.model.arrow, ind.names = FALSE,
          group = as.factor(ps1@sam_data$Days[samples]),
          pch = symbols, pch.size = 3)
```

```{r}
plotVar(final.mbspls.model, var.names = FALSE,
        legend = TRUE, cutoff = 0.5,
        pch = c(0,1,2))
```

```{r}
X.2 = list(Milk_bacteria = X1, 
         Milk_metabolites = X3)

list.keepX = list(Milk_bacteria = rep(25, 2), Milk_metabolites = rep(25,2))
list.keepY = c(rep(25, 2))

final.mbspls.model.circos = block.spls(X = X.2, Y = X2,
                                  ncomp = 2, keepX = list.keepX,
                                  keepY = list.keepY, design = 'full')


## ---- out.width = "90%", fig.cap = "FIGURE 5: Circos plot from multiblock sPLS performed on the gastrulation data The plot represents the correlations greater than 0.8 between variables of different types, represented on the side quadrants", results = FALSE----
tiff("Figures/MBsPLS.tiff",width = 6, height = 12, units = "in", pointsize = 12,res = 400)
par(mfrow = c(2, 1))
circosPlot(final.mbspls.model.circos,group=as.factor(ps1@sam_data$Days)[1:30], cutoff = 0.9,
           Y.name = 'Faecal_bacteria')

circosPlot(final.mbspls.model.circos,group=as.factor(ps1@sam_data$Days)[1:30], cutoff = 0.5,
           Y.name = 'Faecal_bacteria')
dev.off()

pdf("Figures/MBsPLS.pdf",width = 6, height = 12)
par(mfrow = c(2, 1))
circosPlot(final.mbspls.model.circos,group=as.factor(ps1@sam_data$Days)[1:30], cutoff = 0.9,
           Y.name = 'Faecal_bacteria')

circosPlot(final.mbspls.model.circos,group=as.factor(ps1@sam_data$Days)[1:30], cutoff = 0.5,
           Y.name = 'Faecal_bacteria')
dev.off()

```



# Session info


```{r}
sessionInfo()
```
